### 2020-01-23 paper/paper.tex
Working on paper

%%
%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmsmall]{acmart}
\usepackage{booktabs}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}
\acmDOI{1}

%%
%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmMonth{1}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Title Goes Here}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{David Young}
\email{d063y800@ku.edu}

% \authornote{Both authors contributed equally to this research.}
% \email{trovato@corporation.com}
% \orcid{1234-5678-9012}

\author{Andrew Gill}
\email{andygill@ku.edu}

% \authornotemark[1]
% \email{webmaster@marysville-ohio.com}
% \affiliation{%
%   \institution{Institute for Clarity in Documentation}
%   \streetaddress{P.O. Box 1212}
%   \city{Dublin}
%   \state{Ohio}
%   \postcode{43017-6221}
% }

% \author{Lars Th{\o}rv{\"a}ld}
% \affiliation{%
%   \institution{The Th{\o}rv{\"a}ld Group}
%   \streetaddress{1 Th{\o}rv{\"a}ld Circle}
%   \city{Hekla}
%   \country{Iceland}}
% \email{larst@affiliation.org}

% \author{Valerie B\'eranger}
% \affiliation{%
%   \institution{Inria Paris-Rocquencourt}
%   \city{Rocquencourt}
%   \country{France}
% }

% \author{Aparna Patel}
% \affiliation{%
%  \institution{Rajiv Gandhi University}
%  \streetaddress{Rono-Hills}
%  \city{Doimukh}
%  \state{Arunachal Pradesh}
%  \country{India}}

% \author{Huifen Chan}
% \affiliation{%
%   \institution{Tsinghua University}
%   \streetaddress{30 Shuangqing Rd}
%   \city{Haidian Qu}
%   \state{Beijing Shi}
%   \country{China}}

% \author{Charles Palmer}
% \affiliation{%
%   \institution{Palmer Research Laboratories}
%   \streetaddress{8600 Datapoint Drive}
%   \city{San Antonio}
%   \state{Texas}
%   \postcode{78229}}
% \email{cpalmer@prl.com}

% \author{John Smith}
% \affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
% \email{jsmith@affiliation.org}

% \author{Julius P. Kumquat}
% \affiliation{\institution{The Kumquat Consortium}}
% \email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Domain specific languages (DSLs) provide a powerful tool for to abstract over
  a class of problems. In this paper, we introduce a technique for representing
  algebraic data types and pattern matching in a DSL embedded in Haskell. This
  representation is automated through the use of GHC Generics and Template Haskell.
\end{abstract}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011050.10011017</concept_id>
       <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Domain specific languages}

% \keywords{domain specific languages}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
% \begin{CCSXML}
% <ccs2012>
%  <concept>
%   <concept_id>10010520.10010553.10010562</concept_id>
%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
%   <concept_significance>500</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10010520.10010575.10010755</concept_id>
%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
%   <concept_significance>300</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10010520.10010553.10010554</concept_id>
%   <concept_desc>Computer systems organization~Robotics</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10003033.10003083.10003095</concept_id>
%   <concept_desc>Networks~Network reliability</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
% </ccs2012>
% \end{CCSXML}

% \ccsdesc[500]{Computer systems organization~Embedded systems}
% \ccsdesc[300]{Computer systems organization~Redundancy}
% \ccsdesc{Computer systems organization~Robotics}
% \ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
% \keywords{datasets, neural networks, gaze detection, text tagging}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

Embedded domain specific languages (EDSLs) have long been a useful and effective
technique for constructing reusable tools for writing programs for a variety of
different problem domains. Haskell is a language which is particularly
well-suited to EDSLs due to its lazy evaluation, first-class functions and
lexical closures.

[\ldots]

\section{Representing Algebraic Datatypes}

\subsection{E, Rep, RepTy}

[\ldots]

\section{Representing pattern matches}

\subsection{SumMatch}

\subsection{ProdMatch}

\section{Citations and Bibliographies}

% The use of \BibTeX\ for the preparation and formatting of one's
% references is strongly recommended. Authors' names should be complete
% --- use full first names (``Donald E. Knuth'') not initials
% (``D. E. Knuth'') --- and the salient identifying features of a
% reference should be included: title, year, volume, number, pages,
% article DOI, etc.

% The bibliography is included in your source document with these two
% commands, placed just before the \verb|\end{document}| command:
% \begin{verbatim}
%   \bibliographystyle{ACM-Reference-Format}
%   \bibliography{bibfile}
% \end{verbatim}
% where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
% suffix, of the \BibTeX\ file.

% Citations and references are numbered by default. A small number of
% ACM publications have citations and references formatted in the
% ``author year'' style; for these exceptions, please include this
% command in the {\bfseries preamble} (before
% ``\verb|\begin{document}|'') of your \LaTeX\ source:
% \begin{verbatim}
%   \citestyle{acmauthoryear}
% \end{verbatim}

%   Some examples.  A paginated journal article \cite{Abril07}, an
%   enumerated journal article \cite{Cohen07}, a reference to an entire
%   issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
%   monograph/whole book in a series (see 2a in spec. document)
%   \cite{Harel79}, a divisible-book such as an anthology or compilation
%   \cite{Editor00} followed by the same example, however we only output
%   the series if the volume number is given \cite{Editor00a} (so
%   Editor00a's series should NOT be present since it has no vol. no.),
%   a chapter in a divisible book \cite{Spector90}, a chapter in a
%   divisible book in a series \cite{Douglass98}, a multi-volume work as
%   book \cite{Knuth97}, an article in a proceedings (of a conference,
%   symposium, workshop for example) (paginated proceedings article)
%   \cite{Andler79}, a proceedings article with all possible elements
%   \cite{Smith10}, an example of an enumerated proceedings article
%   \cite{VanGundy07}, an informally published work \cite{Harel78}, a
%   doctoral dissertation \cite{Clarkson85}, a master's thesis:
%   \cite{anisi03}, an online document / world wide web resource
%   \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1)
%   \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05} and
%   (Case 3) a patent \cite{JoeScientist001}, work accepted for
%   publication \cite{rous08}, 'YYYYb'-test for prolific author
%   \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
%   contain 'duplicate' DOI and URLs (some SIAM articles)
%   \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
%   multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
%   couple of citations with DOIs:
%   \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
%   citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
%   \cite{R} and \cite{UMassCitations}.

% \section{Acknowledgments}

%Identification of funding sources and other support, and thanks to
%individuals and groups that assisted in the research and the
%preparation of the work should be included in an acknowledgment
%section, which is placed just before the reference section in your
%document.

%This section has a special environment:
%\begin{verbatim}
%  \begin{acks}
%  ...
%  \end{acks}
%\end{verbatim}
%so that the information contained therein can be more easily collected
%during the article metadata extraction phase, and to ensure
%consistency in the spelling of the section heading.

%Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

%\section{Appendices}

%If your work needs an appendix, add it before the
%``\verb|\end{document}|'' command at the conclusion of your source
%document.

%Start the appendix with the ``\verb|appendix|'' command:
%\begin{verbatim}
%  \appendix
%\end{verbatim}
%and note that in the appendix, sections are lettered, not
%numbered. This document has two appendices, demonstrating the section
%and subsection identification method.

%\section{SIGCHI Extended Abstracts}

%The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
%not in Word) produces a landscape-orientation formatted article, with
%a wide left margin. Three environments are available for use with the
%``\verb|sigchi-a|'' template style, and produce formatted output in
%the margin:
%\begin{itemize}
%\item {\verb|sidebar|}:  Place formatted text in the margin.
%\item {\verb|marginfigure|}: Place a figure in the margin.
%\item {\verb|margintable|}: Place a table in the margin.
%\end{itemize}

%%%
%%% The acknowledgments section is defined using the "acks" environment
%%% (and NOT an unnumbered section). This ensures the proper
%%% identification of the section in the article metadata, and the
%%% consistent spelling of the heading.
%\begin{acks}
%To Robert, for the bagels and explaining CMYK and color spaces.
%\end{acks}

%%%
%%% The next two lines define the bibliography style to be used, and
%%% the bibliography file.
%\bibliographystyle{ACM-Reference-Format}
%\bibliography{sample-base}

%%%
%%% If your work has an appendix, this is the place to put it.
%\appendix

%\section{Research Methods}

%\subsection{Part One}

%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
%malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
%sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
%vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
%lacinia dolor. Integer ultricies commodo sem nec semper.

%\subsection{Part Two}

%Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
%ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
%ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
%eros. Vivamus non purus placerat, scelerisque diam eu, cursus
%ante. Etiam aliquam tortor auctor efficitur mattis.

%\section{Online Resources}

%Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
%pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
%enim maximus. Vestibulum gravida massa ut felis suscipit
%congue. Quisque mattis elit a risus ultrices commodo venenatis eget
%dui. Etiam sagittis eleifend elementum.

%Nam interdum magna at lectus dignissim, ac dignissim lorem
%rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
%massa et mattis lacinia.

\end{document}
\endinput

### 2020-01-29 paper/paper.tex
Example for paper

@@
 \usepackage{booktabs}
+\usepackage{listings}
+
+\lstset{language=haskell, deletekeywords={abs}}
 
@@
 
+\subsection{Translation Example} % TODO: Does it make sense for this to be a
+                                 % subsection of the intro?
+
+In the example below, \texttt{example} would get transformed into \texttt{example'}:
+
+\begin{lstlisting}
+import Data.Char (ord)
+
+x :: Either Char Int
+x = Left 'a'
+
+example :: Int
+example =
+  case x of
+    Left  c -> ord c
+    Right i -> i
+
+example' :: Int
+example' =
+  abs (CaseExp (rep x)
+               (SumMatch (OneProdMatch (\c -> ord c))
+                         (OneSumMatch
+                           (OneProdMatch (\i -> i)))))
+\end{lstlisting}
+
 [\ldots]
@@
 \section{Representing Algebraic Datatypes}
+
+[Remove metadata from Generic representation and turn tree-like structure into a list-like structure]
 

### 2020-01-29 src/Case.hs
Attempt to use dependent types to index by constructor name

@@ -11,7 +11,6 @@
 {-# LANGUAGE InstanceSigs #-}
 {-# LANGUAGE ScopedTypeVariables #-}
 {-# LANGUAGE TypeApplications #-}
--- {-# LANGUAGE AllowAmbiguousTypes #-}
 {-# LANGUAGE FlexibleInstances #-}
 {-# LANGUAGE TemplateHaskell #-}
 {-# LANGUAGE DeriveDataTypeable #-}
@@ -24,6 +23,11 @@
 {-# LANGUAGE DefaultSignatures #-}
 {-# LANGUAGE ConstraintKinds #-}
 {-# LANGUAGE DataKinds #-}
+
+{-# LANGUAGE AllowAmbiguousTypes #-}
+
+
+{-# OPTIONS_GHC -ddump-splices #-}
 
 module Case
   where
@@ -60,11 +64,73 @@ import           Data.Bifunctor
 
 import           GHC.TypeLits hiding (Nat)
 
+import           Data.Singletons
+import           Data.Singletons.TH
+
+import           Language.Haskell.TH.Syntax (OccName, NameFlavour)
+
+import           GHC.TypeLits
+import           Data.Type.Bool
+
+import           DSMap
+import           Type.Set
+
+import Data.Char
+
+-- $(genSingletons [''Char])
+
+-- $(genSingletons [''OccName])
+
+-- $(genSingletons [''Name])
+
+nameToSymbol :: Name -> SomeSymbol
+nameToSymbol = someSymbolVal . show
+
 infixl 0 :@
 pattern f :@ x = AppE f x
 
 type Case = (Exp, [Match])
 
+data Nu f where 
+  Nu :: (a -> f a) -> a -> Nu f
+
+data ListF a r where
+  NilF :: ListF a r
+  ConsF :: a -> r -> ListF a r
+  deriving (Functor)
+
+newtype Fix f = Fix {unFix :: f (Fix f)}
+
+type List a = Nu (ListF a)
+
+conv :: Functor f => Nu f -> Fix f
+conv (Nu f x) =
+  Fix (fmap (conv . (\y -> Nu (const y) x) . f) (f x))
+
+fix2list :: Fix (ListF Int) -> [Int]
+fix2list (Fix NilF) = []
+fix2list (Fix (ConsF x xs)) = x : fix2list xs
+
+  -- conv (Nu _ (f x))
+
+  -- Fix $ _ $ f x
+
+-- listConv :: List a -> Fix (ListF a)
+-- listConv (Nu f x) =
[…]

### 2020-01-29 src/DSMap.hs

@@ -1,0 +1,142 @@
+{-# LANGUAGE DataKinds #-}
+{-# LANGUAGE GADTs #-}
+{-# LANGUAGE TypeOperators #-}
+{-# LANGUAGE PolyKinds #-}
+{-# LANGUAGE TypeFamilies #-}
+{-# LANGUAGE KindSignatures #-}
+{-# LANGUAGE UndecidableInstances #-}
+{-# LANGUAGE ScopedTypeVariables #-}
+{-# LANGUAGE PatternSynonyms #-}
+{-# LANGUAGE DeriveGeneric #-}
+
+{-# OPTIONS_GHC -fplugin=Type.Compare.Plugin #-}
+
+module DSMap where
+
+import           Data.Dependent.Map
+
+import           Data.Singletons
+import           Data.Singletons.TypeLits
+
+import           Data.Constraint
+import           Type.Set
+
+import           Data.Functor.Const
+
+import           Data.Typeable
+
+import           GHC.Generics
+import Data.Void
+
+data Test = D | B | C | A deriving (Generic)
+
+-- Constructor names and with their "types"
+type family ConstrDict t :: [(Symbol, *)] where
+  ConstrDict (M1 C ('MetaCons name y z) r p) = '[ '(name, r p) ]
+  ConstrDict (M1 x m r p) = ConstrDict (r p)
+
+
+  -- ConstrDict ((x :+: y) p) = 
+
+
+data SetSing f s where
+  SetSingInsert :: Typeable x => Sing x -> f x -> SetSing f s -> SetSing f (Insert x s)
+  SetSingEmpty :: SetSing f 'Empty
+
+pattern SetSingOne s fx = SetSingInsert s fx SetSingEmpty
+
+setSingLookup :: forall x s f. (Typeable x) => SetSing f s -> Sing x -> Maybe (f x)
+setSingLookup SetSingEmpty _ = Nothing
+setSingLookup (SetSingInsert (s :: Sing t) fx rest) s' =
+  case eqT :: Maybe (x :~: t) of
+    Nothing -> setSingLookup rest s'
+    Just Refl -> Just fx
+
+
+type family SetSingMerge a b where
+  SetSingMerge (SetSing f a) (SetSing f b) = SetSing f (Merge a b)
+
+type family IsSetSing a :: Constraint where
+  IsSetSing (SetSing f a) = ()
+
[…]

### 2020-01-29 src/Effect.hs

@@ -1,0 +1,58 @@
+-- Based on "Monad Transformers and Modular Algebraic Effects" [2019]
+
+{-# LANGUAGE DeriveFunctor #-}
+{-# LANGUAGE TypeOperators #-}
+
+module Effect where
+
+import           Control.Monad
+
+data Free sig x = Var x | Op (sig (Free sig x))
+
+instance Functor sig => Functor (Free sig) where
+  fmap f (Var x) = Var (f x)
+  fmap f (Op s)  = Op (fmap (fmap f) s)
+
+instance Functor sig => Applicative (Free sig) where
+  pure = return
+  (<*>) = ap
+
+instance Functor sig => Monad (Free sig) where
+  return = Var
+  Var x >>= f = f x
+  Op k >>= f = Op (fmap (>>= f) k)
+
+data State s k
+  = Get (s -> k)
+  | Put s (() -> k)
+  deriving (Functor)
+
+get' :: Free (State s) s
+get' = Op (Get return)
+
+put' :: s -> Free (State s) ()
+put' s = Op (Put s return)
+
+incr' :: Free (State Int) ()
+incr' = get' >>= put' . succ
+
+
+fold :: Functor sig => (a -> b) -> (sig b -> b) -> (Free sig a -> b)
+fold gen alg (Var x) = gen x
+fold gen alg (Op op) = alg (fmap (fold gen alg) op)
+
+data (sig1 + sig2) a = Inl (sig1 a) | Inr (sig2 a)
+
+data Void1 k deriving (Functor)
+
+runVoid :: Free Void1 a -> a
+runVoid = fold id undefined
+
+
+(<+>) :: (sig1 b -> b) -> (sig2 b -> b) -> ((sig1 + sig2) b -> b)
+(alg1 <+> alg2) (Inl op) = alg1 op
+(alg1 <+> alg2) (Inr op) = alg2 op
+
+
+
+

### 2020-01-29 src/GPURep.hs

@@ -1,0 +1,4 @@
+module GPURep where
+
+
+

### 2020-01-29 src/Run.hs

@@ -1,0 +1,36 @@
+{-# LANGUAGE KindSignatures #-}
+{-# LANGUAGE TypeSynonymInstances #-}
+{-# LANGUAGE FlexibleInstances #-}
+{-# LANGUAGE PatternSynonyms #-}
+
+module Run where
+
+import           Control.Monad.Reader
+import           Control.Monad.Writer
+
+type Tid = (Int, Int, Int)
+
+runKernel :: KernelM m => (Tid -> m ()) -> m (IO ())
+runKernel = undefined
+
+class KernelM (m :: * -> *)
+
+instance KernelM (Reader r)
+instance Monoid w => KernelM (Writer w)
+
+pattern Phantom x = x
+
+{-# NOINLINE Phantom #-}
+
+-- test :: Maybe a -> Bool
+-- test Nothing = False
+-- test (Just _) = True
+
+test' :: Maybe a -> Bool
+test' (Phantom Nothing) = False
+test' (Phantom (Just _)) = True
+
+{-# RULES "elim-Phantom"
+      forall x. Phantom x = x
+  #-}
+

### 2020-01-29 src/Scott.hs

@@ -8,13 +8,27 @@
 {-# LANGUAGE MultiParamTypeClasses #-}
 {-# LANGUAGE FlexibleInstances #-}
 {-# LANGUAGE InstanceSigs #-}
-{-# LANGUAGE AllowAmbiguousTypes #-} -- XXX: Is this ok?
+{-# LANGUAGE AllowAmbiguousTypes #-} -- XXX: Is this ok? (used for 'scottElim')
+{-# LANGUAGE FunctionalDependencies #-}
 
 module Scott where
 
 import           GHC.Generics
+import           Data.Data
+
 import           Data.Void
 import           Data.Proxy
+
+
+data Example2 = B2 Bool | I2 Int deriving (Generic, Show)
+
+scott_Example2B2 :: ScottRep'' Bool -> ScottRep'' Example2
+scott_Example2B2 bool = ScottRep'' (Proxy, \f _g -> f bool)
+
+scott_Example2I2 :: Int -> ScottRep'' Example2
+scott_Example2I2 int = ScottRep'' (Proxy, \_f g -> g int)
+
+
 
 type family ScottRep rec rep r where
   ScottRep rec (M1 i c f p) r  = ScottRep rec (f p) r
@@ -29,10 +43,21 @@ type family ScottRep rec rep r where
   ScottRep rec ((f :*: g) x) r = (ScottRep rec (f x) r -> ScottRep rec (g x) r -> r) -> r
   ScottRep rec (U1 p) r        = U1 p
 
-type ScottRep' t r =
-  (Proxy r, (ScottRep (ScottRep'' t) (Rep t Void) r))
+-- In the ScottEncode instance for (:+:):
+-- Need: ScottRep (ScottRep'' a) (p Void) r
+-- Have: ScottRep'' pt
+-- Where: Rep pt Void ~ p Void
+-- The "pt" recs need to be replaced with "(ScottRep'' a)" recs, I think...
+type ScottRep' t rep r =
+  (Proxy r, (ScottRep (ScottRep'' t) rep r))
 
-data ScottRep'' t = ScottRep'' (forall r. ScottRep' t r)
+data ScottRep'' t = ScottRep'' (forall r. ScottRep' t (Rep t Void) r)
+
+-- transportScottRep ::
+--   (rep1 -> rep2) ->
+--   ScottRep t rep1 r -> ScottRep t rep2 r
+-- transportScottRep = undefined
+
 
 data Nat = Z | S Nat deriving (Generic, Show)
 
[…]

### 2020-01-30 paper/paper.tex
Paper progress

@@
 \lstset{language=haskell, deletekeywords={abs}}
+
+\newcommand{\ttt}{\texttt}
 
@@
 
-In the example below, \texttt{example} would get transformed into \texttt{example'}:
+In the example below, \ttt{example} would get transformed into \ttt{example'}:
 
@@
 
-[Remove metadata from Generic representation and turn tree-like structure into a list-like structure]
+\subsection{GHC Generics \ttt{Rep}}
 
-\subsection{E, Rep, RepTy}
+[Remove metadata from Generic representation and turn tree-like structure into a (nested) list-like structure]
 
-[\ldots]
+\subsection{\ttt{E}, \ttt{ERep}, \ttt{ERepTy}}
+
+There are three interconnected foundational parts: \ttt{E}, \ttt{ERep} and
+\ttt{ERepTy}. \ttt{E} is the deep embedding of the EDSL (a GADT that encodes
+expressions in the DSL language). \ttt{ERep} is a type class which represents
+all Haskell types which can be represented in the DSL. \ttt{ERepTy} is a type
+family associated to the \ttt{ERep} type class, which represents a "canonical form"
+of the given type. This canonical form can be immediately constructed in the EDSL.
+Canonical form types crucially include \ttt{Either} and \ttt{(,)}, which
+allow all combinations of basic sum types and product types to be encoded. More
+information on this encoding is given in [a different section].
+
+This information is brought into the \ttt{E} type via the constructor
+\ttt{Repped}. The \ttt{E} also contains constructors representing \ttt{Either}
+and \ttt{(,)} values:
+
+\begin{lstlisting}
+data E t where
+  ...
+  Repped :: ERep t => ERepTy t -> E t
+
+  LeftExp :: E a -> E (Either a b)
+  RightExp :: E b -> E (Either a b)
+
+  PairExp :: E a -> E b -> E (a, b)
+  ...
+\end{lstlisting}
+
+\ttt{ERep} and \ttt{ERepTy} provide an interface for transferring values between the EDSL
+world and the Haskell world:
+
+\begin{lstlisting}
+class ERep t where
+  type ERepTy t
+
+  rep :: t -> E t
+  rep = Repped . rep'
+
+  unrep :: ERepTy t -> t
+
+  rep' :: t -> ERepTy t
+\end{lstlisting}
+
+The key algebraic instances mentioned before are as follows:
+
+\begin{lstlisting}
+instance (ERep a, ERep b) => ERep (Either a b) where
+  type ERepTy (Either a b) = Either a b
+
+  rep (Left x) = LeftExp (rep x)
+  rep (Right y) = RightExp (rep y)
+
+  unrep = id
+  rep'  = id
+
+instance (ERep a, ERep b) => ERep (a, b) where
+  type ERepTy (a, b) = (a, b)
+
+  rep (x, y) = PairExp (rep x) (rep y)
+  unrep = id
+  rep'  = id
+\end{lstlisting}
 
@@
 
-\subsection{SumMatch}
+Within the \ttt{E} expression language, a pattern match is represented by the
+\ttt{CaseExp} constructor:
 
-\subsection{ProdMatch}
+\begin{lstlisting}
+data E t where
+  ...
+  CaseExp :: ERep t => E t -> SumMatch (ERepTy t) r -> E r
+  ...
+\end{lstlisting}
+
+A value of type \ttt{SumMatch a b} represents a computation within the EDSL
+which destructures a value of type \ttt{E a} and produces a value of type \ttt{E b}.
+Therefore, \ttt{SumMatch (ERepTy t) r} represents a computation which destructures
+a value of type \ttt{E (ERepTy t)} and produces a value of type \ttt{E b}.
+
+\subsection{\ttt{SumMatch}}
+
+The overall structure of \ttt{SumMatch} is a (heterogeneous) list of \ttt{ProdMatch} values.
+Each item of this list corresponds exactly to one branch in the original \ttt{case} match.
+
+\begin{lstlisting}
+data SumMatch s t where
+  SumMach ::
+    (ERep a, ERep b, ERepTy b ~ b)
+      => ProdMatch a r -> SumMatch b r -> SumMatch (Either a b) r
+
+  EmptyMatch :: SumMatch Void r
+
+  OneSumMatch ::
+    (ERep a, ERep b, ERepTy a ~ a)
+      => ProdMatch a b -> SumMatch a b
+\end{lstlisting}
+
+Note the \ttt{ERepTy a ~ a} constraints. This constraint ensures that the type \ttt{a}
+is already in canonical form (that is, consists entirely of \ttt{Either}, \ttt{(,)} and
+base types).
+
+\subsection{\ttt{ProdMatch}}
+
+\ttt{ProdMatch s t} is equivalent to a function a curried function from \ttt{s} to \ttt{t}.
+Note that \ttt{s} will be a (potentially nested) pair type.
 

### 2020-01-30 src/Case.hs
It makes more sense to use Void for an empty match

@@ -91,7 +91,7 @@ data SumMatch s t where
     (GPURep a, GPURep b, GPURepTy b ~ b)
         => ProdMatch a r -> SumMatch b r -> SumMatch (Either a b) r
 
-  EmptyMatch :: SumMatch () r
+  EmptyMatch :: SumMatch Void r
 
   OneSumMatch :: (GPURep a, GPURep b, GPURepTy a ~ a) => ProdMatch a b -> SumMatch a b
 

### 2020-01-30 paper/paper.tex
More on ProdMatch
Started to describe 'construct'

@@
 \ttt{ProdMatch s t} is equivalent to a function a curried function from \ttt{s} to \ttt{t}.
-Note that \ttt{s} will be a (potentially nested) pair type.
+Note that \ttt{s} will be a (potentially nested) pair type. For example,
+\ttt{ProdMatch (a, (b, c)) r} will be equivalent to \ttt{a -> b -> c -> r}.
+
+\begin{lstlisting}
+data ProdMatch s t where
+  ProdMatch ::
+    (ERep a, ERep b)
+      => (E a -> ProdMatch b r) -> ProdMatch (a, b) r
+
+  OneProdMatch :: ERep a => (E a -> E r) -> ProdMatch a r
+
+  NullaryMatch :: ERep r => E r -> ProdMatch a r
+\end{lstlisting}
+
+\section{Constructing values}
+
+\begin{lstlisting}
+construct ::
+  Construct (a -> b)
+    => (a -> b) -> LiftedFn (a -> b)
+construct = construct' . Construct
+
+-- Example: LiftedFn (a -> b -> c) is: E a -> E b -> E c
+type family LiftedFn t where
+  LiftedFn (a -> b) = E a -> LiftedFn b
+  LiftedFn b = E b
+
+class Construct t where
+  construct' :: E t -> LiftedFn t
+\end{lstlisting}
+
+Example GHCi session, demonstrating \ttt{construct}:
+
+\begin{lstlisting}
+ghci> :t construct (,)
+construct (,)
+  :: (ERep a, ERep b) => E a -> E b -> E (a, b)
+
+ghci> :t construct (,,)
+construct (,,)
+  :: (ERep a, ERep b, ERep c) =>
+     E a -> E b -> E c -> E (a, b, c)
+
+ghci> :t construct (,,) (rep False) (rep 3.1 :: E Double) (rep 1 :: E Int)
+construct (,,) (rep False) (rep 3.1 :: E Double) (rep 1 :: E Int)
+  :: E (Bool, Double, Int)
+\end{lstlisting}
 

### 2020-01-30 paper/paper.tex
This has better formatting

@@
 
-ghci> :t construct (,,) (rep False) (rep 3.1 :: E Double) (rep 1 :: E Int)
-construct (,,) (rep False) (rep 3.1 :: E Double) (rep 1 :: E Int)
-  :: E (Bool, Double, Int)
+ghci> :t construct (,,) (rep False) (rep 3.1 :: E Float) (rep 1 :: E Int)
+construct (,,) (rep False) (rep 3.1 :: E Float) (rep 1 :: E Int)
+  :: E (Bool, Float, Int)
 \end{lstlisting}

### 2020-01-30 paper/paper.tex
Started section on recovering standard match semantics
Basic 'abs' for pattern matching
More of 'abs' definition
More on 'construct' and 'Construct'

@@
 
+\section{Recovering standard semantics for pattern matches}
+
+\begin{lstlisting}
+sumMatchAbs :: ERep s => SumMatch (ERepTy s) t -> s -> t
+...
+
+prodMatchAbs :: ERep s => ProdMatch s t -> s -> t
+...
+
+abs :: E t -> t
+abs (CaseExp x f) = sumMatchAbs f (abs x)
+
+abs (LeftExp x) = Left (abs x)
+abs (RightExp y) = Right (abs y)
+abs (PairExp x y) = (abs x, abs y)
+
+abs (Repped x) = unrep x
+...
+\end{lstlisting}
+
 \section{Constructing values}
+
+\ttt{construct} is designed to bring Haskell constructors into the EDSL world.
 
@@
   construct' :: E t -> LiftedFn t
+\end{lstlisting}
+
+The "helper constructor" \ttt{Construct} is part of the \ttt{E} type:
+
+\begin{lstlisting}
+data E t where
+  ...
+  Construct :: (a -> b) -> E (a -> b)
+  ...
 \end{lstlisting}

### 2020-02-14 src/Plugin.hs
Starting basics of transformation for Core plugin

@@ -1,8 +1,20 @@
+{-# LANGUAGE TemplateHaskellQuotes #-}
+
 module Plugin (plugin) where
 
-import           Expr
+import           Expr hiding (Var, Lit)
+import qualified Expr
+
+import           Data.Data
+import           Data.Generics.Uniplate.Operations
+import qualified Data.Generics.Uniplate.DataOnly as Data
 
 import           GhcPlugins
+
+import           Control.Monad
+
+-- Just for TH Names:
+import qualified Language.Haskell.TH.Syntax as TH
 
 plugin :: Plugin
 plugin =
@@ -15,5 +27,53 @@ install _ coreToDos = do
   return (CoreDoPluginPass "Pattern matching plugin" pass : coreToDos)
 
 pass :: ModGuts -> CoreM ModGuts
-pass m = undefined
+pass guts = do
+  primMap <- forM primMapTH
+                 (\(fromTH, toTH) -> do
+                  fromName <- thNameToGhcName_ fromTH
+                  toName <- thNameToGhcName_ toTH
+
+                  from <- lookupId fromName
+                  to <- lookupId toName
+                  return (from, to))
+
+  bindsOnlyPass (mapM (transformBind primMap)) guts
+
+transformBind :: [(Id, Id)] -> CoreBind -> CoreM CoreBind
+transformBind primMap (NonRec name e) =
+  fmap (NonRec name) (transformExpr primMap e)
+
+transformExpr :: [(Id, Id)] -> Expr Var -> CoreM (Expr Var)
+transformExpr = undefined
+
+-- XXX: The delineation marker probably has to be floated in (or maybe the
+-- transformation can just go through the AST at that marker without
+-- floating in the marker itself)
+
+-- | Transform primitives and constructor/function calls. Skips the
+-- function call transformation on the given 'recName' argument (if
+-- a 'Just').
+transformPrims :: Maybe Var -> [(Id, Id)] -> Expr Var -> Expr Var
+transformPrims recName primMap = Data.transform go
+  where
+    builtin :: Id -> Maybe (Expr Var)
+    builtin v = Var <$> lookup v primMap
+
+    go expr@(Lit x) = undefined
+
+thNameToGhcName_ :: TH.Name -> CoreM Name
+thNameToGhcName_ thName = do
+  nameMaybe <- thNameToGhcName thName
+  case nameMaybe of
[…]

### 2020-02-14 src/Plugin.hs
Helper functions for DSL name lookup

@@ -1,8 +1,10 @@
 {-# LANGUAGE TemplateHaskellQuotes #-}
+{-# LANGUAGE PatternSynonyms #-}
+{-# LANGUAGE TupleSections #-}
 
 module Plugin (plugin) where
 
-import           Expr hiding (Var, Lit)
+import           Expr hiding (Var, Lit, (:@))
 import qualified Expr
 
 import           Data.Data
@@ -13,8 +15,13 @@ import           GhcPlugins
 
 import           Control.Monad
 
+import           Control.Arrow ((&&&), (***))
+
 -- Just for TH Names:
 import qualified Language.Haskell.TH.Syntax as TH
+
+infixl 0 :@
+pattern f :@ x = App f x
 
 plugin :: Plugin
 plugin =
@@ -37,13 +44,20 @@ pass guts = do
                   to <- lookupId toName
                   return (from, to))
 
-  bindsOnlyPass (mapM (transformBind primMap)) guts
+  gpuExprConstrNames <- mapM (\n -> fmap (,n) (thNameToGhcName_ n)) gpuExprConstrNamesTH
 
-transformBind :: [(Id, Id)] -> CoreBind -> CoreM CoreBind
-transformBind primMap (NonRec name e) =
-  fmap (NonRec name) (transformExpr primMap e)
+  gpuExprIdMap <- forM gpuExprConstrNames
+                       (\(name, nameTH) -> do
+                          theId <- lookupId name
+                          return (nameTH, theId))
 
-transformExpr :: [(Id, Id)] -> Expr Var -> CoreM (Expr Var)
+  bindsOnlyPass (mapM (transformBind primMap (Var . getGPUExprIdsFrom gpuExprIdMap))) guts
+
+transformBind :: [(Id, Id)] -> (TH.Name -> Expr Var) -> CoreBind -> CoreM CoreBind
+transformBind primMap dslVar (NonRec name e) =
+  fmap (NonRec name) (transformExpr primMap dslVar e)
+
+transformExpr :: [(Id, Id)] -> (TH.Name -> Expr Var) -> Expr Var -> CoreM (Expr Var)
 transformExpr = undefined
 
 -- XXX: The delineation marker probably has to be floated in (or maybe the
[…]

### 2020-02-14 src/Plugin.hs
Added error message back in

@@ -94,7 +94,7 @@ primMapTH =
 getGPUExprIdsFrom :: [(TH.Name, Id)] -> TH.Name -> Id
 getGPUExprIdsFrom idMap name =
   case lookup name idMap of
-    -- Nothing -> error $ "Cannot find Id for: " ++ show name
+    Nothing -> error $ "Cannot find Id for: " ++ show name
     Just i  -> i
 
 gpuExprConstrNamesTH :: [TH.Name]

### 2020-02-14 src/Plugin.hs
Put in some dictionary building stuff from HERMIT

@@ -1,6 +1,7 @@
 {-# LANGUAGE TemplateHaskellQuotes #-}
 {-# LANGUAGE PatternSynonyms #-}
 {-# LANGUAGE TupleSections #-}
+{-# LANGUAGE CPP #-}
 
 module Plugin (plugin) where
 
@@ -13,9 +14,40 @@ import qualified Data.Generics.Uniplate.DataOnly as Data
 
 import           GhcPlugins
 
+import           InstEnv
+
+import           TcRnTypes
+import           TcSimplify
+import           TcMType
+import           TcRnMonad
+import           TcSMonad
+import           TcEvidence
+import           TcType
+
+import           DsBinds
+import           DsMonad
+
+import           HsBinds
+import           HsDecls
+
+import           CostCentreState
+
+import           Bag
+
+import           Encoding
+
+import           ErrUtils
+
+import           Data.Char
+
 import           Control.Monad
 
-import           Control.Arrow ((&&&), (***))
+import           Control.Arrow ((&&&), (***), first, second)
+
+import           Data.IORef
+
+import qualified Data.Map as Map
+import qualified Data.Set as Set
 
 -- Just for TH Names:
 import qualified Language.Haskell.TH.Syntax as TH
@@ -44,21 +76,21 @@ pass guts = do
                   to <- lookupId toName
                   return (from, to))
 
-  gpuExprConstrNames <- mapM (\n -> fmap (,n) (thNameToGhcName_ n)) gpuExprConstrNamesTH
+  constrNames <- mapM (\n -> fmap (,n) (thNameToGhcName_ n)) constrNamesTH
 
-  gpuExprIdMap <- forM gpuExprConstrNames
+  gpuExprIdMap <- forM constrNames
                        (\(name, nameTH) -> do
                           theId <- lookupId name
                           return (nameTH, theId))
 
-  bindsOnlyPass (mapM (transformBind primMap (Var . getGPUExprIdsFrom gpuExprIdMap))) guts
+  bindsOnlyPass (mapM (transformBind guts (mg_inst_env guts) primMap (getGPUExprIdsFrom gpuExprIdMap))) guts
 
-transformBind :: [(Id, Id)] -> (TH.Name -> Expr Var) -> CoreBind -> CoreM CoreBind
-transformBind primMap dslVar (NonRec name e) =
[…]

### 2020-02-15 src/Plugin.hs
Work on pattern match transformation in plugin

@@ -5,10 +5,10 @@
 
 module Plugin (plugin) where
 
-import           Expr hiding (Var, Lit, (:@))
+import           Expr hiding (Var, Lit, Lam, (:@))
 import qualified Expr
 
-import           Data.Data
+import           Data.Data hiding (TyCon (..), tyConName)
 import           Data.Generics.Uniplate.Operations
 import qualified Data.Generics.Uniplate.DataOnly as Data
 
@@ -16,38 +16,16 @@ import           GhcPlugins
 
 import           InstEnv
 
-import           TcRnTypes
-import           TcSimplify
-import           TcMType
-import           TcRnMonad
+import           FamInstEnv
+
 import           TcSMonad
-import           TcEvidence
-import           TcType
 
-import           DsBinds
-import           DsMonad
-
-import           HsBinds
-import           HsDecls
-
-import           CostCentreState
-
-import           Bag
-
-import           Encoding
-
-import           ErrUtils
-
-import           Data.Char
 
 import           Control.Monad
 
 import           Control.Arrow ((&&&), (***), first, second)
 
-import           Data.IORef
-
-import qualified Data.Map as Map
-import qualified Data.Set as Set
+import           GHCUtils
 
 -- Just for TH Names:
 import qualified Language.Haskell.TH.Syntax as TH
@@ -99,28 +77,36 @@ transformExpr guts primMap lookupVar = undefined
 -- | Transform primitives and constructor/function calls. Skips the
 -- function call transformation on the given 'recName' argument (if
 -- a 'Just').
-transformPrims :: ModGuts -> Maybe Var -> [(Id, Id)] -> (TH.Name -> Var) -> Expr Var -> Expr Var
-transformPrims guts recName primMap lookupVar = Data.transform go
+transformPrims :: ModGuts -> Maybe Var -> [(Id, Id)] -> (TH.Name -> Var) -> Expr Var -> CoreM (Expr Var)
+transformPrims guts recName primMap lookupVar = Data.transformM go
   where
     builtin :: Id -> Maybe (Expr Var)
     builtin v = Var <$> lookup v primMap
 
-    go expr@(Lit x) = Var (lookupVar 'Lit) :@ expr
[…]

### 2020-02-15 src/Deep/Case.hs
Restructure

@@ -69,14 +69,13 @@ import           Data.Complex
 
 import           Data.Constraint (Constraint)
 
-import           Expr
+import           Deep.Expr
 
 infixl 0 :@
 pattern f :@ x = AppE f x
 
 type Case = (Exp, [Match])
 
-data Tagged2 (f :: Maybe k) t x = Tagged2 x
 
 
 
@@ -223,7 +222,7 @@ transformCase0 (CaseE scrutinee matches0@(firstMatch:_)) = do
         Just (tyCons, _) = infoMaybe
 
     return (ConE 'CaseExp :@ scrutinee
-              :@ (AppTypeE (ConE 'SafeSumMatch) (ForallT tyCons [] (foldl AppT (ConT name) (map (VarT . tyVarBndrName) tyCons)))  :@ sumMatches sortedMatches))
+              :@ (ConE 'SafeSumMatch :@ (AppTypeE (ConE 'Proxy) (ConT name))  :@ sumMatches sortedMatches))
   where
     dataCon name = do
       parent <- reify name

### 2020-02-15 src/Deep/Delineate.hs

@@ -1,0 +1,13 @@
+module Deep.Delineate
+  where
+
+import           Deep.Expr
+
+internalize :: GPURep a => GPUExp a -> a
+internalize = gpuAbs
+{-# NOINLINE internalize #-}
+
+externalize :: GPURep a => a -> GPUExp a
+externalize = Construct
+{-# NOINLINE externalize #-}
+

### 2020-02-15 src/Deep/Example.hs

@@ -5,14 +5,16 @@
 {-# LANGUAGE TypeApplications #-}
 {-# LANGUAGE FlexibleInstances #-}
 {-# LANGUAGE StandaloneDeriving #-}
+-- {-# LANGUAGE PolyKinds #-}
+-- {-# LANGUAGE AllowAmbiguousTypes #-}
 
 {-# OPTIONS_GHC -ddump-splices #-}
+{-# OPTIONS_GHC -fprint-explicit-kinds #-}
 
 module Example where
 
 import           Data.List
 
-import           Case
 import           Expr
 
 import           Language.Haskell.TH
@@ -29,6 +31,8 @@ import           Data.Complex
 import           Data.Ratio
 
 import           GHC.Real
+
+import           Data.Proxy
 
 -- import           Debug.Trace
 
@@ -68,17 +72,20 @@ thExample3 = do
     |]
   transformExpr exp
 
-instance GPURep Example'
+instance GPURep Example' where
+  type TypeHead Example' = Example'
 
 
 
 data Example3 = X Int Float deriving (Show, Generic)
 
 instance GPURep Example3 where
+  type TypeHead Example3 = Example3
 
 data Example4 = E1 Int | E2 Float | E3 Bool deriving (Show, Generic)
 
-instance GPURep Example4
+instance GPURep Example4 where
+  type TypeHead Example4 = Example4
 
 thExample4 :: Q Exp
 thExample4 = do
@@ -92,7 +99,8 @@ thExample4 = do
 
 data Example5 = A1 Float Float | A2 Int deriving (Show, Generic)
 
-instance GPURep Example5
+instance GPURep Example5 where
+  type TypeHead Example5 = Example5
 
 thExample5 :: Q Exp
 thExample5 = do
@@ -119,7 +127,8 @@ transformDecTailRec
 
 data IntPair = IntPair Int Int deriving (Show, Generic)
 
-instance GPURep IntPair
+instance GPURep IntPair where
+  type TypeHead IntPair = IntPair
 
 thExample7 :: Q Exp
 thExample7 = do
@@ -134,21 +143,68 @@ thExample7 = do
   return r
 
 
-transformDecTailRec
-  [d|
-  thExample8 :: IntPair -> Int
-  thExample8 p =
-    case p of
-      IntPair x y -> y
+-- transformDecTailRec
+--   [d|
+thExample8 :: IntPair -> Int
+thExample8 p =
+  case p of
[…]

### 2020-02-15 src/Deep/Expr.hs

@@ -7,8 +7,10 @@
 {-# LANGUAGE DeriveGeneric #-}
 {-# LANGUAGE InstanceSigs #-}
 {-# LANGUAGE ScopedTypeVariables #-}
+{-# LANGUAGE TypeOperators #-}
+{-# LANGUAGE FlexibleContexts #-}
 
-module Expr where
+module Deep.Expr where
 
 import           Data.Void
 import           Data.Proxy
@@ -37,21 +39,21 @@ import           Data.Complex
 -- from s -> t, where is a nested pair type, and we use projection
 -- combinators to extract the values inside the function.
 data ProdMatch s t where
-  ProdMatch ::
+  ProdMatch :: forall a b r.
     (GPURep a, GPURep b)
        => (GPUExp a -> ProdMatch b r) -> ProdMatch  (a, b) r
 
-  OneProdMatch :: (GPURep a) => (GPUExp a -> GPUExp r) -> ProdMatch a r
-  NullaryMatch :: GPURep r => GPUExp r -> ProdMatch a r
+  OneProdMatch :: forall a r. (GPURep a) => (GPUExp a -> GPUExp r) -> ProdMatch a r
+  NullaryMatch :: forall a r. GPURep r => GPUExp r -> ProdMatch a r
 
 data SumMatch s t where
-  SumMatch ::
+  SumMatch :: forall a b r.
     (GPURep a, GPURep b, GPURepTy b ~ b)
         => ProdMatch a r -> SumMatch b r -> SumMatch (Either a b) r
 
-  EmptyMatch :: SumMatch Void r
+  EmptyMatch :: forall r. SumMatch Void r
 
-  OneSumMatch :: (GPURep a, GPURep b, GPURepTy a ~ a) => ProdMatch a b -> SumMatch a b
+  OneSumMatch :: forall a b. (GPURep a, GPURep b, GPURepTy a ~ a) => ProdMatch a b -> SumMatch a b
 
 
 -- Done (case ... of A -> x; B -> y)  ==>  (case ... of A -> Done x; B -> Done y)
@@ -61,7 +63,7 @@ data Iter a b
   deriving (Functor, Generic)
 
 data GPUExp t where
-  CaseExp :: (GPURep t) => GPUExp t -> SumMatch (GPURepTy t) r -> GPUExp r
+  CaseExp :: forall t r. (GPURep t) => GPUExp t -> SumMatch (GPURepTy t) r -> GPUExp r
 
   FalseExp :: GPUExp Bool
   TrueExp :: GPUExp Bool
[…]

### 2020-02-15 src/Deep/GHCUtils.hs

@@ -1,0 +1,430 @@
+--
+-- Slightly modified from HERMIT:
+--
+
+{-# LANGUAGE CPP #-}
+{-# LANGUAGE RankNTypes #-}
+
+module GHCUtils where
+
+import           GhcPlugins
+
+import           TcRnTypes
+import           TcSimplify
+import           TcMType
+import           TcRnMonad
+import           TcSMonad
+import           TcEvidence
+import           TcType
+
+import           Finder
+import           LoadIface
+
+import           DsBinds
+import           DsMonad
+
+import           HsBinds
+import           HsDecls
+
+import           CostCentreState
+
+import           Bag
+import           VarSet
+
+import           Encoding
+
+import           ErrUtils
+
+import           Data.IORef
+
+import qualified Data.Map as Map
+import qualified Data.Set as Set
+
+import           Control.Monad
+
+import           Control.Arrow (first, second)
+
+
+import           Data.Char
+import           Data.List
+
+-- | Build a dictionary for the given
+buildDictionary :: ModGuts -> Id -> CoreM (Id, [CoreBind])
+buildDictionary guts evar = do
+    (i, bs) <- runTcM guts $ do
+#if __GLASGOW_HASKELL__ > 710 
+        loc <- getCtLocM (GivenOrigin UnkSkol) Nothing
+#else
+        loc <- getCtLoc $ GivenOrigin UnkSkol
+#endif
+        let predTy = varType evar
+#if __GLASGOW_HASKELL__ > 710 
+            nonC = mkNonCanonical $ CtWanted { ctev_pred = predTy, ctev_dest = EvVarDest evar, ctev_loc = loc }
+            wCs = mkSimpleWC [cc_ev nonC]
+        -- TODO: Make sure solveWanteds is the right function to call.
+        (_wCs', bnds) <- second evBindMapBinds <$> runTcS (solveWanteds wCs)
+#else
+            nonC = mkNonCanonical $ CtWanted { ctev_pred = predTy, ctev_evar = evar, ctev_loc = loc }
+            wCs = mkSimpleWC [nonC]
+        (_wCs', bnds) <- solveWantedsTcM wCs
+#endif
+        -- reportAllUnsolved _wCs' -- this is causing a panic with dictionary instantiation
+                                  -- revist and fix!
+        return (evar, bnds)
+    bnds <- runDsM guts $ dsEvBinds bs
+    return (i,bnds)
+
[…]

### 2020-02-15 src/Plugin.hs

@@ -1,272 +1,0 @@
-{-# LANGUAGE TemplateHaskellQuotes #-}
-{-# LANGUAGE PatternSynonyms #-}
-{-# LANGUAGE TupleSections #-}
-{-# LANGUAGE CPP #-}
-
-module Plugin (plugin) where
-
-import           Expr hiding (Var, Lit, Lam, (:@))
-import qualified Expr
-
-import           Data.Data hiding (TyCon (..), tyConName)
-import           Data.Generics.Uniplate.Operations
-import qualified Data.Generics.Uniplate.DataOnly as Data
-
-import           GhcPlugins
-
-import           InstEnv
-
-import           FamInstEnv
-
-import           TcSMonad
-
-
-import           Control.Monad
-
-import           Control.Arrow ((&&&), (***), first, second)
-
-import           GHCUtils
-
--- Just for TH Names:
-import qualified Language.Haskell.TH.Syntax as TH
-
-infixl 0 :@
-pattern f :@ x = App f x
-
-plugin :: Plugin
-plugin =
-  defaultPlugin
-    { installCoreToDos = install
-    }
-
-install :: [CommandLineOption] -> [CoreToDo] -> CoreM [CoreToDo]
-install _ coreToDos = do
-  return (CoreDoPluginPass "Pattern matching plugin" pass : coreToDos)
-
-pass :: ModGuts -> CoreM ModGuts
-pass guts = do
-  primMap <- forM primMapTH
-                 (\(fromTH, toTH) -> do
-                  fromName <- thNameToGhcName_ fromTH
-                  toName <- thNameToGhcName_ toTH
-
-                  from <- lookupId fromName
-                  to <- lookupId toName
-                  return (from, to))
-
-  constrNames <- mapM (\n -> fmap (,n) (thNameToGhcName_ n)) constrNamesTH
-
-  gpuExprIdMap <- forM constrNames
-                       (\(name, nameTH) -> do
-                          theId <- lookupId name
-                          return (nameTH, theId))
-
-  bindsOnlyPass (mapM (transformBind guts (mg_inst_env guts) primMap (getGPUExprIdsFrom gpuExprIdMap))) guts
-
[…]

### 2020-02-15 src/Plugin/MatchPlugin.hs

@@ -1,0 +1,761 @@
+{-# LANGUAGE TemplateHaskellQuotes #-}
+{-# LANGUAGE PatternSynonyms #-}
+{-# LANGUAGE TupleSections #-}
+{-# LANGUAGE CPP #-}
+{-# LANGUAGE ScopedTypeVariables #-}
+
+module Plugin.MatchPlugin (plugin) where
+
+import           Deep.Expr hiding (Var, Lit, Lam, (:@))
+import qualified Deep.Expr as Expr
+import           Deep.Delineate
+
+import           Data.Data hiding (TyCon (..), tyConName)
+import           Data.Generics.Uniplate.Operations
+import qualified Data.Generics.Uniplate.DataOnly as Data
+
+import           GhcPlugins
+
+import           InstEnv
+
+import           FamInstEnv
+
+import           TcSMonad
+
+
+import           Control.Monad
+
+import           Control.Arrow ((&&&), (***), first, second)
+
+-- import           GHCUtils
+
+-- Just for TH Names:
+import qualified Language.Haskell.TH.Syntax as TH
+
+
+
+import           TcRnTypes
+import           TcSimplify
+import           TcMType
+import           TcRnMonad
+import           TcSMonad
+import           TcEvidence
+import           TcType
+
+import           Finder
+import           LoadIface
+
+import           DsBinds
+import           DsMonad
+
+import           HsBinds
+import           HsDecls
+
+import           CostCentreState
+
+import           Bag
+import           VarSet
+
+import           Encoding
+
+import           ErrUtils
+
+import           Data.IORef
+
+import qualified Data.Map as Map
+import qualified Data.Set as Set
+import           Data.Char
+import           Data.List
+
+infixl 0 :@
+pattern f :@ x = App f x
+
+plugin :: Plugin
+plugin =
+  defaultPlugin
+    { installCoreToDos = install
+    }
+
+install :: [CommandLineOption] -> [CoreToDo] -> CoreM [CoreToDo]
+install _ coreToDos = do
+  return (CoreDoPluginPass "Pattern matching plugin" pass : coreToDos)
+
+pass :: ModGuts -> CoreM ModGuts
[…]

### 2020-02-15 src/Test/PluginExample.hs

@@ -1,0 +1,60 @@
+{-# LANGUAGE DeriveGeneric #-}
+
+{-# OPTIONS_GHC -fplugin=Plugin.MatchPlugin #-}
+
+module Test.PluginExample where
+
+import           Data.List
+
+import           Deep.Expr
+import           Deep.Delineate
+
+import           Data.Char (ord)
+
+import           GHC.Generics
+import           Data.Void
+
+import           Data.Bifunctor
+
+import           Data.Complex
+
+import           Data.Ratio
+
+import           GHC.Real
+
+import           Data.Proxy
+
+-- import           Debug.Trace
+
+-- deriving instance Generic (Ratio a) -- XXX: This instance should probably be in 'base'
+
+data Nat = Z | S Nat deriving (Generic, Show)
+
+data Example = N Nat | B Bool deriving (Generic, Show)
+
+thExample1 :: Int
+thExample1 = do
+  internalize (externalize
+    (case (Right False) :: Either Int Bool of
+      Left x -> x * 2
+      Right y -> fromEnum y))
+
+-- thExample2 :: Q Exp
+-- thExample2 = do
+--   exp <-
+--     [| case (True, the @Int 7) of
+--          (x, y) -> fromEnum x + y
+--     |]
+--   transformExpr exp
+
+-- data Example' = N' Int | B' Bool deriving (Show, Generic)
+
+-- thExample3 :: Q Exp
+-- thExample3 = do
+--   exp <-
+--     [|
+--       case B' False of
+--         N' n -> N' (n+2)
+--         B' b -> B' (not b)
+--     |]
+--   transformExpr exp

### 2020-02-15 paper/paper.tex
Possibly slightly less confusing naming
ConstructAp
Cabal

@@
   algebraic data types and pattern matching in a DSL embedded in Haskell. This
-  representation is automated through the use of GHC Generics and Template Haskell.
+  representation is mechanized through the use of GHC Generics and Template Haskell.
 \end{abstract}
@@
     => (a -> b) -> LiftedFn (a -> b)
-construct = construct' . Construct
+construct = construct' . ConstructE
 
@@
 
-The "helper constructor" \ttt{Construct} is part of the \ttt{E} type:
+The "helper constructors" \ttt{ConstructE} are part of the \ttt{E} type:
 
@@
   ...
-  Construct :: (a -> b) -> E (a -> b)
+  ConstructE :: (a -> b) -> E (a -> b)
+  ConstructAp :: ERep a => E (a -> b) -> E a -> E b
   ...

### 2020-02-15 src/Deep/BB.hs

@@ -6,7 +6,7 @@
 {-# LANGUAGE DeriveGeneric #-}
 {-# LANGUAGE DeriveDataTypeable #-}
 
-module BB where
+module Deep.BB where
 
 import           GHC.Generics
 import           Data.Data

### 2020-02-15 src/Deep/Case.hs

@@ -28,7 +28,7 @@
 {-# LANGUAGE PolyKinds #-}
 {-# LANGUAGE TupleSections #-}
 
-module Case
+module Deep.Case
   where
 
 import           GHC.Generics

### 2020-02-15 src/Deep/Example.hs

@@ -11,11 +11,11 @@
 {-# OPTIONS_GHC -ddump-splices #-}
 {-# OPTIONS_GHC -fprint-explicit-kinds #-}
 
-module Example where
+module Deep.Example where
 
 import           Data.List
 
-import           Expr
+import           Deep.Expr
 
 import           Language.Haskell.TH
 

### 2020-02-15 src/Deep/ExampleMain.hs

@@ -4,7 +4,9 @@
 
 {-# OPTIONS_GHC -ddump-splices #-}
 
-import Example
+module Deep.ExampleMain where
+
+import Deep.Example
 
 import Data.Complex
 import Control.Monad

### 2020-02-15 src/Deep/GHCUtils.hs

@@ -5,7 +5,7 @@
 {-# LANGUAGE CPP #-}
 {-# LANGUAGE RankNTypes #-}
 
-module GHCUtils where
+module Deep.GHCUtils where
 
 import           GhcPlugins
 

### 2020-02-15 src/Deep/Scott.hs

@@ -10,7 +10,7 @@
 {-# LANGUAGE InstanceSigs #-}
 {-# LANGUAGE AllowAmbiguousTypes #-} -- XXX: Is this ok?
 
-module Scott where
+module Deep.Scott where
 
 import           GHC.Generics
 import           Data.Void

### 2020-02-15 src/Deep/TypeError.hs

@@ -6,6 +6,8 @@
 {-# LANGUAGE UndecidableInstances #-}
 -- {-# LANGUAGE AllowAmbiguousTypes #-}
 
+module Deep.TypeError where
+
 import GHC.Types (Constraint, Type)
 import Data.Type.Equality
 
@@ -107,11 +109,11 @@ thIP = Refl
 test2 :: (TypeHead (Tagged2 IntPair IntPair (Int, Int))) :~: IntPair
 test2 = Refl
 
--- test :: Bool
-test_ :: Bool
-test_ = 
-  -- case HRefl :: (TypeHead (Tagged2 (TypeHead IntPair) IntPair (Int, Int))) :~: IntPair of
-    eval (CaseE (LitE (IntPair 1 2)) (SafeSumMatch (Proxy @((TypeHead (Tagged2 (TypeHead IntPair) IntPair (Int, Int))))) (OneSumMatch (ProdMatch (\x -> (OneProdMatch (\y -> LitE True)))))))
+-- -- test :: Bool
+-- test_ :: Bool
+-- test_ = 
+--   -- case HRefl :: (TypeHead (Tagged2 (TypeHead IntPair) IntPair (Int, Int))) :~: IntPair of
+--     eval (CaseE (LitE (IntPair 1 2)) (SafeSumMatch (Proxy @((TypeHead (Tagged2 (TypeHead IntPair) IntPair (Int, Int))))) (OneSumMatch (ProdMatch (\x -> (OneProdMatch (\y -> LitE True)))))))
 
 -- evalTest :: Bool
 -- evalTest = evalSafeSumMatch test (IntPair 1 2)

### 2020-02-15 src/Plugin/MatchPlugin.hs

@@ -103,6 +103,11 @@ pass guts = do
 transformBind :: ModGuts -> InstEnv -> [(Id, Id)] -> (TH.Name -> Var) -> CoreBind -> CoreM CoreBind
 transformBind guts instEnv primMap lookupVar (NonRec name e) =
   fmap (NonRec name) (transformExpr guts Nothing primMap lookupVar e)
+transformBind guts instEnv primMap lookupVar (Rec bnds) = Rec <$> mapM go bnds
+  where
+    go (name, e) =
+      (name,) <$> transformExpr guts (Just name) primMap lookupVar e
+
 -- TODO: Add support for recursive functions
 
 transformExpr :: ModGuts -> Maybe Var -> [(Id, Id)] -> (TH.Name -> Var) -> Expr Var -> CoreM (Expr Var)
@@ -347,6 +352,10 @@ constrNamesTH =
 
   -- Construct --
   ,'construct
+
+  -- Delineate functions --
+  ,'internalize
+  ,'externalize
   ]
 
 

### 2020-02-15 src/Test/Main.hs

@@ -1,0 +1,6 @@
+import           Test.Example
+
+
+main :: IO ()
+main = return ()
+

### 2020-02-15 src/Test/PluginExample.hs

@@ -1,6 +1,6 @@
 {-# LANGUAGE DeriveGeneric #-}
 
-{-# OPTIONS_GHC -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -dcore-lint -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 

### 2020-02-15 src/Plugin/MatchPlugin.hs
Fix automatic partial eval 'bug' and fixed a type error

@@ -98,6 +98,7 @@ pass guts = do
                           theId <- lookupId name
                           return (nameTH, theId))
 
+  -- bindsOnlyPass return guts
   bindsOnlyPass (mapM (transformBind guts (mg_inst_env guts) primMap (getGPUExprIdsFrom gpuExprIdMap))) guts
 
 transformBind :: ModGuts -> InstEnv -> [(Id, Id)] -> (TH.Name -> Var) -> CoreBind -> CoreM CoreBind
@@ -113,7 +114,7 @@ transformBind guts instEnv primMap lookupVar (Rec bnds) = Rec <$> mapM go bnds
 transformExpr :: ModGuts -> Maybe Var -> [(Id, Id)] -> (TH.Name -> Var) -> Expr Var -> CoreM (Expr Var)
 transformExpr guts recName primMap lookupVar = Data.transformM go
   where
-    go (Var f :@ x)
+    go (Var f :@ _ty :@ _dict :@ x)
       | f == lookupVar 'externalize = transformSumMatches guts recName primMap lookupVar x
     go expr = return expr
 
@@ -198,6 +199,7 @@ transformProdMatch guts lookupVar ty0 (altCon@(DataAlt dataAlt), vars0, body) = 
         :@ go pairTyCon repTyCon restTys xs
 
 transformSumMatch :: ModGuts -> (TH.Name -> Var) -> Expr a -> a -> Type -> [Alt a] -> CoreM (Expr a)
+
 transformSumMatch guts lookupVar scrutinee wild ty0 alts@(alt1@(DataAlt dataAlt1, _, _):_) = do
   repTyCon <- findTyCon' guts (varName (lookupVar ''GPURep))
 
@@ -347,6 +349,9 @@ constrNamesTH =
   ,'DoneExp
 
   ,'IfThenElse
+
+  ,''GPURep
+  ,''GPURepTy
 
 
 

### 2020-02-15 src/Test/PluginExample.hs

@@ -1,6 +1,6 @@
 {-# LANGUAGE DeriveGeneric #-}
 
-{-# OPTIONS_GHC -dcore-lint -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -O0 -ddump-simpl -dcore-lint -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 
@@ -32,10 +32,13 @@ data Nat = Z | S Nat deriving (Generic, Show)
 
 data Example = N Nat | B Bool deriving (Generic, Show)
 
+eitherExample :: Either Int Bool
+eitherExample = Right False
+
 thExample1 :: Int
 thExample1 = do
   internalize (externalize
-    (case (Right False) :: Either Int Bool of
+    (case eitherExample of
       Left x -> x * 2
       Right y -> fromEnum y))
 

### 2020-02-15 src/Deep/GHCUtils.hs
Work with named things better

@@ -26,6 +26,8 @@ import           DsMonad
 import           HsBinds
 import           HsDecls
 
+import qualified OccName
+
 import           CostCentreState
 
 import           Bag
@@ -52,13 +54,13 @@ import           Data.List
 buildDictionary :: ModGuts -> Id -> CoreM (Id, [CoreBind])
 buildDictionary guts evar = do
     (i, bs) <- runTcM guts $ do
-#if __GLASGOW_HASKELL__ > 710 
+#if __GLASGOW_HASKELL__ > 710
         loc <- getCtLocM (GivenOrigin UnkSkol) Nothing
 #else
         loc <- getCtLoc $ GivenOrigin UnkSkol
 #endif
         let predTy = varType evar
-#if __GLASGOW_HASKELL__ > 710 
+#if __GLASGOW_HASKELL__ > 710
             nonC = mkNonCanonical $ CtWanted { ctev_pred = predTy, ctev_dest = EvVarDest evar, ctev_loc = loc }
             wCs = mkSimpleWC [cc_ev nonC]
         -- TODO: Make sure solveWanteds is the right function to call.
@@ -306,41 +308,130 @@ mk_dfun_n :: ModGuts -> OccSet
 -- TODO
 mk_dfun_n _ = emptyOccSet
 
+ -- | Possible results from name lookup.
+-- Invariant: One constructor for each NameSpace.
+data Named = NamedId Id
+           | NamedDataCon DataCon
+           | NamedTyCon TyCon
+           | NamedTyVar Var
+
+instance Show Named where
+    show (NamedId _) = "NamedId"
+    show (NamedDataCon _) = "NamedDataCon"
+    show (NamedTyCon _) = "NamedTyCon"
+    show (NamedTyVar _) = "NamedTyVar"
+
+instance NamedThing Named where
+  getOccName (NamedId x) = getOccName x
+  getOccName (NamedDataCon x) = getOccName x
+  getOccName (NamedTyCon x) = getOccName x
+  getOccName (NamedTyVar x) = getOccName x
+
+  getName (NamedId x) = getName x
+  getName (NamedDataCon x) = getName x
+  getName (NamedTyCon x) = getName x
+  getName (NamedTyVar x) = getName x
+
+tyConClassNS :: NameSpace
+tyConClassNS = tcClsName
+
+dataConNS :: NameSpace
+dataConNS = dataName
+
+tyVarNS :: NameSpace
+tyVarNS = tvName
+
+varNS :: NameSpace
+varNS = varNameNS
+
[…]

### 2020-02-15 src/Plugin/MatchPlugin.hs

@@ -6,6 +6,8 @@
 
 module Plugin.MatchPlugin (plugin) where
 
+import           Deep.GHCUtils
+
 import           Deep.Expr hiding (Var, Lit, Lam, (:@))
 import qualified Deep.Expr as Expr
 import           Deep.Delineate
@@ -87,21 +89,21 @@ pass guts = do
                   fromName <- thNameToGhcName_ fromTH
                   toName <- thNameToGhcName_ toTH
 
-                  from <- lookupId fromName
-                  to <- lookupId toName
+                  from <- findId guts fromName emptyVarSet
+                  to <- findId guts toName emptyVarSet
                   return (from, to))
 
   constrNames <- mapM (\n -> fmap (,n) (thNameToGhcName_ n)) constrNamesTH
 
-  gpuExprIdMap <- forM constrNames
+  gpuExprNamedMap <- forM constrNames
                        (\(name, nameTH) -> do
-                          theId <- lookupId name
-                          return (nameTH, theId))
+                          named <- findNamed guts name emptyVarSet
+                          return (nameTH, named))
 
   -- bindsOnlyPass return guts
-  bindsOnlyPass (mapM (transformBind guts (mg_inst_env guts) primMap (getGPUExprIdsFrom gpuExprIdMap))) guts
+  bindsOnlyPass (mapM (transformBind guts (mg_inst_env guts) primMap (getGPUExprNamedsFrom gpuExprNamedMap))) guts
 
-transformBind :: ModGuts -> InstEnv -> [(Id, Id)] -> (TH.Name -> Var) -> CoreBind -> CoreM CoreBind
+transformBind :: ModGuts -> InstEnv -> [(Id, Id)] -> (TH.Name -> Named) -> CoreBind -> CoreM CoreBind
 transformBind guts instEnv primMap lookupVar (NonRec name e) =
   fmap (NonRec name) (transformExpr guts Nothing primMap lookupVar e)
 transformBind guts instEnv primMap lookupVar (Rec bnds) = Rec <$> mapM go bnds
@@ -111,11 +113,15 @@ transformBind guts instEnv primMap lookupVar (Rec bnds) = Rec <$> mapM go bnds
 
 -- TODO: Add support for recursive functions
 
[…]

### 2020-02-24 paper/paper.tex
Expanding on new representation in paper

--- 2020-01-30 paper/paper.tex
@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[acmsmall]{acmart}
+\documentclass[sigplan,screen]{acmart}
 \usepackage{booktabs}
 \usepackage{listings}
+\usepackage{tcolorbox}
 
@@
 
+\newcommand{\typeeq}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}\,}
+
 \newcommand{\ttt}{\texttt}
+
+% \newenvironment<>{todo}[1][]{
+%     \setbeamercolor{block body example}{fg=black,bg=blue}
+%     \setbeamercolor{block title example}{fg=white,bg=red!75!black}
+%     \setbeamertemplate{blocks}[rounded][shadow=false]
+%   \begin{todo}[]}{\end{todo}
+% }
 
@@
 
-%%
-%% These commands are for a JOURNAL article.
-\acmJournal{JACM}
-\acmVolume{1}
-\acmNumber{1}
-\acmArticle{1}
-\acmMonth{1}
+%% These commands are for a PROCEEDINGS abstract or paper.
+\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
+  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
+\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
+  June 03--05, 2018, Woodstock, NY}
+\acmPrice{15.00}
+\acmISBN{978-1-4503-XXXX-X/18/06}
 
@@
 
-% \authornotemark[1]
-% \email{webmaster@marysville-ohio.com}
-% \affiliation{%
-%   \institution{Institute for Clarity in Documentation}
-%   \streetaddress{P.O. Box 1212}
-%   \city{Dublin}
-%   \state{Ohio}
-%   \postcode{43017-6221}
-% }
-
-% \author{Lars Th{\o}rv{\"a}ld}
-% \affiliation{%
-%   \institution{The Th{\o}rv{\"a}ld Group}
-%   \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-%   \city{Hekla}
-%   \country{Iceland}}
-% \email{larst@affiliation.org}
-
-% \author{Valerie B\'eranger}
-% \affiliation{%
-%   \institution{Inria Paris-Rocquencourt}
-%   \city{Rocquencourt}
-%   \country{France}
-% }
-
-% \author{Aparna Patel}
-% \affiliation{%
-%  \institution{Rajiv Gandhi University}
-%  \streetaddress{Rono-Hills}
-%  \city{Doimukh}
-%  \state{Arunachal Pradesh}
-%  \country{India}}
-
-% \author{Huifen Chan}
-% \affiliation{%
-%   \institution{Tsinghua University}
-%   \streetaddress{30 Shuangqing Rd}
-%   \city{Haidian Qu}
-%   \state{Beijing Shi}
-%   \country{China}}
-
-% \author{Charles Palmer}
-% \affiliation{%
-%   \institution{Palmer Research Laboratories}
-%   \streetaddress{8600 Datapoint Drive}
-%   \city{San Antonio}
-%   \state{Texas}
-%   \postcode{78229}}
-% \email{cpalmer@prl.com}
-
-% \author{John Smith}
-% \affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
-% \email{jsmith@affiliation.org}
-
-% \author{Julius P. Kumquat}
-% \affiliation{\institution{The Kumquat Consortium}}
-% \email{jpkumquat@consortium.net}
-
 %%
@@
   algebraic data types and pattern matching in a DSL embedded in Haskell. This
-  representation is automated through the use of GHC Generics and Template Haskell.
+  representation is mechanized through the use of GHC Generics and and GHC Core plugin.
 \end{abstract}
@@
 
+\subsection{Overview}
+
+In this EDSL, the type representing values in the expression language is \ttt{E
+t} and there are two basic functions which translate values to and from this
+expression language with the following type signatures:
+
+\begin{lstlisting}
+rep :: a -> E a
+abs :: E a -> a
+\end{lstlisting}
+
+In this paper, the data constructors of the \ttt{E} type will be described
+incrementally as they are needed.
+
 \subsection{Translation Example} % TODO: Does it make sense for this to be a
@@
 example' =
-  abs (CaseExp (rep x)
-               (SumMatch (OneProdMatch (\c -> ord c))
+  abs (CaseExp
+        (rep x)
+        (SumMatchExp (OneProdMatch
+                       (Lam 1 (Ord (Var 1))))
                      (OneSumMatch
-                           (OneProdMatch (\i -> i)))))
+                       (OneProdMatch
+                         (Lam 2 (Var 2)))))
 \end{lstlisting}
@@
 \section{Representing Algebraic Datatypes}
+
+Briefly, a algebraic type \ttt{T} with an automatically generated \ttt{ERep}
+instance will be given representation in terms of \ttt{Either}, \ttt{(,)},
+\ttt{()} and \ttt{Void}. This "standardized" type representation is
+given by \ttt{ERepTy T}. \ttt{ERepTy} is an type family associated to
+the type class \ttt{ERep}. This type will be called the \textit{canonical type}
+of \ttt{T}.
 
@@
 \begin{lstlisting}
-instance (ERep a, ERep b) => ERep (Either a b) where
+instance (ERep a, ERep b)
+  => ERep (Either a b) where
+
   type ERepTy (Either a b) = Either a b
@@
 
-instance (ERep a, ERep b) => ERep (a, b) where
+instance (ERep a, ERep b)
+  => ERep (a, b) where
+
   type ERepTy (a, b) = (a, b)
@@
   ...
-  CaseExp :: ERep t => E t -> SumMatch (ERepTy t) r -> E r
+  CaseExp ::
+    (ERep t, ERepTy (ERepTy t) ~ ERepTy t)
+      => ERep t
+         -> E (SumMatch (ERepTy t) r)
+         -> E r
   ...
@@
 
-A value of type \ttt{SumMatch a b} represents a computation within the EDSL
+The equality constraint ensures that a canonical type is its own canonical type.
+
+\begin{tcolorbox}
+  \textbf{TODO}: Should we make a synonym for this constraint since it is used in
+multiple places? Like this:
+\begin{lstlisting}
+type CanonicalIsIdem t =
+  ERepTy (ERepTy t) ~ ERepTy t
+\end{lstlisting}
+\end{tcolorbox}
+
+(The above code block is a slight simplification of the actual implementation,
+which has an additional type variable which is used at an "intermediate" point.
+The expanded form is in place to simplify the Core transformation.)
+
+The type \ttt{SumMatch} is defined as
+
+\begin{lstlisting}
+newtype SumMatch a b =
+  MkSumMatch { runSumMatch :: a -> b }
+\end{lstlisting}
+
+For the moment, we will primarily use this type as a type tag and ignore
+the values it can take on.
+
+A value of type \ttt{E (SumMatch a b)} represents a computation within the EDSL
 which destructures a value of type \ttt{E a} and produces a value of type \ttt{E b}.
-Therefore, \ttt{SumMatch (ERepTy t) r} represents a computation which destructures
+Therefore, \ttt{E (SumMatch (ERepTy t) r)} represents a computation which destructures
 a value of type \ttt{E (ERepTy t)} and produces a value of type \ttt{E b}.
 
-\subsection{\ttt{SumMatch}}
+\subsection{\ttt{E (SumMatch a b)}}
 
-The overall structure of \ttt{SumMatch} is a (heterogeneous) list of \ttt{ProdMatch} values.
-Each item of this list corresponds exactly to one branch in the original \ttt{case} match.
+The overall structure of a \ttt{E (SumMatch a b)} value is a (heterogeneous)
+list of \ttt{E (ProdMatch x y)} values.  Each item of this list corresponds
+exactly to one branch in the original \ttt{case} match.
+
+The following constructors generate \ttt{SumMatch}-tagged values in the
+expression language:
 
 \begin{lstlisting}
-data SumMatch s t where
-  SumMach ::
+data E t where
+  ...
+  SumMatchExp ::
     (ERep a, ERep b, ERepTy b ~ b)
-      => ProdMatch a r -> SumMatch b r -> SumMatch (Either a b) r
-
-  EmptyMatch :: SumMatch Void r
+      => E (ProdMatch a r)
+         -> E (SumMatch b r)
+         -> E (SumMatch (Either a b) r)
 
@@
     (ERep a, ERep b, ERepTy a ~ a)
-      => ProdMatch a b -> SumMatch a b
+      => E (ProdMatch a b)
+         -> E (SumMatch a b)
+
+  EmptyMatch ::
+    (ERep b) => E (SumMatch Void b)
+  ...
 \end{lstlisting}
 
-Note the \ttt{ERepTy a ~ a} constraints. This constraint ensures that the type \ttt{a}
-is already in canonical form (that is, consists entirely of \ttt{Either}, \ttt{(,)} and
-base types).
+Note the \ttt{ERepTy a \typeeq a} constraints. This constraint ensures that the type
+\ttt{a} is already in canonical form (that is, consists entirely of
+\ttt{Either}, \ttt{(,)} and base types).
 
-\subsection{\ttt{ProdMatch}}
+\subsection{\ttt{E (ProdMatch s t)}}
 
-\ttt{ProdMatch s t} is equivalent to a function a curried function from \ttt{s} to \ttt{t}.
-Note that \ttt{s} will be a (potentially nested) pair type. For example,
-\ttt{ProdMatch (a, (b, c)) r} will be equivalent to \ttt{a -> b -> c -> r}.
+\ttt{E (ProdMatch x y)} is equivalent to a curried function from \ttt{x} to
+\ttt{y} in the expression language .  Note that \ttt{s} will be a (potentially
+nested) pair type. For example, \ttt{E (ProdMatch (a, (b, c)) r)} will be
+equivalent to \ttt{E (a -> b -> c -> r)}.
 
 \begin{lstlisting}
-data ProdMatch s t where
-  ProdMatch ::
+data E t where
+  ...
+  ProdMatchExp ::
     (ERep a, ERep b)
-      => (E a -> ProdMatch b r) -> ProdMatch (a, b) r
+      => E (a -> ProdMatch b r)
+         -> E (ProdMatch (a, b) r)
 
-  OneProdMatch :: ERep a => (E a -> E r) -> ProdMatch a r
+  NullaryMatch ::
+    (ERep a)
+      => E r -> E (ProdMatch a r)
 
-  NullaryMatch :: ERep r => E r -> ProdMatch a r
+  OneProdMatch ::
+    (ERep a)
+      => E (a -> b) -> E (ProdMatch a b)
+  ...
 \end{lstlisting}
 
-\section{Constructing values}
+The \ttt{ProdMatch} type is defined similarly to the \ttt{SumMatch} type and is
+similiarly used primarily as a type tag:
 
 \begin{lstlisting}
-construct ::
-  Construct (a -> b)
-    => (a -> b) -> LiftedFn (a -> b)
-construct = construct' . Construct
-
--- Example: LiftedFn (a -> b -> c) is: E a -> E b -> E c
-type family LiftedFn t where
-  LiftedFn (a -> b) = E a -> LiftedFn b
-  LiftedFn b = E b
-
-class Construct t where
-  construct' :: E t -> LiftedFn t
+newtype ProdMatch a b =
+  MkProdMatch { runProdMatch :: a -> b }
 \end{lstlisting}
 
-Example GHCi session, demonstrating \ttt{construct}:
+\subsection{Connection to Church encodings}
+TODO: Write this subsection
+
+\subsection{An aside on \ttt{ProdMatch} and \ttt{SumMatch} values}
+
+Though \ttt{ProdMatch} and \ttt{SumMatch} are used throughout this EDSL as type
+tags, they do have values and they are not trivial values. The reason for this
+is that it connects the encoded pattern matches (values from the \ttt{E}) type
+to their semantics. A value of type \ttt{E (SumMatch a b)} is an expression
+which takes in a value of type \ttt{a} (embedded within the expression
+language), internally performs some pattern matching, and produces a value of
+type \ttt{b} (again, embedded within the expression language). This is exactly
+the semantics of a function from \ttt{a} to \ttt{b}. Likewise for \ttt{ProdMatch a b}.
+
+Recall the previously mentioned function
 
 \begin{lstlisting}
-ghci> :t construct (,)
-construct (,)
-  :: (ERep a, ERep b) => E a -> E b -> E (a, b)
+abs :: E a -> a
+\end{lstlisting}
 
-ghci> :t construct (,,)
-construct (,,)
-  :: (ERep a, ERep b, ERep c) =>
-     E a -> E b -> E c -> E (a, b, c)
+Now consider at the type of \ttt{abs} when it is specialized to take \ttt{E
+(SumMatch a b)} values:
 
-ghci> :t construct (,,) (rep False) (rep 3.1 :: E Float) (rep 1 :: E Int)
-construct (,,) (rep False) (rep 3.1 :: E Float) (rep 1 :: E Int)
-  :: E (Bool, Float, Int)
+\begin{lstlisting}
+abs :: E (SumMatch a b) -> SumMatch a b
+\end{lstlisting}
+
+The \ttt{SumMatch a b} value which \ttt{abs} returns is exactly the function
+which pattern matches according to its input value.
+
+Likewise for \ttt{ProdMatch a b} values.
+
+\section{Representing lambdas}
+This representation of pattern matching depends on the ability to bring
+function values into the expression language. This is accomplished
+with the following constructors:
+
+\begin{lstlisting}
+data E t where
+  ...
+  Lam ::
+    (ERep a, Typeable a)
+      => Name a -> E b -> E (a -> b)
+
+  Var :: (Typeable a) => Name a -> E a
+\end{lstlisting}
+
+The \ttt{Typeable} constraints are necessary to lookup correctly typed
+values in the variable binding environment later on.
+
+The \ttt{Name t} type represents a lambda variable identifier together with its
+type \ttt{t}:
+
+\begin{lstlisting}
+newtype Name a = Name Int
+  deriving (Eq, Show)
+
+namesEq :: forall a b.
+  (Typeable a, Typeable b)
+    => Name a -> Name b -> Maybe (a :~: b)
+namesEq (Name n) (Name n') =
+  case eqT :: Maybe (a :~: b) of
+    Just Refl
+      | n == n' -> Just Refl
+      | otherwise -> Nothing
+    Nothing   -> Nothing
+\end{lstlisting}
+
+(Note that the \ttt{ScopedTypeVariables} extension is enabled.)
+
+In the Core transformation, each lambda is given a \ttt{Name} with a globally
+unique \ttt{Int}, sidestepping any name capture issues.
+
+The following datatypes are used to represent a variable binding environment
+of typed names to expression language values:
+
+\begin{lstlisting}
+data EnvMapping where
+  (:=>) :: forall a.
+    Typeable a
+    => Name a -> E a -> EnvMapping
+\end{lstlisting}
+
+This type encodes a single variable binding, with values of the form \ttt{n :=>
+v}, where \ttt{n} is a typed name and \ttt{v} is the value it is bound to.
+
+These bindings are grouped together in the \ttt{Env} type:
+
+\begin{lstlisting}
+newtype Env = Env [EnvMapping]
+
+emptyEnv :: Env
+emptyEnv = Env []
+
+extendEnv :: Env -> EnvMapping -> Env
+extendEnv (Env maps) m = Env (m:maps)
+
+envLookup :: Typeable a
+  => Env -> Name a -> Maybe (E a)
+envLookup (Env maps) = go maps
+  where
+    go []                = Nothing
+
+    go ((n' :=> e):rest) =
+      case namesEq n n' of
+        Just Refl -> Just e
+        Nothing   -> go rest
+\end{lstlisting}
+
+\section{Recovering standard semantics for pattern matches}
+
+\begin{lstlisting}
+abs :: forall t. E t -> t
+abs = absEnv emptyEnv
+
+sumMatchAbs ::
+  (ERepTy (ERepTy s) ~ ERepTy s, ERep s)
+  => Env
+    -> E (SumMatch (ERepTy s) t)
+    -> s
+    -> t
+...
+
+prodMatchAbs :: (ERep s)
+  => Env
+     -> E (ProdMatch s t)
+     -> s
+     -> t
+...
+
+absEnv :: forall t. Env -> E t -> t
+absEnv env (CaseExp x f) =
+  sumMatchAbs f (absEnv env x)
+
+absEnv env (LeftExp x) =
+  Left (absEnv env x)
+
+absEnv env (RightExp y) =
+  Right (absEnv env y)
+
+absEnv env (PairExp x y) =
+  (absEnv env x, absEnv env y)
+
+absEnv env (Repped x) = unrep x
+
+absEnv env env (Var v) =
+  case envLookup env v of
+    Just x -> absEnv env
+    Nothing ->
+      error
+        ("No binding for name "
+          ++ show v)
+
+...
 \end{lstlisting}

### 2020-02-24 src/Deep/Expr.hs

@@ -76,10 +76,10 @@ data Iter a b
   | Done a
   deriving (Functor, Generic)
 
-data Name a = Name Int deriving (Eq, Show)
+newtype Name a = Name Int deriving (Eq, Show)
 
-data EnvMapping =
-  forall a. Typeable a => (Name a) :=> GPUExp a
+data EnvMapping where
+  (:=>) :: forall a. Typeable a => (Name a) -> GPUExp a -> EnvMapping
 
 newtype Env = Env [EnvMapping]
 
@@ -89,11 +89,11 @@ emptyEnv = Env []
 extendEnv :: Env -> EnvMapping -> Env
 extendEnv (Env maps) m = Env (m:maps)
 
-envLookup :: forall a. Typeable a => Env -> Name a -> Maybe (GPUExp a)
-envLookup (Env maps) n@(Name _) = go maps
+envLookup :: Typeable a => Env -> Name a -> Maybe (GPUExp a)
+envLookup (Env maps) n = go maps
   where
     go [] = Nothing
-    go ((n'@(Name _) :=> e):rest) =
+    go ((n' :=> e):rest) =
       case namesEq n n' of
         Just Refl -> Just e
         Nothing   -> go rest
@@ -136,7 +136,6 @@ data GPUExp t where
 
   Repped :: GPURep a => GPURepTy a -> GPUExp a
 
-  -- Lam :: GPURep a => Int -> GPUExp b -> GPUExp (a -> b)
   Lam :: forall a b. (GPURep a, Typeable a) => Name a -> GPUExp b -> GPUExp (a -> b)
   Var :: forall a. Typeable a => Name a -> GPUExp a -- Constructed internally
 
@@ -269,8 +268,8 @@ instance (GPURep a, GPURep b) => GPURep (Either a b) where
   -- type GPURepTy (Either a b) = Either (GPURepTy a) (GPURepTy b)
   type GPURepTy (Either a b) = Either a b
 
-  rep (Left x) = LeftExp (Construct x)
-  rep (Right y) = RightExp (Construct y)
+  rep (Left x) = LeftExp (rep x)
+  rep (Right y) = RightExp (rep y)
 
   rep' = id
   unrep' = id
@@ -288,7 +287,7 @@ instance (GPURep a, GPURep b) => GPURep (a, b) where
   -- type GPURepTy (a, b) =  (GPURepTy a, GPURepTy b)
[…]

### 2020-02-24 src/Test/PluginExample.hs

@@ -1,7 +1,7 @@
 {-# LANGUAGE DeriveGeneric #-}
 {-# LANGUAGE TypeApplications #-}
 
-{-# OPTIONS_GHC -ddump-simpl -O0 -dcore-lint -dcmm-lint -dstg-lint -fcatch-bottoms -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -ddump-simpl -O0 -dcore-lint -dsuppress-all -dcmm-lint -dstg-lint -fcatch-bottoms -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 

### 2020-02-24 paper/paper.tex
Describing Core pluging (not gotten to tail recursion transformation yet)

@@
 \usepackage{tcolorbox}
+\usepackage{tikz}
+\usepackage{tkz-graph}
+\usepackage{stmaryrd}
+\usepackage{amsmath}
+
+% \lstset{basicstyle=\ttfamily}
+\usetikzlibrary{shapes.geometric, arrows}
+
+\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
+\tikzstyle{process} = [rectangle, minimum width=3cm, text width=0.3*\columnwidth, minimum height=1cm, text centered, draw=black, fill=orange!30]
+\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
+\tikzstyle{arrow} = [thick,->,>=stealth]
 
 \lstset{language=haskell, deletekeywords={abs}}
+
+\newcommand{\expr}[1]{\llbracket#1\rrbracket} % Oxford brackets
+\newcommand{\rarr}{\rightarrow}
+\newcommand{\Rarr}{\Rightarrow}
 
@@
 %% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
-  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
-\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-  June 03--05, 2018, Woodstock, NY}
+\acmConference[Haskell '20]{Haskell '20: Proceedings of the 13th
+  ACM SIGPLAN International Symposium on Haskell}{August 27 -- 28, 2020}{New Jersey, United States,}
+\acmBooktitle{Haskell '20: ACM SIGPLAN International Symposium on Haskell,
+  August 27 -- 28, 2020, New Jersey, United States}
 \acmPrice{15.00}
@@
 
-In this EDSL, the type representing values in the expression language is \ttt{E
-t} and there are two basic functions which translate values to and from this
+In this EDSL, the type representing values in the expression language is \ttt{E t}
+and there are two basic functions which translate values to and from this
 expression language with the following type signatures:
@@
 
-Now consider at the type of \ttt{abs} when it is specialized to take \ttt{E
-(SumMatch a b)} values:
+Now consider at the type of \ttt{abs} when it is specialized to take
+\ttt{E (SumMatch a b)} values:
 
@@
 
-This type encodes a single variable binding, with values of the form \ttt{n :=>
-v}, where \ttt{n} is a typed name and \ttt{v} is the value it is bound to.
+This type encodes a single variable binding, with values of the form
+\ttt{n :=> v}, where \ttt{n} is a typed name and \ttt{v} is the value it is bound to.
 
@@
 \end{lstlisting}
+
+\section{Core Plugin}
+
+The Core plugin translates marked expressions. Expressions are marked by
+the \ttt{externalize} function:
+
+\begin{lstlisting}
+externalize :: a -> E a
+\end{lstlisting}
+
+So, \ttt{externalize x} marks the expression \ttt{x}.
+
+If an expression already has an EDSL type (a type of the form \ttt{E a} for some
+\ttt{a}), then it does the marking procedure ignores it and does not wrap it with
+\ttt{externalize}.
+
+\subsection{Overall Control Flow of the Core Plugin}
+
+\begin{tikzpicture}[node distance=2cm]
+\node (start) [startstop] {Core from GHC};
+  \node (findmark) [process, below of=start]{Find marked expression and erase mark};
+    \node (handlecase) [process, below of=findmark]{Mark scrutinee};
+    \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{Encode as EDSL primitive};
+    \node (handleapp) [process, right of=handleprim, xshift=1.6cm]{Wrap with \ttt{Construct} and \ttt{ConstructAp}};
+
+    \node (marksubs1) [process, below of=handleprim]{Mark sub-expressions};
+    \node (marksubs2) [process, below of=handleapp]{Mark sub-expressions};
+
+    \node (onebranch) [process, below of=handlecase]{Transform one branch};
+    \node (prodmatch) [process, below of=onebranch]{Transform one field in product match of current branch};
+    \node (abstractvar) [process, below of=prodmatch]{Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
+    \node (markbody) [process, below of=abstractvar]{Mark branch body};
+    \node (endstep) [process, below of=markbody, right of=markbody, xshift=1.6cm]{End of one transformation step};
+
+  \draw [arrow] (start) -- (findmark);
+  \draw [arrow] (findmark) -- node[anchor=east] {Primitive} (handleprim);
+  \draw [arrow] (findmark) -- node[anchor=west] {Function application/non-prim var} (handleapp);
+  \draw [arrow] (findmark) -- node[anchor=east] {\ttt{case}} (handlecase);
+
+  \draw [arrow] (handleprim) -- (marksubs1);
+
+  \draw [arrow] (handleapp) -- (marksubs2);
+
+  \draw [arrow] (handlecase) -- (onebranch);
+  \draw [arrow] (onebranch) -- (prodmatch);
+  \draw [arrow] (prodmatch) -- (abstractvar);
+  \draw [arrow] (abstractvar) -- (markbody);
+
+  \draw [arrow] (marksubs1) -- (endstep);
+  \draw [arrow] (marksubs2) -- (endstep);
+  \draw [arrow] (markbody) -- (endstep);
+
+  % \draw (prodmatch) edge[bend right, left] node {While fields remain} (abstractvar);
+  % \draw [arrow] (abstractvar.east) to ++(0.5,0) |- node[anchor=west] {While fields remain} (prodmatch.east);
+  \draw [arrow] (abstractvar.west) -| ++(-1.5,0) |- node[anchor=south] {While fields remain} (prodmatch.west);
+
+  \draw [arrow] (markbody.west) -| ++(-3,0) |- node[anchor=east, anchor=south] {While branches remain} (onebranch.west);
+
+  \draw [arrow] (endstep.west) -| ++(-8.5,0) |- node[anchor=west, anchor=south] {While marked expressions remain} (findmark.west);
+
+  % \draw (findmark)  edge[loop left] node {While marked expressions remain} (findmark);
+
+  % \draw (findmark)  edge[loop left] node {While marked expressions remain} (findmark);
+  % \draw (onebranch) edge[loop left] node {While branches remain} (onebranch);
+  % \Loop[label=\(While\, fields\, remain\), labelstyle={left=5pt}](prodmatch);
+  % \path (bodylam) edge [left] (onebranch);
+  % \path (bodylam) edge [loop above] node {test} (onebranch);
+  % \Loop[label=\(While\, remaining\, branches\, exist\), labelstyle={left=5pt}](onebranch)
+  % \draw [arrow] (bodylam) -- (onebranch);
+  % \draw [arrow] (bodylam.east) to  ++ (1.5,0) |- node[anchor=south] {While branches remain} (onebranch);
+  % \draw [arrow] (bodylam.east) to  ++ (1.5,0) |- node[anchor=south] {While branches remain} (onebranch);
+  % \Loop[label=\(While\, branches\, remain\), labelstyle={left=5pt}](onebranch);
+  % \Loop[label=\(While\, marked\, expressions\, remain\), labelstyle={left=5pt}](findmark);
+\end{tikzpicture}
+
+\clearpage
+\subsection{Core Plugin Transformations}
+
+\begin{align*}
+  &C\expr{\ttt{externalize (x + y)}} = \ttt{Add (mark x) (mark y)}\\
+  &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} = \ttt{CaseExp (mark scrutinee) } B\expr{\ttt{\{ ... \}}}\\
+  &\\
+  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} = \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
+  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} = \ttt{OneSumMatchExp }P\expr{\ttt{K x0 ... xN $\rarr$ body}}\\
+  &B\expr{\ttt{\{ \}}} = \ttt{EmptyMatch}\\
+  &\\
+  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} = \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
+  &P\expr{\ttt{x $\rarr$ body}} = \ttt{OneProdMatchExp }A\expr{\texttt{$\lambda$ x $\rarr$ mark body}}\\
+  &P\expr{\ttt{ $\rarr$ body}} = \ttt{NullaryMatch (mark body)}\\
+  &\\
+  &A\expr{\ttt{$\lambda$x $\rarr$ body}} = \ttt{Lam (Name uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{   (where \texttt{uniq} is a globally unique identifier)}
+\end{align*}
 

### 2020-02-24 paper/paper.tex
Name types in transformation equation

@@
   &\\
-  &A\expr{\ttt{$\lambda$x $\rarr$ body}} = \ttt{Lam (Name uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{   (where \texttt{uniq} is a globally unique identifier)}
+  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} = \ttt{Lam (Name @a uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{   (where \texttt{uniq} is a globally unique identifier)}
 \end{align*}

### 2020-02-24 paper/paper.tex
Typo in transformation equation

--- 2020-02-24 paper/paper.tex
@@
   &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} = \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
-  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} = \ttt{OneSumMatchExp }P\expr{\ttt{K x0 ... xN $\rarr$ body}}\\
+  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} = \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
   &B\expr{\ttt{\{ \}}} = \ttt{EmptyMatch}\\
@@
   &\\
-  &A\expr{\ttt{$\lambda$x $\rarr$ body}} = \ttt{Lam (Name uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{   (where \texttt{uniq} is a globally unique identifier)}
+  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} = \ttt{Lam (Name @a uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{\;\;\;\;\;(where \texttt{uniq} is a globally unique identifier)}
 \end{align*}

### 2020-02-24 paper/paper.tex
Info about 'mark'

@@
   &\\
-  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} = \ttt{Lam (Name @a uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{\;\;\;\;\;(where \texttt{uniq} is a globally unique identifier)}
+  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} = \ttt{Lam (Name @a uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{\;\;\;\;\;(where \texttt{uniq} is a globally unique identifier)}\\
+  &\\
+  &\ttt{mark (x :: a)} =
+    \begin{cases}
+      \ttt{externalize x} & \text{if}\; \nexists t, a = E\; t\\
+      \ttt{x} & \text{if}\; \exists t, a = E\; t
+    \end{cases}
 \end{align*}

### 2020-02-24 paper/paper.tex
Different notation for 'mark'

@@
 \begin{align*}
-  &C\expr{\ttt{externalize (x + y)}} = \ttt{Add (mark x) (mark y)}\\
-  &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} = \ttt{CaseExp (mark scrutinee) } B\expr{\ttt{\{ ... \}}}\\
+  &C\expr{\ttt{externalize (x + y)}} = \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
+  &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} = \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
   &\\
@@
   &P\expr{\ttt{x0 x1 ... $\rarr$ body}} = \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
-  &P\expr{\ttt{x $\rarr$ body}} = \ttt{OneProdMatchExp }A\expr{\texttt{$\lambda$ x $\rarr$ mark body}}\\
-  &P\expr{\ttt{ $\rarr$ body}} = \ttt{NullaryMatch (mark body)}\\
+  &P\expr{\ttt{x $\rarr$ body}} = \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ $M\expr{\ttt{body}}$}}\\
+  &P\expr{\ttt{ $\rarr$ body}} = \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
   &\\
-  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} = \ttt{Lam (Name @a uniq) (mark body$[\ttt{x} \mapsto \ttt{Var uniq}]$)}\text{\;\;\;\;\;(where \texttt{uniq} is a globally unique identifier)}\\
+  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} = \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
   &\\
-  &\ttt{mark (x :: a)} =
+  &M\expr{\ttt{(x :: a)}} =
     \begin{cases}

### 2020-02-24 paper/paper.tex
Use type equality symbol
Brief discussion of transformation termination

@@
 \clearpage
-\subsection{Core Plugin Transformations}
+\subsection{Syntactic Transformations of the Core Plugin}
 
@@
   &P\expr{\ttt{x0 x1 ... $\rarr$ body}} = \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
-  &P\expr{\ttt{x $\rarr$ body}} = \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ $M\expr{\ttt{body}}$}}\\
+  &P\expr{\ttt{x $\rarr$ body}} = \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
   &P\expr{\ttt{ $\rarr$ body}} = \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
@@
     \begin{cases}
-      \ttt{externalize x} & \text{if}\; \nexists t, a = E\; t\\
-      \ttt{x} & \text{if}\; \exists t, a = E\; t
+      \ttt{externalize x} & \text{if}\; \nexists t, a\; \typeeq E\; t\\
+      \ttt{x} & \text{if}\; \exists t, a\; \typeeq E\; t
     \end{cases}
 \end{align*}
+
+Note that the total number of marks introduced is upper bounded by the number of
+subexpressions in the original Core given to the transformation by GHC and each
+full transformation step (that is, $C$) eliminates one mark. Therefore, the
+transformation will terminate.
 

### 2020-02-24 paper/paper.tex
Add tail recursion step to diagram and fix minor code formatting issue

@@
-%%
 %% This is file `sample-acmsmall.tex',
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[deletekeywords={Ord}]
 import Data.Char (ord)
@@
 \node (start) [startstop] {Core from GHC};
-  \node (findmark) [process, below of=start]{Find marked expression and erase mark};
+  \node (tailrec) [process, below of=start]{Transform tail recursion (if applicable)};
+  \node (findmark) [process, below of=tailrec]{Find marked expression and erase mark};
     \node (handlecase) [process, below of=findmark]{Mark scrutinee};
@@
 
-  \draw [arrow] (start) -- (findmark);
+  \draw [arrow] (start) -- (tailrec);
+  \draw [arrow] (tailrec) -- (findmark);
   \draw [arrow] (findmark) -- node[anchor=east] {Primitive} (handleprim);

### 2020-02-24 src/Plugin/MatchPlugin.hs

@@ -321,7 +321,7 @@ transformPrims0 guts recName primMap exprVars e = {- transformLams guts mark <=<
           falseId <- lift $ findIdTH guts 'False
           trueId <- lift $ findIdTH guts 'True
           iteId <- lift $ findIdTH guts 'IfThenElse
-     
+
           boolTyCon <- lift $ findTyConTH guts ''Bool
 
           let (DataAlt d, _, _) : _ = alts
@@ -410,7 +410,7 @@ transformPrims0 guts recName primMap exprVars e = {- transformLams guts mark <=<
 
             litId <- lift $ findIdTH guts 'Expr.Lit
             return (Var litId :@ Type doubleTy :@ numDict :@ expr)
-        
+
           | "F#" <- occNameString (occName f) = do
             numTyCon <- lift $ findTyConTH guts ''Num
             floatTyCon <- lift $ findTyConTH guts ''Float
@@ -422,9 +422,9 @@ transformPrims0 guts recName primMap exprVars e = {- transformLams guts mark <=<
             return (Var litId :@ Type floatTy :@ numDict :@ expr)
 
         go expr@(Var f :@ x)
-          | not (isTypeArg x) && 
+          | not (isTypeArg x) &&
             not (isDerivedOccName (occName f)) && last (occNameString (occName f)) /= '#',
-            Just (aTy, bTy) <- splitFunTy_maybe (varType f) = 
+            Just (aTy, bTy) <- splitFunTy_maybe (varType f) =
               whenNotExprTyped guts expr $ do
 
             repTyCon <- lift $ findTyConTH guts ''GPURep
@@ -1007,7 +1007,7 @@ topNormaliseType' guts ty = (secondM (collapseRepTys guts) =<<) $ do
 
   runTcM guts . fmap fst . runTcS $ do
     famInstEnvs <- getFamInstEnvs
-    
+
     case topNormaliseType_maybe famInstEnvs ty of
       Nothing -> return (normaliseType famInstEnvs Nominal ty)
       Just (co, ty')  ->
@@ -1091,7 +1091,7 @@ splitTypeApps_maybe (lhs0 :@ x) =
 
 splitTypeApps_maybe _ = Nothing
 
-   
+
 
 findIdTH :: ModGuts -> TH.Name -> CoreM Id
 findIdTH guts thName = do

### 2020-02-24 paper/paper.tex
Improved wording

@@
 Briefly, a algebraic type \ttt{T} with an automatically generated \ttt{ERep}
-instance will be given representation in terms of \ttt{Either}, \ttt{(,)},
+instance is given representation in terms of \ttt{Either}, \ttt{(,)},
 \ttt{()} and \ttt{Void}. This "standardized" type representation is
@@
 \ttt{E (ProdMatch x y)} is equivalent to a curried function from \ttt{x} to
-\ttt{y} in the expression language .  Note that \ttt{s} will be a (potentially
-nested) pair type. For example, \ttt{E (ProdMatch (a, (b, c)) r)} will be
+\ttt{y} in the expression language .  Note that \ttt{s} is a (potentially
+nested) pair type. For example, \ttt{E (ProdMatch (a, (b, c)) r)} is
 equivalent to \ttt{E (a -> b -> c -> r)}.

### 2020-02-24 paper/paper.tex
Expanded on tail rec step in flowchart
Attempt at tail recursion equations

@@
 \node (start) [startstop] {Core from GHC};
-  \node (tailrec) [process, below of=start]{Transform tail recursion (if applicable)};
+  \node (tailrec) [process, below of=start]{Transform tail recursion of definitions with a mark (when applicable)};
   \node (findmark) [process, below of=tailrec]{Find marked expression and erase mark};
@@
 \begin{align*}
-  &C\expr{\ttt{externalize (x + y)}} = \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
-  &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} = \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
+  D&\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}\\
+      &:= \begin{cases}
+        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter $A\expr{\lambda \ttt{x} \rarr \ttt{case s of $T_{\ttt{f}}\expr{\ttt{\{ ...
+\}}}$}}$)}}$} & \text{if \ttt{f} occurs in \ttt{\{ ...
+\}}}\\
+        \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
+      \end{cases}\\
+\end{align*}
+
+\begin{align*}
+  T_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} := \\
+    &:=
+      \begin{cases}
+        \ttt{\{ K x0 ... xN} \rarr \ttt{body}[\ttt{f} \mapsto \ttt{Step}]; T_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs in \ttt{body}}\\
+        \ttt{\{ K x0 ... xN $\rarr$ Done body; ... \}} & \text{otherwise}
+      \end{cases}\\
+  T_f&\expr{\ttt{\{ \}}} := \ttt{\{ \}}\\
+\end{align*}
+\begin{align*}
+  &C^{*}\expr{\ttt{x}} :=
+    \begin{cases}
+      x & \text{if x has no marked subexpressions}\\
+      C^{*}\expr{C\expr{\ttt{externalize y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\
+    \end{cases}\\
+\end{align*}
+\begin{align*}
+  &C\expr{\ttt{externalize (runIter x)}} := \ttt{TailRec $M\expr{\ttt{x}}$}\\
+  &C\expr{\ttt{externalize (x + y)}} := \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
+  &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} := \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
   &\\
-  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} = \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
-  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} = \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
+  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} := \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
+  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} := \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
   &B\expr{\ttt{\{ \}}} = \ttt{EmptyMatch}\\
   &\\
-  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} = \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
-  &P\expr{\ttt{x $\rarr$ body}} = \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
-  &P\expr{\ttt{ $\rarr$ body}} = \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
+  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} := \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
+  &P\expr{\ttt{x $\rarr$ body}} := \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
+  &P\expr{\ttt{ $\rarr$ body}} := \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
   &\\
-  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} = \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
+  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} := \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
   &\\
-  &M\expr{\ttt{(x :: a)}} =
+  &M\expr{\ttt{(x :: a)}} :=
     \begin{cases}

### 2020-02-24 paper/paper.tex
Slight formatting improvement
More minor formatting improvements

@@
       \begin{cases}
-        \ttt{\{ K x0 ... xN} \rarr \ttt{body}[\ttt{f} \mapsto \ttt{Step}]; T_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs in \ttt{body}}\\
+        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs in \ttt{body}}\\
         \ttt{\{ K x0 ... xN $\rarr$ Done body; ... \}} & \text{otherwise}
@@
     \begin{cases}
-      x & \text{if x has no marked subexpressions}\\
+      \ttt{x} & \text{if \ttt{x} has no marked subexpressions}\\
       C^{*}\expr{C\expr{\ttt{externalize y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\

### 2020-02-24 paper/paper.tex
Basics of tail recursion representation in expression language
Better wording
Format 'TODO'

@@
 \subsection{Connection to Church encodings}
-TODO: Write this subsection
+\begin{tcolorbox}
+  \textbf{TODO}: Write this subsection
+\end{tcolorbox}
 
@@
 Likewise for \ttt{ProdMatch a b} values.
+
+\section{Representing tail recursion}
+\begin{tcolorbox}
+  \textbf{TODO}: Expand this section
+\end{tcolorbox}
+
+\begin{lstlisting}
+data Iter a b = Step b | Done a
+  deriving (Functor, Generic)
+
+runIter :: (ERep a, ERep b)
+  => (a -> Iter b a)
+     -> (a -> b)
+runIter f = go
+  where
+    go x =
+      case f x of
+        Done r  -> r
+        Step x' -> go x'
+\end{lstlisting}
+
+In the \ttt{E} type, this is used to represent tail recursion by
+the following constructors:
+
+\begin{lstlisting}
+data E t where
+  ...
+  StepExp :: E b -> E (Iter a b)
+  DoneExp :: E a -> E (Iter a b)
+
+  TailRec :: (ERep a, ERep b)
+               => E (b -> Iter a b)
+                  -> E (b -> a)
+  ...
+\end{lstlisting}
 

### 2020-02-24 paper/paper.tex
formatting
absEnv interp. of tail rec and lambdas (also fixed width formatting for lstlisting)

@@
 
-\lstset{language=haskell, deletekeywords={abs}}
+\lstset{language=haskell, deletekeywords={abs},basicstyle=\ttfamily}
 
@@
 
+absEnv env (TailRec f) = \ x ->
+  case absEnv env f x of
+    Step x' -> absEnv env (TailRec f) x'
+    Done r  -> r
+
+absEnv env (Var v) =
+  case envLookup env v of
+    Just x -> absEnv env
+    Nothing ->
+      error
+        ("No binding for name "
+          ++ show v)
+
+absEnv env (Lam (name :: Name a)
+                (body :: E b)) =
+  \ (arg :: a) ->
+    let go :: forall x. E x -> E x
+        go expr@(Var name2 :: E a') =
+          case namesEq name name2 of
+            Just Refl -> rep arg
+            _ ->
+              case envLookup env name2 of
+                Just v  -> v
+                Nothing -> expr
+
+        -- [... traverse rest of expression
+        --      in go ...]
+    in
+    absEnv (extendEnv env
+                      (name :=> rep arg))
+           (go body)
+
 absEnv env (LeftExp x) =
@@
 absEnv env (Repped x) = unrep x
-
-absEnv env env (Var v) =
-  case envLookup env v of
-    Just x -> absEnv env
-    Nothing ->
-      error
-        ("No binding for name "
-          ++ show v)
 
@@
 
+\clearpage
 \subsection{Overall Control Flow of the Core Plugin}

### 2020-02-24 src/Deep/Expr.hs

@@ -483,12 +483,11 @@ gpuAbsEnv env (Lam (name :: Name a) (body :: GPUExp b)) = \(arg :: a) ->
 
     let go :: forall x. GPUExp x -> GPUExp x
         go expr@(Var name2 :: GPUExp t') =
-          case eqT :: Maybe (t' :~: a) of
-            Just Refl
-              | name == name2 -> rep arg
-            _   ->
+          case namesEq name name2 of
+            Just Refl -> rep arg
+            _ ->
               case envLookup env name2 of
-                Just v -> v
+                Just v  -> v
                 Nothing -> expr
 
         go (CaseExp x f) = CaseExp (go x) (go f)

### 2020-02-24 src/Plugin/MatchPlugin.hs
Attempt to expose lambda transformation and improve intermediate type-safety of tail recursion transformation

@@ -338,14 +338,14 @@ transformPrims0 guts recName primMap exprVars e = {- transformLams guts mark <=<
           else
             transformSumMatch guts mark scrutinee wild ty alts
 
-        go expr@(Lam v body) = do
-          expTyCon <- lift $ findTyConTH guts ''GPUExp
+        -- go expr@(Lam v body) = do
+        --   expTyCon <- lift $ findTyConTH guts ''GPUExp
 
-          let ty = mkTyConApp expTyCon [varType v]
+        --   let ty = mkTyConApp expTyCon [varType v]
 
-          bodyMarked <- mark (replaceVarType v ty body)
+        --   bodyMarked <- mark (replaceVarType v ty body)
 
-          return (Lam (setVarType v ty) bodyMarked)
+        --   return (Lam (setVarType v ty) bodyMarked)
 
         -- Numeric literals
         go expr@(Lit x :@ ty@(Type{}) :@ dict)
@@ -361,6 +361,8 @@ transformPrims0 guts recName primMap exprVars e = {- transformLams guts mark <=<
         --   | isLam v, isDict dict1, isDict dict2 = do
         --       markedBody <- mark body
         --       return (Var v:@ tyA :@ tyB :@ dict1 :@ dict2 :@ name :@ markedBody)
+
+        go expr@(Lam v e) = abstractOver guts v e
 
         go expr@(Var v)
           | False <- isFunTy (varType v) = whenNotExprTyped guts expr $ do
@@ -743,7 +745,8 @@ transformTailRec guts recVar e0 = do
               resultTyDict <- lift $ buildDictionaryT guts (mkTyConApp repTyCon [resultTy])
               tyDict <- lift $ buildDictionaryT guts (mkTyConApp repTyCon [ty])
 
-              theLam <- abstractOver guts lamV x'
+              -- theLam <- abstractOver guts lamV x'
+              let theLam = Lam lamV x'
 
               return (Var f :@ fTyArg :@ fDict :@
                         (Var runIterId :@ Type resultTy :@ Type ty :@ resultTyDict :@ tyDict :@ theLam))

### 2020-02-24 paper/paper.tex
formatting
Typo in equation
Paper equations now reflect modified tail recursion transformation with increased intermediate type-safety
Added note about the tail recursion transformation being independent of the rest of the EDSL transformations and 'E' type
Handle lambdas properly (including those generated from tail rec transformation)
Section on the structure of the paper

@@
 incrementally as they are needed.
+
+\subsection{Structure of the Paper}
+
+\begin{itemize}
+  \item In section 3, an EDSL representation of pattern matching will be
+    introduced.
+  \item Sections 4 and 5 will discuss techniques for representing tail recursion
+    and lambdas (respectively) in an EDSL. The tail recursion transformation
+    does not depend on the rest of the EDSL and can be used separately from it.
+  \item Section 7 describes the general control flow and specific syntactic
+    transformations of the Core plugin.
+\end{itemize}
 
@@
       &:= \begin{cases}
-        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter $A\expr{\lambda \ttt{x} \rarr \ttt{case s of $T_{\ttt{f}}\expr{\ttt{\{ ...
-\}}}$}}$)}}$} & \text{if \ttt{f} occurs in \ttt{\{ ...
-\}}}\\
+        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of $T_{\ttt{f}}\expr{\ttt{\{ ... \}}}$}$))}}$} & \text{if \ttt{f} occurs in \ttt{\{ ... \}}}\\
         \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
@@
       \end{cases}\\
-  T_f&\expr{\ttt{\{ \}}} := \ttt{\{ \}}\\
+  T_{\ttt{f}}&\expr{\ttt{\{ \}}} := \ttt{\{ \}}\\
 \end{align*}
+
+(Note that the tail recursion transformation given by $D$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
+As a result, the tail recursion transformation can be used on its own.)
+
 \begin{align*}
@@
   &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} := \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
+  &C\expr{\ttt{externalize ($\lambda$x $\rarr$ body)}} := A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
   &\\
@@
   &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} := \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
-  &B\expr{\ttt{\{ \}}} = \ttt{EmptyMatch}\\
+  &B\expr{\ttt{\{ \}}} := \ttt{EmptyMatch}\\
   &\\
@@
 
+\clearpage
 \section{Citations and Bibliographies}
+
+\begin{tcolorbox}
+  \textbf{TODO}: Fill in
+\end{tcolorbox}
 

### 2020-02-25 paper/paper.tex
Move a 'TODO' into generated output
Add 'TODO' note about Oxford brackets
Improved wording of Oxford bracket 'TODO'
Describe section 6 in 'Structure' section

@@
     does not depend on the rest of the EDSL and can be used separately from it.
+  \item Section 6 provides an implementation of \ttt{abs} which interprets an EDSL
+    expression under the "standard" semantics (that is, applying this implementation
+    to the result of the Core transformation will result in a program that behaves exactly
+    the same as the original, untransformed, program).
   \item Section 7 describes the general control flow and specific syntactic
@@
 
-\subsection{Translation Example} % TODO: Does it make sense for this to be a
-                                 % subsection of the intro?
+\subsection{Translation Example} % TODO: 
+\begin{tcolorbox}
+  \textbf{TODO}: Does it make sense for this to be a subsection of the intro?
+\end{tcolorbox}
 
@@
 
+\begin{tcolorbox}
+  \textbf{TODO}: Does this subsection use Oxford brackets in an appropriate way?
+\end{tcolorbox}
+
 \begin{align*}
@@
 \end{align*}
+
+\clearpage
 

### 2020-02-25 paper/paper.tex
Improved example

@@
 
-In the example below, \ttt{example} would get transformed into \ttt{example'}:
+In the example below, the Core plugin transforms \ttt{example} into \ttt{example'}:
 
@@
 example =
-  case x of
+  internalize (externalize
+    (case x of
       Left  c -> ord c
-    Right i -> i
+      Right i -> i))
 

### 2020-02-25 paper/paper.tex
formatting
Added info about Construct/ConstructAp to equations

@@
 In the example below, the Core plugin transforms \ttt{example} into \ttt{example'}:
-
 \begin{lstlisting}[deletekeywords={Ord}]
@@
 x = Left 'a'
-
 example :: Int
@@
 \end{align*}
-
 (Note that the tail recursion transformation given by $D$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
 As a result, the tail recursion transformation can be used on its own.)
-
 \begin{align*}
@@
   &C\expr{\ttt{externalize ($\lambda$x $\rarr$ body)}} := A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
+  &C\expr{\ttt{externalize (f x)}} := \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
+  &C\expr{\ttt{externalize x}} := \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
   &\\

### 2020-02-25 paper/paper.tex
Slight clarification

@@
         \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
-      \end{cases}\\
+      \end{cases}
 \end{align*}
@@
     \begin{cases}
-      \ttt{x} & \text{if \ttt{x} has no marked subexpressions}\\
+      \ttt{x} & \text{if \ttt{x} has no subexpressions marked with \ttt{externalize}}\\
       C^{*}\expr{C\expr{\ttt{externalize y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\

### 2020-02-25 paper/paper.tex
More accurately describe tail rec transformation

@@
 \end{tcolorbox}
-
 \begin{align*}
@@
 \begin{align*}
-  T_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} := \\
+  T_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
+    &:= % \ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \} \}}\\
+      \begin{cases}
+        \ttt{\{ K x0 ... xN $\rarr$ case s of $T_{\ttt{f}}\expr{\ttt{\{ ...$_A$ \}}}$; $T_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}} & \text{if \ttt{f} occurs in \ttt{\{ ...$_A$ \}}}\\
+        \ttt{\{ K x0 ... xN $\rarr$ Done (case s of \{ ...$_A$ \}); $T_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}\}$} & \text{otherwise}
+      \end{cases}\\
+  T_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
     &:=
@@
 \end{align*}
-(Note that the tail recursion transformation given by $D$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
-As a result, the tail recursion transformation can be used on its own.)
 \begin{align*}
@@
 \clearpage
+
+(Note that the tail recursion transformation given by $D$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
+As a result, the tail recursion transformation can be used on its own.)
 

### 2020-02-25 paper/paper.tex
Specify that conditionals are on free occurrences

@@
       &:= \begin{cases}
-        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of $T_{\ttt{f}}\expr{\ttt{\{ ... \}}}$}$))}}$} & \text{if \ttt{f} occurs in \ttt{\{ ... \}}}\\
+        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of $T_{\ttt{f}}\expr{\ttt{\{ ...
+\}}}$}$))}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ...
+\}}}\\
         \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
@@
       \begin{cases}
-        \ttt{\{ K x0 ... xN $\rarr$ case s of $T_{\ttt{f}}\expr{\ttt{\{ ...$_A$ \}}}$; $T_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}} & \text{if \ttt{f} occurs in \ttt{\{ ...$_A$ \}}}\\
+        \ttt{\{ K x0 ... xN $\rarr$ case s of $T_{\ttt{f}}\expr{\ttt{\{ ...$_A$ \}}}$; $T_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}} & \text{if \ttt{f} occurs free in \ttt{\{ ...$_A$ \}}}\\
         \ttt{\{ K x0 ... xN $\rarr$ Done (case s of \{ ...$_A$ \}); $T_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}\}$} & \text{otherwise}
@@
       \begin{cases}
-        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs in \ttt{body}}\\
+        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body}}\\
         \ttt{\{ K x0 ... xN $\rarr$ Done body; ... \}} & \text{otherwise}

### 2020-02-25 paper/paper.tex
Better wording/structure

@@
 
-(Note that the tail recursion transformation given by $D$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
-As a result, the tail recursion transformation can be used on its own.)
+Note that:
+\begin{itemize}
+  \item The tail recursion transformation given by $D$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
+As a result, the tail recursion transformation can be used on its own.
 
-Note that the total number of marks introduced is upper bounded by the number of
+  \item The total number of marks introduced is upper bounded by the number of
 subexpressions in the original Core given to the transformation by GHC and each
@@
 transformation will terminate.
+\end{itemize}
 

### 2020-02-25 paper/paper.tex
Is the tail rec transformation easier to understand with mutually recursive rewrites?
Improved wording in flowchart

@@
 \node (start) [startstop] {Core from GHC};
-  \node (tailrec) [process, below of=start]{Transform tail recursion of definitions with a mark (when applicable)};
+  \node (tailrec) [process, below of=start]{Transform tail recursion of definitions that have a mark (when applicable)};
   \node (findmark) [process, below of=tailrec]{Find marked expression and erase mark};
@@
 \end{tcolorbox}
+
 \begin{align*}
@@
       &:= \begin{cases}
-        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of $T_{\ttt{f}}\expr{\ttt{\{ ...
+        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ...
 \}}}$}$))}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ...
@@
 \begin{align*}
-  T_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
-    &:= % \ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \} \}}\\
-      \begin{cases}
-        \ttt{\{ K x0 ... xN $\rarr$ case s of $T_{\ttt{f}}\expr{\ttt{\{ ...$_A$ \}}}$; $T_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}} & \text{if \ttt{f} occurs free in \ttt{\{ ...$_A$ \}}}\\
-        \ttt{\{ K x0 ... xN $\rarr$ Done (case s of \{ ...$_A$ \}); $T_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}\}$} & \text{otherwise}
-      \end{cases}\\
-  T_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
+  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
+    &:=
+        \ttt{\{ K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}}\\
+  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
     &:=
       \begin{cases}
-        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body}}\\
+        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body}}\\
         \ttt{\{ K x0 ... xN $\rarr$ Done body; ... \}} & \text{otherwise}
       \end{cases}\\
-  T_{\ttt{f}}&\expr{\ttt{\{ \}}} := \ttt{\{ \}}\\
+  T'_{\ttt{f}}&\expr{\ttt{\{ \}}} := \ttt{\{ \}}\\
+  T_{\ttt{f}}&\expr{\ttt{case s of \{ ... \}}}\\
+    &:=
+      \begin{cases}
+        \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
+        \ttt{Done (case s of \{ ... \})} & \text{otherwise}
+      \end{cases}
 \end{align*}

### 2020-02-25 paper/paper.tex
Improved wording in description of marking
Slight change to where marks are removed and add corresponding transformation names to flowchart

@@
 If an expression already has an EDSL type (a type of the form \ttt{E a} for some
-\ttt{a}), then it does the marking procedure ignores it and does not wrap it with
-\ttt{externalize}.
+\ttt{a}), then the marking procedure ignores it and does not wrap it with
+\ttt{externalize} (see $M$ in Section 7.2).
 
@@
 
+In the following flowchart, the names in parentheses in a node refers to the names of transformations from the following section (Section 7.2).
+
 \begin{tikzpicture}[node distance=2cm]
 \node (start) [startstop] {Core from GHC};
-  \node (tailrec) [process, below of=start]{Transform tail recursion of definitions that have a mark (when applicable)};
-  \node (findmark) [process, below of=tailrec]{Find marked expression and erase mark};
-    \node (handlecase) [process, below of=findmark]{Mark scrutinee};
-    \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{Encode as EDSL primitive};
-    \node (handleapp) [process, right of=handleprim, xshift=1.6cm]{Wrap with \ttt{Construct} and \ttt{ConstructAp}};
+  \node (tailrec) [process, below of=start, yshift=-0.5cm]{($D, T, T'$) Transform tail recursion of definitions that have a mark (when applicable)};
+  \node (findmark) [process, below of=tailrec, yshift=-0.5cm]{($C^*$) Find marked expression and erase mark};
+    \node (handlecase) [process, below of=findmark]{($C$) Mark scrutinee};
+    \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{($C$) Encode as EDSL primitive};
+    \node (handleapp) [process, right of=handleprim, xshift=1.6cm]{($C$) Wrap with \ttt{Construct} and \ttt{ConstructAp}};
 
-    \node (marksubs1) [process, below of=handleprim]{Mark sub-expressions};
-    \node (marksubs2) [process, below of=handleapp]{Mark sub-expressions};
+    \node (marksubs1) [process, below of=handleprim]{($C$) Mark sub-expressions};
+    \node (marksubs2) [process, below of=handleapp]{($C$) Mark sub-expressions};
 
-    \node (onebranch) [process, below of=handlecase]{Transform one branch};
-    \node (prodmatch) [process, below of=onebranch]{Transform one field in product match of current branch};
-    \node (abstractvar) [process, below of=prodmatch]{Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
-    \node (markbody) [process, below of=abstractvar]{Mark branch body};
+    \node (onebranch) [process, below of=handlecase]{($B$) Transform one branch};
+    \node (prodmatch) [process, below of=onebranch]{($P$) Transform one field in product match of current branch};
+    \node (abstractvar) [process, below of=prodmatch]{($P$) Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
+    \node (markbody) [process, below of=abstractvar]{($P$) Mark branch body};
     \node (endstep) [process, below of=markbody, right of=markbody, xshift=1.6cm]{End of one transformation step};
@@
       \ttt{x} & \text{if \ttt{x} has no subexpressions marked with \ttt{externalize}}\\
-      C^{*}\expr{C\expr{\ttt{externalize y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\
+      C^{*}\expr{C\expr{\ttt{y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\
     \end{cases}\\
@@
 \begin{align*}
-  &C\expr{\ttt{externalize (runIter x)}} := \ttt{TailRec $M\expr{\ttt{x}}$}\\
-  &C\expr{\ttt{externalize (x + y)}} := \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
-  &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} := \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
-  &C\expr{\ttt{externalize ($\lambda$x $\rarr$ body)}} := A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
-  &C\expr{\ttt{externalize (f x)}} := \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
-  &C\expr{\ttt{externalize x}} := \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
+  &C\expr{\ttt{runIter x}} := \ttt{TailRec $M\expr{\ttt{x}}$}\\
+  &C\expr{\ttt{x + y}} := \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
+  &C\expr{\ttt{case scrutinee of \{ ... \}}} := \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
+  &C\expr{\ttt{($\lambda$x $\rarr$ body)}} := A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
+  &C\expr{\ttt{(f x)}} := \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
+  &C\expr{\ttt{x}} := \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
   &\\

### 2020-02-25 paper/paper.tex
Switched Oxford brackets to parens
Using a slightly more common rewrite notation

@@
 
-\newcommand{\expr}[1]{\llbracket#1\rrbracket} % Oxford brackets
+\newcommand{\expr}[1]{(#1)} % Expression quasiquote
 \newcommand{\rarr}{\rightarrow}
 \newcommand{\Rarr}{\Rightarrow}
+\newcommand{\rewrites}{\Longrightarrow}
 
@@
 
-\begin{tcolorbox}
-  \textbf{TODO}: Does this subsection use Oxford brackets in an appropriate way?
-\end{tcolorbox}
+% \begin{tcolorbox}
+%   \textbf{TODO}: Does this subsection use Oxford brackets in an appropriate way?
+% \end{tcolorbox}
 
@@
   D&\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}\\
-      &:= \begin{cases}
+      &\rewrites \begin{cases}
         \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ...
@@
   T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
-    &:=
+    &\rewrites
         \ttt{\{ K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}}\\
   T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
-    &:=
+    &\rewrites
       \begin{cases}
@@
       \end{cases}\\
-  T'_{\ttt{f}}&\expr{\ttt{\{ \}}} := \ttt{\{ \}}\\
+  T'_{\ttt{f}}&\expr{\ttt{\{ \}}} \rewrites \ttt{\{ \}}\\
   T_{\ttt{f}}&\expr{\ttt{case s of \{ ... \}}}\\
-    &:=
+    &\rewrites
       \begin{cases}
@@
 \begin{align*}
-  &C^{*}\expr{\ttt{x}} :=
+  &C^{*}\expr{\ttt{x}} \rewrites
     \begin{cases}
@@
 \begin{align*}
-  &C\expr{\ttt{runIter x}} := \ttt{TailRec $M\expr{\ttt{x}}$}\\
-  &C\expr{\ttt{x + y}} := \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
-  &C\expr{\ttt{case scrutinee of \{ ... \}}} := \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
-  &C\expr{\ttt{($\lambda$x $\rarr$ body)}} := A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
-  &C\expr{\ttt{(f x)}} := \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
-  &C\expr{\ttt{x}} := \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
+  &C\expr{\ttt{runIter x}} \rewrites \ttt{TailRec $M\expr{\ttt{x}}$}\\
+  &C\expr{\ttt{x + y}} \rewrites \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
+  &C\expr{\ttt{case scrutinee of \{ ... \}}} \rewrites \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
+  &C\expr{\ttt{($\lambda$x $\rarr$ body)}} \rewrites A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
+  &C\expr{\ttt{(f x)}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
+  &C\expr{\ttt{x}} \rewrites \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
   &\\
-  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} := \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
-  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} := \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
-  &B\expr{\ttt{\{ \}}} := \ttt{EmptyMatch}\\
+  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} \rewrites \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
+  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} \rewrites \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
+  &B\expr{\ttt{\{ \}}} \rewrites \ttt{EmptyMatch}\\
   &\\
-  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} := \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
-  &P\expr{\ttt{x $\rarr$ body}} := \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
-  &P\expr{\ttt{ $\rarr$ body}} := \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
+  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} \rewrites \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
+  &P\expr{\ttt{x $\rarr$ body}} \rewrites \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
+  &P\expr{\ttt{ $\rarr$ body}} \rewrites \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
   &\\
-  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} := \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
+  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} \rewrites \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
   &\\
-  &M\expr{\ttt{(x :: a)}} :=
+  &M\expr{\ttt{(x :: a)}} \rewrites
     \begin{cases}

### 2020-02-25 paper/paper.tex
Restored 'TODO' about Oxford brackets
Expand introduction with motivation, add list of contributions and remove 'Structure of paper' section

@@
 
-Embedded domain specific languages (EDSLs) have long been a useful and effective
-technique for constructing reusable tools for writing programs for a variety of
+Embedded domain specific languages (EDSLs) have long been an effective
+technique for constructing reusable tools for working in a variety of
 different problem domains. Haskell is a language which is particularly
@@
 
+Despite these advantages Haskell provides for creating EDSLs, there are a few
+constructs which have often proved illusive to embed in an EDSL. One prominent
+example is that of a \ttt{case} expression. Pattern matching is a convenient way
+to implement control flow structures and to view components of data structures,
+so it is frequently a desirable feature for many EDSLs.
+
+Additionally, lambdas can also be difficult to implement in EDSLs. A major reason
+for this is that control flow constructs which are typically \textit{outside} the EDSL,
+such as pattern matching and tail recursion, can make it challenging to "look inside"
+the lambda.
+
+The following contributions are made by this paper:
+
+\begin{itemize}
+\item A representation of pattern matching in a Haskell EDSL (Section 3)
+\item Representations of tail recursion (Section 4) and lambdas (Section 5) in
+  the EDSL. We will see that lambdas become easier to implement once pattern
+  matching and tail recursion already have EDSL representations.
+\item A GHC Core plugin which transforms a subset of standard Haskell code
+  into this EDSL for pattern matching, extended with tail recursion,
+  lambdas and primitive operations (Section 7)
+\item An implementation of a interpreter for the "standard" semantics of the EDSL (Section 6).
+\end{itemize}
+
 \subsection{Overview}
@@
 
-\subsection{Structure of the Paper}
-
-\begin{itemize}
-  \item In section 3, an EDSL representation of pattern matching will be
-    introduced.
-  \item Sections 4 and 5 will discuss techniques for representing tail recursion
-    and lambdas (respectively) in an EDSL. The tail recursion transformation
-    does not depend on the rest of the EDSL and can be used separately from it.
-  \item Section 6 provides an implementation of \ttt{abs} which interprets an EDSL
-    expression under the "standard" semantics (that is, applying this implementation
-    to the result of the Core transformation will result in a program that behaves exactly
-    the same as the original, untransformed, program).
-  \item Section 7 describes the general control flow and specific syntactic
-    transformations of the Core plugin.
-\end{itemize}
-
-\subsection{Translation Example} % TODO: 
+\subsection{Translation Example}
 \begin{tcolorbox}
@@
 abs :: E (SumMatch a b) -> SumMatch a b
+\end{lstlisting}
+
+If we postcompose with \ttt{runSumMatch}, we get:
+
+\begin{lstlisting}
+runSumMatch . abs :: E (SumMatch a b) -> (a -> b)
 \end{lstlisting}
@@
 \subsection{Syntactic Transformations of the Core Plugin}
-
-% \begin{tcolorbox}
-%   \textbf{TODO}: Does this subsection use Oxford brackets in an appropriate way?
-% \end{tcolorbox}
-
+\begin{tcolorbox}
+  \textbf{TODO}: Should we use Oxford brackets here?
+Maybe a bit unusual, but we are operating syntactically and mapping from \ttt{a} to \ttt{E a} (could \ttt{E a} be considered an "intermediary" semantic domain?).
+\end{tcolorbox}
+\vspace*{-0.6cm}
 \begin{align*}
@@
 \begin{itemize}
-  \item The tail recursion transformation given by $D$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
+  \item The tail recursion transformation given by $D$, $T'_{\ttt{f}}$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
 As a result, the tail recursion transformation can be used on its own.

### 2020-02-25 paper/paper.tex
Slight rewording and restructuring

@@
 % \renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+\renewcommand{\shortauthors}{Young and Gill}
 
@@
 Despite these advantages Haskell provides for creating EDSLs, there are a few
-constructs which have often proved illusive to embed in an EDSL. One prominent
-example is that of a \ttt{case} expression. Pattern matching is a convenient way
-to implement control flow structures and to view components of data structures,
-so it is frequently a desirable feature for many EDSLs.
+constructs for which representations have proven illusive. One prominent example
+is that of a \ttt{case} expression. Pattern matching is a convenient way to
+implement control flow structures and to view components of data structures, so
+it is frequently a desirable feature for many EDSLs.
 
@@
 x = Left 'a'
+
 example :: Int
@@
         (SumMatchExp (OneProdMatch
-                       (Lam 1 (Ord (Var 1))))
+                       (Lam 1
+                         (Ord (Var 1))))
                      (OneSumMatch
                        (OneProdMatch
-                         (Lam 2 (Var 2)))))
+                         (Lam 2
+                           (Var 2)))))
 \end{lstlisting}
@@
 \begin{tcolorbox}
-  \textbf{TODO}: Expand this section
+  \textbf{TODO}: Expand the explanation in this section
 \end{tcolorbox}
@@
 The \ttt{Name t} type represents a lambda variable identifier together with its
-type \ttt{t}:
+type \ttt{t} (Note that the \ttt{ScopedTypeVariables} extension is enabled.):
 
@@
 \end{lstlisting}
-
-(Note that the \ttt{ScopedTypeVariables} extension is enabled.)
 

### 2020-02-25 src/Deep/Expr.hs
instance GPURep ()

@@ -351,6 +351,11 @@ instance GPURep (U1 p) where
   rep' U1 = ()
   unrep' () = U1
 
+instance GPURep () where
+  type GPURepTy () = ()
+
+  rep' = id
+  unrep' = id
 
 class GenericRep repr where
     type GenericRepTy repr

### 2020-02-25 src/Plugin/MatchPlugin.hs
Fix for NullaryMatch transformation

@@ -503,7 +503,7 @@ transformProdMatch guts mark resultTy ty0_ (altCon@(DataAlt dataAlt), vars0, bod
     go body pairTyCon repTyCon (ty1:_) [] = do
       nullaryMatchId <- lift $ findIdTH guts 'NullaryMatch
 
-      resultTyDict <- lift $ buildDictionaryT guts (mkTyConApp repTyCon [resultTy])
+      resultTyDict <- lift $ buildDictionaryT guts (mkTyConApp repTyCon [ty1])
 
       return (Var nullaryMatchId :@ Type ty1 :@ Type resultTy :@ resultTyDict :@ body)
 

### 2020-02-25 src/Test/PluginExample.hs
Paper examples. Also, recursive types work now due to the 'delayed GPURepTy' implemented to fix another issue

@@ -1,5 +1,6 @@
 {-# LANGUAGE DeriveGeneric #-}
 {-# LANGUAGE TypeApplications #-}
+{-# LANGUAGE GADTs #-}
 
 {-# OPTIONS_GHC -ddump-simpl -O0 -dcore-lint -dsuppress-all -dcmm-lint -dstg-lint -fcatch-bottoms -fplugin=Plugin.MatchPlugin #-}
 
@@ -147,6 +148,27 @@ example7_ x = IntPair 1 2
 --       False -> 3 * 2
 --       True -> 3 * 3))
 
+data ComplexPair where
+  ComplexPair :: Complex Double
+                 -> Complex Double
+                 -> ComplexPair
+  deriving (Generic, Eq, Show)
+
+instance GPURep ComplexPair
+
+sumComplexPair :: ComplexPair -> Complex Double
+sumComplexPair t =
+  internalize (externalize
+    (case t of
+      ComplexPair a b -> a + b))
+
+realSum :: ComplexPair -> Double
+realSum p =
+  internalize (externalize
+    (case p of
+      ComplexPair (a :+ _) (b :+ _) ->
+        a + b))
+
 mandelbrot_nextZ :: (Complex Double, Complex Double) -> Complex Double
 mandelbrot_nextZ t =
   internalize (externalize
@@ -172,9 +194,37 @@ mandelbrot_point c =
   internalize (externalize
     (mandelbrot_helper (0, c, 0)))
 
+data IntList = Nil | Cons Int IntList
+  deriving (Generic, Show)
+
+instance GPURep IntList
+
+isEmpty :: IntList -> Bool
+isEmpty t =
+  internalize (externalize
+    (case t of
+      Nil -> True
+      Cons x xs -> False))
+
+intListLength_helper :: (Int, IntList) -> Int
+intListLength_helper p =
+  internalize (externalize
+    (case p of
+      (acc, t) ->
+        case t of
+          Nil -> acc
+          Cons _ xs -> intListLength_helper (acc+1, xs)))
+
+intListLength :: IntList -> Int
+intListLength t = intListLength_helper (0, t)
+
 main :: IO ()
 main =
-  putStrLn mandelbrotTestAscii
+  print (isEmpty Nil, isEmpty (Cons 1 Nil)
+        ,intListLength (Cons 10 (Cons 100 (Cons 1000 (Cons 10000 Nil))))
+        )
[…]

### 2020-02-25 paper/paper.tex
Explanation for 'delayed ERepTy', with examples including a recursive example

@@
 
+Only the "outermost" type will be deconstructed into these fundamental building
+blocks, and further deconstruction can take place on these pieces later on. For
+example, consider this type:
+
+\begin{lstlisting}
+data ComplexPair where
+  ComplexPair :: Complex Double
+                 -> Complex Double
+                 -> ComplexPair
+  deriving (Generic, Eq, Show)
+
+instance ERep ComplexPair
+\end{lstlisting}
+
+Note that the instance definition is automatically generated from the
+\ttt{Generic} instance.
+
+Given this code, \ttt{ERep ComplexPair \typeeq (Complex Double, Complex Double)}. Here
+is an example that demonstrates why this is more useful than if it *fully* deconstructed
+each type into \ttt{Either}s and \ttt{(,)}:
+
+\begin{lstlisting}
+sumComplexPair ::
+  ComplexPair -> Complex Double
+sumComplexPair p =
+  internalize (externalize
+    (case p of
+      ComplexPair a b -> a + b))
+\end{lstlisting}
+
+If \ttt{ERep ComplexPair} was fully deconstructed into
+\ttt{((Double, Double), (Double, Double))}, we would need a \ttt{Num}
+instance for \ttt{(Double, Double)}.
+What we really want is to use the \ttt{Num} instance that already
+exists for \ttt{Complex Double}, and this is what preserving this type
+information allows us to do.
+
+It is important to note that we are still able to further deconstruct this type
+with further pattern matches, since \ttt{ERep (Complex Double) \typeeq (Double, Double)}:
+
+\begin{lstlisting}
+realSum :: ComplexPair -> Double
+realSum p =
+  internalize (externalize
+    (case p of
+      ComplexPair a b ->
+        case a of
+          a_real :+ _ ->
+            case b of
+              b_real :+ _ ->
+                a_real + b_real))
+\end{lstlisting}
+
+\begin{tcolorbox}
+  \textbf{TODO}: Should we use a single nested pattern match in the above block? Both
+  are functional (and equivalent) from a technical perspective.
+\end{tcolorbox}
+
+An additional benefit of this is that recursive types require no special handling. For
+example, consider:
+
+\begin{lstlisting}
+data IntList = Nil | Cons Int IntList
+  deriving (Generic, Show)
+
+instance ERep IntList
+\end{lstlisting}
+
+Note that \ttt{ERepTy IntList \typeeq Either () (Int, IntList)}. If \ttt{ERepTy}
+attempted to "fully deconstruct" \ttt{IntList}, it would send the compiler
+into an infinite loop.
+
+This allows us to implement functions on such recursive types:
+
+\begin{lstlisting}
+isEmpty :: IntList -> Bool
+isEmpty t =
+  internalize (externalize
+    (case t of
+      Nil -> True
+      Cons x xs -> False))
+
+intListSum_helper :: (Int, IntList) -> Int
+intListSum_helper p =
+  internalize (externalize
+    (case p of
+       (acc, t) ->
+         case t of
+           Nil -> acc
+           Cons x xs ->
+             intListSum_helper (x+acc, xs)))
+
+intListSum :: IntList -> Int
+intListSum t =
+  internalize (externalize
+    (intListSum_helper (0, t)))
+\end{lstlisting}
+
 \subsection{GHC Generics \ttt{Rep}}
@@
 \ttt{ERep} and \ttt{ERepTy} provide an interface for transferring values between the EDSL
-world and the Haskell world:
+expression language and the source Haskell language:
 

### 2020-02-25 src/Test/PluginExample.hs
intListSum example

@@ -216,12 +216,28 @@ intListLength_helper p =
           Cons _ xs -> intListLength_helper (acc+1, xs)))
 
 intListLength :: IntList -> Int
-intListLength t = intListLength_helper (0, t)
+intListLength t = internalize (externalize (intListLength_helper (0, t)))
+
+intListSum_helper :: (Int, IntList) -> Int
+intListSum_helper p =
+  internalize (externalize
+    (case p of
+       (acc, t) ->
+         case t of
+           Nil -> acc
+           Cons x xs -> intListSum_helper (x+acc, xs)))
+
+intListSum :: IntList -> Int
+intListSum t =
+  internalize (externalize (intListSum_helper (0, t)))
 
 main :: IO ()
 main =
+  let intList = Cons 10 (Cons 100 (Cons 1000 (Cons 10000 Nil)))
+  in
   print (isEmpty Nil, isEmpty (Cons 1 Nil)
-        ,intListLength (Cons 10 (Cons 100 (Cons 1000 (Cons 10000 Nil))))
+        ,intListLength intList
+        ,intListSum intList
         )
   -- print $ realSum (ComplexPair (2 :+ 100) (3 :+ 10000))
   -- putStrLn mandelbrotTestAscii

### 2020-02-25 paper/paper.tex
Better formatting for type equality symbol
Improved some wording

@@
 
-\newcommand{\typeeq}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}\,}
+\newcommand{\typeeq}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}\,\;}
 
@@
 Briefly, a algebraic type \ttt{T} with an automatically generated \ttt{ERep}
-instance is given representation in terms of \ttt{Either}, \ttt{(,)},
-\ttt{()} and \ttt{Void}. This "standardized" type representation is
-given by \ttt{ERepTy T}. \ttt{ERepTy} is an type family associated to
-the type class \ttt{ERep}. This type will be called the \textit{canonical type}
-of \ttt{T}.
+instance is given a representation in terms of \ttt{Either}, \ttt{(,)}, \ttt{()}
+and \ttt{Void}.  This "standardized" type representation is given by \ttt{ERepTy
+T}. \ttt{ERepTy} is an type family associated to the type class \ttt{ERep}. This
+type will be called the \textit{canonical type} of \ttt{T}.
 
@@
 Given this code, \ttt{ERep ComplexPair \typeeq (Complex Double, Complex Double)}. Here
-is an example that demonstrates why this is more useful than if it *fully* deconstructed
-each type into \ttt{Either}s and \ttt{(,)}:
+is an example that demonstrates why this is more useful than if it \textit{fully} deconstructed
+each type into \ttt{Either}, \ttt{(,)} and \ttt{()}:
 
@@
 
-If \ttt{ERep ComplexPair} was fully deconstructed into
-\ttt{((Double, Double), (Double, Double))}, we would need a \ttt{Num}
-instance for \ttt{(Double, Double)}.
-What we really want is to use the \ttt{Num} instance that already
-exists for \ttt{Complex Double}, and this is what preserving this type
-information allows us to do.
+If \ttt{ERep ComplexPair} were fully deconstructed into \ttt{((Double, Double),
+(Double, Double))}, we would need a \ttt{Num} instance for \ttt{(Double,
+Double)}.  What we really want is to use the fact that a \ttt{Num} instance
+already exists for \ttt{Complex Double}. This is exactly what preserving this
+type information allows us to do.
 
@@
 \node (start) [startstop] {Core from GHC};
-  \node (tailrec) [process, below of=start, yshift=-0.5cm]{($D, T, T'$) Transform tail recursion of definitions that have a mark (when applicable)};
+  \node (tailrec) [process, below of=start, yshift=-0.5cm]{($D, T_{\ttt{f}}, T'_{\ttt{f}}$) Transform tail recursion of definitions that have a mark (when applicable)};
   \node (findmark) [process, below of=tailrec, yshift=-0.5cm]{($C^*$) Find marked expression and erase mark};

### 2020-02-25 paper/paper.tex
More forward references to rewrite rules in flowchart

@@
 
-    \node (marksubs1) [process, below of=handleprim]{($C$) Mark sub-expressions};
-    \node (marksubs2) [process, below of=handleapp]{($C$) Mark sub-expressions};
+    \node (marksubs1) [process, below of=handleprim]{($C, M$) Mark sub-expressions};
+    \node (marksubs2) [process, below of=handleapp]{($C, M$) Mark sub-expressions};
 
@@
     \node (prodmatch) [process, below of=onebranch]{($P$) Transform one field in product match of current branch};
-    \node (abstractvar) [process, below of=prodmatch]{($P$) Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
-    \node (markbody) [process, below of=abstractvar]{($P$) Mark branch body};
+    \node (abstractvar) [process, below of=prodmatch]{($P, A$) Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
+    \node (markbody) [process, below of=abstractvar]{($P, M$) Mark branch body};
     \node (endstep) [process, below of=markbody, right of=markbody, xshift=1.6cm]{End of one transformation step};

### 2020-02-25 paper/paper.tex
An additional forward reference in flowchart and improved spacing

@@
   \node (findmark) [process, below of=tailrec, yshift=-0.5cm]{($C^*$) Find marked expression and erase mark};
-    \node (handlecase) [process, below of=findmark]{($C$) Mark scrutinee};
+    \node (handlecase) [process, below of=findmark]{($C, M$) Mark scrutinee};
     \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{($C$) Encode as EDSL primitive};
@@
     \node (prodmatch) [process, below of=onebranch]{($P$) Transform one field in product match of current branch};
-    \node (abstractvar) [process, below of=prodmatch]{($P, A$) Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
+    \node (abstractvar) [process, below of=prodmatch, yshift=-0.5cm]{($P, A$) Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
     \node (markbody) [process, below of=abstractvar]{($P, M$) Mark branch body};

### 2020-02-25 paper/paper.tex
Made 'todo' environment

@@
 \newcommand{\ttt}{\texttt}
+
+\newenvironment{todo}
+  {\begin{tcolorbox}
+   \textbf{TODO}:
+  }
+  {\end{tcolorbox}
+  }
 
@@
 \subsection{Translation Example}
-\begin{tcolorbox}
-  \textbf{TODO}: Does it make sense for this to be a subsection of the intro?
-\end{tcolorbox}
+\begin{todo}
+  Does it make sense for this to be a subsection of the intro?
+\end{todo}
 
@@
 
-\begin{tcolorbox}
-  \textbf{TODO}: Should we use a single nested pattern match in the above block? Both
+\begin{todo}
+  Should we use a single nested pattern match in the above block? Both
   are functional (and equivalent) from a technical perspective.
-\end{tcolorbox}
+\end{todo}
 
@@
 
-\begin{tcolorbox}
-  \textbf{TODO}: Should we make a synonym for this constraint since it is used in
+\begin{todo}
+  Should we make a synonym for this constraint since it is used in
 multiple places? Like this:
@@
 \end{lstlisting}
-\end{tcolorbox}
+\end{todo}
 
@@
 \subsection{Connection to Church encodings}
-\begin{tcolorbox}
-  \textbf{TODO}: Write this subsection
-\end{tcolorbox}
+\begin{todo}
+  Write this subsection
+\end{todo}
 
@@
 \section{Representing tail recursion}
-\begin{tcolorbox}
-  \textbf{TODO}: Expand the explanation in this section
-\end{tcolorbox}
+\begin{todo}
+  Expand the explanation in this section
+\end{todo}
 
@@
 \subsection{Syntactic Transformations of the Core Plugin}
-\begin{tcolorbox}
-  \textbf{TODO}: Should we use Oxford brackets here?
-Maybe a bit unusual, but we are operating syntactically and mapping from \ttt{a} to \ttt{E a} (could \ttt{E a} be considered an "intermediary" semantic domain?).
-\end{tcolorbox}
+\begin{todo}
+  Should we use Oxford brackets here? Maybe a bit unusual, but we are operating syntactically and mapping from \ttt{a} to \ttt{E a} (could \ttt{E a} be considered an "intermediary" semantic domain?).
+\end{todo}
 \vspace*{-0.6cm}
@@
 
-\begin{tcolorbox}
-  \textbf{TODO}: Fill in
-\end{tcolorbox}
+\begin{todo}
+  Fill in
+\end{todo}
 

### 2020-02-25 src/Deep/GHCUtils.hs
Remove debug prints

@@ -112,7 +112,6 @@ buildDictionaryT guts = \ ty0 -> do
             Just (tyCon, tyArgs) -> do
                 tyArgs' <- mapM (normaliseType' guts) tyArgs
                 let r = mkTyConApp tyCon tyArgs'
-                liftIO $ putStrLn $ "r = " ++ showPpr dflags tyArgs'
                 return r
             Nothing -> return ty0
 
@@ -146,8 +145,8 @@ tryToFillCoHoles guts bind =
     go :: CoreExpr -> CoreM CoreExpr
     go expr@(Coercion (HoleCo coHole@(CoercionHole {ch_co_var}))) = do
       dflags <- getDynFlags
-      liftIO $ putStrLn ("got to a co hole named " ++ showPpr dflags ch_co_var)
-      liftIO $ putStrLn ("    with type " ++ showPpr dflags (varType ch_co_var))
+      -- liftIO $ putStrLn ("got to a co hole named " ++ showPpr dflags ch_co_var)
+      -- liftIO $ putStrLn ("    with type " ++ showPpr dflags (varType ch_co_var))
       case varType ch_co_var of
         CoercionTy co -> do
           let mkCo r t = mkReflCo r t
@@ -173,8 +172,6 @@ tryToFillCoHoles guts bind =
           return expr
         TyConApp tyCon lst@(_:_:ty:_) -> do
           return (Coercion (mkReflCo Nominal ty))
-          -- liftIO $ putStrLn ("    TyConApp (" ++ showPpr dflags tyCon ++ ") (" ++ showPpr dflags xs ++ ")")
-          -- return expr
         _ -> do
           liftIO $ putStrLn "    cannot fill (outer)"
           return expr

### 2020-02-25 src/Plugin/MatchPlugin.hs

@@ -284,8 +284,6 @@ mark0 guts x = do
           repTyCon <- lift $ findTyConTH guts ''GPURep
           dflags <- getDynFlags
 
-          liftIO $ putStrLn $ "mark0: xTy = " ++ showPpr dflags xTy
-          liftIO $ putStrLn $ "mark0: x   = {{" ++ showPpr dflags x ++ "}}"
           dict <- lift $ buildDictionaryT guts (mkTyConApp repTyCon [xTy])
 
           return (Var externalizeName :@ Type xTy :@ dict :@ x)
@@ -315,7 +313,6 @@ transformPrims0 guts recName primMap exprVars e = {- transformLams guts mark <=<
 
         go :: Expr Var -> MatchM (Expr Var)
         go (Case scrutinee wild ty alts) = do
-          liftIO $ putStrLn "saw a case"
           dflags <- getDynFlags
 
           falseId <- lift $ findIdTH guts 'False
@@ -485,7 +482,6 @@ transformProdMatch guts mark resultTy ty0_ (altCon@(DataAlt dataAlt), vars0, bod
   repTyCon <- lift $ findTyConTH guts ''GPURep
 
   ty0 <- lift $ (unwrapExpType guts <=< unwrapExpType guts) ty0_
-  liftIO $ putStrLn $ "ty0 = " ++ showPpr dflags ty0
 
   repType <- lift $ computeRepType guts ty0
 
@@ -667,8 +663,6 @@ transformSumMatch guts mark scrutinee wild resultTy alts@(alt1@(DataAlt dataAlt1
       ty1' <- lift $ repTyUnwrap guts ty1
       restTy' <- lift $ repTyUnwrap guts restTy
 
-      liftIO $ putStrLn $ "coPair = " ++ showPpr dflags coPair
-
       return (Var sumMatchId
                 :@ Type ty1'
                 :@ Type restTy'
@@ -825,8 +819,6 @@ transformLams guts mark e0 = Data.transformM go e0
           argTy' <- lift $ unwrapExpType guts (varType arg)
           bodyTy' <- lift $ unwrapExpType guts (exprType body')
 
-          -- liftIO $ putStrLn $ "(argTy', bodyTy') = " ++ showPpr dflags (argTy', bodyTy')
-
           iHash <- lift $ findIdTH guts 'GHC.Types.I#
           -- intTy <- findTypeTH guts ''Int#
 
[…]

### 2020-02-25 paper/paper.tex
Slightly expanded a couple things

@@
                  -> ComplexPair
-  deriving (Generic, Eq, Show)
+  deriving (Generic, Show)
 
@@
 
-[Remove metadata from Generic representation and turn tree-like structure into a (nested) list-like structure]
+\begin{todo}
+  This section is to talk about the details of GHC's generic \ttt{Rep} representation
+  and how it is converted to the representation in this paper. Should this section
+  exist?
+\end{todo}
 
@@
 Canonical form types crucially include \ttt{Either} and \ttt{(,)}, which
-allow all combinations of basic sum types and product types to be encoded. More
-information on this encoding is given in [a different section].
+allow all combinations of basic sum types and product types to be encoded, in the
+manner described at the beginning of Section 2.
 

### 2020-02-25 src/Deep/Expr.hs
Abstracted out traversal from lambda eval part of gpuAbs

@@ -11,6 +11,7 @@
 {-# LANGUAGE FlexibleContexts #-}
 {-# LANGUAGE GeneralizedNewtypeDeriving #-}
 {-# LANGUAGE DataKinds #-}
+{-# LANGUAGE RankNTypes #-}
 
 {-# OPTIONS_GHC -Wall #-}
 
@@ -147,7 +148,7 @@ data GPUExp t where
   Sub :: Num a => GPUExp a -> GPUExp a -> GPUExp a
   Mul :: Num a => GPUExp a -> GPUExp a -> GPUExp a
 
-  Twice :: GPUExp (a -> a) -> GPUExp a -> GPUExp a
+  -- Twice :: GPUExp (a -> a) -> GPUExp a -> GPUExp a
 
   FromEnum :: Enum a => GPUExp a -> GPUExp Int
   FromIntegral :: (Integral a, Num b) => GPUExp a -> GPUExp b
@@ -176,6 +177,49 @@ data GPUExp t where
 
   Construct :: a -> GPUExp a
   ConstructAp :: forall a b. (GPURep a) => GPUExp (a -> b) -> GPUExp a -> GPUExp b
+
+transformE :: (forall r. GPUExp r -> GPUExp r) -> GPUExp a -> GPUExp a
+transformE tr0 = tr
+  where
+    tr :: forall s. GPUExp s -> GPUExp s
+    tr = go . tr0
+
+    go :: forall s. GPUExp s -> GPUExp s
+    go (CaseExp x f) = CaseExp (tr x) (tr f)
+    go (SumMatchExp x y) = SumMatchExp (tr x) (tr y)
+    go (ProdMatchExp f) = ProdMatchExp (tr f)
+    go (NullaryMatch x) = NullaryMatch (tr x)
+    go (OneSumMatch p) = OneSumMatch (tr p)
+    go (OneProdMatch f) = OneProdMatch (tr f)
+    go EmptyMatch = EmptyMatch
+    go FalseExp = FalseExp
+    go TrueExp = TrueExp
+    go e@(Repped _) = e
+    go (Lam n x) = Lam n (tr x)
+    go e@(Var _) = e
+    go e@(Lit _) = e
+    go (Add x y) = Add (tr x) (tr y)
+    go (Sub x y) = Sub (tr x) (tr y)
+    go (Mul x y) = Mul (tr x) (tr y)
+    -- go (Twice f x) = Twice
[…]

### 2020-02-25 src/Test/PluginExample.hs

@@ -152,7 +152,7 @@ data ComplexPair where
   ComplexPair :: Complex Double
                  -> Complex Double
                  -> ComplexPair
-  deriving (Generic, Eq, Show)
+  deriving (Generic, Show)
 
 instance GPURep ComplexPair
 
@@ -174,6 +174,7 @@ mandelbrot_nextZ t =
   internalize (externalize
     (case t of
       (c, z) -> (z*z) + c))
+{-# NOINLINE mandelbrot_nextZ #-}
 
 mandelbrot_helper :: (Int, Complex Double, Complex Double) -> Maybe Int
 mandelbrot_helper t =
@@ -188,6 +189,7 @@ mandelbrot_helper t =
                 if (real*real) + (imag*imag) > 4
                   then Just iters
                   else mandelbrot_helper (iters+1, c, mandelbrot_nextZ (c, z))))
+{-# NOINLINE mandelbrot_helper #-}
 
 mandelbrot_point :: Complex Double -> Maybe Int
 mandelbrot_point c =
@@ -214,6 +216,7 @@ intListLength_helper p =
         case t of
           Nil -> acc
           Cons _ xs -> intListLength_helper (acc+1, xs)))
+{-# NOINLINE intListLength_helper #-}
 
 intListLength :: IntList -> Int
 intListLength t = internalize (externalize (intListLength_helper (0, t)))
@@ -226,21 +229,21 @@ intListSum_helper p =
          case t of
            Nil -> acc
            Cons x xs -> intListSum_helper (x+acc, xs)))
+{-# NOINLINE intListSum_helper #-}
 
 intListSum :: IntList -> Int
 intListSum t =
   internalize (externalize (intListSum_helper (0, t)))
 
 main :: IO ()
-main =
+main = do
   let intList = Cons 10 (Cons 100 (Cons 1000 (Cons 10000 Nil)))
-  in
   print (isEmpty Nil, isEmpty (Cons 1 Nil)
         ,intListLength intList
         ,intListSum intList
         )
-  -- print $ realSum (ComplexPair (2 :+ 100) (3 :+ 10000))
-  -- putStrLn mandelbrotTestAscii
+  print $ realSum (ComplexPair (2 :+ 100) (3 :+ 10000))
+  putStrLn mandelbrotTestAscii
 
 
 

### 2020-02-26 paper/paper.tex
Fixed typo in rewrite rule in paper
Fixed another typo in paper rewrite transformation
Modified wording (also, formatting)

@@
 is that of a \ttt{case} expression. Pattern matching is a convenient way to
-implement control flow structures and to view components of data structures, so
-it is frequently a desirable feature for many EDSLs.
+implement control flow structures and to inspect data structures, so it is
+frequently a desirable feature for many EDSLs.
 
@@
 
-In this EDSL, the type representing values in the expression language is \ttt{E t}
-and there are two basic functions which translate values to and from this
-expression language with the following type signatures:
+In this EDSL, the type representing values in the expression language is \ttt{E
+t} and there are two basic functions which translate values to and from this
+expression language with the following type signatures (omitting type class
+constraints for the moment):
 
@@
 
-[\ldots]
-
 \section{Representing Algebraic Datatypes}
@@
 \begin{lstlisting}
-runSumMatch . abs :: E (SumMatch a b) -> (a -> b)
+runSumMatch . abs
+  :: E (SumMatch a b) -> (a -> b)
 \end{lstlisting}
@@
       \begin{cases}
-        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body}}\\
-        \ttt{\{ K x0 ... xN $\rarr$ Done body; ... \}} & \text{otherwise}
+        \ttt{\{ K x0 ... xN $\rarr$ body0}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body0}}\\
+        \ttt{\{ K x0 ... xN $\rarr$ Done body0; $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}} \}$} & \text{otherwise}
       \end{cases}\\

### 2020-02-26 paper/paper.tex
Fixed typo

@@
 
-Given this code, \ttt{ERep ComplexPair \typeeq (Complex Double, Complex Double)}. Here
+Given this code, \ttt{ERepTy ComplexPair \typeeq (Complex Double, Complex Double)}. Here
 is an example that demonstrates why this is more useful than if it \textit{fully} deconstructed
@@
 
-If \ttt{ERep ComplexPair} were fully deconstructed into \ttt{((Double, Double),
+If \ttt{ERepTy ComplexPair} were fully deconstructed into \ttt{((Double, Double),
 (Double, Double))}, we would need a \ttt{Num} instance for \ttt{(Double,
@@
 It is important to note that we are still able to further deconstruct this type
-with further pattern matches, since \ttt{ERep (Complex Double) \typeeq (Double, Double)}:
+with further pattern matches, since \ttt{ERepTy (Complex Double) \typeeq (Double, Double)}:
 

### 2020-02-29 paper/paper.tex
Change the way brackets work in rewrites in paper

@@
 \newcommand{\rewrites}{\Longrightarrow}
+
+% TODO: Use these:
+\newcommand{\exprt}[1]{\expr{\ttt{#1}}}
+\newcommand{\tdots}{\ttt{...}}
+\newcommand{\exprdots}{\expr{\tdots}}
 
@@
       &\rewrites \begin{cases}
-        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ...
-\}}}$}$))}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ...
-\}}}\\
+        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}))}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
         \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
@@
 \begin{align*}
-  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
+  T'_{\ttt{f}}&\expr{\ttt{K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$}}\\
     &\rewrites
-        \ttt{\{ K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}}\\
-  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
+        \ttt{K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{ ...$_B$ }}$}\\
+  T'_{\ttt{f}}&\expr{\ttt{K x0 ... xN $\rarr$ body0; ... }}\\
     &\rewrites
       \begin{cases}
-        \ttt{\{ K x0 ... xN $\rarr$ body0}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body0}}\\
-        \ttt{\{ K x0 ... xN $\rarr$ Done body0; $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}} \}$} & \text{otherwise}
+        \ttt{K x0 ... xN $\rarr$ body0}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{ ... }} & \text{if \ttt{f} occurs free in \ttt{body0}}\\
+        \ttt{K x0 ... xN $\rarr$ Done body0; $T'_{\ttt{f}}\expr{\ttt{ ... }}$} & \text{otherwise}
       \end{cases}\\
-  T'_{\ttt{f}}&\expr{\ttt{\{ \}}} \rewrites \ttt{\{ \}}\\
+  T'_{\ttt{f}}&\expr{\varepsilon} \rewrites \varepsilon\;\;\; \text{(Note: $\varepsilon$ represents the empty string)}\\
   T_{\ttt{f}}&\expr{\ttt{case s of \{ ... \}}}\\
@@
       \begin{cases}
-        \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
+        \ttt{case s of \{ $T'_{\ttt{f}}\expr{\ttt{ ... }}$ \}} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
         \ttt{Done (case s of \{ ... \})} & \text{otherwise}
@@
   &C\expr{\ttt{case scrutinee of \{ ... \}}} \rewrites \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
-  &C\expr{\ttt{($\lambda$x $\rarr$ body)}} \rewrites A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
-  &C\expr{\ttt{(f x)}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
+  &C\expr{\ttt{$\lambda$x $\rarr$ body}} \rewrites A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
+  &C\expr{\ttt{f x}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
   &C\expr{\ttt{x}} \rewrites \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
@@
   &\\
-  &M\expr{\ttt{(x :: a)}} \rewrites
+  &M\expr{\ttt{x :: a}} \rewrites
     \begin{cases}

### 2020-03-01 paper/paper.tex
Rework some braces and add more ellipsis subscripts

@@
 \begin{align*}
-  T'_{\ttt{f}}&\expr{\ttt{K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$}}\\
+  T'_{\ttt{f}}&\expr{\ttt{K x0 ...$_U$ xN $\rarr$ case s of \{ ...$_V$ \}; ...$_W$}}\\
     &\rewrites
-        \ttt{K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{ ...$_B$ }}$}\\
-  T'_{\ttt{f}}&\expr{\ttt{K x0 ... xN $\rarr$ body0; ... }}\\
+        \ttt{K x0 ...$_U$ xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_V$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{ ...$_W$ }}$}\\
+  T'_{\ttt{f}}&\expr{\ttt{K x0 ...$_U$ xN $\rarr$ body0; ...$_V$ }}\\
     &\rewrites
       \begin{cases}
-        \ttt{K x0 ... xN $\rarr$ body0}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{ ... }} & \text{if \ttt{f} occurs free in \ttt{body0}}\\
-        \ttt{K x0 ... xN $\rarr$ Done body0; $T'_{\ttt{f}}\expr{\ttt{ ... }}$} & \text{otherwise}
+        \ttt{K x0 ...$_U$ xN $\rarr$ body0}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{ ...$_V$ }} & \text{if \ttt{f} occurs free in \ttt{body0}}\\
+        \ttt{K x0 ...$_U$ xN $\rarr$ Done body0; $T'_{\ttt{f}}\expr{\ttt{ ...$_V$ }}$} & \text{otherwise}
       \end{cases}\\
@@
   &C\expr{\ttt{x + y}} \rewrites \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
-  &C\expr{\ttt{case scrutinee of \{ ... \}}} \rewrites \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
+  &C\expr{\ttt{case scrutinee of \{ ... \}}} \rewrites \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{ ... }}\\
   &C\expr{\ttt{$\lambda$x $\rarr$ body}} \rewrites A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
@@
   &\\
-  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} \rewrites \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
-  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} \rewrites \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
-  &B\expr{\ttt{\{ \}}} \rewrites \ttt{EmptyMatch}\\
+  &B\expr{\ttt{K x0 ...$_U$ xN $\rarr$ body0; ...$_V$}} \rewrites \ttt{SumMatchExp }P\expr{\ttt{x0 ...$_U$ xN $\rarr$ body0}}\; B\expr{\ttt{ ...$_V$ }}\\
+  &B\expr{\ttt{K x0 ... xN $\rarr$ body}} \rewrites \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
+  &B\expr{\ttt{$\varepsilon$}} \rewrites \ttt{EmptyMatch}\\
   &\\

### 2020-03-01 src/CodeGen/C.hs
Starting work on C codegen backend

@@ -1,0 +1,127 @@
+{-# LANGUAGE ExistentialQuantification #-}
+{-# LANGUAGE FlexibleContexts #-}
+{-# LANGUAGE GADTs #-}
+
+module CodeGen.C
+  where
+
+import           Deep.Expr
+
+import           Data.Typeable
+
+import           Data.Semigroup
+import           Control.Monad.Writer
+import           Data.List
+
+type CCode = String
+type CName = String
+
+data SomeName = forall a. Typeable a => SomeName (Name a)
+
+data Closure a b
+  = Closure
+    { closure_fvs :: [SomeName] -- | Free variables
+    , closure_name :: Name a
+    , closure_body :: GPUExp b
+    }
+
+getVarCount :: GPUExp a -> Int
+getVarCount = getMax . execWriter . go
+  where
+    -- TODO: Abstract this traversal pattern
+    go :: forall s. GPUExp s -> Writer (Max Int) ()
+    go (Var (Name n)) = tell (Max n)
+
+    go (CaseExp x y) = go x *> go y
+    go (ProdMatchExp x) = go x
+    go (SumMatchExp x y) = go x *> go y
+    go (NullaryMatch x) = go x
+    go (OneSumMatch x) = go x
+    go (OneProdMatch x) = go x
+    go EmptyMatch = pure ()
+    go FalseExp = pure ()
+    go TrueExp = pure ()
+    go (Repped {}) = pure ()
+    go (Lam n x) = go x
+    go (Lit {}) = pure ()
+    go (Add x y) = go x *> go y
+    go (Sub x y) = go x *> go y
+    go (Mul x y) = go x *> go y
+    go (FromEnum x) = go x
+    go (FromIntegral x) = go x
+    go (Sqrt x) = go x
+    go (Equal x y) = go x *> go y
+    go (Lte x y) = go x *> go y
+    go (Gt x y) = go x *> go y
+    go (Not x) = go x
+    go (LeftExp x) = go x
+    go (RightExp x) = go x
+    go (PairExp x y) = go x *> go y
+    go (FstExp x) = go x
+    go (SndExp x) = go x
+    go (StepExp x) = go x
+    go (DoneExp x) = go x
+    go (IfThenElse x y z) = go x *> go y *> go z
+    go (TailRec x) = go x
+    go (Construct {}) = pure ()
+    go (ConstructAp x y) = go x *> go y
[…]

### 2020-03-08 src/Deep/Expr.hs
Some progress on rep-unrep elimination (and cast-float-over-app transformation, in the process)

@@ -358,7 +358,7 @@ class GPURep t where
   rep :: t -> GPUExp t
 
   default rep :: (GPURep (GPURepTy t)) => t -> GPUExp t
-  rep = ConstructRep . construct
+  rep x = ConstructRep (construct x)
   {-# INLINABLE rep #-}
 
   rep' :: t -> GPURepTy t

### 2020-03-08 src/Deep/GHCUtils.hs

@@ -706,7 +706,7 @@ isDictVar :: Var -> Bool
 isDictVar v =
   let str = occNameString (occName v)
   in
-  length str >= 2 && (take 2 str == "$f" || take 2 str == "$d" || take 2 str == "C:")
+  length str >= 2 && (take 2 str == "$f" || take 2 str == "$d" || take 2 str == "$w" || take 2 str == "C:")
 
 isDict :: CoreExpr -> Bool
 isDict (Var v) = isDictVar v
@@ -723,12 +723,63 @@ isDictNotClass :: DynFlags -> CoreExpr -> Bool
 isDictNotClass dflags e =
   isDict e && not (isClassVar e)
 
+isDictNotClassApp :: DynFlags -> CoreExpr -> Bool
+isDictNotClassApp dflags e@(App _ _) =
+  case collectArgs e of
+    (fn, _) -> let r = isDictNotClass dflags fn in trace ("isDictNotClassApp: " ++ showPpr dflags (r, fn))  r
+isDictNotClassApp dflags e = False
+
 -- | For use with Uniplate. Example:
 -- transform (applyWhen predicate fn) expr
 applyWhen :: (a -> Bool) -> (a -> a) -> a -> a
 applyWhen p f x
   | p x       = f x
   | otherwise = x
+
+-- | Given a function name, a transformation and a Core expression,
+-- transform the Core expression when it is an application where
+-- the function has the same name as the given function name.
+targetTheFn :: Id -> (CoreExpr -> CoreExpr) -> CoreExpr -> CoreExpr
+targetTheFn fn t e@(App _ _)
+  | (Var fn', args) <- collectArgs e
+  = --trace ("targetTheFn: " ++ show (occNameString (occName fn), occNameString (occName fn'))) $
+        if fn == fn'
+          then mkApps (t (Var fn')) args
+          else e
+targetTheFn fn t e = e
+
+
+targetTheFnMaybe :: Maybe Id -> (CoreExpr -> CoreExpr) -> CoreExpr -> CoreExpr
+targetTheFnMaybe Nothing   _ e = e
+targetTheFnMaybe (Just fn) t e = targetTheFn fn t e
+
+maybeTransform :: Maybe a -> (a -> b -> b) -> b -> b
+maybeTransform Nothing  _ y = y
+maybeTransform (Just x) f y = f x y
+
+-- | Target application expressions that have an argument which is an
[…]

### 2020-03-08 src/Plugin/MatchPlugin.hs

@@ -807,7 +807,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                 repDictExpr <- lift $ buildDictionaryT guts (mkTyConApp repTyCon [exprType expr])
                 repDictRepTy <- lift $ buildDictionaryT guts (mkTyConApp repTyCon [mkTyConApp repTyTyCon [exprType expr]])
                 repId <- lift $ findIdTH guts 'rep
-                -- constructFnId <- lift $ findIdTH guts 'construct
+                constructFnId <- lift $ findIdTH guts 'construct
 
                 let constr' = mkApps (Var constr) (tys ++ dicts)
                 let constr'Type = exprType constr'
@@ -823,12 +823,41 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 let repUnfolding = getUnfolding guts dflags repId
 
+                let constructUnfolding = getUnfolding guts dflags constructFnId
+                    (constructFn0, _) =
+                        betaReduceAll
+                          constructUnfolding
+                          [ Type (exprType expr)
+                          , repDictExpr
+                          ]
+                    constructFn1 = caseInline dflags $ Data.transform letNonRecSubst constructFn0
+                    (constructFnE, _) = collectArgs constructFn1
+
+                let constructFn =
+                        case constructFnE of
+                          Var f -> Just f
+                          _ -> Nothing
+
+
+                liftIO $ putStrLn $ "~~~~~~~~constructFn = " ++ showPpr dflags constructFn
+                -- liftIO $ putStrLn $ "~~~~~~~~constructFn1 = " ++ showPpr dflags constructFn
+
                 let (fn, restArgs) = betaReduceAll
                                        repUnfolding
                                          [ Type (exprType expr)
                                          , repDictExpr
                                          -- , (mkApps constr' args)
                                          ]
+
+                let repFn1 = caseInline dflags $ Data.transform letNonRecSubst fn
+                    (repFnE, _) = collectArgs repFn1
+
+                    repFn =
+                      case repFnE of
[…]

### 2020-03-08 paper/paper.tex
Add 'not for redistribution' notice and adjust flowchart label
Possible GHC bug (somewhere, a call to (ga .) is failing and I never call it in this project). Probably related to the cast-float-over-app transformationgp

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Title Goes Here}
+\title{Title Goes Here\\
+\large{Not For Redistribution}}
 
@@
   into this EDSL for pattern matching, extended with tail recursion,
-  lambdas and primitive operations (Section 7)
+  lambdas and primitive operations (Section 8)
 \item An implementation of a interpreter for the "standard" semantics of the EDSL (Section 6).
@@
 
+\begin{todo}
+  Cite "Rewriting a Shallow DSL using a GHC Compiler Extension"
+  (2017; Grebe, Young, Gill)
+\end{todo}
+
 \section{Representing lambdas}
@@
 
+\section{C Backend}
+\begin{todo}
+  Write this section
+\end{todo}
+
 \section{Core Plugin}
@@
 \ttt{a}), then the marking procedure ignores it and does not wrap it with
-\ttt{externalize} (see $M$ in Section 7.2).
+\ttt{externalize} (see $M$ in Section 8.2).
 
@@
 
-In the following flowchart, the names in parentheses in a node refers to the names of transformations from the following section (Section 7.2).
+In the following flowchart, the names in parentheses in a node refers to the names of transformations from the following section (Section 8.2).
 
@@
   \draw [arrow] (findmark) -- node[anchor=east] {Primitive} (handleprim);
-  \draw [arrow] (findmark) -- node[anchor=west] {Function application/non-prim var} (handleapp);
+  \draw [arrow] (findmark) -- node[anchor=west, yshift=0.1cm] {Function application/non-prim var} (handleapp);
   \draw [arrow] (findmark) -- node[anchor=east] {\ttt{case}} (handlecase);
@@
 \clearpage
+\section{Related Work}
+\begin{todo}
+  Cite and compare/contrast with pattern matching technique described in section
+  3.3 of "Unembedding Domain-Specific Languages" (2009; Atkey, Lindley, Yallop)
+\end{todo}
+
 \section{Citations and Bibliographies}

### 2020-03-08 src/Deep/GHCUtils.hs

@@ -822,26 +822,32 @@ simpleOptExpr' e = do
 
 castFloatAppEither :: DynFlags -> CoreExpr -> Either String CoreExpr
 castFloatAppEither dflags (App (Cast e1 co) e2) =
-  let (coA, coB) = decomposeFunCo (coercionRole co) co
-  in -- TODO: This seems too easy?
-  Right $ Cast (App e1 (Cast e2 coA)) coB
 
-    -- trace ("castFloatAppEither: co = " ++ showPpr dflags co) $
-    --    case co of
-    --         TyConAppCo _r t [c1, c2] ->
-    --             if isFunTyCon t
-    --               then trace "castFloatApp firing" $ return $ Cast (App e1 (Cast e2 (SymCo c1))) c2
-    --               else Left "caseFloatAppEither"
--- -- #if __GLASGOW_HASKELL__ > 710
--- --             ForAllCo{} -> Left "castFloatAppR: ForAllCo TODO"
--- -- #else
-    --         ForAllCo t _ c2 -> -- TODO: Does this work as expected with the newer GHC API?
-    --             case e2 of
-    --               Type x' -> trace "caseFloatApp forallco" $
-    --                 return (Cast (App e1 e2) (CoreSubst.substCo (CoreSubst.extendTvSubst emptySubst t x') c2))
-    --               _ -> Left "caseFloatAppEither"
--- -- #endif
-    --         _ -> trace ("caseFloatApp failed on: " ++ showPpr dflags co) $ Left "castFloatApp"
+    trace ("castFloatAppEither: co = " ++ showPpr dflags co) $
+       case co of
+
+            TyConAppCo _r t [c1, c2] ->
+                if isFunTyCon t
+                  then trace "castFloatApp firing" $ return $ Cast (App e1 (Cast e2 (SymCo c1))) c2
+                  else Left "caseFloatAppEither"
+
+-- #if __GLASGOW_HASKELL__ > 710
+--             ForAllCo{} -> Left "castFloatAppR: ForAllCo TODO"
+-- #else
+
+            ForAllCo t _ c2 -> -- TODO: Does this work as expected with the newer GHC API?
+                case e2 of
+                  Type x' -> trace "caseFloatApp forallco" $
+                    return (Cast (App e1 e2) (CoreSubst.substCo (CoreSubst.extendTvSubst emptySubst t x') c2))
+                  _ -> Left "caseFloatAppEither"
+
+-- #endif
+            _ ->
[…]

### 2020-03-08 src/Deep/GHCUtils.hs
Circumventing possible GHC bug and also rep-unrep elimination seems to happen more. However, we now get a coercion from-type mismatch in Core Lint

@@ -11,6 +11,8 @@ module Deep.GHCUtils where
 
 import           GhcPlugins hiding (freeVarsBind)
 
+import           Pair
+
 import           CoreSubst
 import           Unique
 import           PrelNames
@@ -826,27 +828,29 @@ castFloatAppEither dflags (App (Cast e1 co) e2) =
     trace ("castFloatAppEither: co = " ++ showPpr dflags co) $
        case co of
 
-            TyConAppCo _r t [c1, c2] ->
-                if isFunTyCon t
-                  then trace "castFloatApp firing" $ return $ Cast (App e1 (Cast e2 (SymCo c1))) c2
-                  else Left "caseFloatAppEither"
+            -- TyConAppCo _r t [c1, c2] ->
+            --     if isFunTyCon t
+            --       then trace "castFloatApp firing" $ return $ Cast (App e1 (Cast e2 (modifyRole (SymCo c1)))) (modifyRole c2)
+            --       else Left "caseFloatAppEither"
 
 -- #if __GLASGOW_HASKELL__ > 710
 --             ForAllCo{} -> Left "castFloatAppR: ForAllCo TODO"
 -- #else
 
-            ForAllCo t _ c2 -> -- TODO: Does this work as expected with the newer GHC API?
+            ForAllCo t kc c2 -> -- TODO: Does this work as expected with the newer GHC API?
                 case e2 of
-                  Type x' -> trace "caseFloatApp forallco" $
-                    return (Cast (App e1 e2) (CoreSubst.substCo (CoreSubst.extendTvSubst emptySubst t x') c2))
+                  Type x' -> trace ("caseFloatApp forallco: { " ++ showPpr dflags (t, kc, c2) ++ "\n}\n") $
+                    return (Cast (App e1 e2) (CoreSubst.substCo (CoreSubst.extendTvSubst emptySubst t x') (modifyRole c2)))
                   _ -> Left "caseFloatAppEither"
 
 -- #endif
             _ ->
-                let (coA, coB) = decomposeFunCo (coercionRole co) co
-                in -- TODO: This seems too easy?
-                Right $ Cast (App e1 e2) coA
+              case decomposeFunCo_maybe (coercionRole co) co of
[…]

### 2020-03-08 src/Deep/GHCUtils.hs
Passes Core Lint now

@@ -825,7 +825,7 @@ simpleOptExpr' e = do
 castFloatAppEither :: DynFlags -> CoreExpr -> Either String CoreExpr
 castFloatAppEither dflags (App (Cast e1 co) e2) =
 
-    trace ("castFloatAppEither: co = " ++ showPpr dflags co) $
+    --trace ("castFloatAppEither: co = " ++ showPpr dflags co) $
        case co of
 
             -- TyConAppCo _r t [c1, c2] ->
@@ -839,14 +839,15 @@ castFloatAppEither dflags (App (Cast e1 co) e2) =
 
             ForAllCo t kc c2 -> -- TODO: Does this work as expected with the newer GHC API?
                 case e2 of
-                  Type x' -> trace ("caseFloatApp forallco: { " ++ showPpr dflags (t, kc, c2) ++ "\n}\n") $
+                  Type x' -> --trace ("caseFloatApp forallco: { " ++ showPpr dflags (t, kc, c2) ++ "\n}\n") $
+                    -- return (Cast (App e1 e2) (CoreSubst.substCo (CoreSubst.extendTvSubst emptySubst t x') (modifyRole c2)))
                     return (Cast (App e1 e2) (CoreSubst.substCo (CoreSubst.extendTvSubst emptySubst t x') (modifyRole c2)))
                   _ -> Left "caseFloatAppEither"
 
 -- #endif
             _ ->
               case decomposeFunCo_maybe (coercionRole co) co of
-                Just (coA, coB) -> Right $ Cast (App e1 (Cast e2 (mkSymCo (modifyRole coA)))) (modifyRole coB)
+                Just (coA, coB) -> Right $ Cast (App e1 (Cast e2 (mkSymCo (modifyRole coA)))) (mkSymCo (modifyRole coB))
                 Nothing -> Left "castFloatAppEither: decomposeFunCo_maybe gave Nothing"
                 -- Right $ Cast (App e1 (Cast e2 coA)) coB
   where

### 2020-03-08 src/Plugin/MatchPlugin.hs

@@ -782,10 +782,12 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
         -- Handle constructors that are not builtins
         go expr@(_ :@ _)
-          | Just (tys, dicts, constr, args) <- getConstructorApp_maybe expr,
+          | Just (tys, dicts, constr, args0) <- getConstructorApp_maybe expr,
             Nothing <- builtin constr,
             not (hasExprTy expr),
             not (isDict (Var constr)) = do
+
+            let args = reverse args0  -- TODO: Why is this necessary?
 
             dflags <- getDynFlags
             constructRepId <- lift $ findIdTH guts 'ConstructRep
@@ -796,8 +798,8 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 liftIO $ putStrLn ""
                 liftIO $ putStrLn $ "Handling constructor: " ++ showPpr dflags constr
-                liftIO $ putStrLn $ "  ==> with args:      " ++ showPpr dflags args
-                liftIO $ putStrLn ""
+                -- liftIO $ putStrLn $ "  ==> with args:      " ++ showPpr dflags args
+                -- liftIO $ putStrLn ""
 
                 repTyCon <- lift $ findTyConTH guts ''GPURep
                 repTyTyCon <- lift $ findTyConTH guts ''GPURepTy
@@ -812,14 +814,9 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                 let constr' = mkApps (Var constr) (tys ++ dicts)
                 let constr'Type = exprType constr'
 
-                -- markedConstr' <- mark constr'
                 markedArgs <- mapM mark args
 
                 tc@(co, ty') <- lift $ computeRepTypeCo guts (exprType expr)
-
-                -- liftIO $ putStrLn ""
-                -- liftIO $ putStrLn $ "+++++++++ tc = " ++ showPpr dflags tc
-                -- liftIO $ putStrLn ""
 
                 let repUnfolding = getUnfolding guts dflags repId
 
@@ -839,9 +836,6 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                           _ -> Nothing
 
 
-                liftIO $ putStrLn $ "~~~~~~~~constructFn = " ++ showPpr dflags constructFn
-                -- liftIO $ putStrLn $ "~~~~~~~~constructFn1 = " ++ showPpr dflags constructFn
-
[…]

### 2020-03-08 src/Deep/Expr.hs
Fixed non-exhaustive pattern match in gpuAbsEnv. Now this example seems to run correctly with the newly modified Core transformation

@@ -15,7 +15,7 @@
 {-# LANGUAGE DeriveTraversable #-}
 {-# LANGUAGE StandaloneDeriving #-}
 
--- {-# OPTIONS_GHC -Wall #-}
+{-# OPTIONS_GHC -Wall #-}
 
 module Deep.Expr where
 
@@ -691,6 +691,9 @@ gpuAbsEnv env (Lam (name :: Name a) (body :: GPUExp b)) = \(arg :: a) ->
     in
 
     gpuAbsEnv (extendEnv env (name :=> rep arg)) $ go body
+
+gpuAbsEnv env (App f x) = gpuAbsEnv env f (gpuAbsEnv env x)
+gpuAbsEnv env (CharLit c) = c
 
 class Canonical t where
   type GenericOp t :: (* -> *) -> (* -> *) -> * -> *

### 2020-03-08 src/Plugin/MatchPlugin.hs
Remove all generation of 'Construct' constructor from Core plugin

@@ -299,11 +299,6 @@ transformExpr :: ModGuts -> Var -> Maybe Var -> [(Id, CoreExpr)] -> Expr Var -> 
 transformExpr guts currName recNameM primMap =
   {- transformApps guts <=< -} untilNothingM (transformExprMaybe guts currName recNameM primMap)
 
-constructExpr :: ModGuts -> Id -> MatchM CoreExpr
-constructExpr guts fId = do
-  constructDC <- lift $ findIdTH guts 'Construct
-
-  return (Var constructDC :@ Type (varType fId) :@ Var fId)
 
 -- XXX: The delineation marker probably has to be floated in (or maybe the
 -- transformation can just go through the AST at that marker without
@@ -633,10 +628,14 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
         go expr@(Lam v e) = abstractOver guts v e
 
         go expr@(Var v)
-          | not (isFunTy (varType v))
+          | not (isFunTy (varType v))  -- TODO: Can this be removed?
           , not (hasExprTy expr) = do
-              constructId <- lift $ findIdTH guts 'Construct
-              return (Var constructId :@ Type (varType v) :@ expr)
+              repId <- lift $ findIdTH guts 'rep
+
+              repTyCon <- lift $ findTyConTH guts ''GPURep
+              repDict <- lift $ buildDictionaryT guts (mkTyConApp repTyCon [varType v])
+
+              return (Var repId :@ Type (varType v) :@ repDict :@ expr)
 
         -- go expr@(Var v :@ Type t)
         --   | isConstructor v = do
@@ -1278,7 +1277,7 @@ transformTailRec guts recVar e0 = do
 
           return (Var f :@ Type newTy :@ newDict :@ x')
         else return expr
-    go1 ty _ e = return e
+    go1 ty lamV e = go2 ty lamV e
 
     go2 :: Type -> Var -> CoreExpr -> MatchM CoreExpr
     go2 ty lamV expr@(Var f :@ fTyArg :@ fDict :@ x) = do

### 2020-03-08 src/Test/PluginExample.hs

@@ -237,6 +237,9 @@ intListSum :: IntList -> Int
 intListSum t =
   internalize (externalize (intListSum_helper (0, t)))
 
+intListSumE :: IntList -> GPUExp Int
+intListSumE t = externalize (intListSum_helper (0, t))
+
 example_x :: Either Char Int
 example_x = Left 'a'
 {-# INLINE example_x #-}

### 2020-03-08 src/CodeGen/C.hs
intListSum example works all the way through 'regular Haskell' to generating correct C code

@@ -9,6 +9,8 @@
 {-# LANGUAGE TypeOperators #-}
 {-# LANGUAGE DefaultSignatures #-}
 
+{-# OPTIONS_GHC -Wall #-}
+
 module CodeGen.C
   where
 
@@ -488,6 +490,22 @@ genExp (App lamExp@(Lam (Name n) _) x) resultName = do
 
 genExp (App (TailRec lam) x) resultName = genExp (App lam x) resultName
 
+genExp (App _ _) _ = error "genExp: App"
+
+-- genExp (App f x) resultName = do
+--   rName <- freshCName
+--   xName <- freshCName
+
+--   rCode <- genExp f fName
+--   xCode <- genExp x xName
+
+--   callCode <- callClosure 
+
+--   return $ unlines
+--     [ "var_t " <> fName <> ";"
+--     , "var_t " <> xName <> ";"
+--     , 
+
 genExp (Lit x :: GPUExp t) resultName =
   return $ unlines
     [ resultName <> ".value = " <> "malloc(sizeof(" <> ty <> "));"
@@ -577,6 +595,21 @@ genExp (Mul x y) resultName = do
     , xCode
     , yCode
     , "MATH_OP(*, " <> resultName <> ", " <> xName <> ", " <> yName <> ");"
+    ]
+
+genExp (Add x y) resultName = do
+  xName <- freshCName
+  yName <- freshCName
+
+  xCode <- genExp x xName
+  yCode <- genExp y yName
+
+  return $ unlines
+    [ "var_t " <> xName <> ";"
+    , "var_t " <> yName <> ";"
+    , xCode
+    , yCode
+    , "MATH_OP(+, " <> resultName <> ", " <> xName <> ", " <> yName <> ");"
     ]
 
 genExp (Equal x y) resultName = do
@@ -677,6 +710,33 @@ genExp UnitExp resultName =
     , resultName <> ".value = 0;"
     ]
 
+genExp (CharLit c) resultName =
+  return $ unlines
+    [ resultName <> ".tag = EXPR_CHAR;"
+    , "*(char*)(" <> resultName <> ".value) = " <> show c <> ";"
+    ]
+
+genExp (Ord x) resultName = do
+  xName <- freshCName
+
+  xCode <- genExp x xName
+
+  return $ unlines
+    [ "var_t " <> xName <> ";"
+    , xCode
+    , ""
+    , resultName <> ".tag = EXPR_INT;"
[…]

### 2020-03-08 src/Test/PluginExample.hs

@@ -2,7 +2,7 @@
 {-# LANGUAGE TypeApplications #-}
 {-# LANGUAGE GADTs #-}
 
-{-# OPTIONS_GHC -fexpose-all-unfoldings -ddump-simpl -O0 -dcore-lint -dsuppress-all -dcmm-lint -dstg-lint -fcatch-bottoms -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -ddump-simpl -fexpose-all-unfoldings -O0 -dcore-lint -dsuppress-all -dcmm-lint -dstg-lint -fcatch-bottoms -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 
@@ -253,14 +253,14 @@ example =
 {-# NOINLINE example #-}
 
 main :: IO ()
-main = --putStrLn (genProgram cTest6)
-    do
+main = do
   let intList = Cons 10 (Cons 100 (Cons 1000 (Cons 10000 Nil)))
-  print (isEmpty Nil, isEmpty (Cons 1 Nil)
-        ,intListLength intList
-        ,intListSum intList
-        ,example
-        )
+  putStrLn (genProgram (intListSumE intList)) --putStrLn (genProgram cTest6)
+  -- print (isEmpty Nil, isEmpty (Cons 1 Nil)
+  --       ,intListLength intList
+  --       ,intListSum intList
+  --       ,example
+  --       )
   -- print $ realSum (ComplexPair (2 :+ 100) (3 :+ 10000))
   -- putStrLn mandelbrotTestAscii
 

### 2020-03-08 test.c

@@ -33,7 +33,7 @@ typedef struct closure_t {
   var_t (*fn)(var_t, struct closure_t*);
 } closure_t;
 
-var_t vars[2];
+var_t vars[23];
 
 #define MATH_OP(op, result, a, b)\
   do {\
@@ -62,12 +62,14 @@ bool isIterTag(var_type_tag tag) {
   return tag == EXPR_STEP || tag == EXPR_DONE;
 }
 
-var_t lam_0(var_t, struct closure_t*);
-var_t lam_1(var_t, struct closure_t*);
-var_t lam_2(var_t, struct closure_t*);
+var_t lam_19(var_t, struct closure_t*);
+var_t lam_21(var_t, struct closure_t*);
+var_t lam_20(var_t, struct closure_t*);
+var_t lam_23(var_t, struct closure_t*);
+var_t lam_22(var_t, struct closure_t*);
 
 
-var_t lam_0(var_t arg, struct closure_t* self) {
+var_t lam_19(var_t arg, struct closure_t* self) {
   var_t x1;
 x1.tag = EXPR_STEP;
 x1.value = malloc(sizeof(var_t));
@@ -82,7 +84,7 @@ x2 = x1;
 
 closure_t x3;
 x3.fv_env = malloc(sizeof(var_t)*1);
-x3.fn = &lam_2;
+x3.fn = &lam_20;
 x3.fv_env[0] = ((var_t*)x2.value)[0];
 
 
@@ -98,11 +100,11 @@ x1.value = x4.value;
   return x1;
 }
 
-var_t lam_1(var_t arg, struct closure_t* self) {
+var_t lam_21(var_t arg, struct closure_t* self) {
   var_t x5;
 closure_t* x6 = malloc(sizeof(closure_t));
 (*x6).fv_env = malloc(sizeof(var_t)*1);
-(*x6).fn = &lam_2;
+(*x6).fn = &lam_20;
 (*x6).fv_env[0] = arg;
 
 
@@ -113,125 +115,201 @@ x5.value = (void*)x6;
   return x5;
 }
 
-var_t lam_2(var_t arg, struct closure_t* self) {
+var_t lam_20(var_t arg, struct closure_t* self) {
   var_t x8;
 var_t x9;
-var_t x10;
-var_t x11;
-
-var_t x12;
[…]

### 2020-03-22 paper/paper.tex
Some updates to the paper

@@
 \title{Title Goes Here\\
-\large{Not For Redistribution}}
+\large{Not For Distribution}}
 
@@
 \begin{abstract}
-  Domain specific languages (DSLs) provide a powerful tool for to abstract over
+  Domain specific languages (DSLs) provide a powerful tool for abstracting over
   a class of problems. In this paper, we introduce a technique for representing
@@
 \item A representation of pattern matching in a Haskell EDSL (Section 3)
-\item Representations of tail recursion (Section 4) and lambdas (Section 5) in
-  the EDSL. We will see that lambdas become easier to implement once pattern
-  matching and tail recursion already have EDSL representations.
+\item A representation of lambdas (Section 5) in the EDSL, taking advantage of
+  the fact that both pattern matching and tail recursion are already captured in
+  the EDSL.
 \item A GHC Core plugin which transforms a subset of standard Haskell code
@@
 \item An implementation of a interpreter for the "standard" semantics of the EDSL (Section 6).
+\item An outline of a C backend for this EDSL (Section 7).
 \end{itemize}
@@
 abs :: E a -> a
+\end{lstlisting}
+
+Additionally, the following functions mark the parts of the code which the Core
+plugin will target:
+
+\begin{lstlisting}
+internalize :: E a -> a
+externalize :: a -> E a
 \end{lstlisting}
@@
   abs (CaseExp
-        (rep x)
+        (LeftExp 'a')
         (SumMatchExp (OneProdMatch
@@
 
+\begin{todo}
+  Look into \ttt{generic-sop} and compare then cite corresponding paper.
+\end{todo}
+
 Briefly, a algebraic type \ttt{T} with an automatically generated \ttt{ERep}
@@
 already exists for \ttt{Complex Double}. This is exactly what preserving this
-type information allows us to do.
+type information allows us to do. We can later use this preserved type
+information in backends, through the corresponding \ttt{Typeable} instances (for
+instance, to provide special support for arithmetic on complex numbers).
 
@@
            Cons x xs ->
-             intListSum_helper (x+acc, xs)))
+             intListSum_helper
+               (x+acc, xs)))
 
@@
 This information is brought into the \ttt{E} type via the constructor
-\ttt{Repped}. The \ttt{E} also contains constructors representing \ttt{Either}
+\ttt{ConstructRep}. The \ttt{E} also contains constructors representing \ttt{Either}
 and \ttt{(,)} values:
@@
   ...
-  Repped :: ERep t => ERepTy t -> E t
+  ConstructRep ::
+    (Typeable a, ERep a)
+    => E (ERepTy a) -> E a
 
@@
 \begin{lstlisting}
-class ERep t where
+class Typeable t => ERep t where
   type ERepTy t
 
+  construct :: t -> E (ERepTy t)
+
   rep :: t -> E t
-  rep = Repped . rep'
-
-  unrep :: ERepTy t -> t
+
+  default rep ::
+    (ERep (ERepTy t))
+    => t -> E t
+  rep x = ConstructRep (construct x)
+  {-# INLINABLE rep #-}
+
+  unrep' :: ERepTy t -> t
 
@@
 
-  unrep = id
+  unrep' = id
   rep'   = id
@@
   rep (x, y) = PairExp (rep x) (rep y)
-  unrep = id
+  unrep' = id
   rep'   = id
@@
 The overall structure of a \ttt{E (SumMatch a b)} value is a (heterogeneous)
-list of \ttt{E (ProdMatch x y)} values.  Each item of this list corresponds
+list of \ttt{E (ProdMatch x b)} values.  Each item of this list corresponds
 exactly to one branch in the original \ttt{case} match.
@@
 
-absEnv env (Repped x) = unrep x
+absEnv env (ConstructRep x) =
+  unrep' (absEnv env x)
 
@@
 
-In the following flowchart, the names in parentheses in a node refers to the names of transformations from the following section (Section 8.2).
+In the following flowchart, the names in parentheses in a node refers to the names of transformations from the following section (Section 8.2). Note that
+calls to \ttt{unrep :: E a -> a} are only used internally and will not exist in the final result of the transformation.
 
@@
     \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{($C$) Encode as EDSL primitive};
-    \node (handleapp) [process, right of=handleprim, xshift=1.6cm]{($C$) Wrap with \ttt{Construct} and \ttt{ConstructAp}};
+    \node (handleconstr) [process, right of=handleprim, xshift=1.6cm]{($C$) Wrap with \ttt{construct}};
 
     \node (marksubs1) [process, below of=handleprim]{($C, M$) Mark sub-expressions};
-    \node (marksubs2) [process, below of=handleapp]{($C, M$) Mark sub-expressions};
+    \node (marksubs2) [process, below of=handleconstr]{($C, M$) Mark sub-expressions and wrap with \ttt{unrep}};
+    \node (betareduce) [process, below of=marksubs2]{Beta reduce \ttt{(construct ...)} expression and simplify};
+    \node (repunrep) [process, below of=betareduce]{Replace expressions of the form \ttt{(rep (unrep x))} with \ttt{x}};
 
@@
   \draw [arrow] (findmark) -- node[anchor=east] {Primitive} (handleprim);
-  \draw [arrow] (findmark) -- node[anchor=west, yshift=0.1cm] {Function application/non-prim var} (handleapp);
+  \draw [arrow] (findmark) -- node[anchor=west, yshift=0.1cm] {Constructor/constructor application} (handleconstr);
   \draw [arrow] (findmark) -- node[anchor=east] {\ttt{case}} (handlecase);
@@
 
-  \draw [arrow] (handleapp) -- (marksubs2);
+  \draw [arrow] (handleconstr) -- (marksubs2);
 
@@
 
+  \draw [arrow] (marksubs2) -- (betareduce);
+  \draw [arrow] (betareduce) -- (repunrep);
+
   \draw [arrow] (marksubs1) -- (endstep);
-  \draw [arrow] (marksubs2) -- (endstep);
+  \draw [arrow] (repunrep) -- (endstep);
   \draw [arrow] (markbody) -- (endstep);
@@
 \subsection{Syntactic Transformations of the Core Plugin}
-\begin{todo}
-  Should we use Oxford brackets here? Maybe a bit unusual, but we are operating syntactically and mapping from \ttt{a} to \ttt{E a} (could \ttt{E a} be considered an "intermediary" semantic domain?).
-\end{todo}
-\vspace*{-0.6cm}
+
+% \begin{todo}
+%   Should we use Oxford brackets here?
+Maybe a bit unusual, but we are operating syntactically and mapping from \ttt{a} to \ttt{E a} (could \ttt{E a} be considered an "intermediary" semantic domain?).
+% \end{todo}
+% \vspace*{-0.6cm}
+
 \begin{align*}
@@
   &C\expr{\ttt{$\lambda$x $\rarr$ body}} \rewrites A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
-  &C\expr{\ttt{f x}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
-  &C\expr{\ttt{x}} \rewrites \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
+  &C\expr{\ttt{K x0 ... xN }} \rewrites \ttt{construct (K (unrep $M\expr{\ttt{x0}}$) ... (unrep $M\expr{\ttt{xN}}$))}\;\;\;\;\text{(Where \ttt{K} is a constructor)}\\
+  % &C\expr{\ttt{f x}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
+  % &C\expr{\ttt{x}} \rewrites \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
   &\\

### 2020-04-06 paper/paper.tex
Added a bit about the tail recursion transformation

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Title Goes Here\\
+\title{First Class Pattern Matching in a Deeply Embedded Domain Specific Language\\
 \large{Not For Distribution}}
@@
 the lambda.
+\begin{todo}
+  Add brief summary of relationship to lambdas
+\end{todo}
 
@@
 
-In this paper, the data constructors of the \ttt{E} type will be described
-incrementally as they are needed.
+The \ttt{E} type is the type of expressions in the EDSL.  In this paper, the
+data constructors of this type will be described incrementally as they
+are needed.
 
@@
 nested) pair type. For example, \ttt{E (ProdMatch (a, (b, c)) r)} is
-equivalent to \ttt{E (a -> b -> c -> r)}.
+equivalent to \ttt{E (a -> E (b -> E (c -> E r)))}.
 
@@
 
+Tail recursion is given a direct-style representation using a simple sum type.
+
+Consider a tail recursive function of the type \ttt{f :: a -> b}. Each recursive
+call can be seen as a simple "update" of the values of the arguments, since
+these calls are all in tail position. This is why tail recursive functions
+can be easily compiled to simple loops or conditional jumps.
+
+We can take advantage of this view by transforming a tail recursive function
+\ttt{f :: a -> b} into a function \ttt{f' :: a -> Iter b a}. \ttt{Iter b a} is a
+type which can either correspond to an argument "update" (in the sense mentioned
+previously) or a final result. This type is implemented as a sum of the types
+\ttt{a} and \ttt{b}.
+
 \begin{lstlisting}
@@
   deriving (Functor, Generic)
-
+\end{lstlisting}
+
+To use the new \ttt{f'} function, we repeatedly call it until it gives a
+\ttt{Done} value. If it gives a \ttt{Step} value, we pass the the value wrapped
+in the \ttt{Step} back into \ttt{f'} and continue.
+
+The function \ttt{runIter} provides the standard semantics for executing such a
+function representing a tail recursive function:
+
+\begin{lstlisting}
 runIter :: (ERep a, ERep b)
@@
 
+This technique can be contrasted with the more traditional trampolining
+technique for implementing tail recursion. Conventional trampolining uses a
+sum type of the result type and a thunk with the code necessary to continue
+execution.
+
+\begin{todo}
+  Cite "Trampolined Style" (S. E. Gantz, D. P. Friedman, M. Ward; 1999)
+\end{todo}
+
+In the technique presented here, we do not need to allocate closures and we
+actually do not to use higher-order functions at all.
+
 In the \ttt{E} type, this is used to represent tail recursion by
@@
 \section{Recovering standard semantics for pattern matches}
+
+The standard semantics for the EDSL is given by \ttt{abs}.
 
@@
 
-So, \ttt{externalize x} marks the expression \ttt{x}.
+For example, \ttt{externalize x} marks the expression \ttt{x}.
 

### 2020-04-06 paper/paper.tex
Restructured section 6 a bit and started to flesh it out

@@
 
-In the technique presented here, we do not need to allocate closures and we
-actually do not to use higher-order functions at all.
+In the technique presented here, we do not need to allocate thunks or closures.
+We actually do not to use higher-order functions at all in this tail recursion
+representation.
 
@@
 
-The standard semantics for the EDSL is given by \ttt{abs}.
+The standard semantics for the EDSL is given by \ttt{abs}. Two helper
+functions, \ttt{sumMatchAbs} and \ttt{prodMatchAbs}, are used to provide
+the standard semantics for matches on sum types and matches on product
+types, respectively. These helper functions are based on the mechanism
+described in section 3.4.
 
@@
 ...
-
-absEnv :: forall t. Env -> E t -> t
-absEnv env (CaseExp x f) =
-  sumMatchAbs f (absEnv env x)
-
-absEnv env (TailRec f) = \ x ->
-  case absEnv env f x of
-    Step x' -> absEnv env (TailRec f) x'
-    Done r  -> r
-
-absEnv env (Var v) =
-  case envLookup env v of
-    Just x -> absEnv env
-    Nothing ->
-      error
-        ("No binding for name "
-          ++ show v)
-
-absEnv env (Lam (name :: Name a)
-                (body :: E b)) =
-  \ (arg :: a) ->
-    let go :: forall x. E x -> E x
-        go expr@(Var name2 :: E a') =
-          case namesEq name name2 of
-            Just Refl -> rep arg
-            _ ->
-              case envLookup env name2 of
-                Just v  -> v
-                Nothing -> expr
-
-        -- [... traverse rest of expression
-        --      in go ...]
-    in
-    absEnv (extendEnv env
-                      (name :=> rep arg))
-           (go body)
-
-absEnv env (LeftExp x) =
-  Left (absEnv env x)
-
-absEnv env (RightExp y) =
-  Right (absEnv env y)
-
-absEnv env (PairExp x y) =
-  (absEnv env x, absEnv env y)
-
-absEnv env (ConstructRep x) =
-  unrep' (absEnv env x)
-
-...
-\end{lstlisting}
+\end{lstlisting}
+
+See Appendix A for more implementation details of \ttt{absEnv}.
 
@@
 
+\clearpage
+
+\appendix
+\section{Additional Code for Standard Haskell Semantics Implementation}
+\begin{lstlisting}
+absEnv :: forall t. Env -> E t -> t
+absEnv env (CaseExp x f) =
+  sumMatchAbs f (absEnv env x)
+
+absEnv env (TailRec f) = \ x ->
+  case absEnv env f x of
+    Step x' -> absEnv env (TailRec f) x'
+    Done r  -> r
+
+absEnv env (Var v) =
+  case envLookup env v of
+    Just x -> absEnv env
+    Nothing ->
+      error
+        ("No binding for name "
+          ++ show v)
+
+absEnv env (Lam (name :: Name a)
+                (body :: E b)) =
+  \ (arg :: a) ->
+    let go :: forall x. E x -> E x
+        go expr@(Var name2 :: E a') =
+          case namesEq name name2 of
+            Just Refl -> rep arg
+            _ ->
+              case envLookup env name2 of
+                Just v  -> v
+                Nothing -> expr
+
+        -- [... traverse rest of expression
+        --      in go ...]
+    in
+    absEnv (extendEnv env
+                      (name :=> rep arg))
+           (go body)
+
+absEnv env (LeftExp x) =
+  Left (absEnv env x)
+
+absEnv env (RightExp y) =
+  Right (absEnv env y)
+
+absEnv env (PairExp x y) =
+  (absEnv env x, absEnv env y)
+
+absEnv env (ConstructRep x) =
+  unrep' (absEnv env x)
+
+...
+\end{lstlisting}
+
+\clearpage
+
 \section{Citations and Bibliographies}

### 2020-04-21 src/CodeGen/C.hs
Work on 'semantic tags'

@@ -39,6 +39,9 @@ import           Data.Ord
 
 import           Data.Complex
 
+import           CodeGen.C.Types
+import           CodeGen.C.Helper
+
 import Debug.Trace
 
 nub' :: Ord a => [a] -> [a]
@@ -50,8 +53,6 @@ deleteAll x (y:ys)
   | y == x    =     deleteAll x ys
   | otherwise = y : deleteAll x ys
 
-type CCode = String
-type CName = String
 
 data SomeName = forall a. Typeable a => SomeName (Name a)
 
@@ -364,13 +365,18 @@ prelude varCount =
     , ", EXPR_STEP"
     , ", EXPR_DONE"
     , ", EXPR_UNBOXED"
-    , ", EXPR_COMPLEX     // Complex numbers"
     , "} var_type_tag;"
+    , ""
+    , "typedef enum semantic_type_tag {"
+    , "  NO_SEMANTIC_TAG"
+    , ", EXPR_COMPLEX     // Complex numbers"
+    , "} semantic_type_tag;"
     , ""
     , "struct closure_t;"
     , ""
     , "typedef struct var_t {"
     , "  var_type_tag tag;"
+    , "  semantic_type_tag semantic_tag;  // For optional additional information about the \"un-deconstructed\" type (for instance, it could be the case that tag=EXPR_PAIR, but specific_tag=EXPR_COMPLEX)"
     , "  void* value;"
     , "} var_t;"
     , ""
@@ -392,19 +398,10 @@ prelude varCount =
     , "#define GET_COMPARE_OP_GT >"
     , "#define GET_COMPARE_OP_EQUAL =="
     , ""
-    , "  // Remove any additional wrappers (for instance, a EXPR_COMPLEX wraps a EXPR_PAIR"
-    , "#define UNWRAP(result, x)\\"
-    , "  do {\\"
-    , "    if ((x).tag == EXPR_COMPLEX) {\\"
-    , "      result = *((var_t*)((x).value));\\"
-    , "    } else {\\"
-    , "      result = x;\\"
-    , "    }\\"
-    , "  } while (0);"
-    , ""
     , "#define MATH_OP(op, result, a, b)\\"
     , "  do {\\"
     , "    assert((a).tag == (b).tag);\\"
[…]

### 2020-04-21 src/CodeGen/C/Helper.hs

@@ -1,0 +1,44 @@
+module CodeGen.C.Helper
+  where
+
+import           CodeGen.C.Types
+
+import           Data.List
+
+stmt :: CCode -> CCode
+stmt = (<> ";")
+
+cCall :: CName -> [CCode] -> CCode
+cCall fnName args =
+  fnName <> "(" <> intercalate ", " args <> ")"
+
+derefPtr :: CCode -> CCode
+derefPtr x = "*(" <> x <> ")"
+
+cCast :: CCode -> CCode -> CCode
+cCast toType x = "((" <> toType <> ")" <> x <> ")"
+
+cIndex :: CCode -> Int -> CCode
+cIndex x i = "(" <> x <> ")[" <> show i <> "]"
+
+(!) :: CCode -> Int -> CCode
+(!) = cIndex
+
+(=:) :: CCode -> CCode -> CCode
+x =: y = stmt (x <> " = " <> y)
+
+cDecl :: CName -> CName -> CCode
+cDecl cType v = stmt (cType <> " " <> v)
+
+(#) :: CCode -> CName -> CCode
+x # fieldName = x <> "." <> fieldName
+
+macroDef :: CName -> [CName] -> [CCode] -> CCode
+macroDef name args theLines =
+  unlines
+    [ "#define " <> name <> "(" <> intercalate ", " args <> ")\\"
+    , "  do {\\"
+    , unlines . map ("    "++) . map (++"\\") $ theLines
+    , "  } while (0);"
+    ]
+

### 2020-04-21 src/CodeGen/C/Types.hs

@@ -1,0 +1,5 @@
+module CodeGen.C.Types where
+
+type CCode = String
+type CName = String
+

### 2020-04-21 paper/paper.tex
Some advice for readers
Formatting

@@
 
+Readers who are primarily interested only in the encoding of pattern matching itself can direct
+most of their attention to sections 2, 3 and 6.
+
 \subsection{Overview}
@@
 The \ttt{Name t} type represents a lambda variable identifier together with its
-type \ttt{t} (Note that the \ttt{ScopedTypeVariables} extension is enabled.):
+type \ttt{t} (Note that the \ttt{ScopedTypeVariables} extension is enabled):
 

### 2020-04-21 src/Deep/GHCUtils.hs
In-progress bug fix for constructor handling

@@ -78,6 +78,8 @@ import           Data.List
 import           Data.Generics.Uniplate.Operations
 import qualified Data.Generics.Uniplate.DataOnly as Data
 
+import           Data.Data (Data)
+
 import Debug.Trace
 
 -- | Build a dictionary for the given
@@ -693,8 +695,14 @@ getUnfolding guts dflags i =
 
 -- | Transform scrutinee of a case. If not a 'case', leave it alone.
 onScrutinee :: (CoreExpr -> CoreExpr) -> CoreExpr -> CoreExpr
-onScrutinee f (Case scrutinee wild ty alts) = Case (f scrutinee) wild ty alts
-onScrutinee _ e = e
+onScrutinee f = maybeApply (onScrutinee_maybe f)
+-- onScrutinee f (Case scrutinee wild ty alts) = Case (f scrutinee) wild ty alts
+-- onScrutinee _ e = e
+
+onScrutinee_maybe :: (CoreExpr -> CoreExpr) -> CoreExpr -> Maybe CoreExpr
+onScrutinee_maybe f (Case scrutinee wild ty alts) = Just $ Case (f scrutinee) wild ty alts
+onScrutinee_maybe _ e = Nothing
+
 
 -- | Transform the function position of a collection of Apps:
 -- f :@ x0 :@ x1 :@ ... :@ xN    ==>    (t f) :@ x0 :@ x1 :@ ... :@ xN
@@ -704,11 +712,18 @@ onAppFun t e@(App _ _) =
   in mkApps (t f) args
 onAppFun t e = e
 
+onAppFunId :: (Id -> CoreExpr) -> CoreExpr -> CoreExpr
+onAppFunId t e = onAppFun go e
+  where
+    go (Var f) = t f
+    go _ = e
+onAppFunId _ e = e
+
 isDictVar :: Var -> Bool
 isDictVar v =
   let str = occNameString (occName v)
   in
-  length str >= 2 && (take 2 str == "$f" || take 2 str == "$d" || take 2 str == "$w" || take 2 str == "C:")
+  length str >= 2 && (take 2 str == "$f" || take 2 str == "$d" || take 2 str == "$w" || take 2 str == "$c" || take 2 str == "C:")
 
 isDict :: CoreExpr -> Bool
 isDict (Var v) = isDictVar v
@@ -742,22 +757,52 @@ applyWhen p f x
 -- transform the Core expression when it is an application where
[…]

### 2020-04-21 src/Plugin/MatchPlugin.hs

@@ -672,7 +672,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                typeType <- lift $ findTypeTH guts ''Kind.Type
                typeableDictA <- lift $ buildDictionaryT guts (mkTyConApp typeableTyCon [typeType, tyA'])
                typeableDictB <- lift $ buildDictionaryT guts (mkTyConApp typeableTyCon [typeType, tyB'])
-              
+
                fromIntegralConstrId <- lift $ findIdTH guts 'FromIntegral
 
                return (Var fromIntegralConstrId :@ tyA :@ tyB :@ typeableDictA :@ typeableDictB :@ dictA :@ dictB :@ markedX)
@@ -787,7 +787,9 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 let repUnfolding = getUnfolding guts dflags repId
 
-                let constructUnfolding = getUnfolding guts dflags constructFnId
+                let getUnfolding' = getUnfolding guts dflags
+
+                let constructUnfolding = getUnfolding' constructFnId
                     (constructFn0, _) =
                         betaReduceAll
                           constructUnfolding
@@ -802,6 +804,8 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                           Var f -> Just f
                           _ -> Nothing
 
+
+                liftIO $ putStrLn $ "repUnfolding = {" ++ showPpr dflags repUnfolding ++ "}"
 
                 let (fn, restArgs) = betaReduceAll
                                        repUnfolding
@@ -831,13 +835,18 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 let (Var dictFn, dictArgs) = collectArgs e''
 
-                let simplE' =  Data.transform betaReduce $ Data.transform (targetTheFnMaybe constructFn (Data.transform (tryUnfoldAndReduceDict guts dflags) . unfoldAndReduceDict guts dflags)) $ Data.transform letNonRecSubst $ Data.transform betaReduce $
+                let reduceConstruct arg
[…]

### 2020-04-21 src/Test/PluginExample.hs

@@ -8,7 +8,7 @@
 -- {-# LANGUAGE StandaloneDeriving #-}
 -- {-# LANGUAGE FlexibleInstances #-}
 
-{-# OPTIONS_GHC -O0 -fno-worker-wrapper -fno-strictness -fno-unbox-small-strict-fields -fno-unbox-strict-fields -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-coercions -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 
@@ -137,20 +137,20 @@ example7_ x = IntPair 1 2
 --           then y
 --           else x))
 
--- example8 :: IntPair -> Int
--- example8 p =
---   internalize (externalize
---     (case p of
---       IntPair x y -> y))
+example8 :: IntPair -> Int
+example8 p =
+  internalize (externalize
+    (case p of
+      IntPair x y -> y))
 
--- example9 :: IntPair -> Int
--- example9 p =
---   internalize (externalize
---     (case p of
---       IntPair x y ->
---         if x == 0
---           then example8 p
---           else example9 (IntPair (x-1) (x*y))))
+example9 :: IntPair -> Int
+example9 p =
+  internalize (externalize
+    (case p of
+      IntPair x y ->
+        if x == 0
+          then example8 p
+          else example9 (IntPair (x-1) (x*y))))
 
 -- floatTest :: Float
 -- floatTest =
@@ -289,14 +289,18 @@ doubleListSumE t = externalize (doubleListSum_helper (0, t))
 
 -- deriving instance Generic1 FinBitTree
 
-nonterm_test0 :: Int -> Int
-nonterm_test0 0 = 0
-nonterm_test0 n = (n-1)
+-- nonterm_test0 :: Int -> Int
+-- nonterm_test0 0 = 0
+-- nonterm_test0 n = nonterm_test1 (n-1)
 
-nonterm_test :: Int -> Int
-nonterm_test x =
-  internalize (externalize
-    (nonterm_test0 4))
+-- nonterm_test1 :: Int -> Int
+-- nonterm_test1 0 = 1
+-- nonterm_test1 n = nonterm_test0 (n-1)
+
+-- nonterm_test :: Int -> Int
[…]

### 2020-04-22 src/Deep/GHCUtils.hs
Seems to almost fix the issue, except for coercion related problems

@@ -695,13 +695,20 @@ getUnfolding guts dflags i =
 
 -- | Transform scrutinee of a case. If not a 'case', leave it alone.
 onScrutinee :: (CoreExpr -> CoreExpr) -> CoreExpr -> CoreExpr
-onScrutinee f = maybeApply (onScrutinee_maybe f)
+onScrutinee f = maybeApply (onScrutinee_maybe (Just . f))
 -- onScrutinee f (Case scrutinee wild ty alts) = Case (f scrutinee) wild ty alts
 -- onScrutinee _ e = e
 
-onScrutinee_maybe :: (CoreExpr -> CoreExpr) -> CoreExpr -> Maybe CoreExpr
-onScrutinee_maybe f (Case scrutinee wild ty alts) = Just $ Case (f scrutinee) wild ty alts
+onScrutinee_maybe :: (CoreExpr -> Maybe CoreExpr) -> CoreExpr -> Maybe CoreExpr
+onScrutinee_maybe f (Case scrutinee wild ty alts) = do
+    s' <- f scrutinee
+    Just $ Case s' wild ty alts
 onScrutinee_maybe _ e = Nothing
+
+whenVarHasClassName_maybe :: Name -> (CoreExpr -> CoreExpr) -> CoreExpr -> Maybe CoreExpr
+whenVarHasClassName_maybe cName f e@(Var v)
+  | Just c <- isClassOpId_maybe v, className c == cName = Just (f e)
+whenVarHasClassName_maybe _ _ e = Nothing
 
 
 -- | Transform the function position of a collection of Apps:
@@ -785,6 +792,24 @@ targetTheFnId fn t e =
   case targetTheFnId_maybe fn t e of
     Just e' -> e'
     _ -> e
+
+targetTheFnApp_maybe :: Id -> (CoreExpr -> CoreExpr) -> CoreExpr -> Maybe CoreExpr
+targetTheFnApp_maybe fn t e@(App _ _)
+  | (Var fn', _) <- collectArgs e, fn' == fn =
+      Just (t e)
+targetTheFnApp_maybe _ _ _ = Nothing
+
+targetTheFnApp :: Id -> (CoreExpr -> CoreExpr) -> CoreExpr -> CoreExpr
+targetTheFnApp fn t e =
+  case targetTheFnApp_maybe fn t e of
+    Just e' -> e'
+    _ -> e
+
+-- | Apply to f in (((...((f x0 x1 ... xN) y0) ...) yM) y(M-1))
[…]

### 2020-04-22 src/Plugin/MatchPlugin.hs

@@ -867,7 +867,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                 let newExpr''2
                       = -- Data.transform betaReduce $ Data.transform (maybeApply (fmap (Data.transform (caseInline dflags) . betaReduce . Data.transform (caseInline dflags)) . unfoldTheFn_maybe guts dflags constructFnId))
                         Data.transform betaReduce $ Data.transform (tryCastFloatApp dflags) $ caseInlineT dflags $ Data.transform betaReduce $ Data.transform (maybeApply (fmap betaReduce . unfoldTheFn_maybe guts dflags constructFnId))
-                        $ Data.transform betaReduce $ targetLastAppArg (Data.transform letNonRecSubst . Data.transform caseInlineDefault . betaReduce . onAppFun (tryUnfoldAndReduceDict guts dflags)) $ Data.transform (maybeApply (fmap (Data.transform letNonRecSubst . tryUnfoldAndReduceDict guts dflags . caseInlineT dflags) . onScrutinee_maybe (tryUnfoldAndReduceDict guts dflags)))
+                        $ Data.transform betaReduce $ targetLastAppArg (Data.transform letNonRecSubst . Data.transform caseInlineDefault . betaReduce . onAppFun (tryUnfoldAndReduceDict guts dflags)) $ Data.transform (maybeApply (fmap (Data.transform letNonRecSubst . tryUnfoldAndReduceDict guts dflags . caseInlineT dflags) . onScrutinee_maybe (unfoldAndReduceDict_maybe guts dflags)))
                         $ targetAppArgs betaReduce
                         $ betaReduce
                         $ targetAppArgs betaReduce
@@ -877,13 +877,65 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 fromFnId <- lift $ findIdTH guts 'from
 
-                let newExpr''
-                        = --Data.transform (targetTheFn fromFnId (tryUnfoldAndReduceDict guts dflags)) --(unfoldTheFn guts dflags fromFnId)
-                         -- (caseInline dflags)
-                            Data.transform (maybeApply (fmap (onAppFun (tryUnfoldAndReduceDict guts dflags)) . onScrutinee_maybe (tryUnfoldAndReduceDict guts dflags)))
-                          $ Data.transform betaReduce
[…]

### 2020-04-23 paper/paper.tex
EDSL for C code gen in paper appendix

@@
 \end{todo}
+
+The C backend makes use of a small EDSL for generating C code. See
+Appendix B for implementation details of this ESDL.
 
@@
 
+\section{EDSL for C code generation}
+\begin{lstlisting}
+type CCode = String
+type CName = String
+
+stmt :: CCode -> CCode
+stmt = (<> ";")
+
+cCall :: CName -> [CCode] -> CCode
+cCall fnName args =
+  fnName <> "(" <> intercalate ", " args <> ")"
+
+derefPtr :: CCode -> CCode
+derefPtr x = "*(" <> x <> ")"
+
+cCast :: CCode -> CCode -> CCode
+cCast toType x =
+  "((" <> toType <> ")" <> x <> ")"
+
+cIndex :: CCode -> Int -> CCode
+cIndex x i = "(" <> x <> ")[" <> show i <> "]"
+
+(!) :: CCode -> Int -> CCode
+(!) = cIndex
+
+(=:) :: CCode -> CCode -> CCode
+x =: y = stmt (x <> " = " <> y)
+
+cDecl :: CName -> CName -> CCode
+cDecl cType v = stmt (cType <> " " <> v)
+
+(#) :: CCode -> CName -> CCode
+x # fieldName = x <> "." <> fieldName
+
+macroDef :: CName -> [CName] -> [CCode] -> CCode
+macroDef name args theLines =
+  unlines
+    [ "#define " <> name
+              <> "(" <>
+                intercalate ", " args
+              <> ")\\"
+    , "  do {\\"
+    , unlines
+        . map ("    "++)
+        . map (++"\\")
+        $ theLines
+    , "  } while (0);"
+    ]
+\end{lstlisting}
+
 \clearpage

### 2020-04-23 paper/paper.tex
Some C backend details

@@
 \begin{todo}
-  Write this section
+  Expand this section
 \end{todo}
@@
 Appendix B for implementation details of this ESDL.
+
+This backend is based on a globally unique name generator. Similar to
+SSA form, a new C variable is created for each Haskell subexpression.
+It differs from SSA, however, in that these variables can be assigned multiple
+times. In particular, a variable can be assigned in multiple branches of
+an \ttt{if} statement or inside of a \ttt{while} loop. This simplifies
+the implementation, as $\phi$ nodes are not needed.
+
+The following \ttt{struct}s provide the C representation for Haskell expressions:
+
+\begin{lstlisting}
+typedef enum var_type_tag {
+  EXPR_INT
+, EXPR_FLOAT
+, EXPR_DOUBLE
+, EXPR_CHAR
+, EXPR_BOOL
+, EXPR_CLOSURE
+, EXPR_EITHER_LEFT
+, EXPR_EITHER_RIGHT
+, EXPR_PAIR
+, EXPR_UNIT
+, EXPR_STEP
+, EXPR_DONE
+, EXPR_UNBOXED
+} var_type_tag;
+
+typedef enum semantic_type_tag {
+  NO_SEMANTIC_TAG
+, EXPR_COMPLEX     // Complex numbers
+} semantic_type_tag;
+
+struct closure_t;
+
+typedef struct var_t {
+  var_type_tag tag;
+  semantic_type_tag semantic_tag;
+  void* value;
+} var_t;
+
+typedef struct closure_t {
+  var_t* fv_env;
+  var_t (*fn)(var_t, struct closure_t*);
+} closure_t;
+\end{lstlisting}
+
+The \ttt{semantic\_tag} allows type information preserved by the \ttt{Typeable} constraints
+in the EDSL to be brought over to the generated C code. In our implementation, this is
+used to distinguish between pairs of \ttt{Double}s and \ttt{Complex Double} values. In the
+case of a \ttt{Complex Double}, the \ttt{var\_t} would have a \ttt{tag} field with the value
+\ttt{EXPR\_PAIR} and a \ttt{semantic\_tag} field with the value \ttt{EXPR\_COMPLEX}.
 

### 2020-04-23 src/Deep/GHCUtils.hs
Some progress on moving coercions properly

@@ -988,6 +988,20 @@ tryCastFloatApp dflags e =
     Left _ -> e
     Right e' -> e'
 
+combineCasts_maybe :: CoreExpr -> Maybe CoreExpr
+combineCasts_maybe (Cast (Cast e coA) coB) = Just $ Cast e (mkTransCo coA coB)
+combineCasts_maybe _ = Nothing
+
+combineCasts :: CoreExpr -> CoreExpr
+combineCasts e =
+  case combineCasts_maybe e of
+    Just e' -> e'
+    _ -> e
+
+targetCastM :: Applicative m => (CoreExpr -> m CoreExpr) -> CoreExpr -> m CoreExpr
+targetCastM t e@(Cast _ _) = t e
+targetCastM _ e = pure e
+
 -- | Build a CoreExpr for a DFunUnfolding
 dFunExpr :: Unfolding -> CoreExpr
 dFunExpr dunf@(DFunUnfolding {}) = mkCoreLams (df_bndrs dunf) $ mkCoreConApps (df_con dunf) (df_args dunf)

### 2020-04-23 src/Plugin/MatchPlugin.hs

@@ -918,6 +918,8 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 let newExpr''6
                       =
+                          Data.transform combineCasts $
+
                           Data.transform (caseInline dflags) $
                           Data.transform betaReduce $
                           Data.transform (maybeApply (fmap (tryUnfoldAndReduceDict guts dflags) . caseInline_maybe dflags)) $
@@ -935,7 +937,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                           Data.transform (maybeApply (unfoldTheFn_maybe guts dflags constructFnId)) $
                           newExpr''5
 
-                newExpr'' <- Data.transformM (elimRepUnrep guts) $ newExpr''6
+                newExpr'' <- Data.transformM (targetCastM (elimRepUnrep guts)) $ newExpr''6
 
                 -- case newExpr'' of
                 --   App{} ->

### 2020-04-23 src/Test/PluginExample.hs

@@ -8,7 +8,7 @@
 -- {-# LANGUAGE StandaloneDeriving #-}
 -- {-# LANGUAGE FlexibleInstances #-}
 
-{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -dno-suppress-coercions -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 

### 2020-05-13 src/Deep/GHCUtils.hs
Some generalization and more coercion manipulation. Still getting Core lint errors

@@ -1318,6 +1318,7 @@ betaReduceAll e@(Cast (Lam v body) co) (a:as) =
     Nothing ->
         -- TODO: Make sure this does, in fact, mean it's either a constraint or a forall
         -- TODO: Make sure this works in those cases ^
+        -- TODO: Should there be a Cast here?
       betaReduceAll (substCoreExpr v a body) as
       -- error $ "betaReduceAll: " ++ showSDocUnsafe (ppr (coercionKind co))
       -- (e, as)

### 2020-05-13 src/Plugin/MatchPlugin.hs

@@ -49,6 +49,7 @@ import qualified Data.Kind as Kind
 import           Control.Monad
 import           Control.Monad.Writer hiding (pass, Alt)
 import           Control.Monad.State
+import           Control.Applicative
 
 import           Data.Foldable
 
@@ -788,6 +789,9 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                 constructFnId <- lift $ findIdTH guts 'construct
                 fromId <- lift $ findIdTH guts 'from
 
+                let tryUnfoldAndReduceDict' = tryUnfoldAndReduceDict guts dflags
+                let unfoldAndReduceDict_maybe' = unfoldAndReduceDict_maybe guts dflags
+
                 let constr' = mkApps (Var constr) (tys ++ dicts)
                 let constr'Type = exprType constr'
 
@@ -817,11 +821,15 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 -- liftIO $ putStrLn $ "repUnfolding = {" ++ showPpr dflags repUnfolding ++ "}"
 
-                let (fn, restArgs) = betaReduceAll
+                let (fn0, restArgs) = betaReduceAll
                                        repUnfolding
                                          [ Type (exprType expr)
                                          , repDictExpr
                                          ]
+
+                let unfoldAndReduceDictT = Data.transform tryUnfoldAndReduceDict'
+
+                let fn = caseInline dflags $ onScrutinee (betaReduce . onAppFun tryUnfoldAndReduceDict') $ Data.transform letNonRecSubst fn0
 
                 let repFn1 = caseInline dflags $ Data.transform letNonRecSubst fn
                     (repFnE, _) = collectArgs repFn1
@@ -831,15 +839,19 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                         Var f -> Just f
                         _ -> Nothing
 
-                let expr2' = Data.transform letNonRecSubst fn
-
                 let e' =
-                         onScrutinee (unfoldAndReduceDict guts dflags)
+                         -- onScrutinee (Data.transform tryUnfoldAndReduceDict') $
+                         onScrutinee (tryUnfoldAndReduceDict guts dflags)
+                         --     $ tryUnfoldAndReduceDict'
[…]

### 2020-05-13 src/Test/PluginExample.hs

@@ -8,7 +8,7 @@
 -- {-# LANGUAGE StandaloneDeriving #-}
 -- {-# LANGUAGE FlexibleInstances #-}
 
-{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -dno-suppress-type-applications -fprint-explicit-coercions -fprint-equality-relations -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -dno-suppress-coercions -dno-suppress-type-applications -fprint-equality-relations -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 

### 2020-05-13 src/Deep/GHCUtils.hs
Possibly fixed out-of-scope coercion type variable error

@@ -157,10 +157,16 @@ normaliseTypeCo guts ty = normaliseTypeCo_role guts Nominal ty
 
 
 normaliseTypeCo_role :: ModGuts -> Role -> Type -> CoreM (Coercion, Type)
-normaliseTypeCo_role guts role ty =
-  runTcM guts . fmap fst . runTcS $ do
+normaliseTypeCo_role guts role ty = do
+  (r, env) <- runTcM guts . runTcS $ do
     famInstEnvs <- getFamInstEnvs
     return (normaliseType famInstEnvs role ty)
+
+  when (not (isEmptyEvBindMap env)) $ traceM $ "normalise env(" ++ showSDocUnsafe (ppr ty) ++ ") = {" ++ showSDocUnsafe (ppr env) ++ "}"
+
+  -- traceM $ "normalise coercion: " ++ showSDocUnsafe (ppr (fst r))
+
+  return r
 
 
 

### 2020-05-13 src/Plugin/MatchPlugin.hs

@@ -273,6 +273,18 @@ returnChanged x = do
   changed
   return x
 
+-- -- TODO: Figure out why this is needed to mitigate an out-of-scope type variable issue in a coercion
+-- removeExpVarOnlyCoercion :: DynFlags -> Coercion -> Coercion
+-- removeExpVarOnlyCoercion dflags = Data.transform go
+--   where
+--     go co
+--       | Just (ty, role) <- isReflCo_maybe co
+--       , Just (tyF, tyArg) <- splitAppTy_maybe ty
+--       , isTyVarTy tyArg
+--         = trace ("coercion type var: " ++ showPpr dflags tyArg) co
+--     go co = co
+
+
 -- | 'Nothing' indicates that no changes were made
 transformExprMaybe :: ModGuts -> Var -> Maybe Var -> [(Id, CoreExpr)] -> Expr Var -> MatchM (Maybe (Expr Var))
 transformExprMaybe guts currName recName primMap e = do
@@ -295,8 +307,10 @@ transformExprMaybe guts currName recName primMap e = do
     go expr = return expr
 
 transformExpr :: ModGuts -> Var -> Maybe Var -> [(Id, CoreExpr)] -> Expr Var -> MatchM (Expr Var)
-transformExpr guts currName recNameM primMap =
-  {- transformApps guts <=< -} untilNothingM (transformExprMaybe guts currName recNameM primMap)
+transformExpr guts currName recNameM primMap e = do
+  dflags <- lift getDynFlags
+  {- Data.transform (onCoercion (removeExpVarOnlyCoercion dflags)) <$> -}
+  untilNothingM (transformExprMaybe guts currName recNameM primMap) e
 
 
 -- XXX: The delineation marker probably has to be floated in (or maybe the
@@ -976,7 +990,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 let (constructRepFn, [Type ty, dict1, dict2, constructRepArg]) = collectArgs newExpr'''
 
-                traceM $ "constructRepArg = {" ++ showPpr dflags constructRepArg ++ "}"
+                -- traceM $ "constructRepArg = {" ++ showPpr dflags constructRepArg ++ "}"
 
                     -- Normalize types
                 constructRepArg' <- lift $ Data.transformM (maybeApplyM (onCast_maybe
@@ -1007,18 +1021,19 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
[…]

### 2020-05-13 paper/paper.tex
Bibliography/citations

@@
 
+\citestyle{acmauthoryear}
+
 %%
@@
 %% end of the preamble, start of the body of the document source.
+
 \begin{document}
@@
 
+Any data type with a \ttt{Generic} instance
+
 \subsection{\ttt{E}, \ttt{ERep}, \ttt{ERepTy}}
@@
 
-Tail recursion is given a direct-style representation using a simple sum type.
+Tail recursion is given a direct-style representation using a simple sum type,
+using the technique described in~\cite{Grebe:2017:RSD:3136040.3136048}.
 
@@
 sum type of the result type and a thunk with the code necessary to continue
-execution.
-
-\begin{todo}
-  Cite "Trampolined Style" (S. E. Gantz, D. P. Friedman, M. Ward; 1999)
-\end{todo}
+execution.~\cite{Ganz:99:Trampolined}
 
@@
 \end{lstlisting}
-
-\begin{todo}
-  Cite "Rewriting a Shallow DSL using a GHC Compiler Extension"
-  (2017; Grebe, Young, Gill)
-\end{todo}
 
@@
 
-\section{Citations and Bibliographies}
-
-\begin{todo}
-  Fill in
-\end{todo}
+% \section{Citations and Bibliographies}
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{paper}
+
+% \begin{todo}
+%   Fill in
+% \end{todo}
 

### 2020-05-13 paper/paper.tex
Discuss connection to GHC generics, rearrange appendices

@@
 
+\subsection{Motivation}
+
 \subsection{Overview}
@@
 
-intListSum_helper :: (Int, IntList) -> Int
-intListSum_helper p =
+intListSum :: (Int, IntList) -> Int
+intListSum p =
   internalize (externalize
@@
            Cons x xs ->
-             intListSum_helper
+             intListSum
                (x+acc, xs)))
-
-intListSum :: IntList -> Int
-intListSum t =
-  internalize (externalize
-    (intListSum_helper (0, t)))
-\end{lstlisting}
-
-\subsection{GHC Generics \ttt{Rep}}
-
-\begin{todo}
-  This section is to talk about the details of GHC's generic \ttt{Rep} representation
-  and how it is converted to the representation in this paper. Should this section
-  exist?
-\end{todo}
-
-Any data type with a \ttt{Generic} instance
+\end{lstlisting}
 
@@
 
+With GHC Generics, any data type \ttt{T} with a \ttt{Generic} instance has a
+corresponding \ttt{Rep T} type, which gives a generic representation of \ttt{T}.
+Conversion between values of this type and values of the original type \ttt{T} is
+given by the functions \ttt{to} and \ttt{from}. The generic representation \ttt{Rep T}
+can be traversed by functions which operate solely on the structure of the data type.
+This generic representation contains additional metadata which we do not need.
+However, we can automatically generate a \ttt{ERep} instance for any type which
+has a \ttt{Generic} instance.
+
+As \ttt{Generic} instances are automatically generated, this provides a simple mechanism
+to automaticly generate \ttt{ERep} instances.
+
 This information is brought into the \ttt{E} type via the constructor
@@
   EXPR_INT
-, EXPR_FLOAT
-, EXPR_DOUBLE
-, EXPR_CHAR
-, EXPR_BOOL
-, EXPR_CLOSURE
-, EXPR_EITHER_LEFT
-, EXPR_EITHER_RIGHT
 , EXPR_PAIR
+...
 , EXPR_UNIT
-, EXPR_STEP
-, EXPR_DONE
-, EXPR_UNBOXED
 } var_type_tag;
@@
   Cite and compare/contrast with pattern matching technique described in section
-  3.3 of "Unembedding Domain-Specific Languages" (2009; Atkey, Lindley, Yallop)
+  3.3 of "Unembedding Domain-Specific Languages" (2009; Atkey, Lindley, Yallop)~\cite{Atkey:09:Unembedding}
 \end{todo}
 
-\clearpage
+% \section{Citations and Bibliographies}
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{paper}
 
@@
 \end{lstlisting}
-
-\clearpage
-
-% \section{Citations and Bibliographies}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
 % \begin{todo}

### 2020-05-13 paper/paper.tex
Updated paper todos

@@
 \begin{todo}
-  Add brief summary of relationship to lambdas
+  Maybe add brief summary of relationship to lambdas?
 \end{todo}
@@
 
+\begin{todo}
+  Write
+\end{todo}
+
 \subsection{Overview}
@@
 \begin{todo}
-  Expand the explanation in this section
+  Is the explanation in this section sufficient since we cite the Haskino paper,
+  which describes how the tail recursion transformation was used in that project?
 \end{todo}
@@
 
-\begin{todo}
-  Cite "Rewriting a Shallow DSL using a GHC Compiler Extension"
-  (2017; Grebe, Young, Gill)
-\end{todo}
-
 \section{Representing lambdas}
@@
 \begin{todo}
-  Expand this section
+  Should this section be expanded?
 \end{todo}

### 2020-05-13 paper/paper.tex
Some comparison to Atkey, at al. 2009

@@
 
+A similar EDSL-oriented representation of pattern matching is given in
+~\cite[Section~3.3]{Atkey:09:Unembedding}. In that paper, patterns were given their own
+representation which allows for compound patterns. This is useful in the context
+of that work, as the programmer works directly with these patterns.
+
+In the present paper, however, the representation is generated automatically by
+a compiler plugin. As a result of GHC's desugared Core (which does not have
+compound patterns), there is no need to directly express compound patterns at
+the representation-level.
+
 % \section{Citations and Bibliographies}

### 2020-05-14 src/Deep/GHCUtils.hs
A slight simplification (same error)

@@ -767,6 +767,10 @@ onTypeM :: Applicative m => (Type -> m Type) -> CoreExpr -> m CoreExpr
 onTypeM f (Type ty) = Type <$> f ty
 onTypeM _ e = pure e
 
+onType :: (Type -> Type) -> CoreExpr -> CoreExpr
+onType f (Type ty) = Type $ f ty
+onType _ e = e
+
 onCoercion :: (Coercion -> Coercion) -> CoreExpr -> CoreExpr
 onCoercion f (Coercion co) = Coercion (f co)
 onCoercion _ e = e
@@ -1320,12 +1324,24 @@ betaReduceAll e@(Cast (Lam v body) co) (a:as) =
   case splitFunCo_maybe co of
     Just (coA, coB) ->
       let (remaining, args) = betaReduceAll (Lam (setVarType v (coercionRKind coA)) (Cast body coB)) (Cast a coA:as)
-      in (remaining, as)
+      in (remaining, args)
     Nothing ->
         -- TODO: Make sure this does, in fact, mean it's either a constraint or a forall
         -- TODO: Make sure this works in those cases ^
         -- TODO: Should there be a Cast here?
-      betaReduceAll (substCoreExpr v a body) as
+      trace ("betaReduceAll: coercion kind = " ++ showSDocUnsafe (ppr (coercionKind co))) $
+      case splitForAllCo_maybe co of
+        Just (tyVar, coA, coB) ->
+          betaReduceAll (Lam (setVarType v (coercionRKind coA)) (Cast body coB)) (Cast a coA:as)
+        _ -> --error $ "betaReduceAll: " ++ showSDocUnsafe (ppr (coercionLKind co, coercionRKind co))
+          let (coA, coB) = decomposeFunCo (coercionRole co) co
+          in
+            betaReduceAll (Lam (setVarType v (coercionRKind coA)) (Cast body coB)) (Cast a coA:as)
+          -- case a of
+          --   Type ty ->
+          --     let substTyVar = substTy (extendTvSubst emptyTCvSubst tyVar a)
+          --     betaReduceAll (substCoreExpr v a (Data.transform (onType tySubst) body)) as
+          --   _ -> error "betaReduceAll"
       -- error $ "betaReduceAll: " ++ showSDocUnsafe (ppr (coercionKind co))
       -- (e, as)
 

### 2020-05-14 src/Plugin/MatchPlugin.hs

@@ -910,7 +910,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                                                                -- $ Data.transform (onAppWhen (appFunFromModule internalTypeableModule) (simpleOptExpr dflags))
                                                                -- $ Data.transform (unfoldAndBetaReduce guts dflags (idIsFrom internalTypeableModule))
                                                                -- $ occurAnalyseExpr
-                                                               $ simpleOptExpr dflags
+                                                               -- $ simpleOptExpr dflags
                                                                $ onVar getUnfolding'
                                                                $ e)
                                                       . Data.transform (caseInline dflags)
@@ -944,7 +944,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                 let newExpr'0 = untilNothing elimConstruct newExpr
                 let newExpr'
                       = Data.transform (caseInline dflags)
-                        $ Data.transform betaReduce 
+                        $ Data.transform betaReduce
                         $ Data.transform (maybeApply (combineCasts_maybe dflags)) $ newExpr'0
 
                 -- error (showPpr dflags (Data.transform(unfoldAndBetaReduce guts dflags (idIsFrom internalTypeableModule)) newExpr'0))
@@ -956,24 +956,26 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                 castExpId <- lift $ findIdTH guts 'CastExp
 
 
-                let newExpr'''
-                      = Data.transform
-                          (maybeApply
-                            (upOneLevel_maybe
-                              (fixExtCast dflags expType castExpId)
-                              (\e0 ->
-                                let r = getAny . execWriter $
-                                          (descendIntoCastsM
-                                            (\e ->
-                                              case (e, collectArgs e) of
-                                                (App {}, (Var fn, _)) -> do
-                                                  if fn == externalizeId
-                                                    then tell (Any True)
-                                                    else return ()
-                                                  return e
[…]

### 2020-05-14 src/Deep/GHCUtils.hs
Trying to preserve accurate coercions more in the transformation. Back to out-of-scope coercion type variable. No longer any calls to simpleOptExpr

@@ -169,6 +169,11 @@ normaliseTypeCo_role guts role ty = do
   return r
 
 
+normaliseCoercion :: ModGuts -> Coercion -> CoreM Coercion
+normaliseCoercion guts co = do
+  (co', ty) <- normaliseTypeCo_role guts (coercionRole co) (coercionRKind co)
+  return (mkTransCo co co')
+
 
 {-
 -- TODO: This is a stop-gap measure. Try to figure out why some of the
@@ -774,6 +779,10 @@ onType _ e = e
 onCoercion :: (Coercion -> Coercion) -> CoreExpr -> CoreExpr
 onCoercion f (Coercion co) = Coercion (f co)
 onCoercion _ e = e
+
+onCoercionM :: Applicative m => (Coercion -> m Coercion) -> CoreExpr -> m CoreExpr
+onCoercionM f (Coercion co) = Coercion <$> f co
+onCoercionM _ e = pure e
 
 removeCasts :: CoreExpr -> CoreExpr
 removeCasts (Cast e _) = removeCasts e

### 2020-05-14 src/Plugin/MatchPlugin.hs

@@ -941,7 +941,8 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
                 -- TODO: Make sure this recursively calls elimConstruct
                 -- properly (enough times)
-                let newExpr'0 = untilNothing elimConstruct newExpr
+                -- let newExpr'0 = untilNothing elimConstruct newExpr
+                newExpr'0 <- {- lift . (Data.transformM (onCoercionM (normaliseCoercion guts)) <=< Data.transformM (onTypeM (normaliseType' guts))) $ -} return $ untilNothing elimConstruct newExpr
                 let newExpr'
                       = Data.transform (caseInline dflags)
                         $ Data.transform betaReduce
@@ -1076,6 +1077,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                 -- traceM $ "co'' = {" ++ showPpr dflags co'' ++ "}"
                 -- traceM $ "mkSymCo co' = {" ++ showPpr dflags (mkSymCo co') ++ "}"
                 -- error "debug"
+                -- Data.transformM (fixConstructorCast guts dflags) constructedResult
                 return constructedResult
 
 
@@ -1152,30 +1154,29 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 -- normaliseConstructorType guts dflags e0@(Var v)
 --   | Just _ <- isDataConId_maybe fn
 --     = do
-        
 
 fixConstructorCast :: ModGuts -> DynFlags -> CoreExpr -> MatchM CoreExpr
-fixConstructorCast guts dflags e0@(Cast (app@(App {})) co0)
-  | (Var fn, args) <- collectArgs app
-  -- , trace ("might fix: " ++ showPpr dflags fn) True
-  , Just _ <- isDataConId_maybe fn
-  = do
-       let ty = coercionLKind co0
+-- fixConstructorCast guts dflags e0@(Cast (app@(App {})) co0)
+--   | (Var fn, args) <- collectArgs app
+--   -- , trace ("might fix: " ++ showPpr dflags fn) True
+--   , Just _ <- isDataConId_maybe fn
+--   = do
+--        let ty = coercionLKind co0
 
-       expTyCon <- lift $ findTyConTH guts ''GPUExp
-       repTyTyCon <- lift $ findTyConTH guts ''GPURepTy
-       case splitTyConApp_maybe ty of
[…]

### 2020-05-20 src/Deep/GHCUtils.hs
I think this is close to working. Need to not transform dictionary arguments to ConstructRep, etc

@@ -777,6 +777,13 @@ onAppFun = maybeApply . onAppFun_maybe . (Just .)
 onAppFunId :: (Id -> CoreExpr) -> CoreExpr -> CoreExpr
 onAppFunId = maybeApply . onAppFunId_maybe . (Just .)
 
+onCase_maybe :: (CoreExpr -> Maybe CoreExpr) -> CoreExpr -> Maybe CoreExpr
+onCase_maybe t e@(Case {}) = t e
+onCase_maybe _ _ = Nothing
+
+onCase :: (CoreExpr -> CoreExpr) -> CoreExpr -> CoreExpr
+onCase t = maybeApply (Just . t)
+
 onTypeM :: Applicative m => (Type -> m Type) -> CoreExpr -> m CoreExpr
 onTypeM f (Type ty) = Type <$> f ty
 onTypeM _ e = pure e
@@ -1289,6 +1296,11 @@ onVarWhen p t (Var v)
   | p v = t v
 onVarWhen _ _ e = e
 
+onVarWhen_maybe :: (Id -> Bool) -> (Id -> Maybe CoreExpr) -> CoreExpr -> Maybe CoreExpr
+onVarWhen_maybe p t (Var v)
+  | p v = t v
+onVarWhen_maybe _ _ _ = Nothing
+
 onAppWhen :: (CoreExpr -> Bool) -> (CoreExpr -> CoreExpr) -> CoreExpr -> CoreExpr
 onAppWhen p t e@(App {})
   | p e = t e
@@ -1394,6 +1406,19 @@ whenId :: Bool -> (a -> a) -> a -> a
 whenId b f x
   | b = f x
   | otherwise = x
+
+transformMaybe :: Data a => (a -> Maybe a) -> a -> Maybe a
+transformMaybe f e0 =
+  let (r, Any madeProgress) = runWriter (Data.transformM go e0)
+  in
+    if madeProgress
+      then Just r
+      else Nothing
+  where
+    go e =
+      case f e of
+        Nothing -> return e
+        Just e' -> tell (Any True) >> return e'
 
 -- | Beta-reduce as many lambda-binders as possible.
 betaReduceAll :: CoreExpr -> [CoreExpr] -> (CoreExpr, [CoreExpr])

### 2020-05-20 src/Plugin/MatchPlugin.hs

@@ -102,6 +102,7 @@ import           Data.Char
 import           Data.List
 
 import           CoreOpt
+import           Id
 
 -- import           Data.Typeable.Internal
 
@@ -991,12 +992,51 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                 externalizeId <- lift $ findIdTH guts 'externalize
                 castExpId <- lift $ findIdTH guts 'CastExp
 
-                let elimConstruct = Data.transform (caseInline dflags)
-                                  . Data.transform betaReduce
-                                  . Data.transform (replaceVarId constructFnId (getUnfolding' constructFnId))
+                let elimConstruct = fmap (Data.transform (caseInline dflags)
+                                           . Data.transform betaReduce)
+                                    . transformMaybe (replaceVarId_maybe constructFnId (getUnfolding' constructFnId))
+
+                let elimTheFn fnId = upOneLevel_maybe (Just
+                                                      . Data.transform (onAppFun (Data.transform betaReduce
+                                                                                 . Data.transform letNonRecSubst
+                                                                                 . betaReduce
+                                                                                 . onAppFun tryUnfoldAndReduceDict'
+                                                                                 . Data.transform letNonRecSubst
+                                                                                 . tryUnfoldAndReduceDict')))
+                                          (upOneLevel_maybe (Just
+                                                            . Data.transform (caseInline dflags)
+                                                            . Data.transform (onScrutinee tryUnfoldAndReduceDict')
+                                                            . Data.transform betaReduce)
+                                              (upOneLevel_maybe (Just)
+                                                (fmap (Data.transform (caseInline dflags)
+                                                          . Data.transform betaReduce)
+                                                  . transformMaybe (replaceVarId_maybe fnId (getUnfolding' fnId)))))
 
 
-                let newExpr''' = Data.transform elimConstruct newExpr''
+                let newExpr'''0 =
+                      Data.transform betaReduce $
+                      Data.transform (combineCasts dflags) $
+                      Data.transform (maybeApply (elimTheFn constructFnId)) $
+                      Data.transform (caseInline dflags) $
+                      Data.transform betaReduce $
[…]

### 2020-05-20 paper/paper.tex
Update paper todo
paper: Added a couple CCS tags and some description of motivation of the research

@@
 %% complete the rights form.
+
 \setcopyright{acmcopyright}
@@
 %% These commands are for a PROCEEDINGS abstract or paper.
+
 \acmConference[Haskell '20]{Haskell '20: Proceedings of the 13th
@@
    <concept>
+       <concept_id>10011007.10011006.10011041.10011046</concept_id>
+       <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10011007.10011006.10011041.10011047</concept_id>
+       <concept_desc>Software and its engineering~Source code generation</concept_desc>
+       <concept_significance>300</concept_significance>
+       </concept>
+   <concept>
        <concept_id>10011007.10011006.10011050.10011017</concept_id>
@@
 
+\ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
+\ccsdesc[300]{Software and its engineering~Source code generation}
 \ccsdesc[500]{Software and its engineering~Domain specific languages}
+
+% \begin{CCSXML}
+% <ccs2012>
+%    <concept>
+%        <concept_id>10011007.10011006.10011050.10011017</concept_id>
+%        <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
+%        <concept_significance>500</concept_significance>
+%        </concept>
+%  </ccs2012>
+% \end{CCSXML}
+
+% \ccsdesc[500]{Software and its engineering~Domain specific languages}
 
@@
 Readers who are primarily interested only in the encoding of pattern matching itself can direct
-most of their attention to sections 2, 3 and 6.
+most of their attention to Sections 2, 3 and 6. Most of the technical aspects of
+the transformation performed by the Core plugin are described in Section 8.
 
@@
 \begin{todo}
-  Write
+  Expand?
+\end{todo}
+
+A major benefit of the Haskell language is that you can often derive an
+implementation from a specification, either partially or fully. When it is not
+possible to fully derive an implementation from a specification, it is often
+easier to check an implementation against a specification, due in part to the
+management of side effects in Haskell. This is more difficult to do in
+imperative languages like C or CUDA (as they provide no restrictions on where
+side effects can occur).
+
+Using Haskell EDSLs that allow the use of powerful language features like
+pattern matching, these advantages can be brought to other domains where it is
+currently either difficult or impossible to use Haskell.
+
+\begin{todo}
+  Cite examples. Conal Elliott's last two papers (the one about convolutions and
+  the latest one about automatic differentiation), would probably be good
+  examples of this in Haskell.
+
+  A paper which describes a Haskell implementation of a specification given as
+  an operational semantics would also be a good example.
 \end{todo}
@@
 \begin{todo}
-  Cite and compare/contrast with pattern matching technique described in section
-  3.3 of "Unembedding Domain-Specific Languages" (2009; Atkey, Lindley, Yallop)~\cite{Atkey:09:Unembedding}
+  Does this need more fleshing out?
 \end{todo}
@@
 
-\section{EDSL for C code generation}
+\section{A small EDSL for C code generation used internally by the C backend}
 \begin{lstlisting}

### 2020-05-20 paper/paper.tex
paper: Reworded some of the motivation, added TODO citations for last two Elliott papers (which both use homomorphic specification)

@@
 \documentclass[sigplan,screen]{acmart}
+\usepackage[switch]{lineno}
 \usepackage{booktabs}
@@
 \tikzstyle{arrow} = [thick,->,>=stealth]
+
+% \renewcommand{\linenumberfont}{\color{blue}}
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{red}}
 
@@
 
+\linenumbers
 \section{Introduction}
@@
 
-A major benefit of the Haskell language is that you can often derive an
-implementation from a specification, either partially or fully. When it is not
-possible to fully derive an implementation from a specification, it is often
-easier to check an implementation against a specification, due in part to the
-management of side effects in Haskell. This is more difficult to do in
-imperative languages like C or CUDA (as they provide no restrictions on where
-side effects can occur).
-
-Using Haskell EDSLs that allow the use of powerful language features like
-pattern matching, these advantages can be brought to other domains where it is
-currently either difficult or impossible to use Haskell.
+A major benefit of the Haskell language is that it is frequently possible derive
+a Haskell implementation from a specification, either partially or fully. When
+it is not possible to fully derive an implementation from a specification, it is
+often easier to check a Haskell implementation against a specification than it
+is in many imperative languages due, in part, to the management of side effects
+in Haskell. This is more difficult to do using an imperative language such as C,
+as they provide no restrictions on where side effects can occur.
+
+Haskell EDSLs that support powerful language features like pattern matching
+allow these advantages to be brought to problem domains where it is currently
+either difficult or impossible to use a language that has these benefits.
 
 \begin{todo}
-  Cite examples. Conal Elliott's last two papers (the one about convolutions and
-  the latest one about automatic differentiation), would probably be good
-  examples of this in Haskell.
-
-  A paper which describes a Haskell implementation of a specification given as
-  an operational semantics would also be a good example.
+  Cite examples where Haskell implementations have been derived specifications.
+  Conal Elliott's latest two papers (as of May 2020),
+  \cite{Elliott-2018-ad-icfp} and \cite{Elliott2019-convolution-extended}, would
+  be good examples of this.
+
+  A paper which describes a Haskell implementation derived from an operational
+  semantics would also be a good example.
 \end{todo}

### 2020-05-21 src/Deep/GHCUtils.hs
Getting closer to a correct constructor transformation. I think all that's left is fixing the coercions

@@ -90,7 +90,7 @@ import qualified Data.Generics.Uniplate.DataOnly as Data
 
 import           Data.Data (Data)
 
-import           Data.Maybe (fromMaybe, isNothing)
+import           Data.Maybe (fromMaybe, isNothing, isJust)
 
 import Debug.Trace
 
@@ -934,7 +934,13 @@ maybeTransform3 f (Just x) y = f x y
 -- TODO: Try to generalize this to work with any `Data` type (maybe use
 -- Uniplate?).
 upOneLevel_maybe :: (CoreExpr -> Maybe CoreExpr) -> (CoreExpr -> Maybe CoreExpr) -> CoreExpr -> Maybe CoreExpr
-upOneLevel_maybe p q (App f0 x0) = do
+upOneLevel_maybe p q e@(App f0 x0) = do
+  -- XXX: Ignores class variable applications
+
+  let (fnE, args) = collectArgs e
+
+  guard (not (isClassVar fnE))
+
   let f_M = q f0
       x_M = q x0
 
@@ -1259,9 +1265,9 @@ combineCasts_maybe dflags origE@(Cast (Cast e coA) coB) =
         Just (coA', coB') ->
           if eqType (coercionLKind coB') (coercionRKind coB')
             then Just $ Cast e coA'
-            else trace ("combineCasts_maybe: coB' = {" ++ showPpr dflags coB' ++ "}") $
-                 trace ("^-------------------------->" ++ showPpr dflags (coercionLKind coB', coercionRKind coB')) $
-                 trace ("===========================>" ++ showPpr dflags (getCoVar_maybe coB')) $
+            else --trace ("combineCasts_maybe: coB' = {" ++ showPpr dflags coB' ++ "}") $
+                 --trace ("^-------------------------->" ++ showPpr dflags (coercionLKind coB', coercionRKind coB')) $
+                 --trace ("===========================>" ++ showPpr dflags (getCoVar_maybe coB')) $
                  -- trace ("===========================>" ++ showPpr dflags (buildCoercion (coercionLKind coB') (coercionRKind coB'))) $
                  Just $ Cast e (mkTransCo coA' coB')
         Nothing ->
@@ -1419,6 +1425,128 @@ transformMaybe f e0 =
       case f e of
         Nothing -> return e
         Just e' -> tell (Any True) >> return e'
+
[…]

### 2020-05-21 src/Plugin/MatchPlugin.hs

@@ -396,6 +396,15 @@ isConstructor v
       DataConWrapId {} -> True
       _ -> trace ("is not constructor: " ++ occNameString (occName v)) False
 
+getConstructorTyCon_maybe :: Var -> Maybe TyCon
+getConstructorTyCon_maybe v
+  | isId v =
+    case idDetails v of
+      DataConWorkId dc -> Just $ dataConTyCon dc
+      DataConWorkId dc -> Just $ dataConTyCon dc
+      _ -> Nothing
+getConstructorTyCon_maybe _ = Nothing
+
 builtinConstructor :: Var -> Bool
 builtinConstructor v =
     "I#" == occNameString (occName v)
@@ -408,29 +417,30 @@ getConstructorApp_maybe ::
     -> Maybe
         ( [Expr Var]  -- Type arguments
         , [Expr Var]  -- Dictionary arguments
+        , TyCon       -- The TyCon of the type of the constructor
         , Var         -- Constructor name
         , [Expr Var]) -- Constructor arguments
 getConstructorApp_maybe e = 
   case go e of
     -- Just (_, _, _, []) -> Nothing -- TODO: Do we want this?
-    (Just (_, _, constr, _))
+    (Just (_, _, _, constr, _))
       | builtinConstructor constr -> Nothing
     r -> r
   where
     go (Var v)
-      | isConstructor v = Just ([], [], v, [])
+      | Just tyCon <- getConstructorTyCon_maybe v = Just ([], [], tyCon, v, [])
     go (App f x)
       | isTypeArg x = do
-          (tys, dicts, constr, args) <- go f
-          return (x:tys, dicts, constr, args)
+          (tys, dicts, tyCon, constr, args) <- go f
+          return (x:tys, dicts, tyCon, constr, args)
 
       | isCoArg x || isDict x = do
-          (tys, dicts, constr, args) <- go f
-          return (tys, x:dicts, constr, args)
+          (tys, dicts, tyCon, constr, args) <- go f
+          return (tys, x:dicts, tyCon, constr, args)
 
       | otherwise             = do
-          (tys, dicts, constr, args) <- go f
-          return (tys, dicts, constr, x:args)
+          (tys, dicts, tyCon, constr, args) <- go f
+          return (tys, dicts, tyCon, constr, x:args)
 
     go _ = Nothing
 
@@ -776,7 +786,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
 
         -- Handle constructors that are not builtins
         go expr@(_ :@ _)
[…]

### 2020-05-21 src/Test/PluginExample.hs

@@ -8,7 +8,7 @@
 -- {-# LANGUAGE StandaloneDeriving #-}
 -- {-# LANGUAGE FlexibleInstances #-}
 
-{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -dno-suppress-type-applications -fprint-equality-relations -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -fprint-equality-relations -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 

### 2020-05-21 src/Deep/GHCUtils.hs
Seem to have fixed some of the coercion problems

@@ -150,6 +150,13 @@ buildDictionaryT guts = \ ty0 -> do
                 [NonRec v e] | i == v -> e -- the common case that we would have gotten a single non-recursive let
                 _ -> mkCoreLets bnds (varToCoreExpr i)
 
+buildCo :: HasCallStack => ModGuts -> Type -> Type -> CoreM CoreExpr
+buildCo guts tyA tyB = do
+  let k = tcTypeKind tyA
+      ty = mkTyConApp eqTyCon [k, tyA, tyB]
+
+  buildDictionaryT guts ty
+
 normaliseType' :: ModGuts -> Type -> CoreM Type
 normaliseType' guts = fmap snd . normaliseTypeCo guts
 
@@ -1271,8 +1278,8 @@ combineCasts_maybe dflags origE@(Cast (Cast e coA) coB) =
                  -- trace ("===========================>" ++ showPpr dflags (buildCoercion (coercionLKind coB') (coercionRKind coB'))) $
                  Just $ Cast e (mkTransCo coA' coB')
         Nothing ->
-          error $ "combineCasts_maybe: *** coercions do not match ***: " ++ showPpr dflags (coercionKind coA, coercionKind coB)
-          -- trace "combineCasts_maybe: *** coercions do not match ***" $ Just e
+          -- error $ "combineCasts_maybe: *** coercions do not match ***: " ++ showPpr dflags (coercionKind coA, coercionKind coB)
+          trace "combineCasts_maybe: *** coercions do not match ***" $ Just origE
 combineCasts_maybe _ _ = Nothing
 
 combineCasts :: DynFlags -> CoreExpr -> CoreExpr

### 2020-05-21 src/Plugin/MatchPlugin.hs

@@ -762,6 +762,14 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
               markedX <- mark x
               return (b :@ tyA :@ tyB :@ markedX)
 
+        -- go expr@(Var f :@ Type tyA :@ (Var g :@ tyB :@ dict :@ x))
+        --   | not (hasExprTy expr)
+        --   , isDict dict
+        --   , Nothing <- getConstructorApp_maybe expr = do
+        --       unrepId <- lift $ findIdTH guts 'unrep
+        --       when (f == unrepId) $ traceM "found unrep"
+        --       return expr
+
         go expr@(Var f :@ x)
           | Just b <- builtin f
           , not (isDict x)
@@ -969,7 +977,7 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                                   $ Data.transform (caseInline dflags) $ betaReduce (simplE' :@ mkApps constr' unreppedArgs)
                     -- traceM $ "simplE'' = " ++ showPpr dflags simplE''
 
-                    let (simplE''Fn, simplE''Args0) = collectArgs simplE''
+                    let (simplE''Fn, simplE''Args0@(Type tyArg:_)) = collectArgs simplE''
                         simplE''Arg0 = last simplE''Args0
 
                         simplE''Arg = untilNothing (elimTheFn constructFnId) simplE''Arg0
@@ -1027,8 +1035,19 @@ transformPrims0 guts currName recName primMap exprVars e = {- transformLams guts
                     -- traceM $ "newExpr'0 = {" ++ showPpr dflags newExpr'0 ++ "}"
                     -- traceM $ "newExpr''' = {" ++ showPpr dflags newExpr''' ++ "}"
 
+                    let toTy = mkTyConApp expTyCon [mkTyConApp repTyTyCon [tyArg]]
 
-                    constructedResult <- return $ prepareResult newExpr'''
+                    -- newCo <- lift $ buildCo guts (exprType newExpr''') toTy
+                    (co1, _) <- lift $ normaliseTypeCo_role guts Representational toTy
+                    (co2, _) <- lift $ normaliseTypeCo_role guts Representational (exprType newExpr''')
+                    let co3 = mkTransCo co2 (mkSymCo co1)
[…]

### 2020-05-21 src/Test/PluginExample.hs

@@ -8,7 +8,7 @@
 -- {-# LANGUAGE StandaloneDeriving #-}
 -- {-# LANGUAGE FlexibleInstances #-}
 
-{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -fprint-equality-relations -fplugin=Plugin.MatchPlugin #-}
+{-# OPTIONS_GHC -O0 -Wtype-defaults -fexpose-all-unfoldings -dcore-lint -dsuppress-all -dno-suppress-coercions -dno-suppress-type-applications -fprint-equality-relations -fplugin=Plugin.MatchPlugin #-}
 
 module Test.PluginExample where
 

### 2020-05-21 src/Plugin/MatchPlugin.hs
Improved type safety of 'abstractOver' and added some elimination of 'unrep (Var ...)' when appropriate

@@ -356,13 +356,56 @@ isVar _ = False
 
 -- Mark for further transformation
 mark0 :: HasCallStack => ModGuts -> Expr Var -> MatchM (Expr Var)
-mark0 _ e@(Coercion {}) = return e
+-- mark0 _ e@(Coercion {}) = return e
+mark0 guts e@(Var f :@ Type ty :@ dict :@ x) = do
+  externalizeName <- lift $ findIdTH guts 'externalize
+
+  if f == externalizeName
+    then do
+      r <- tryMarkUnrepVar guts x
+      case r of
+        Just x' -> return x'
+        Nothing -> return e
+    else do
+      r <- tryMarkUnrepVar guts e
+      case r of
+        Just x' -> return x'
+        Nothing -> markIt guts x
+
 mark0 guts x = do
+  r <- tryMarkUnrepVar guts x
+  case r of
+    Just x' -> return x'
+    Nothing -> markIt guts x
+
+tryMarkUnrepVar :: HasCallStack => ModGuts -> Expr Var -> MatchM (Maybe (Expr Var))
+tryMarkUnrepVar guts e@(Var f :@ Type tyA :@ var@(Var g :@ Type tyB :@ _dict :@ _)) = do
+  unrepId <- lift $ findIdTH guts 'unrep
+  varConId <- lift $ findIdTH guts 'Expr.Var
+
+  dflags <- getDynFlags
+
+  if f == unrepId && g == varConId
+    then do
+      -- traceM $ "=== mark0 ===> " ++ showPpr dflags var
+      return $ Just var
+    else do
+      -- traceM $ "=== mark0 ===> " ++ showPpr dflags var
+      return Nothing
+
+tryMarkUnrepVar guts e = Just <$> markIt guts e
+
+-- | Mark without special treatment for any 'unrep (Var ...)' expressions
+markIt :: HasCallStack => ModGuts -> Expr Var -> MatchM (Expr Var)
+markIt guts x = do
   dflags <- getDynFlags
   expTyCon <- lift $ findTyConTH guts ''GPUExp
 
+  -- traceM $ "markIt: " ++ showPpr dflags x
+
   eType <- hasExprType guts x
-  -- liftIO $ putStrLn $ "mark0 exprType = " ++ showPpr dflags (exprType x)
+  -- traceM $ "markIt x hasExprType: " ++ showPpr dflags eType
+  -- liftIO $ putStrLn $ "markIt exprType = " ++ showPpr dflags (exprType x)
   if eType
[…]

### 2020-07-11 paper/paper.tex
Fixed typo
Format paper for GPCE; add Mark to author list

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigplan,screen]{acmart}
+\documentclass[sigplan,anonymous,review]{acmart}
 \usepackage[switch]{lineno}
@@
 % \renewcommand{\linenumberfont}{\color{blue}}
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{red}}
+
+% \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{red}}
 
@@
 
-\acmConference[Haskell '20]{Haskell '20: Proceedings of the 13th
-  ACM SIGPLAN International Symposium on Haskell}{August 27 -- 28, 2020}{New Jersey, United States,}
-\acmBooktitle{Haskell '20: ACM SIGPLAN International Symposium on Haskell,
-  August 27 -- 28, 2020, New Jersey, United States}
+\acmConference[GPCE '20]{Haskell '20: Proceedings of the 13th
+  ACM SIGPLAN International Symposium on Haskell}{November 15 -- 20, 2020}{Chicago, Illinois, United States}
+\acmBooktitle{Proceedings of the 19th ACM SIGPLAN International
+Conference on Generative Programming: Concepts and Experiences (GPCE '20)
+  November 15 -- 20, 2020, Chicago, Illinois, United States}
 \acmPrice{15.00}
@@
 
+\author{Mark Grebe}
+\email{grebe@ucmo.edu}
+
 \author{Andrew Gill}
@@
 % \renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-\renewcommand{\shortauthors}{Young and Gill}
+\renewcommand{\shortauthors}{Young, Grebe and Gill}
 
@@
   algebraic data types and pattern matching in a DSL embedded in Haskell. This
-  representation is mechanized through the use of GHC Generics and and GHC Core plugin.
+  representation is mechanized through the use of GHC Generics and a GHC Core plugin.
 \end{abstract}
@@
 
-\linenumbers
 \section{Introduction}

### 2020-07-11 paper/paper.tex
Update conference name

@@
 
-\acmConference[GPCE '20]{Haskell '20: Proceedings of the 13th
+\acmConference[GPCE '20]{GPCE '20: Proceedings of the 13th
   ACM SIGPLAN International Symposium on Haskell}{November 15 -- 20, 2020}{Chicago, Illinois, United States}

### 2020-07-19 paper/paper.tex
Church encoding section

@@
 \begin{todo}
-  Write this subsection
+  Work on this subsection
+\end{todo}
+
+Algebraic datatypes can be represented as lambda calculus terms using
+\textit{Church encodings}. For example, the pair constructor can be
+represented as
+
+\begin{lstlisting}
+church_mkPair
+  :: a -> b -> (a -> b -> r) -> r
+church_mkPair = \ x y f -> f x y
+\end{lstlisting}
+
+with the eliminators
+
+\begin{lstlisting}
+church_fst :: ((a -> b -> a) -> r) -> r
+church_fst = \ p -> p (\ x y -> x)
+
+church_snd :: ((a -> b -> b) -> r) -> r
+church_snd = \ p -> p (\ x y -> y)
+\end{lstlisting}
+
+Coproducts can be represented with the constructors
+
+\begin{lstlisting}
+church_mkLeft
+  :: a -> (a -> r) -> (b -> r) -> r
+church_mkLeft = \ x f g -> f x
+
+church_mkRight
+  :: b -> (a -> r) -> (b -> r) -> r
+church_mkRight = \ y f g -> g y
+\end{lstlisting}
+
+and the eliminator
+
+\begin{lstlisting}
+church_eitherElim
+  :: (a -> b -> r) -> a -> b -> r
+church_eitherElim = \ e f g -> e f g
+\end{lstlisting}
+
+In this representation:
+
+\begin{itemize}
+  \item A pair of type \ttt{A} and type \ttt{B} is represented with a value of type
+\ttt{forall r. (A -> B -> r) -> r}.
+
+  \item A coproduct of type \ttt{A} and type \ttt{B} is represented with a value of type
+\ttt{forall r. (A -> r) -> (B -> r) -> r}.
+\end{itemize}
+
+More generally, an algebraic datatype can be thought of as a sum of products
+(that is, a coproduct of products).
+
+\begin{todo}
+  Maybe it could make sense to essentially derive the pattern matching representation
+  from this Church encoding. If so, this could be described around here.
 \end{todo}

### 2020-07-20 paper/paper.tex
More on the relationship with Church encodings

@@
 \end{itemize}
+
+Both of those representations represent the type as a function that performs the
+corresponding case analysis. A deep embedding for the operations
+\ttt{church\_mkLeft} and \\ \ttt{church\_mkRight} would look
+like:
+
+\begin{lstlisting}
+data ChurchCoproduct r where
+  Church_MkLeft
+    :: a -> (a -> r) -> (b -> r) ->
+       ChurchCoproduct r
+  Church_MkRight
+    :: b -> (a -> r) -> (b -> r) ->
+       ChurchCoproduct r
+
+church_coproductEval
+  :: ChurchCoproduct r -> r
+church_coproductEval (Church_MkLeft x f g)
+  = f x
+church_coproductEval (Church_MkRight y f g)
+  = g y
+\end{lstlisting}
+
+If we factor out the data contained in these coproducts so that it is separated from the
+case analysis functions, we can obtain
+
+\begin{lstlisting}
+data ChurchCoproductMatch t r where
+  ChurchCoproductMatch
+    :: (a -> r) -> (b -> r) ->
+       ChurchCoproductMatch (Either a b) r
+
+church_coproductMatchEval
+  :: Either a b ->
+     ChurchCoproductMatch (Either a b) r ->
+     r
+church_coproductMatchEval
+    (Left x)
+    (ChurchCoproductMatch f g)
+  = f x
+church_coproductMatchEval
+    (Right y)
+    (ChurchCoproductMatch f g)
+  = g y
+\end{lstlisting}
 

### 2020-07-20 paper/paper.tex
Similarity to functions [Church encoding section in paper]

@@
     (ChurchCoproductMatch f g)
+  = g y
+\end{lstlisting}
+
+Note that a value of type \ttt{ChurchCoproductMatch (Either a b) r} is very similar to
+a function from \ttt{Either a b} to \ttt{r}, with \ttt{church\_coproductMatchEval} allowing
+us to apply it to an argument. We can make this fact more explicit in the types with the
+change:
+
+\begin{lstlisting}
+data ChurchCoproductMatch t where
+  ChurchCoproductMatch
+    :: (a -> r) -> (b -> r) ->
+       ChurchCoproductMatch
+          (Either a b -> r)
+
+church_coproductMatchEval
+  :: ChurchCoproductMatch (t -> r) ->
+     t -> r
+church_coproductMatchEval
+    (ChurchCoproductMatch f g)
+    (Left x)
+  = f x
+church_coproductMatchEval
+    (ChurchCoproductMatch f g)
+    (Right y)
   = g y

### 2020-07-20 paper/paper.tex
Expanding on Church encoding derivation idea in Church encoding section of paper

@@
 
-More generally, an algebraic datatype can be thought of as a sum of products
-(that is, a coproduct of products).
-
-\begin{todo}
-  Maybe it could make sense to essentially derive the pattern matching representation
-  from this Church encoding. If so, this could be described around here.
-\end{todo}
+The type signature of \ttt{church\_coProductMatchEval} can be rewritten in the same form as a
+typical deep embedding evaluation function by moving the second argument into the right-hand
+side of the definition:
+
+\begin{lstlisting}
+church_coproductMatchEval
+  :: ChurchCoproductMatch t -> t
+
+church_coproductMatchEval
+    (ChurchCoproductMatch f g)
+  = \ c ->
+      case c of
+        Left x  -> f x
+        Right y -> g y
+\end{lstlisting}
+
+An embedding for product matches can be similarly derived from the Church encoding for product
+types:
+
+\begin{lstlisting}
+data ChurchProductMatch t where
+  ChurchProductMatch
+    :: (a -> b -> r) ->
+       ChurchProductMatch ((a, b) -> r)
+
+church_productMatchEval
+  :: ChurchProductMatch t -> t
+
+church_productMatchEval
+    (ChurchProductMatch f)
+  = \ p ->
+      case p of
+        (x, y) -> f x y
+\end{lstlisting}
+
+As we will have another notion of function types embedded in our language, we define two
+types to distinguish between these. We define two of them so that we can also distinguish
+between coproduct matches and product matches.
+
+\begin{lstlisting}
+newtype CoproductMatch a r
+  = CoproductMatch (a -> r)
+
+newtype ProductMatch a r
+  = ProductMatch (a -> r)
+
+data ChurchCoproductMatch t where
+  ChurchCoproductMatch
+    :: (a -> r) -> (b -> r) ->
+       ChurchCoproductMatch
+         (CoproductMatch (Either a b) r)
+
+church_coproductMatchEval
+  :: ChurchCoproductMatch t -> t
+
+church_coproductMatchEval
+    (ChurchCoproductMatch f g)
+  = CoproductMatch (\ c ->
+      case c of
+        Left x  -> f x
+        Right y -> g y)
+\end{lstlisting}
+
+The definition for product follows similarly, using \\ \ttt{ProductMatch}.
+
+Here we make the observation that algebraic datatypes are structured as coproducts of products. With
+this information in hand, we can combine \ttt{ChurchCoproductMatch} and \ttt{ChurchProductMatch}.
+
+\begin{lstlisting}
+newtype CoproductMatch a r
+  = CoproductMatch
+      { runCoproductMatch :: a -> r }
+
+newtype ProductMatch a r
+  = ProductMatch
+      { runProductMatch :: a -> r }
+
+data ChurchMatch t where
+  ChurchCoproductMatch
+    :: ChurchMatch (ProductMatch a r) ->
+       ChurchMatch (CoproductMatch b r) ->
+       ChurchMatch
+         (CoproductMatch (Either a b) r)
+
+  ChurchProductMatch
+    :: ChurchMatch (a -> ProductMatch b r)
+       -> ChurchMatch (ProductMatch (a, b) r)
+
+  ChurchOneCoproductMatch
+    :: ChurchMatch (ProductMatch a r) ->
+       ChurchMatch (CoproductMatch a r)
+
+  ChurchOneProductMatch
+    :: (a -> r) ->
+       ChurchMatch (ProductMatch a r)
+
+  ChurchNullaryProductMatch
+    :: r -> ChurchMatch (ProductMatch a r)
+
+  ChurchLam
+    :: (a -> b) -> ChurchMatch (a -> b)
+
+church_matchEval :: ChurchMatch t -> t
+church_matchEval
+    (ChurchCoproductMatch f g) =
+  CoproductMatch (\ c ->
+    case c of
+      Left x  ->
+        runProductMatch
+          (church_matchEval f) x
+      Right y ->
+        runCoproductMatch
+          (church_matchEval g) y)
+
+church_matchEval
+    (ChurchProductMatch f) =
+  ProductMatch (\ p ->
+    case p of
+      (x, y) ->
+        runProductMatch
+          (church_matchEval f x) y)
+
+church_matchEval
+    (ChurchOneCoproductMatch f) =
+  CoproductMatch (\ x ->
+    runProductMatch
+      (church_matchEval f) x)
+
+church_matchEval
+    (ChurchOneProductMatch f) =
+  ProductMatch (\ x -> f x)
+
+church_matchEval
+    (ChurchNullaryProductMatch x) =
+  ProductMatch (\ _ -> x)
+
+church_matchEval
+    (ChurchLam f) = \ x -> f x
+\end{lstlisting}
+
+% \begin{lstlisting}
+% data ChurchMatch t where
+%   ChurchCoproductMatch
+%     :: ChurchMatch ((a, b) -> r) -> ChurchMatch (Either b c -> r) ->
+%        ChurchCoproductMatch (Either
+% \end{lstlisting}
+
+% \begin{todo}
+%   Maybe it could make sense to essentially derive the pattern matching representation
+%   from this Church encoding. If so, this could be described around here.
+% \end{todo}
 

### 2020-07-20 paper/paper.tex
Formatting
Added a couple of TODOs in the Church encoding section

@@
     :: ChurchMatch (a -> ProductMatch b r)
-       -> ChurchMatch (ProductMatch (a, b) r)
+       -> ChurchMatch
+            (ProductMatch (a, b) r)
 
@@
 \end{lstlisting}
+
+\begin{todo}
+  Explicitly mention how the constructs in this last part
+  relate to their counterparts in the main part of the paper.
+\end{todo}
+
+\begin{todo}
+  Figure out how recursive types fit into this framework. Church encodings can
+  handle recursive types and so can our representation. This is also the case
+  for Scott encodings, etc, and is the difference between Church and Scott
+  encodings.
+
+  Does our way of handling recursive types correspond to how any of
+  those encodings handle recursive types?
+\end{todo}
 

### 2020-07-20 paper/paper.tex
Update title
Next draft of abstract

--- 2020-07-20 paper/paper.tex
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{First Class Pattern Matching in a Deeply Embedded Domain Specific Language\\
+\title{On Adding Pattern Matching to Haskell-based
+  Deeply Embedded Domain Specific Languages\\
 \large{Not For Distribution}}
@@
 \begin{abstract}
+  %Say what problem is
+  The data flow in deeply embedded domain specific langauges  falls out naturally.
+
+  Capturing control flow in embbed
+  is critical for 
+  Using a deep embedding of 
+ Programs that write programs 
+  Sometimes,  
+Embedded Domain Specific Languages are the killer app of functional languages.
+%Say why it is an interesting problem
+Languages with reification of code fragments allows
+for high-level abstractions that directly support
+diverse applications incuding machine learning and hardware description generation.
+% 
+  Haskell in particual has been used for many embeded code generators, such
+as Lava, SunRoof, Conal work, CCC, AD, ... (finally tagless) diagrams?
+%Say what our solution achieves
+In this paper, we expand the capabilies of traditional,
+Haskell-based, deep embeddings with a compiler extension
+for capturing the useful abstraction of pattern matching. 
+%Here is why the world will be a better place
+With this new support, the Haskell that we capture for expressing
+embededed-based solutions can be cleaner, more Haskell idiomatic,
+and more declaritive in nature.
+
   Domain specific languages (DSLs) provide a powerful tool for abstracting over
-  a class of problems. In this paper, we introduce a technique for representing
-  algebraic data types and pattern matching in a DSL embedded in Haskell. This
+  a class of problems.
+  This
   representation is mechanized through the use of GHC Generics and a GHC Core plugin.

### 2020-07-20 paper/paper.tex
Work on abstract

@@
   %Say what problem is
-  The data flow in deeply embedded domain specific langauges  falls out naturally.
-
-  Capturing control flow in embbed
+
+  Capturing control flow in an embedded domain specific language is crucial for
+  writing programs in a "natural" style. This enables the programmer to use the
+  constructs that are already builtin to the base language, such as pattern
+  matching and recursion.
+
+  In addition to control flow, when these features are reified, the data flow in
+  a deeply embedded domain specific langauge also falls out naturally.
+
   is critical for 
@@
 Embedded Domain Specific Languages are the killer app of functional languages.
+
 %Say why it is an interesting problem
-Languages with reification of code fragments allows
+
+In embedded domain specific languages which do not capture control flow
+constructs, workarounds must be employed which will be less familiar
+than the builtin control flow mechanisms of the host language. When
+control flow mechanisms are represented in the embedded language,
+a much more natural style of programming is offered to the programmer.
+
+Languages with reification of code fragments allow
 for high-level abstractions that directly support
@@
 % 
-  Haskell in particual has been used for many embeded code generators, such
-as Lava, SunRoof, Conal work, CCC, AD, ... (finally tagless) diagrams?
+  Haskell in particular has been used for many embedded code generators, such
+  as Lava, SunRoof, Conal work, CCC, AD, ... (finally tagless) diagrams? Feldspar (note: Feldspar has some form of \ttt{switch} primitive)?
+
 %Say what our solution achieves
+
 In this paper, we expand the capabilies of traditional,
 Haskell-based, deep embeddings with a compiler extension
-for capturing the useful abstraction of pattern matching. 
+for capturing the useful abstraction of pattern matching. This is accomplished
+using a deeply embedded representation of the pattern match construct.
+
 %Here is why the world will be a better place
+
 With this new support, the Haskell that we capture for expressing
-embededed-based solutions can be cleaner, more Haskell idiomatic,
-and more declaritive in nature.
+embedded-based solutions can be cleaner, more Haskell idiomatic,
+and more declarative in nature.
 

### 2020-07-20 paper/paper.tex
Abstract for reviewers

--- 2020-07-20 paper/paper.tex
@@
 \begin{abstract}
-  %Say what problem is
-  The data flow in deeply embedded domain specific langauges  falls out naturally.
-
-  Capturing control flow in embbed
-  is critical for 
-  Using a deep embedding of 
- Programs that write programs 
-  Sometimes,  
-Embedded Domain Specific Languages are the killer app of functional languages.
-%Say why it is an interesting problem
-Languages with reification of code fragments allows
-for high-level abstractions that directly support
-diverse applications incuding machine learning and hardware description generation.
-% 
-  Haskell in particual has been used for many embeded code generators, such
-as Lava, SunRoof, Conal work, CCC, AD, ... (finally tagless) diagrams?
-%Say what our solution achieves
+% Say what problem is
+  Capturing control flow is the Achilles heel of Haskell-based
+  deeply embedding Domain Specific Languages,
+  clunkly workarounds must be employed which will be less familiar
+  than the builtin control flow mechanisms of the host language.  
+% Say why it is an interesting problem
+  However, capturing control flow in an embedded domain specific language is crucial for
+  writing programs in a natural style, and enables the programmer to use the
+  constructs that are already builtin to the base language, such as pattern
+  matching and recursion.
+% Say what our solution achieves
   In this paper, we expand the capabilies of traditional,
   Haskell-based, deep embeddings with a compiler extension
-for capturing the useful abstraction of pattern matching. 
-%Here is why the world will be a better place
+  for reifying conditionals and pattern matching.
+% Here is why the world will be a better place
   With this new support, the Haskell that we capture for expressing
-embededed-based solutions can be cleaner, more Haskell idiomatic,
-and more declaritive in nature.
-
-  Domain specific languages (DSLs) provide a powerful tool for abstracting over
-  a class of problems.
-  This
-  representation is mechanized through the use of GHC Generics and a GHC Core plugin.
+  deeply embedded domain specific languges can be cleaner, more Haskell idiomatic,
+  and more declarative in nature.
+
 \end{abstract}

### 2020-07-21 paper/paper.tex
Some small changes in the abstract

@@
   Capturing control flow is the Achilles heel of Haskell-based
-  deeply embedding Domain Specific Languages,
-  clunkly workarounds must be employed which will be less familiar
-  than the builtin control flow mechanisms of the host language.  
+  deeply embedded domain specific languages. Clunky, awkward workarounds must be
+  employed which will be less familiar than the builtin control flow mechanisms
+  of the host language.
+
 % Say why it is an interesting problem
@@
 % Say what our solution achieves
-  In this paper, we expand the capabilies of traditional,
+
+  In this paper, we expand the capabilities of traditional,
   Haskell-based, deep embeddings with a compiler extension
@@
   With this new support, the Haskell that we capture for expressing
-  deeply embedded domain specific languges can be cleaner, more Haskell idiomatic,
+  deeply embedded domain specific languages can be cleaner, more Haskell idiomatic,
   and more declarative in nature.

### 2020-07-21 paper/paper.tex
Another refactor

@@
   Capturing control flow is the Achilles heel of Haskell-based
-  deeply embedded domain specific languages. Clunky, awkward workarounds must be
-  employed which will be less familiar than the builtin control flow mechanisms
-  of the host language.
+  deeply embedded domain specific languages.
+  Rather than use
+  the builtin control flow mechanisms, artifical control flow combinators
+  are used instead.
 
 % Say why it is an interesting problem
-  However, capturing control flow in an embedded domain specific language is crucial for
-  writing programs in a natural style, and enables the programmer to use the
+  However, capturing traditional control flow in an deeply embedded domain specific language
+  would allow for writing programs in a natural style, and enables the programmer to use the
   constructs that are already builtin to the base language, such as pattern
@@
   With this new support, the Haskell that we capture for expressing
-  deeply embedded domain specific languages can be cleaner, more Haskell idiomatic,
+  deeply embedded domain specific languages can be cleaner, Haskell-idiomatic,
   and more declarative in nature.

### 2020-07-21 paper/paper.tex
Small wibble

@@
   However, capturing traditional control flow in an deeply embedded domain specific language
-  would allow for writing programs in a natural style, and enables the programmer to use the
+  would support the writing of programs in a natural style by allowing the programmer to use the
   constructs that are already builtin to the base language, such as pattern

### 2020-07-22 paper/paper.tex
spelling
Started expansion of related work section

@@
   Rather than use
-  the builtin control flow mechanisms, artifical control flow combinators
+  the builtin control flow mechanisms, artificial control flow combinators
   are used instead.
@@
 \subsection{Overview}
+\label{sec:Overview}
 
@@
 
+There is other recent work using deep embeddings in functional languages for 
+system development.  One example is the Ivory language \cite{Elliott2015-Ivory} 
+which provides a deeply embedded DSL for use in programming high assurance
+systems,  However, it's syntax is typical of a deep EDSL and requires additional 
+keywords and structures above idiomatic Haskell.
+
+The Feldspar project~\cite{Axelsson:10:Feldspar,Axelsson,Svenningsson:13:Combining}
+is a Haskell embedding of a monadic interface
+that targets C, and focuses on high-performance. Interestingly, this work
+also attempt to make use of both deep and shallow embeddings inside a single
+implementation. Both Feldspar and and our work use some form
+of monadic reification technology~\cite{Persson:11:MonadicDSL,Svenningsson:13:Compositional,Sculthorpe:13:ConstrainedMonad}.
+
+Svenningsson and Axelsson~\cite{Svenningsson:13:Combining}
+explored combining deep and shallow embedding.  They used
+a deep embedding as a low level language, then extended
+the deep embedding with a shallow embedding written on top
+of it. Haskell type classes were used to minimize the effort of adding
+new features to the language.
+
+Elliott developed GHC plugins~\cite{github:lambda-ccc}\cite{github:reification-rules}
+for compiling Haskell to hardware~\cite{github:Elliott:Talk:2015},
+using worker-wrapper style transformations equivalent to the \verb|abs| and
+\verb|rep| transformations described in Section \ref{sec:Overview}. These plugins were later generalized to enable additional interpretations~\cite{Elliott:2017}.
+
 % \section{Citations and Bibliographies}

### 2020-07-22 paper/paper.tex
wibble

@@
 that targets C, and focuses on high-performance. Interestingly, this work
-also attempt to make use of both deep and shallow embeddings inside a single
-implementation. Both Feldspar and and our work use some form
+also attempts to make use of both deep and shallow embeddings inside a single
+implementation. Both Feldspar and our work use some form
 of monadic reification technology~\cite{Persson:11:MonadicDSL,Svenningsson:13:Compositional,Sculthorpe:13:ConstrainedMonad}.

### 2020-07-22 paper/paper.tex
Table showing the relationship between things from the Church encoding section and the things from the rest of the paper

@@
 \begin{todo}
-  Work on this subsection
+  Work on this subsection. Maybe this shouldn't be a separate section at all and
+  maybe its contents should be incorporated throughout the rest of the paper.
 \end{todo}
@@
 
+Comparing this back to our larger encoding, there is the correspondence: \\
+
+\begin{tabular}{|l|l|}
+  \hline
+  \ttt{ChurchCoproductMatch} & \ttt{SumMatchExp} \\
+  \ttt{ChurchProductMach} & \ttt{ProdMatchExp} \\
+  \ttt{CoproductMatch} & \ttt{SumMatch} \\
+  \ttt{ProductMatch} & \ttt{ProdMatch} \\
+  \ttt{ChurchLam} & \ttt{Lam} \\
+  \hline
+\end{tabular} \\
+
 \begin{todo}
@@
   those encodings handle recursive types?
+
+  Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
 \end{todo}

### 2020-07-22 paper/paper.tex
Added a TODO note about an alternate implementation of 'runIter'

@@
 \end{lstlisting}
+
+\begin{todo}
+  \ttt{runIter} has the following alternate implementation. Would this be
+  useful to use somewhere?
+
+  \begin{lstlisting}
+  runIter f = getDone . fix (f >=>)
+
+  getDone :: Iter b a -> b
+  getDone (Done x) = x
+  getDone _ = error "getDone"
+  \end{lstlisting}
+
+  Maybe this could be useful in some proofs or derivations somewhere (in this
+  paper or elsewhere)?
+
+  Note that \verb|error| is never called under any circumstance in this
+  implementation of \verb|runIter|.
+\end{todo}
 

### 2020-07-22 paper/paper.tex
Clarification in TODO about alternate implementation of 'runIter'
Remove highlighting on => part of >=>
[paper] Include 'unrep' in lambda transformation

@@
 
-  \begin{lstlisting}
+  \begin{lstlisting}[otherkeywords={}]
   runIter f = getDone . fix (f >=>)
@@
   \end{lstlisting}
+
+  where \verb|Iter| is given a standard \verb|Monad| instance (essentially the
+  same as the \verb|Either| \verb|Monad| instance). \\
 
@@
   &\\
-  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} \rewrites \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
+  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} \rewrites \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{unrep (Var @a uniq)}]$}}$}\\
+  &\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
   &\\

### 2020-07-22 paper/paper.tex
[paper] unrep elim in marking rewrite

@@
   &\\
+  &M\expr{\ttt{unrep x}} \rewrites \ttt{x}\\
   &M\expr{\ttt{x :: a}} \rewrites

### 2020-07-22 paper/paper.tex
[paper] function application rewrite

--- 2020-07-22 paper/paper.tex
@@
   &C\expr{\ttt{K x0 ... xN }} \rewrites \ttt{construct (K (unrep $M\expr{\ttt{x0}}$) ... (unrep $M\expr{\ttt{xN}}$))}\;\;\;\;\text{(Where \ttt{K} is a constructor)}\\
+  &C\expr{\ttt{f x}} \rewrites \ttt{App $M\expr{f}$ $M\expr{x}$}\\
   % &C\expr{\ttt{f x}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
@@
   &\\
-  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} \rewrites \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
+  &P\expr{\ttt{x0 x1 ... xN $\rarr$ body}} \rewrites \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... xN $\rarr$ body}}$}}\\
   &P\expr{\ttt{x $\rarr$ body}} \rewrites \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\

### 2020-07-24 paper/paper.tex
Added Scala papers+ to Related work section

@@
 
-The Feldspar project~\cite{Axelsson:10:Feldspar,Axelsson,Svenningsson:13:Combining}
+The Feldspar project~\cite{Axelsson:10:Feldspar,Svenningsson:13:Combining}
 is a Haskell embedding of a monadic interface
-that targets C, and focuses on high-performance. Interestingly, this work
-also attempts to make use of both deep and shallow embeddings inside a single
-implementation. Both Feldspar and our work use some form
-of monadic reification technology~\cite{Persson:11:MonadicDSL,Svenningsson:13:Compositional,Sculthorpe:13:ConstrainedMonad}.
+that targets C, and focuses on high-performance.  Both Feldspar and our work use 
+some form of monadic reification 
+technology~\cite{Persson:11:MonadicDSL,Svenningsson:13:Compositional,Sculthorpe:13:ConstrainedMonad}.
 
@@
 
+Yin-Yang~\cite{Jovanovic:2014} provides a framework for DSL
+embedding in Scala which uses Scala macros to provide the 
+translation from a shallow to deep embedding.  Yin-Yang
+goes beyond the translation by also providing autogeneration
+of the deep DSL from the shallow DSL.  The focus of
+Yin-Yang is in generalizing the shallow to deep transformations,
+and does not include recursive transformations.
+
+Forge~\cite{Sujeeth:2013} is a Scala based meta-EDSL framework
+which can generate both shallow and deep embeddings from
+a single EDSL specification.  Embeddings generated by Forge
+use abstract \verb|Rep| types, analogous to our EDSL's 
+\verb|E| types.  Their shallow embedding is generated
+as a pure Scala library, while the deeply embedded version
+is generated as an EDSL using the Delite~\cite{Sujeeth:2014}
+framework.
+
 Elliott developed GHC plugins~\cite{github:lambda-ccc}\cite{github:reification-rules}
 for compiling Haskell to hardware~\cite{github:Elliott:Talk:2015},
-using worker-wrapper style transformations equivalent to the \verb|abs| and
+using worker-wrapper style transformations~\cite{Gill:09:WW}
+equivalent to the \verb|abs| and
 \verb|rep| transformations described in Section \ref{sec:Overview}. These plugins were later generalized to enable additional interpretations~\cite{Elliott:2017}.

### 2020-07-24 paper/paper.tex
Improve a type signature in Church encoding section
TODO about nested pattern match -> note in paper
Conal citation in 'Motivation' and more explaination of tail recursion transformation

@@
 such as pattern matching and tail recursion, can make it challenging to "look inside"
-the lambda.
+the lambda. When these control structures are reified within the EDSL, it is
+easier to also reify lambdas, as you can simply apply them to a dummy argument
+to access their body (with the dummy value substituted for the argument).
 \begin{todo}
@@
 A major benefit of the Haskell language is that it is frequently possible derive
-a Haskell implementation from a specification, either partially or fully. When
-it is not possible to fully derive an implementation from a specification, it is
-often easier to check a Haskell implementation against a specification than it
-is in many imperative languages due, in part, to the management of side effects
-in Haskell. This is more difficult to do using an imperative language such as C,
-as they provide no restrictions on where side effects can occur.
+a Haskell implementation from a specification, either partially or fully.
+Examples of this technique include the work in \cite{Elliott-2018-ad-icfp} as
+well as the work in \cite{Elliott2019-convolution-extended}.  When it is not
+possible to fully derive an implementation from a specification, it is often
+easier to check a Haskell implementation against a specification than it is in
+many imperative languages due, in part, to the management of side effects in
+Haskell. This is more difficult to do using an imperative language such as C, as
+they provide no restrictions on where side effects can occur.
 
@@
   Cite examples where Haskell implementations have been derived specifications.
-  Conal Elliott's latest two papers (as of May 2020),
-  \cite{Elliott-2018-ad-icfp} and \cite{Elliott2019-convolution-extended}, would
-  be good examples of this.
+Conal Elliott's latest two papers (as of May 2020), \cite{Elliott-2018-ad-icfp} and \cite{Elliott2019-convolution-extended}, would be good examples of this.
+(This is now cited)
 
@@
 
-\begin{todo}
-  Should we use a single nested pattern match in the above block? Both
-  are functional (and equivalent) from a technical perspective.
-\end{todo}
+Note that the above pattern matches can be written as a single nested pattern
+match. Both forms compile down to the same Core representation.
 
@@
   :: a -> b -> (a -> b -> r) -> r
-church_mkPair = \ x y f -> f x y
+church_mkPair x y = \ f -> f x y
 \end{lstlisting}
@@
 church_fst :: ((a -> b -> a) -> r) -> r
-church_fst = \ p -> p (\ x y -> x)
+church_fst p = p (\ x y -> x)
 
 church_snd :: ((a -> b -> b) -> r) -> r
-church_snd = \ p -> p (\ x y -> y)
+church_snd p = p (\ x y -> y)
 \end{lstlisting}
@@
   :: a -> (a -> r) -> (b -> r) -> r
-church_mkLeft = \ x f g -> f x
+church_mkLeft x = \ f g -> f x
 
@@
   :: b -> (a -> r) -> (b -> r) -> r
-church_mkRight = \ y f g -> g y
+church_mkRight y = \ f g -> g y
 \end{lstlisting}
@@
 church_eitherElim
-  :: (a -> b -> r) -> a -> b -> r
-church_eitherElim = \ e f g -> e f g
+  :: ((a -> r) -> (b -> r) -> r) -> r
+church_eitherElim e = \ f g -> e f g
 \end{lstlisting}
@@
 \begin{todo}
-  Explicitly mention how the constructs in this last part
-  relate to their counterparts in the main part of the paper.
-\end{todo}
-
-\begin{todo}
   Figure out how recursive types fit into this framework. Church encodings can
@@
 \section{Representing tail recursion}
-\begin{todo}
-  Is the explanation in this section sufficient since we cite the Haskino paper,
-  which describes how the tail recursion transformation was used in that project?
-\end{todo}
-
 Tail recursion is given a direct-style representation using a simple sum type,
@@
 previously) or a final result. This type is implemented as a sum of the types
-\ttt{a} and \ttt{b}.
+\ttt{a} and \ttt{b}. Recursive calls are transformed to \ttt{Step} applications
+and non-recursive branches are wrapped in \ttt{Done} applications.
 

### 2020-07-25 paper/paper.tex
First cut at control flow diagram
Restructuring description of plugin

@@
 
+In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses in a node refers
+to the names of transformations given in the rewrite rules in Figure~\ref{fig:Rewrites}. Calls
+to \ttt{unrep :: E a -> a} are only used internally and will not exist in
+the final result of the transformation.
+
+Note that:
+\begin{itemize}
+  \item The tail recursion transformation given by $D$, $T'_{\ttt{f}}$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
+As a result, the tail recursion transformation can be used on its own.
+
+  \item The total number of marks introduced is upper bounded by the number of
+subexpressions in the original Core given to the transformation by GHC and each
+full transformation step (that is, $C$) eliminates one mark. Therefore, the
+transformation will terminate.
+\end{itemize}
+
 \clearpage
-\subsection{Overall Control Flow of the Core Plugin}
-
-In the following flowchart, the names in parentheses in a node refers to the names of transformations from the following section (Section 8.2). Note that
-calls to \ttt{unrep :: E a -> a} are only used internally and will not exist in the final result of the transformation.
-
-\begin{tikzpicture}[node distance=2cm]
-\node (start) [startstop] {Core from GHC};
-  \node (tailrec) [process, below of=start, yshift=-0.5cm]{($D, T_{\ttt{f}}, T'_{\ttt{f}}$) Transform tail recursion of definitions that have a mark (when applicable)};
-  \node (findmark) [process, below of=tailrec, yshift=-0.5cm]{($C^*$) Find marked expression and erase mark};
-    \node (handlecase) [process, below of=findmark]{($C, M$) Mark scrutinee};
-    \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{($C$) Encode as EDSL primitive};
-    \node (handleconstr) [process, right of=handleprim, xshift=1.6cm]{($C$) Wrap with \ttt{construct}};
-
-    \node (marksubs1) [process, below of=handleprim]{($C, M$) Mark sub-expressions};
-    \node (marksubs2) [process, below of=handleconstr]{($C, M$) Mark sub-expressions and wrap with \ttt{unrep}};
-    \node (betareduce) [process, below of=marksubs2]{Beta reduce \ttt{(construct ...)} expression and simplify};
-    \node (repunrep) [process, below of=betareduce]{Replace expressions of the form \ttt{(rep (unrep x))} with \ttt{x}};
-
-    \node (onebranch) [process, below of=handlecase]{($B$) Transform one branch};
-    \node (prodmatch) [process, below of=onebranch]{($P$) Transform one field in product match of current branch};
-    \node (abstractvar) [process, below of=prodmatch, yshift=-0.5cm]{($P, A$) Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
-    \node (markbody) [process, below of=abstractvar]{($P, M$) Mark branch body};
-    \node (endstep) [process, below of=markbody, right of=markbody, xshift=1.6cm]{End of one transformation step};
-
-  \draw [arrow] (start) -- (tailrec);
-  \draw [arrow] (tailrec) -- (findmark);
-  \draw [arrow] (findmark) -- node[anchor=east] {Primitive} (handleprim);
-  \draw [arrow] (findmark) -- node[anchor=west, yshift=0.1cm] {Constructor/constructor application} (handleconstr);
-  \draw [arrow] (findmark) -- node[anchor=east] {\ttt{case}} (handlecase);
-
-  \draw [arrow] (handleprim) -- (marksubs1);
-
-  \draw [arrow] (handleconstr) -- (marksubs2);
-
-  \draw [arrow] (handlecase) -- (onebranch);
-  \draw [arrow] (onebranch) -- (prodmatch);
-  \draw [arrow] (prodmatch) -- (abstractvar);
-  \draw [arrow] (abstractvar) -- (markbody);
-
-  \draw [arrow] (marksubs2) -- (betareduce);
-  \draw [arrow] (betareduce) -- (repunrep);
-
-  \draw [arrow] (marksubs1) -- (endstep);
-  \draw [arrow] (repunrep) -- (endstep);
-  \draw [arrow] (markbody) -- (endstep);
-
-  % \draw (prodmatch) edge[bend right, left] node {While fields remain} (abstractvar);
-  % \draw [arrow] (abstractvar.east) to ++(0.5,0) |- node[anchor=west] {While fields remain} (prodmatch.east);
-  \draw [arrow] (abstractvar.west) -| ++(-1.5,0) |- node[anchor=south] {While fields remain} (prodmatch.west);
-
-  \draw [arrow] (markbody.west) -| ++(-3,0) |- node[anchor=east, anchor=south] {While branches remain} (onebranch.west);
-
-  \draw [arrow] (endstep.west) -| ++(-8.5,0) |- node[anchor=west, anchor=south] {While marked expressions remain} (findmark.west);
-
-  % \draw (findmark)  edge[loop left] node {While marked expressions remain} (findmark);
-
-  % \draw (findmark)  edge[loop left] node {While marked expressions remain} (findmark);
-  % \draw (onebranch) edge[loop left] node {While branches remain} (onebranch);
-  % \Loop[label=\(While\, fields\, remain\), labelstyle={left=5pt}](prodmatch);
-  % \path (bodylam) edge [left] (onebranch);
-  % \path (bodylam) edge [loop above] node {test} (onebranch);
-  % \Loop[label=\(While\, remaining\, branches\, exist\), labelstyle={left=5pt}](onebranch)
-  % \draw [arrow] (bodylam) -- (onebranch);
-  % \draw [arrow] (bodylam.east) to  ++ (1.5,0) |- node[anchor=south] {While branches remain} (onebranch);
-  % \draw [arrow] (bodylam.east) to  ++ (1.5,0) |- node[anchor=south] {While branches remain} (onebranch);
-  % \Loop[label=\(While\, branches\, remain\), labelstyle={left=5pt}](onebranch);
-  % \Loop[label=\(While\, marked\, expressions\, remain\), labelstyle={left=5pt}](findmark);
-\end{tikzpicture}
+\begin{figure*}
+  \centering 
+  \resizebox{2.0\columnwidth}{10cm}{\includegraphics{Diagrams/Flowchart.pdf}}
+   \caption{Control Flow of the Core Plugin}
+   \label{fig:CorePlugin}
+\end{figure*}
 
 \clearpage
-\subsection{Syntactic Transformations of the Core Plugin}
+
+% \subsection{Syntactic Transformations of the Core Plugin}
 
@@
 
+\begin{figure}[hp]
+  \centering
 \begin{align*}
@@
 
-\clearpage
-
-Note that:
-\begin{itemize}
-  \item The tail recursion transformation given by $D$, $T'_{\ttt{f}}$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type.
-As a result, the tail recursion transformation can be used on its own.
-
-  \item The total number of marks introduced is upper bounded by the number of
-subexpressions in the original Core given to the transformation by GHC and each
-full transformation step (that is, $C$) eliminates one mark. Therefore, the
-transformation will terminate.
-\end{itemize}
+\captionsetup{justification=centering}
+\caption{Rewrite rules}
+\label{fig:Rewrites}
+\end{figure}
+
+% \clearpage
 

### 2020-07-26 paper/paper.tex
Rework and shorten Church encoding section

@@
 
-with the eliminators
-
-\begin{lstlisting}
-church_fst :: ((a -> b -> a) -> r) -> r
-church_fst p = p (\ x y -> x)
-
-church_snd :: ((a -> b -> b) -> r) -> r
-church_snd p = p (\ x y -> y)
-\end{lstlisting}
-
 Coproducts can be represented with the constructors
@@
 
-and the eliminator
-
-\begin{lstlisting}
-church_eitherElim
-  :: ((a -> r) -> (b -> r) -> r) -> r
-church_eitherElim e = \ f g -> e f g
-\end{lstlisting}
-
 In this representation:
@@
 Both of those representations represent the type as a function that performs the
-corresponding case analysis. A deep embedding for the operations
-\ttt{church\_mkLeft} and \\ \ttt{church\_mkRight} would look
-like:
-
-\begin{lstlisting}
-data ChurchCoproduct r where
-  Church_MkLeft
+corresponding case analysis. A deep embedding of the operations
+\ttt{churchMkPair}, \ttt{church\_mkLeft} and \ttt{church\_mkRight} is given
+by the type
+
+\begin{lstlisting}
+data Church r where
+  ChurchPair
+    :: a -> b -> (a -> b -> r) ->
+       Church r
+  ChurchLeft
     :: a -> (a -> r) -> (b -> r) ->
-       ChurchCoproduct r
-  Church_MkRight
+       Church r
+  ChurchRight
     :: b -> (a -> r) -> (b -> r) ->
-       ChurchCoproduct r
-
-church_coproductEval
-  :: ChurchCoproduct r -> r
-church_coproductEval (Church_MkLeft x f g)
-  = f x
-church_coproductEval (Church_MkRight y f g)
-  = g y
-\end{lstlisting}
-
-If we factor out the data contained in these coproducts so that it is separated from the
-case analysis functions, we can obtain
-
-\begin{lstlisting}
-data ChurchCoproductMatch t r where
-  ChurchCoproductMatch
+       Church r
+\end{lstlisting}
+
+with the evaluation function
+
+\begin{lstlisting}
+churchEval :: Church r -> r
+churchEval (ChurchPair x y f) = f x y
+churchEval (ChurchLeft x f g) = f x
+churchEval (ChurchRight y f g) = g y
+\end{lstlisting}
+
+However, we only want to keep track of the functions used for elimination. By moving
+the values actually being contained in the products and coproducts out to the
+evaluation function, we arrive at
+
+\begin{lstlisting}
+data Church' t r where
+  ChurchPair'
+    :: (a -> b -> r) ->
+       Church' (a, b) r
+
+  ChurchEither'
     :: (a -> r) -> (b -> r) ->
-       ChurchCoproductMatch (Either a b) r
-
-church_coproductMatchEval
-  :: Either a b ->
-     ChurchCoproductMatch (Either a b) r ->
-     r
-church_coproductMatchEval
-    (Left x)
-    (ChurchCoproductMatch f g)
-  = f x
-church_coproductMatchEval
-    (Right y)
-    (ChurchCoproductMatch f g)
-  = g y
-\end{lstlisting}
-
-Note that a value of type \ttt{ChurchCoproductMatch (Either a b) r} is very similar to
-a function from \ttt{Either a b} to \ttt{r}, with \ttt{church\_coproductMatchEval} allowing
-us to apply it to an argument. We can make this fact more explicit in the types with the
-change:
-
-\begin{lstlisting}
-data ChurchCoproductMatch t where
-  ChurchCoproductMatch
-    :: (a -> r) -> (b -> r) ->
-       ChurchCoproductMatch
-          (Either a b -> r)
-
-church_coproductMatchEval
-  :: ChurchCoproductMatch (t -> r) ->
-     t -> r
-church_coproductMatchEval
-    (ChurchCoproductMatch f g)
-    (Left x)
-  = f x
-church_coproductMatchEval
-    (ChurchCoproductMatch f g)
-    (Right y)
-  = g y
-\end{lstlisting}
-
-The type signature of \ttt{church\_coProductMatchEval} can be rewritten in the same form as a
-typical deep embedding evaluation function by moving the second argument into the right-hand
-side of the definition:
-
-\begin{lstlisting}
-church_coproductMatchEval
-  :: ChurchCoproductMatch t -> t
-
-church_coproductMatchEval
-    (ChurchCoproductMatch f g)
-  = \ c ->
-      case c of
+       Church' (Either a b) r
+
+churchEval' :: Church' t r -> t -> r
+churchEval' (ChurchPair' f) (x, y) = f x y
+churchEval' (ChurchEither' f g) e =
+  case e of
     Left  x -> f x
@@
 \end{lstlisting}
-
-An embedding for product matches can be similarly derived from the Church encoding for product
-types:
-
-\begin{lstlisting}
-data ChurchProductMatch t where
-  ChurchProductMatch
-    :: (a -> b -> r) ->
-       ChurchProductMatch ((a, b) -> r)
-
-church_productMatchEval
-  :: ChurchProductMatch t -> t
-
-church_productMatchEval
-    (ChurchProductMatch f)
-  = \ p ->
-      case p of
-        (x, y) -> f x y
-\end{lstlisting}
-
-As we will have another notion of function types embedded in our language, we define two
-types to distinguish between these. We define two of them so that we can also distinguish
-between coproduct matches and product matches.
-
-\begin{lstlisting}
-newtype CoproductMatch a r
-  = CoproductMatch (a -> r)
-
-newtype ProductMatch a r
-  = ProductMatch (a -> r)
-
-data ChurchCoproductMatch t where
-  ChurchCoproductMatch
-    :: (a -> r) -> (b -> r) ->
-       ChurchCoproductMatch
-         (CoproductMatch (Either a b) r)
-
-church_coproductMatchEval
-  :: ChurchCoproductMatch t -> t
-
-church_coproductMatchEval
-    (ChurchCoproductMatch f g)
-  = CoproductMatch (\ c ->
-      case c of
-        Left x  -> f x
-        Right y -> g y)
-\end{lstlisting}
-
-The definition for product follows similarly, using \\ \ttt{ProductMatch}.
-
-Here we make the observation that algebraic datatypes are structured as coproducts of products. With
-this information in hand, we can combine \ttt{ChurchCoproductMatch} and \ttt{ChurchProductMatch}.
-
-\begin{lstlisting}
-newtype CoproductMatch a r
-  = CoproductMatch
-      { runCoproductMatch :: a -> r }
-
-newtype ProductMatch a r
-  = ProductMatch
-      { runProductMatch :: a -> r }
-
-data ChurchMatch t where
-  ChurchCoproductMatch
-    :: ChurchMatch (ProductMatch a r) ->
-       ChurchMatch (CoproductMatch b r) ->
-       ChurchMatch
-         (CoproductMatch (Either a b) r)
-
-  ChurchProductMatch
-    :: ChurchMatch (a -> ProductMatch b r)
-       -> ChurchMatch
-            (ProductMatch (a, b) r)
-
-  ChurchOneCoproductMatch
-    :: ChurchMatch (ProductMatch a r) ->
-       ChurchMatch (CoproductMatch a r)
-
-  ChurchOneProductMatch
-    :: (a -> r) ->
-       ChurchMatch (ProductMatch a r)
-
-  ChurchNullaryProductMatch
-    :: r -> ChurchMatch (ProductMatch a r)
-
-  ChurchLam
-    :: (a -> b) -> ChurchMatch (a -> b)
-
-church_matchEval :: ChurchMatch t -> t
-church_matchEval
-    (ChurchCoproductMatch f g) =
-  CoproductMatch (\ c ->
-    case c of
-      Left x  ->
-        runProductMatch
-          (church_matchEval f) x
-      Right y ->
-        runCoproductMatch
-          (church_matchEval g) y)
-
-church_matchEval
-    (ChurchProductMatch f) =
-  ProductMatch (\ p ->
-    case p of
-      (x, y) ->
-        runProductMatch
-          (church_matchEval f x) y)
-
-church_matchEval
-    (ChurchOneCoproductMatch f) =
-  CoproductMatch (\ x ->
-    runProductMatch
-      (church_matchEval f) x)
-
-church_matchEval
-    (ChurchOneProductMatch f) =
-  ProductMatch (\ x -> f x)
-
-church_matchEval
-    (ChurchNullaryProductMatch x) =
-  ProductMatch (\ _ -> x)
-
-church_matchEval
-    (ChurchLam f) = \ x -> f x
-\end{lstlisting}
-
-Comparing this back to our larger encoding, there is the correspondence: \\
-
-\begin{tabular}{|l|l|}
-  \hline
-  \ttt{ChurchCoproductMatch} & \ttt{SumMatchExp} \\
-  \ttt{ChurchProductMach} & \ttt{ProdMatchExp} \\
-  \ttt{CoproductMatch} & \ttt{SumMatch} \\
-  \ttt{ProductMatch} & \ttt{ProdMatch} \\
-  \ttt{ChurchLam} & \ttt{Lam} \\
-  \hline
-\end{tabular} \\
 

### 2020-07-26 paper/paper.tex
Can now remove todos from output PDF by commenting out one line of LaTeX
Tweak of sentence in related work

@@
 
+\newcommand{\showtodos}{}  % Comment out to hide todos
+
 \newenvironment{todo}
-  {\begin{tcolorbox}
-   \textbf{TODO}:
-  }
-  {\end{tcolorbox}
+  {\ifthenelse{\isundefined{\showtodos}}{\comment}{\begin{tcolorbox}
+    \textbf{TODO}:}}
+  {\ifthenelse{\isundefined{\showtodos}}{\endcomment}{\end{tcolorbox}}
   }
@@
 is a Haskell embedding of a monadic interface
-that targets C, and focuses on high-performance.  Both Feldspar and our work use 
-some form of monadic reification 
-technology~\cite{Persson:11:MonadicDSL,Svenningsson:13:Compositional,Sculthorpe:13:ConstrainedMonad}.
+that targets C, and focuses on high-performance.  Both Feldspar and our work
+mix deep and shallow language constructs ~\cite{Persson:11:MonadicDSL,Svenningsson:13:Compositional,Sculthorpe:13:ConstrainedMonad}.
 

### 2020-07-27 paper/paper.tex
Added recursion example

@@
 In the technique presented here, we do not need to allocate thunks or closures.
-We actually do not to use higher-order functions at all in this tail recursion
+We actually do not need to use higher-order functions at all in this tail recursion
 representation.
 
-In the \ttt{E} type, this is used to represent tail recursion by
-the following constructors:
+In the \ttt{E} type of our DSL, these types of data structures are used to represent tail 
+recursion by using the following constructors.
 
@@
   ...
+\end{lstlisting}
+
+The process of transforming a tail recursive function in the DSL to a 
+representation which could be passed into the C backend and used
+to generate a traditional iterative loop goes as follows.
+
+Starting with the tail recursive function  \ttt{isEven}:
+
+\begin{lstlisting}
+isEven :: Int -> Bool
+isEven x =
+    case x == 0 of
+      True -> True
+      False ->
+        case x == 1 of
+          True -> False
+          False -> isEven (x - 2)
+\end{lstlisting}
+
+We translate this to the expression language, and then apply two of
+the constructors we have defined.  The first of these is an initial 
+\ttt{DoneExp} which we then wrap with a \ttt{TailRec} to provide the
+representation of the iteration execution:
+
+\begin{lstlisting}
+isEven' :: E Int -> E (Int -> Bool)
+isEven' x =
+  TailRec ( DoneExp (
+    IfThenElseExp (Equal x (rep 0))
+      TrueExp
+      (IfThenElseExp (Equal x (rep 1))
+        FalseExp
+        (IsEven' (Sub x (rep 2))))
+\end{lstlisting}
+
+The transformation then proceeds by pushing the DoneExp constructor through
+the conditionals, and the recursive call is transformed into a StepExp constructor.
+This leaves us the the transformed example:
+
+\begin{lstlisting}
+isEven'' :: E Int -> E (Int -> Bool)
+isEven'' x =
+  TailRec (
+    IfThenElseExp (Equal x (rep 0))
+      (DoneExp TrueExp)
+      (IfThenElseExp (Equal x (rep 1))
+        (DoneExp FalseExp)
+        (StepExp (Sub x (rep 2))))
 \end{lstlisting}

### 2020-07-29 paper/paper.tex
A small table connecting Church encoding section back to the rest of the paper

--- 2020-07-26 paper/paper.tex
@@
 \subsection{Connection to Church encodings}
-\begin{todo}
-  Work on this subsection. Maybe this shouldn't be a separate section at all and
-  maybe its contents should be incorporated throughout the rest of the paper.
-\end{todo}
+% \begin{todo}
+%   Work on this subsection. Maybe this shouldn't be a separate section at all and
+%   maybe its contents should be incorporated throughout the rest of the paper.
+% \end{todo}
 
@@
 \end{lstlisting}
+
+The \verb|Church'| type forms a large part of the foundation of the pattern
+matching aspect of the \verb|E| type. A correspondence between the two
+is given in the table below.\\
+
+{\small
+\noindent\begin{tabular}{|p{3.4cm}|p{4.4cm}|}
+  \hline
+  \ttt{Church' (a, b) r} & \ttt{E (SumMatch (a, b) r} \\
+  \hline
+  \ttt{Church' (Either a b) r} & \ttt{E (ProdMatch (Either a b) r)}\\
+  \hline
+\end{tabular}
+}
 

### 2020-07-30 paper/paper.tex
wibble

@@
         FalseExp
-        (IsEven' (Sub x (rep 2))))
+        (isEven' (Sub x (rep 2))))
 \end{lstlisting}

### 2020-07-30 paper/paper.tex
Clean up some less relevant TODOs and minor formatting

@@
 \subsection{Translation Example}
-\begin{todo}
-  Does it make sense for this to be a subsection of the intro?
-\end{todo}
+
+% \begin{todo}
+%   Does it make sense for this to be a subsection of the intro?
+% \end{todo}
 
@@
 
-\begin{todo}
-  Look into \ttt{generic-sop} and compare then cite corresponding paper.
-\end{todo}
+% \begin{todo}
+%   Look into \ttt{generic-sop} and compare then cite corresponding paper.
+% \end{todo}
 
@@
 
-\begin{todo}
-  Should we make a synonym for this constraint since it is used in
-multiple places? Like this:
-\begin{lstlisting}
-type CanonicalIsIdem t =
-  ERepTy (ERepTy t) ~ ERepTy t
-\end{lstlisting}
-\end{todo}
+% \begin{todo}
+%   Should we make a synonym for this constraint since it is used in
+% multiple places? Like this:
+% \begin{lstlisting}
+% type CanonicalIsIdem t =
+%   ERepTy (ERepTy t) ~ ERepTy t
+% \end{lstlisting}
+% \end{todo}
 
@@
 The \verb|Church'| type forms a large part of the foundation of the pattern
-matching aspect of the \verb|E| type. A correspondence between the two
+matching aspect of the \verb|E| type. A rough correspondence between the two
 is given in the table below.\\
@@
 \section{Related Work}
-\begin{todo}
-  Does this need more fleshing out?
-\end{todo}
+
+% \begin{todo}
+%   Does this need more fleshing out?
+% \end{todo}
 
@@
 cCall fnName args =
-  fnName <> "(" <> intercalate ", " args <> ")"
+  fnName
+  <> "(" <> intercalate ", " args <> ")"
 
@@
 cIndex :: CCode -> Int -> CCode
-cIndex x i = "(" <> x <> ")[" <> show i <> "]"
+cIndex x i =
+  "(" <> x <> ")[" <> show i <> "]"
 
@@
 
-macroDef :: CName -> [CName] -> [CCode] -> CCode
+macroDef
+  :: CName -> [CName] -> [CCode]
+     -> CCode
 macroDef name args theLines =

### 2020-07-30 paper/paper.tex
Added citation to [Andy Gill 2014] paper in the context of capturing lambdas

@@
 the lambda. When these control structures are reified within the EDSL, it is
-easier to also reify lambdas, as you can simply apply them to a dummy argument
-to access their body (with the dummy value substituted for the argument).
-\begin{todo}
-  Maybe add brief summary of relationship to lambdas?
-\end{todo}
+easier to also reify lambdas using the dummy argument technique described in~\cite{Gill:2014:DSLSynth}.
+
+% \begin{todo}
+%   Maybe add brief summary of relationship to lambdas?
+% \end{todo}
 

### 2020-07-30 paper/paper.tex
Clean up a couple of todos and use 'tiny' style for code in appendices

@@
 \begin{todo}
-  Cite examples where Haskell implementations have been derived specifications.
+  % Cite examples where Haskell implementations have been derived specifications.
 Conal Elliott's latest two papers (as of May 2020), \cite{Elliott-2018-ad-icfp} and \cite{Elliott2019-convolution-extended}, would be good examples of this.
@@
 
-\begin{todo}
-  Figure out how recursive types fit into this framework. Church encodings can
-  handle recursive types and so can our representation. This is also the case
-  for Scott encodings, etc, and is the difference between Church and Scott
-  encodings.
-
-  Does our way of handling recursive types correspond to how any of
-  those encodings handle recursive types?
-
-  Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
-\end{todo}
+% \begin{todo}
+%   Figure out how recursive types fit into this framework. Church encodings can
+%   handle recursive types and so can our representation. This is also the case
+%   for Scott encodings, etc, and is the difference between Church and Scott
+%   encodings.
+
+%   Does our way of handling recursive types correspond to how any of
+%   those encodings handle recursive types?
+
+%   Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
+% \end{todo}
 
@@
 
-\begin{todo}
-  \ttt{runIter} has the following alternate implementation. Would this be
-  useful to use somewhere?
-
-  \begin{lstlisting}[otherkeywords={}]
-  runIter f = getDone . fix (f >=>)
-
-  getDone :: Iter b a -> b
-  getDone (Done x) = x
-  getDone _ = error "getDone"
-  \end{lstlisting}
-
-  where \verb|Iter| is given a standard \verb|Monad| instance (essentially the
-  same as the \verb|Either| \verb|Monad| instance). \\
-
-  Maybe this could be useful in some proofs or derivations somewhere (in this
-  paper or elsewhere)?
-
-  Note that \verb|error| is never called under any circumstance in this
-  implementation of \verb|runIter|.
-\end{todo}
+% \begin{todo}
+%   \ttt{runIter} has the following alternate implementation. Would this be
+%   useful to use somewhere?
+
+%   \begin{lstlisting}[otherkeywords={}]
+%   runIter f = getDone . fix (f >=>)
+
+%   getDone :: Iter b a -> b
+%   getDone (Done x) = x
+%   getDone _ = error "getDone"
+%   \end{lstlisting}
+
+%   where \verb|Iter| is given a standard \verb|Monad| instance (essentially the
+%   same as the \verb|Either| \verb|Monad| instance). \\
+
+%   Maybe this could be useful in some proofs or derivations somewhere (in this
+%   paper or elsewhere)?
+
+%   Note that \verb|error| is never called under any circumstance in this
+%   implementation of \verb|runIter|.
+% \end{todo}
 
@@
 \section{Additional Code for Standard Haskell Semantics Implementation}
-\begin{lstlisting}
+\begin{lstlisting}[basicstyle=\tiny]
 absEnv :: forall t. Env -> E t -> t
@@
 
+\vfill\break
+
 \section{A small EDSL for C code generation used internally by the C backend}
-\begin{lstlisting}
+\begin{lstlisting}[basicstyle=\tiny]
 type CCode = String

### 2020-07-30 paper/paper.tex
Wording
Work on some technical details of 'isEven' tail recursion example

@@
 
-Note that the above pattern matches can be written as a single nested pattern
+The above pattern matches could also be written as a single nested pattern
 match. Both forms compile down to the same Core representation.
@@
 \ttt{E (ProdMatch x y)} is equivalent to a curried function from \ttt{x} to
-\ttt{y} in the expression language .  Note that \ttt{s} is a (potentially
+\ttt{y} in the expression language.  \ttt{s} is a (potentially
 nested) pair type. For example, \ttt{E (ProdMatch (a, (b, c)) r)} is
@@
 isEven x =
+  internalize (externalize (
     case x == 0 of
@@
           True -> False
-          False -> isEven (x - 2)
-\end{lstlisting}
-
-We translate this to the expression language, and then apply two of
-the constructors we have defined.  The first of these is an initial 
-\ttt{DoneExp} which we then wrap with a \ttt{TailRec} to provide the
-representation of the iteration execution:
-
-\begin{lstlisting}
-isEven' :: E Int -> E (Int -> Bool)
-isEven' x =
-  TailRec ( DoneExp (
-    IfThenElseExp (Equal x (rep 0))
-      TrueExp
-      (IfThenElseExp (Equal x (rep 1))
-        FalseExp
-        (isEven' (Sub x (rep 2))))
-\end{lstlisting}
-
-The transformation then proceeds by pushing the DoneExp constructor through
-the conditionals, and the recursive call is transformed into a StepExp constructor.
-This leaves us the the transformed example:
-
-\begin{lstlisting}
-isEven'' :: E Int -> E (Int -> Bool)
-isEven'' x =
-  TailRec (
-    IfThenElseExp (Equal x (rep 0))
+          False -> isEven (x - 2)))
+\end{lstlisting}
+
+% We translate this to the expression language, and then apply two of
+% the constructors we have defined.  The first of these is an initial 
+% \ttt{Done} which we then wrap with a \ttt{runIter} to provide the
+% representation of the iteration execution.
+% The transformation then proceeds by pushing the DoneExp constructor through
+% the conditionals, and the recursive call is transformed into a StepExp constructor.
+% This leaves us the the transformed example:
+
+We first translate this to use the \verb|Iter| type. The lambda is brought
+inside of the \ttt{internalize (externalize ...)} call and wrapped in a call to
+\ttt{runIter}. This inner expression is then eta expanded. Finally, each \verb|case|
+branch which does not contain a recursive call is wrapped in a \verb|Done| and
+every recursive call is substitute for an application of the \verb|Step| constructor.
+
+In the following code, \verb|eta| is a fresh variable.
+
+\begin{lstlisting}
+isEven' :: Int -> Bool
+isEven' =
+  internalize (externalize
+    (\ eta ->
+      runIter (
+        \ x ->
+          case x == 0 of
+            True -> Done True
+            False ->
+              case x == 1 of
+                True -> Done False
+                False -> Step (x - 2))
+        eta))
+\end{lstlisting}
+
+After this initial tail recursion transformation, the rest of the plugin brings this
+result into the expression language. \verb|TailRec| is substituted for \verb|runIter|,
+\verb|DoneExp| is substituted for \verb|Done| and \verb|StepExp| is substituted for \verb|Step|.
+Next, the pattern matching transformation occurs, producing the final result:
+
+\break
+
+\begin{lstlisting}
+isEven'' :: Int -> Bool
+isEven'' =
+  internalize (
+    (Lam (Var 1)
+      (App
+        (TailRec (Lam (Var 2)
+          (IfThenElseExp (Equal (Var 2)
+                                (rep 0))
             (DoneExp TrueExp)
-      (IfThenElseExp (Equal x (rep 1))
+            (IfThenElseExp (Equal (Var 1)
+                                  (rep 1))
               (DoneExp FalseExp)
-        (StepExp (Sub x (rep 2))))
+              (StepExp (Sub (Var 2)
+                            (rep 2)))))))
+        (Var 1))))
 \end{lstlisting}

### 2020-07-30 paper/paper.tex
Briefly tie the C backend back to the rest of the transformation system using a previously mentioned example

@@
 
-See Appendix A for more implementation details of \ttt{absEnv}.
+See Appendix A for implementation details of \ttt{absEnv}.
 
@@
 \ttt{EXPR\_PAIR} and a \ttt{semantic\_tag} field with the value \ttt{EXPR\_COMPLEX}.
+
+A C code generation function \ttt{genC} is implemented
+
+\begin{lstlisting}
+genC :: E a -> String
+...
+\end{lstlisting}
+
+To use this, for example, with the previously mentioned \ttt{isEven} implementation we
+can write:
+
+\begin{lstlisting}
+isEvenE :: E (Int -> Bool)
+isEvenE =
+  externalize (
+    case x == 0 of
+      True -> True
+      False ->
+        case x == 1 of
+          True -> False
+          False -> isEvenE (x - 2))
+
+isEvenC :: String
+isEvenC = genC isEvenE
+\end{lstlisting}
 

### 2020-07-30 paper/paper.tex
Better section references

@@
 \begin{itemize}
-\item A representation of pattern matching in a Haskell EDSL (Section 3)
-\item A representation of lambdas (Section 5) in the EDSL, taking advantage of
+  \item A representation of pattern matching in a Haskell EDSL (Section~\ref{sec:PatRep})
+  \item A representation of lambdas (Section~\ref{sec:LamRep}) in the EDSL, taking advantage of
   the fact that both pattern matching and tail recursion are already captured in
@@
   into this EDSL for pattern matching, extended with tail recursion,
-  lambdas and primitive operations (Section 8)
-\item An implementation of a interpreter for the "standard" semantics of the EDSL (Section 6).
-\item An outline of a C backend for this EDSL (Section 7).
+  lambdas and primitive operations (Section~\ref{sec:CorePlugin})
+\item An implementation of a interpreter for the "standard" semantics of the EDSL (Section~\ref{sec:StdSemantics}).
+\item An outline of a C backend for this EDSL (Section~\ref{sec:CBackend}).
 \end{itemize}
@@
 Readers who are primarily interested only in the encoding of pattern matching itself can direct
-most of their attention to Sections 2, 3 and 6. Most of the technical aspects of
-the transformation performed by the Core plugin are described in Section 8.
+most of their attention to Sections~\ref{sec:ADTRep},~\ref{sec:PatRep} and~\ref{sec:StdSemantics}. Most of the technical aspects of
+the transformation performed by the Core plugin are described in Section~\ref{sec:CorePlugin}.
 
@@
 \section{Representing Algebraic Datatypes}
+\label{sec:ADTRep}
 
@@
 allow all combinations of basic sum types and product types to be encoded, in the
-manner described at the beginning of Section 2.
+manner described at the beginning of Section~\ref{sec:ADTRep}.
 
@@
 \section{Representing pattern matches}
+\label{sec:PatRep}
 
@@
 \subsection{An aside on \ttt{ProdMatch} and \ttt{SumMatch} values}
+\label{sec:MatchAside}
 
@@
 \section{Representing lambdas}
+\label{sec:LamRep}
+
 This representation of pattern matching depends on the ability to bring
@@
 \section{Recovering standard semantics for pattern matches}
+\label{sec:StdSemantics}
 
@@
 types, respectively. These helper functions are based on the mechanism
-described in section 3.4.
+described in Section~\ref{sec:MatchAside}.
 
@@
 \section{C Backend}
+\label{sec:CBackend}
+
 \begin{todo}
@@
 \section{Core Plugin}
+\label{sec:CorePlugin}
 
@@
 \ttt{a}), then the marking procedure ignores it and does not wrap it with
-\ttt{externalize} (see $M$ in Section 8.2).
+\ttt{externalize} (see $M$ in Figure~\ref{fig:Rewrites}).
 
@@
 equivalent to the \verb|abs| and
-\verb|rep| transformations described in Section \ref{sec:Overview}. These plugins were later generalized to enable additional interpretations~\cite{Elliott:2017}.
+\verb|rep| transformations described in Section~\ref{sec:Overview}. These plugins were later generalized to enable additional interpretations~\cite{Elliott:2017}.
 

### 2020-07-31 paper/paper.tex
Removed a widow
wibble
Making the abstract one paragraph
Note the removal of the word traditional (for spacing only).

@@
   are used instead.
-
 % Say why it is an interesting problem
@@
 % Say what our solution achieves
-
-  In this paper, we expand the capabilities of traditional,
-  Haskell-based, deep embeddings with a compiler extension
+  In this paper, we expand the capabilities of
+  Haskell-based deep embeddings with a compiler extension
   for reifying conditionals and pattern matching.
@@
   construct :: t -> E (ERepTy t)
-
   rep :: t -> E t
@@
   unrep' :: ERepTy t -> t
-
   rep' :: t -> ERepTy t
@@
   rep (Right y) = RightExp (rep y)
-
   unrep' = id
@@
 
-\break
+% \break
 
@@
 
-A C code generation function \ttt{genC} is implemented
+A C code generation function \ttt{genC} is implemented:
 

### 2020-07-31 paper/paper.tex
Fix citation to correct work

@@
 the lambda. When these control structures are reified within the EDSL, it is
-easier to also reify lambdas using the dummy argument technique described in~\cite{Gill:2014:DSLSynth}.
+easier to also reify lambdas using the dummy argument technique described in~\cite{Elliott:03:CompileDSEL-JFP}.
 

### 2020-07-31 paper/paper.tex
Adding cite to the two DSL references
spacing for section 1.1

@@
   }
+\newenvironment{todont}
+               {\comment}
+               {\endcomment}
 
@@
 
-\begin{todo}
+\begin{todont}
   Expand?
-\end{todo}
+\end{todont}
 
@@
 
+\newpage
 Haskell EDSLs that support powerful language features like pattern matching
@@
 
+The basis of deep embeddings in functional languages is well understood.
+\cite{Elliott:03:CompileDSEL-JFP} explains the basic idea, and \cite{Gill:2014:DSLSynth}
+gives an acceessable and more recent account of the area.
+
 A similar EDSL-oriented representation of pattern matching is given in

### 2020-07-31 paper/paper.tex
Changed Figure 1 to be full page
Fix a small technical detail

--- 2020-07-31 paper/paper.tex
@@
   are used instead.
+
 % Say why it is an interesting problem
@@
 % Say what our solution achieves
-  In this paper, we expand the capabilities of
-  Haskell-based deep embeddings with a compiler extension
+
+  In this paper, we expand the capabilities of traditional,
+  Haskell-based, deep embeddings with a compiler extension
   for reifying conditionals and pattern matching.
@@
       &\rewrites \begin{cases}
-        \ttt{f = internalize $C^{*}\expr{\ttt{externalize (runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}))}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
+        \ttt{f = internalize $C^{*}\expr{\ttt{externalize ($\lambda \eta \rarr$ runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}) $\eta$)}}$} & \text{\parbox{2.5cm}{if \ttt{f} occurs free in \ttt{\{ ...
+\}}, where $\eta$ is a fresh name}}\\
+        \\
         \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}

### 2020-07-31 paper/paper.tex

--- 2020-07-31 paper/paper.tex
@@
   \centering 
-  \resizebox{2.0\columnwidth}{10cm}{\includegraphics{Diagrams/Flowchart.pdf}}
+  \resizebox{2.0\columnwidth}{20cm}{\includegraphics{Diagrams/Flowchart.pdf}}
    \caption{Control Flow of the Core Plugin}

### 2020-07-31 paper/paper.tex
Properly deal with internalized and non-internalized definitions in rewrite rules

@@
 \begin{align*}
-  D&\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}\\
-      &\rewrites \begin{cases}
-        \ttt{f = internalize $C^{*}\expr{\ttt{externalize ($\lambda \eta \rarr$ runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}) $\eta$)}}$} & \text{\parbox{2.5cm}{if \ttt{f} occurs free in \ttt{\{ ...
+  &D\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}
+      \rewrites I(D(\ttt{f x = externalize (case s of \{ ... \})}))\\
+  &D\expr{\ttt{f x = externalize (case s of \{ ... \})}}\\
+      &\quad\rewrites \begin{cases}
+        \ttt{f = $C^{*}\expr{\ttt{externalize ($\lambda \eta \rarr$ runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}) $\eta$)}}$} & \text{\parbox{5cm}{if \ttt{f} occurs free in \ttt{\{ ...
 \}}, where $\eta$ is a fresh name}}\\
         \\
-        \ttt{f x = internalize $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
-      \end{cases}
-\end{align*}
-
-\begin{align*}
-  T'_{\ttt{f}}&\expr{\ttt{K x0 ...$_U$ xN $\rarr$ case s of \{ ...$_V$ \}; ...$_W$}}\\
-    &\rewrites
+        \ttt{f x = $C^{*}\expr{\ttt{externalize (case s of \{ ... \})}}$} & \text{otherwise}
+      \end{cases}\\
+  &I\expr{\ttt{f x = externalize e}} \rewrites \ttt{f x = internalize (externalize e)}\\
+  &I\expr{\ttt{f = externalize e}} \rewrites \ttt{f = internalize (externalize e)}\\
+  &T'_{\ttt{f}}\expr{\ttt{K x0 ...$_U$ xN $\rarr$ case s of \{ ...$_V$ \}; ...$_W$}}\\
+    &\quad\rewrites
         \ttt{K x0 ...$_U$ xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_V$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{ ...$_W$ }}$}\\
-  T'_{\ttt{f}}&\expr{\ttt{K x0 ...$_U$ xN $\rarr$ body0; ...$_V$ }}\\
-    &\rewrites
+  &T'_{\ttt{f}}\expr{\ttt{K x0 ...$_U$ xN $\rarr$ body0; ...$_V$ }}\\
+    &\quad\rewrites
       \begin{cases}
@@
       \end{cases}\\
-  T'_{\ttt{f}}&\expr{\varepsilon} \rewrites \varepsilon\;\;\; \text{(Note: $\varepsilon$ represents the empty string)}\\
-  T_{\ttt{f}}&\expr{\ttt{case s of \{ ... \}}}\\
-    &\rewrites
+  &T'_{\ttt{f}}\expr{\varepsilon} \rewrites \varepsilon\;\;\; \text{(Note: $\varepsilon$ represents the empty string)}\\
+  &T_{\ttt{f}}\expr{\ttt{case s of \{ ... \}}}\\
+    &\quad\rewrites
       \begin{cases}
@@
         \ttt{Done (case s of \{ ... \})} & \text{otherwise}
-      \end{cases}
-\end{align*}
-\begin{align*}
+      \end{cases}\\
   &C^{*}\expr{\ttt{x}} \rewrites
@@
     \end{cases}\\
-\end{align*}
-\begin{align*}
   &C\expr{\ttt{runIter x}} \rewrites \ttt{TailRec $M\expr{\ttt{x}}$}\\
@@
   % &C\expr{\ttt{x}} \rewrites \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
-  &\\
   &B\expr{\ttt{K x0 ...$_U$ xN $\rarr$ body0; ...$_V$}} \rewrites \ttt{SumMatchExp }P\expr{\ttt{x0 ...$_U$ xN $\rarr$ body0}}\; B\expr{\ttt{ ...$_V$ }}\\
@@
   &B\expr{\ttt{$\varepsilon$}} \rewrites \ttt{EmptyMatch}\\
-  &\\
   &P\expr{\ttt{x0 x1 ... xN $\rarr$ body}} \rewrites \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... xN $\rarr$ body}}$}}\\
@@
   &P\expr{\ttt{ $\rarr$ body}} \rewrites \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
-  &\\
   &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} \rewrites \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{unrep (Var @a uniq)}]$}}$}\\
   &\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
-  &\\
   &M\expr{\ttt{unrep x}} \rewrites \ttt{x}\\

### 2020-08-01 paper/paper.tex
Center figure 2 caption
Improved TailRec rewrite
wibble
cleaning up page 2.
reformat for style
Cite Boehm 1985, minor formatting

@@
 
-\begin{todo}
+\begin{todont}
   % Cite examples where Haskell implementations have been derived specifications.
@@
   semantics would also be a good example.
-\end{todo}
+\end{todont}
 
@@
     (case p of
-      ComplexPair a b ->
-        case a of
-          a_real :+ _ ->
-            case b of
+      ComplexPair a b -> case a of
+          a_real :+ _ -> case b of
               b_real :+ _ ->
@@
   CaseExp ::
-    (ERep t, ERepTy (ERepTy t) ~ ERepTy t)
+    (ERep t,
+     ERepTy (ERepTy t) ~ ERepTy t)
       => ERep t
@@
 
+\newpage
 \begin{lstlisting}
@@
 
-\ttt{E (ProdMatch x y)} is equivalent to a curried function from \ttt{x} to
-\ttt{y} in the expression language.  \ttt{s} is a (potentially
+\ttt{E (ProdMatch s t)} is equivalent to a curried function from \ttt{s} to
+\ttt{t} in the expression language.  \ttt{s} is a (potentially
 nested) pair type. For example, \ttt{E (ProdMatch (a, (b, c)) r)} is
@@
 Algebraic datatypes can be represented as lambda calculus terms using
-\textit{Church encodings}. For example, the pair constructor can be
+\textit{Church encodings}.~\cite{Boehm:1985:AutoSynth} For example, the pair constructor can be
 represented as
@@
       True -> True
-      False ->
-        case x == 1 of
+      False -> case x == 1 of
           True -> False
@@
             True -> Done True
-            False ->
-              case x == 1 of
+            False -> case x == 1 of
                 True -> Done False
@@
   (Typeable a, Typeable b)
-    => Name a -> Name b -> Maybe (a :~: b)
+   => Name a
+   -> Name b
+   -> Maybe (a :~: b)
 namesEq (Name n) (Name n') =
@@
 These bindings are grouped together in the \ttt{Env} type:
-
+\newpage
 \begin{lstlisting}
@@
 
+\newpage
 \section{C Backend}
@@
 
-\begin{todo}
+\begin{todont}
   Should this section be expanded?
-\end{todo}
+\end{todont}
 
@@
 can write:
-
+\newpage
 \begin{lstlisting}
@@
 
-\begin{figure}[hp]
+\begin{figure*}[hp]
   \centering
@@
     \end{cases}\\
-  &C\expr{\ttt{runIter x}} \rewrites \ttt{TailRec $M\expr{\ttt{x}}$}\\
+  &C\expr{\ttt{runIter f x}} \rewrites \ttt{App (TailRec $M\expr{\ttt{f}}$) $M\expr{\ttt{x}}$}\\
   &C\expr{\ttt{x + y}} \rewrites \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
@@
 \label{fig:Rewrites}
-\end{figure}
+\end{figure*}
 
@@
 
-\vfill\break
+% \vfill\break
 

### 2020-08-01 paper/paper.tex
wibble

@@
 which provides a deeply embedded DSL for use in programming high assurance
-systems,  However, it's syntax is typical of a deep EDSL and requires additional
+systems,  However, its syntax is typical of a deep EDSL and requires additional
 keywords and structures above idiomatic Haskell.

### 2020-08-01 paper/paper.tex
wibble
Modified placeholder DOI and ISBN
Move note about \varepsilon in rewrite rules

@@
 \acmYear{2020}
-\acmDOI{1}
+\acmDOI{XX.XXXX/XXXXXXX}
 
@@
 \acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
+\acmISBN{XXX-X-XXXX-XXXX-X/XX/XX}
 
@@
 Therefore, \ttt{E (SumMatch (ERepTy t) r)} represents a computation which destructures
-a value of type \ttt{E (ERepTy t)} and produces a value of type \ttt{E b}.
+a value of type \ttt{E (ERepTy t)} and produces a value of type \ttt{E r}.
 
@@
 
-\ttt{E (ProdMatch s t)} is equivalent to a curried function from \ttt{s} to
-\ttt{t} in the expression language.  \ttt{s} is a (potentially
+\ttt{E (ProdMatch s t)} is equivalent to a function from \ttt{s} to
+\ttt{t} in the expression language where \ttt{s} is a (potentially
 nested) pair type. For example, \ttt{E (ProdMatch (a, (b, c)) r)} is
@@
 In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses in a node refers
-to the names of transformations given in the rewrite rules in Figure~\ref{fig:Rewrites}. Calls
+to the names of transformations given in the rewrite rules in Figure~\ref{fig:Rewrites}. In Figure~\ref{fig:Rewrites},
+$\varepsilon$ represents an empty term (in the context of the term rewriting system used in that Figure). Calls
 to \ttt{unrep :: E a -> a} are only used internally and will not exist in
@@
       \end{cases}\\
+  &\\
   &I\expr{\ttt{f x = externalize e}} \rewrites \ttt{f x = internalize (externalize e)}\\
@@
       \end{cases}\\
-  &T'_{\ttt{f}}\expr{\varepsilon} \rewrites \varepsilon\;\;\; \text{(Note: $\varepsilon$ represents the empty string)}\\
+  &T'_{\ttt{f}}\expr{\varepsilon} \rewrites \varepsilon\\
   &T_{\ttt{f}}\expr{\ttt{case s of \{ ... \}}}\\

### 2020-08-01 paper/paper.tex
Different wording in tail rec description
wibble

@@
 In the technique presented here, we do not need to allocate thunks or closures.
-We actually do not need to use higher-order functions at all in this tail recursion
-representation.
-
-In the \ttt{E} type of our DSL, these types of data structures are used to represent tail 
-recursion by using the following constructors.
+We do not need to use higher-order functions for this representation, other than \ttt{runIter} itself.
+
+In the \ttt{E} type of our DSL, this form of tail recursion representation is
+given a deep embedding by the following constructors:
 
@@
 
-We first translate this to use the \verb|Iter| type. The lambda is brought
+We first transform this to use the \verb|Iter| type. The lambda is brought
 inside of the \ttt{internalize (externalize ...)} call and wrapped in a call to
@@
 branch which does not contain a recursive call is wrapped in a \verb|Done| and
-every recursive call is substitute for an application of the \verb|Step| constructor.
+every recursive call is transformed into an application of the \verb|Step| constructor.
 

### 2020-08-01 paper/paper.tex
Citation for SSA
Better comparision to SSA
wibble

@@
 
-This backend is based on a globally unique name generator. Similar to
-SSA form, a new C variable is created for each Haskell subexpression.
+This backend is based on a globally unique name generator. A new C variable is created for each Haskell subexpression.
+The result appears similar to static single assignment (SSA) form~\cite{Rosen:1988:SSA}.
 It differs from SSA, however, in that these variables can be assigned multiple
@@
 
-In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses in a node refers
+In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses in a node refer
 to the names of transformations given in the rewrite rules in Figure~\ref{fig:Rewrites}. In Figure~\ref{fig:Rewrites},

### 2020-09-24 paper/paper.tex
Removed 'not for distribution'
Short title
Citations for Bird and Gibbons

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{On Adding Pattern Matching to Haskell-based
-  Deeply Embedded Domain Specific Languages\\
-\large{Not For Distribution}}
+\title[On Adding Pattern Matching to Haskell-based EDSLs]{On Adding Pattern Matching to Haskell-based
+  Deeply Embedded Domain Specific Languages}
 
@@
 A major benefit of the Haskell language is that it is frequently possible derive
-a Haskell implementation from a specification, either partially or fully.
-Examples of this technique include the work in \cite{Elliott-2018-ad-icfp} as
-well as the work in \cite{Elliott2019-convolution-extended}.  When it is not
-possible to fully derive an implementation from a specification, it is often
-easier to check a Haskell implementation against a specification than it is in
-many imperative languages due, in part, to the management of side effects in
-Haskell. This is more difficult to do using an imperative language such as C, as
-they provide no restrictions on where side effects can occur.
+a Haskell implementation from a specification, either partially or fully.  This
+technique is used extensively, for example, in \cite{Bird:2010:Pearls} and
+\cite{Gibbons:2002:Calc}.  Recent examples of this technique include the work in
+\cite{Elliott-2018-ad-icfp} as well as the work in
+\cite{Elliott2019-convolution-extended}.  When it is not possible to fully
+derive an implementation from a specification, it is often easier to check a
+Haskell implementation against a specification than it is in many imperative
+languages due, in part, to the management of side effects in Haskell. This is
+more difficult to do using an imperative language such as C, as they provide no
+restrictions on where side effects can occur.
 

### 2020-09-24 paper/paper.tex
Started on conclusion

@@
 \subsection{Connection to Church encodings}
+\label{sec:Church}
 % \begin{todo}
@@
 \clearpage
+
+\section{Conclusion}
+
+\begin{todo}
+Should some of this go into the introduction?
+\end{todo}
+
+In this paper:
+\begin{itemize}
+  \item We address the goal of translating Haskell pattern matching into an
+    EDSL. An example usage of this mechanism is given in Section
+    \ref{sec:CBackend} in the form of a transformation of a subset of Haskell
+    into an EDSL that is then used to generate C code, effectively compiling
+    that subset of Haskell into C.
+
+  \item A representation of pattern matching via a deep embedding of Church
+    encodings of products and coproducts presented in Section \ref{sec:Church}.
+    This encoding relies on having a uniform representation of algebraic data
+    types.
+
+  \item To provide the uniform representation of algebraic data types mentioned
+    in the previous point, we describe a mechanism for automatically encoding
+    ADTs as compositions of \verb|(,)|, \verb|Either| and \verb|()| using
+    \verb|ERep| and \verb|ERepTy| in Section \ref{sec:ADTRep}.
+
+  \item A GHC plugin is outlined in Section \ref{sec:CorePlugin} which provides automation for the translation of
+    Haskell code into these EDSL representations.
+
+\end{itemize}
+
+\begin{todo}
+Should we have a diagram somewhere in the paper to show how these pieces all fit together?
+\end{todo}
+
 \section{Related Work}

### 2020-09-29 paper/paper.tex
Add some description of the C backend towards the beginning of the paper

@@
 
+In this paper, a C backend is given as an example. The Core plugin will
+translate a subset of Haskell to the EDSL we have specified. The C backend will
+then take expressions written in this EDSL and translate them to C code. This C
+code can be compiled by a standard C compiler. In effect, the subset of Haskell
+represented in the EDSL is compiled to C.
+
 Readers who are primarily interested only in the encoding of pattern matching itself can direct
@@
 either difficult or impossible to use a language that has these benefits.
+
+A C backend is described in this paper. However, other backends could be constructed,
+targeting different architectures.
 

### 2020-09-29 paper/paper.tex
Remove comparison to SSA

@@
 
-In this paper, a C backend is given as an example. The Core plugin will
+In this paper, a C backend is given as an example in Section \ref{sec:CBackend}. The Core plugin will
 translate a subset of Haskell to the EDSL we have specified. The C backend will
@@
 
-A C backend is described in this paper. However, other backends could be constructed,
-targeting different architectures.
+A C backend is described in Section \ref{sec:CBackend}. However, other backends
+could be constructed, targeting different architectures.
 
@@
 Appendix B for implementation details of this ESDL.
-
-This backend is based on a globally unique name generator. A new C variable is created for each Haskell subexpression.
-The result appears similar to static single assignment (SSA) form~\cite{Rosen:1988:SSA}.
-It differs from SSA, however, in that these variables can be assigned multiple
-times. In particular, a variable can be assigned in multiple branches of
-an \ttt{if} statement or inside of a \ttt{while} loop. This simplifies
-the implementation, as $\phi$ nodes are not needed.
 

### 2020-09-29 paper/paper.tex
More information about the standard interpretation

@@
 ...
+\end{lstlisting}
+
+Recall that the internal representations of \verb|SumMatch a b| and \verb|ProdMatch a b|
+are functions of type \verb|a -> b|. These types are unwrapped by the functions
+
+\begin{lstlisting}
+runSumMatch :: SumMatch a b -> (a -> b)
+runProdMatch :: ProdMatch a b -> (a -> b)
+\end{lstlisting}
+
+Given this, the \verb|ProdMatchExp| case of the \verb|prodMatchAbs| function can be
+implemented as follows:
+
+\begin{lstlisting}
+prodMatchAbs env (ProdMatchExp f) =
+  \ pair ->
+    case pair of
+      (x, y) ->
+        runProdMatch (absEnv env f x) y
 \end{lstlisting}

### 2020-09-29 paper/paper.tex
Add a couple of TODOs and go into more detail in the first example in the paper

@@
 
+\begin{todo}
+  Make it more clear how to integrate these concepts with other EDSLs. Maybe we should
+  add a feature to the underlying implementation that allows an existing EDSL to be
+  "enriched" with pattern matching support.
+\end{todo}
+
 The following contributions are made by this paper:
@@
 
-In the example below, the Core plugin transforms \ttt{example} into \ttt{example'}:
+In the example below, the Core plugin transforms \verb|example| into
+\verb|example'|. Note that \verb|example'| is an EDSL expression which is then
+used by the EDSL interpreter function \verb|abs| to produce the final result,
+\verb|example''|:
+
 \begin{lstlisting}[deletekeywords={Ord}]
@@
 example =
-  internalize (externalize
+  externalize
     (case x of
       Left  c -> ord c
-      Right i -> i))
-
-example' :: Int
+      Right i -> i)
+
+example' :: E Int
 example' =
-  abs (CaseExp
+  CaseExp
         (LeftExp 'a')
@@
                            (Var 2)))))
+
+example'' :: Int
+example'' = abs example'
 \end{lstlisting}
@@
 
-% \begin{todo}
-%   Figure out how recursive types fit into this framework. Church encodings can
-%   handle recursive types and so can our representation. This is also the case
-%   for Scott encodings, etc, and is the difference between Church and Scott
-%   encodings.
-
-%   Does our way of handling recursive types correspond to how any of
-%   those encodings handle recursive types?
-
-%   Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
-% \end{todo}
+\begin{todo}
+  Figure out how recursive types fit into this framework. Church encodings can
+  handle recursive types and so can our representation. This is also the case
+  for Scott encodings, etc, and is the difference between Church and Scott
+  encodings.
+
+  Does our way of handling recursive types correspond to how any of
+  those encodings handle recursive types?
+
+  Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
+\end{todo}
 

### 2020-10-03 paper/paper.tex
Convert to PADL format (LNCS)

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigplan,anonymous,review]{acmart}
-\usepackage[switch]{lineno}
+\RequirePackage{amsmath,amssymb}
+% \RequirePackage{stmaryrd}
+\documentclass[runningheads, a4paper]{llncs}
+% \usepackage{biblatex}
+% \usepackage[switch]{lineno}
 \usepackage{booktabs}
@@
 \usepackage{stmaryrd}
-\usepackage{amsmath}
+% \usepackage{amsmath}
 
@@
 
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
-\acmDOI{XX.XXXX/XXXXXXX}
+% \setcopyright{acmcopyright}
+% \copyrightyear{2020}
+% \acmYear{2020}
+% \acmDOI{XX.XXXX/XXXXXXX}
 
@@
 
-\acmConference[GPCE '20]{GPCE '20: Proceedings of the 13th
-  ACM SIGPLAN International Symposium on Haskell}{November 15 -- 20, 2020}{Chicago, Illinois, United States}
-\acmBooktitle{Proceedings of the 19th ACM SIGPLAN International
-Conference on Generative Programming: Concepts and Experiences (GPCE '20)
-  November 15 -- 20, 2020, Chicago, Illinois, United States}
-\acmPrice{15.00}
-\acmISBN{XXX-X-XXXX-XXXX-X/XX/XX}
-
-\citestyle{acmauthoryear}
+% \acmConference[GPCE '20]{GPCE '20: Proceedings of the 13th
+%   ACM SIGPLAN International Symposium on Haskell}{November 15 -- 20, 2020}{Chicago, Illinois, United States}
+% \acmBooktitle{Proceedings of the 19th ACM SIGPLAN International
+% Conference on Generative Programming: Concepts and Experiences (GPCE '20)
+%   November 15 -- 20, 2020, Chicago, Illinois, United States}
+% \acmPrice{15.00}
+% \acmISBN{XXX-X-XXXX-XXXX-X/XX/XX}
+
+% \citestyle{acmauthoryear}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title[On Adding Pattern Matching to Haskell-based EDSLs]{On Adding Pattern Matching to Haskell-based
+\title{On Adding Pattern Matching to Haskell-based
   Deeply Embedded Domain Specific Languages}
+
+\titlerunning{On Adding Pattern Matching to Haskell-based EDSLs}
 
@@
 %% used to denote shared contribution to the research.
-\author{David Young}
-\email{d063y800@ku.edu}
-
+\author{David Young\inst{1} \and Mark Grebe\inst{2} \and Andrew Gill\inst{1}}
+\institute{University of Kansas \and University of Central Missouri}
+
+\maketitle
 % \authornote{Both authors contributed equally to this research.}
@@
 
-\author{Mark Grebe}
-\email{grebe@ucmo.edu}
-
-\author{Andrew Gill}
-\email{andygill@ku.edu}
+% \author{Mark Grebe}
+% \email{grebe@ucmo.edu}
+
+% \author{Andrew Gill}
+% \email{andygill@ku.edu}
 
@@
 % \renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-\renewcommand{\shortauthors}{Young, Grebe and Gill}
+
+% \renewcommand{\shortauthors}{Young, Grebe and Gill}
 
@@
 
-\begin{CCSXML}
-<ccs2012>
-   <concept>
-       <concept_id>10011007.10011006.10011041.10011046</concept_id>
-       <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10011007.10011006.10011041.10011047</concept_id>
-       <concept_desc>Software and its engineering~Source code generation</concept_desc>
-       <concept_significance>300</concept_significance>
-       </concept>
-   <concept>
-       <concept_id>10011007.10011006.10011050.10011017</concept_id>
-       <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
-       <concept_significance>500</concept_significance>
-       </concept>
- </ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
-\ccsdesc[300]{Software and its engineering~Source code generation}
-\ccsdesc[500]{Software and its engineering~Domain specific languages}
+% \begin{CCSXML}
+% <ccs2012>
+%    <concept>
+%        <concept_id>10011007.10011006.10011041.10011046</concept_id>
+%        <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
+%        <concept_significance>300</concept_significance>
+%        </concept>
+%    <concept>
+%        <concept_id>10011007.10011006.10011041.10011047</concept_id>
+%        <concept_desc>Software and its engineering~Source code generation</concept_desc>
+%        <concept_significance>300</concept_significance>
+%        </concept>
+%    <concept>
+%        <concept_id>10011007.10011006.10011050.10011017</concept_id>
+%        <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
+%        <concept_significance>500</concept_significance>
+%        </concept>
+%  </ccs2012>
+% \end{CCSXML}
+
+% \ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
+% \ccsdesc[300]{Software and its engineering~Source code generation}
+% \ccsdesc[500]{Software and its engineering~Domain specific languages}
 
@@
 %% information and builds the first part of the formatted document.
-\maketitle
 
@@
 
-\newpage
+% \newpage
 Haskell EDSLs that support powerful language features like pattern matching
@@
 
-\newpage
+% \newpage
 \begin{lstlisting}
@@
 
-  Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
+  % Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
 \end{todo}
@@
 These bindings are grouped together in the \ttt{Env} type:
-\newpage
+% \newpage
 \begin{lstlisting}
@@
 
-\newpage
+% \newpage
 \section{C Backend}
@@
 can write:
-\newpage
+% \newpage
 \begin{lstlisting}
@@
 
-\clearpage
+% \clearpage
 \begin{figure*}
   \centering 
-  \resizebox{2.0\columnwidth}{20cm}{\includegraphics{Diagrams/Flowchart.pdf}}
+  \resizebox{0.8\columnwidth}{10cm}{\includegraphics{Diagrams/Flowchart.pdf}}
+  % \includegraphics{Diagrams/Flowchart.pdf}
    \caption{Control Flow of the Core Plugin}
@@
 
-\clearpage
+% \clearpage
 
@@
 \begin{align*}
-  &D\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}
-      \rewrites I(D(\ttt{f x = externalize (case s of \{ ... \})}))\\
+  &D\expr{\ttt{f x = internalize (externalize (case s of \{ ... \}))}}\\
+      &\quad\rewrites I(D(\ttt{f x = externalize (case s of \{ ... \})}))\\
   &D\expr{\ttt{f x = externalize (case s of \{ ... \})}}\\
@@
 
-\captionsetup{justification=centering}
+% \captionsetup{justification=centering}
 \caption{Rewrite rules}
@@
 
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{splncs04}
 \bibliography{paper}

### 2020-10-08 paper/paper.tex
Line numbers
update latex
Reformat some example code, comment-out appendixes, move Conclusion to after Related Work

@@
 % \usepackage{biblatex}
-% \usepackage[switch]{lineno}
+\usepackage{lineno}
 \usepackage{booktabs}
@@
 
-% \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{red}}
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{red}}
 
@@
 
-\newcommand{\showtodos}{}  % Comment out to hide todos
+% \newcommand{\showtodos}{}  % Comment out to hide todos
 
@@
 %% article.
+\linenumbers
 \begin{abstract}
@@
 \begin{lstlisting}[deletekeywords={Ord}]
-import Data.Char (ord)
-
 x :: Either Char Int
@@
 example =
-  externalize
-    (case x of
-      Left  c -> ord c
-      Right i -> i)
+  externalize (case x of Left c -> ord c; Right i -> i)
 
 example' :: E Int
-example' =
-  CaseExp
-        (LeftExp 'a')
-        (SumMatchExp (OneProdMatch
-                       (Lam 1
-                         (Ord (Var 1))))
-                     (OneSumMatch
-                       (OneProdMatch
-                         (Lam 2
-                           (Var 2)))))
+example' = CaseExp (LeftExp 'a') (SumMatchExp
+  (OneProdMatch (Lam 1 (Ord (Var 1))))
+  (OneSumMatch (OneProdMatch (Lam 2 (Var 2)))))
 
@@
 
-\section{Conclusion}
-
-\begin{todo}
-Should some of this go into the introduction?
-\end{todo}
-
-In this paper:
-\begin{itemize}
-  \item We address the goal of translating Haskell pattern matching into an
-    EDSL. An example usage of this mechanism is given in Section
-    \ref{sec:CBackend} in the form of a transformation of a subset of Haskell
-    into an EDSL that is then used to generate C code, effectively compiling
-    that subset of Haskell into C.
-
-  \item A representation of pattern matching via a deep embedding of Church
-    encodings of products and coproducts presented in Section \ref{sec:Church}.
-    This encoding relies on having a uniform representation of algebraic data
-    types.
-
-  \item To provide the uniform representation of algebraic data types mentioned
-    in the previous point, we describe a mechanism for automatically encoding
-    ADTs as compositions of \verb|(,)|, \verb|Either| and \verb|()| using
-    \verb|ERep| and \verb|ERepTy| in Section \ref{sec:ADTRep}.
-
-  \item A GHC plugin is outlined in Section \ref{sec:CorePlugin} which provides automation for the translation of
-    Haskell code into these EDSL representations.
-
-\end{itemize}
-
-\begin{todo}
-Should we have a diagram somewhere in the paper to show how these pieces all fit together?
-\end{todo}
-
 \section{Related Work}
@@
 
+\section{Conclusion}
+
+\begin{todo}
+Should some of this go into the introduction?
+\end{todo}
+
+In this paper:
+\begin{itemize}
+  \item We address the goal of translating Haskell pattern matching into an
+    EDSL. An example usage of this mechanism is given in Section
+    \ref{sec:CBackend} in the form of a transformation of a subset of Haskell
+    into an EDSL that is then used to generate C code, effectively compiling
+    that subset of Haskell into C.
+
+  \item A representation of pattern matching via a deep embedding of Church
+    encodings of products and coproducts presented in Section \ref{sec:Church}.
+    This encoding relies on having a uniform representation of algebraic data
+    types.
+
+  \item To provide the uniform representation of algebraic data types mentioned
+    in the previous point, we describe a mechanism for automatically encoding
+    ADTs as compositions of \verb|(,)|, \verb|Either| and \verb|()| using
+    \verb|ERep| and \verb|ERepTy| in Section \ref{sec:ADTRep}.
+
+  \item A GHC plugin is outlined in Section \ref{sec:CorePlugin} which provides automation for the translation of
+    Haskell code into these EDSL representations.
+
+\end{itemize}
+
+\begin{todo}
+Should we have a diagram somewhere in the paper to show how these pieces all fit together?
+\end{todo}
 % \section{Citations and Bibliographies}
@@
 
-\appendix
-\section{Additional Code for Standard Haskell Semantics Implementation}
-\begin{lstlisting}[basicstyle=\tiny]
-absEnv :: forall t. Env -> E t -> t
-absEnv env (CaseExp x f) =
-  sumMatchAbs f (absEnv env x)
-
-absEnv env (TailRec f) = \ x ->
-  case absEnv env f x of
-    Step x' -> absEnv env (TailRec f) x'
-    Done r  -> r
-
-absEnv env (Var v) =
-  case envLookup env v of
-    Just x -> absEnv env
-    Nothing ->
-      error
-        ("No binding for name "
-          ++ show v)
-
-absEnv env (Lam (name :: Name a)
-                (body :: E b)) =
-  \ (arg :: a) ->
-    let go :: forall x. E x -> E x
-        go expr@(Var name2 :: E a') =
-          case namesEq name name2 of
-            Just Refl -> rep arg
-            _ ->
-              case envLookup env name2 of
-                Just v  -> v
-                Nothing -> expr
-
-        -- [... traverse rest of expression
-        --      in go ...]
-    in
-    absEnv (extendEnv env
-                      (name :=> rep arg))
-           (go body)
-
-absEnv env (LeftExp x) =
-  Left (absEnv env x)
-
-absEnv env (RightExp y) =
-  Right (absEnv env y)
-
-absEnv env (PairExp x y) =
-  (absEnv env x, absEnv env y)
-
-absEnv env (ConstructRep x) =
-  unrep' (absEnv env x)
-
-...
-\end{lstlisting}
-
-% \vfill\break
-
-\section{A small EDSL for C code generation used internally by the C backend}
-\begin{lstlisting}[basicstyle=\tiny]
-type CCode = String
-type CName = String
-
-stmt :: CCode -> CCode
-stmt = (<> ";")
-
-cCall :: CName -> [CCode] -> CCode
-cCall fnName args =
-  fnName
-  <> "(" <> intercalate ", " args <> ")"
-
-derefPtr :: CCode -> CCode
-derefPtr x = "*(" <> x <> ")"
-
-cCast :: CCode -> CCode -> CCode
-cCast toType x =
-  "((" <> toType <> ")" <> x <> ")"
-
-cIndex :: CCode -> Int -> CCode
-cIndex x i =
-  "(" <> x <> ")[" <> show i <> "]"
-
-(!) :: CCode -> Int -> CCode
-(!) = cIndex
-
-(=:) :: CCode -> CCode -> CCode
-x =: y = stmt (x <> " = " <> y)
-
-cDecl :: CName -> CName -> CCode
-cDecl cType v = stmt (cType <> " " <> v)
-
-(#) :: CCode -> CName -> CCode
-x # fieldName = x <> "." <> fieldName
-
-macroDef
-  :: CName -> [CName] -> [CCode]
-     -> CCode
-macroDef name args theLines =
-  unlines
-    [ "#define " <> name
-              <> "(" <>
-                intercalate ", " args
-              <> ")\\"
-    , "  do {\\"
-    , unlines
-        . map ("    "++)
-        . map (++"\\")
-        $ theLines
-    , "  } while (0);"
-    ]
-\end{lstlisting}
-% \begin{todo}
-%   Fill in
-% \end{todo}
+% \appendix
+% \section{Additional Code for Standard Haskell Semantics Implementation}
+% \begin{lstlisting}[basicstyle=\tiny]
+% absEnv :: forall t. Env -> E t -> t
+% absEnv env (CaseExp x f) =
+%   sumMatchAbs f (absEnv env x)
+
+% absEnv env (TailRec f) = \ x ->
+%   case absEnv env f x of
+%     Step x' -> absEnv env (TailRec f) x'
+%     Done r  -> r
+
+% absEnv env (Var v) =
+%   case envLookup env v of
+%     Just x -> absEnv env
+%     Nothing ->
+%       error
+%         ("No binding for name "
+%           ++ show v)
+
+% absEnv env (Lam (name :: Name a)
+%                 (body :: E b)) =
+%   \ (arg :: a) ->
+%     let go :: forall x. E x -> E x
+%         go expr@(Var name2 :: E a') =
+%           case namesEq name name2 of
+%             Just Refl -> rep arg
+%             _ ->
+%               case envLookup env name2 of
+%                 Just v  -> v
+%                 Nothing -> expr
+
+%         -- [... traverse rest of expression
+%         --      in go ...]
+%     in
+%     absEnv (extendEnv env
+%                       (name :=> rep arg))
+%            (go body)
+
+% absEnv env (LeftExp x) =
+%   Left (absEnv env x)
+
+% absEnv env (RightExp y) =
+%   Right (absEnv env y)
+
+% absEnv env (PairExp x y) =
+%   (absEnv env x, absEnv env y)
+
+% absEnv env (ConstructRep x) =
+%   unrep' (absEnv env x)
+
+% ...
+% \end{lstlisting}
+
+% % \vfill\break
+
+% \section{A small EDSL for C code generation used internally by the C backend}
+% \begin{lstlisting}[basicstyle=\tiny]
+% type CCode = String
+% type CName = String
+
+% stmt :: CCode -> CCode
+% stmt = (<> ";")
+
+% cCall :: CName -> [CCode] -> CCode
+% cCall fnName args =
+%   fnName
+%   <> "(" <> intercalate ", " args <> ")"
+
+% derefPtr :: CCode -> CCode
+% derefPtr x = "*(" <> x <> ")"
+
+% cCast :: CCode -> CCode -> CCode
+% cCast toType x =
+%   "((" <> toType <> ")" <> x <> ")"
+
+% cIndex :: CCode -> Int -> CCode
+% cIndex x i =
+%   "(" <> x <> ")[" <> show i <> "]"
+
+% (!) :: CCode -> Int -> CCode
+% (!) = cIndex
+
+% (=:) :: CCode -> CCode -> CCode
+% x =: y = stmt (x <> " = " <> y)
+
+% cDecl :: CName -> CName -> CCode
+% cDecl cType v = stmt (cType <> " " <> v)
+
+% (#) :: CCode -> CName -> CCode
+% x # fieldName = x <> "." <> fieldName
+
+% macroDef
+%   :: CName -> [CName] -> [CCode]
+%      -> CCode
+% macroDef name args theLines =
+%   unlines
+%     [ "#define " <> name
+%               <> "(" <>
+%                 intercalate ", " args
+%               <> ")\\"
+%     , "  do {\\"
+%     , unlines
+%         . map ("    "++)
+%         . map (++"\\")
+%         $ theLines
+%     , "  } while (0);"
+%     ]
+% \end{lstlisting}
 

### 2020-10-08 paper/paper.tex
Shrunk figures, but they are appearing out of ordergp
Figures in correct order

@@
 \usepackage{stmaryrd}
+\usepackage{wrapfig}
 % \usepackage{amsmath}
@@
 
+\begin{wrapfigure}{R}{0.5\linewidth}
+  % \centering
+  \resizebox{0.5\columnwidth}{7cm}{\includegraphics{Diagrams/Flowchart.pdf}}
+  % \includegraphics{Diagrams/Flowchart.pdf}
+   \caption{Control Flow of the Core Plugin}
+   \label{fig:CorePlugin}
+\end{wrapfigure}%
+
 In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses in a node refer
@@
 % \clearpage
-\begin{figure*}
-  \centering 
-  \resizebox{0.8\columnwidth}{10cm}{\includegraphics{Diagrams/Flowchart.pdf}}
-  % \includegraphics{Diagrams/Flowchart.pdf}
-   \caption{Control Flow of the Core Plugin}
-   \label{fig:CorePlugin}
-\end{figure*}
 
@@
 
-\begin{figure*}[hp]
+% \begin{wrapfigure}{r}{0.8\linewidth}
+  \begin{wrapfigure}{R}{1.2\linewidth}
   \centering
+\resizebox{1\linewidth}{!}{
+  \begin{minipage}{\linewidth}
 \begin{align*}
@@
       &\quad\rewrites \begin{cases}
-        \ttt{f = $C^{*}\expr{\ttt{externalize ($\lambda \eta \rarr$ runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}) $\eta$)}}$} & \text{\parbox{5cm}{if \ttt{f} occurs free in \ttt{\{ ...
+        \ttt{f = $C^{*}\expr{\ttt{externalize ($\lambda \eta \rarr$ runIter ($\lambda \ttt{x} \rarr \ttt{case s of \{ $T'_{\ttt{f}}\exprdots$}$ \}) $\eta$)}}$} & \text{\parbox{2.4cm}{if \ttt{f} occurs free in \ttt{\{ ...
 \}}, where $\eta$ is a fresh name}}\\
@@
   &C\expr{\ttt{$\lambda$x $\rarr$ body}} \rewrites A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
-  &C\expr{\ttt{K x0 ... xN }} \rewrites \ttt{construct (K (unrep $M\expr{\ttt{x0}}$) ... (unrep $M\expr{\ttt{xN}}$))}\;\;\;\;\text{(Where \ttt{K} is a constructor)}\\
+  &C\expr{\ttt{K x0 ... xN }}\\
+    &\quad\rewrites \ttt{construct (K (unrep $M\expr{\ttt{x0}}$) ... (unrep $M\expr{\ttt{xN}}$))}\;\;\;\;\text{(Where \ttt{K} is a constructor)}\\
   &C\expr{\ttt{f x}} \rewrites \ttt{App $M\expr{f}$ $M\expr{x}$}\\
@@
 \end{align*}
+  \end{minipage}
+}%
+% }}
 
@@
 \label{fig:Rewrites}
-\end{figure*}
+% \end{wrapfigure}
+\end{wrapfigure}
 

### 2020-10-08 paper/paper.tex
Comment out Church encoding and C backend sections

@@
 
-\subsection{Connection to Church encodings}
-\label{sec:Church}
-% \begin{todo}
-%   Work on this subsection. Maybe this shouldn't be a separate section at all and
-%   maybe its contents should be incorporated throughout the rest of the paper.
-% \end{todo}
-
-Algebraic datatypes can be represented as lambda calculus terms using
-\textit{Church encodings}.~\cite{Boehm:1985:AutoSynth} For example, the pair constructor can be
-represented as
-
-\begin{lstlisting}
-church_mkPair
-  :: a -> b -> (a -> b -> r) -> r
-church_mkPair x y = \ f -> f x y
-\end{lstlisting}
-
-Coproducts can be represented with the constructors
-
-\begin{lstlisting}
-church_mkLeft
-  :: a -> (a -> r) -> (b -> r) -> r
-church_mkLeft x = \ f g -> f x
-
-church_mkRight
-  :: b -> (a -> r) -> (b -> r) -> r
-church_mkRight y = \ f g -> g y
-\end{lstlisting}
-
-In this representation:
-
-\begin{itemize}
-  \item A pair of type \ttt{A} and type \ttt{B} is represented with a value of type
-\ttt{forall r. (A -> B -> r) -> r}.
-
-  \item A coproduct of type \ttt{A} and type \ttt{B} is represented with a value of type
-\ttt{forall r. (A -> r) -> (B -> r) -> r}.
-\end{itemize}
-
-Both of those representations represent the type as a function that performs the
-corresponding case analysis. A deep embedding of the operations
-\ttt{churchMkPair}, \ttt{church\_mkLeft} and \ttt{church\_mkRight} is given
-by the type
-
-\begin{lstlisting}
-data Church r where
-  ChurchPair
-    :: a -> b -> (a -> b -> r) ->
-       Church r
-  ChurchLeft
-    :: a -> (a -> r) -> (b -> r) ->
-       Church r
-  ChurchRight
-    :: b -> (a -> r) -> (b -> r) ->
-       Church r
-\end{lstlisting}
-
-with the evaluation function
-
-\begin{lstlisting}
-churchEval :: Church r -> r
-churchEval (ChurchPair x y f) = f x y
-churchEval (ChurchLeft x f g) = f x
-churchEval (ChurchRight y f g) = g y
-\end{lstlisting}
-
-However, we only want to keep track of the functions used for elimination. By moving
-the values actually being contained in the products and coproducts out to the
-evaluation function, we arrive at
-
-\begin{lstlisting}
-data Church' t r where
-  ChurchPair'
-    :: (a -> b -> r) ->
-       Church' (a, b) r
-
-  ChurchEither'
-    :: (a -> r) -> (b -> r) ->
-       Church' (Either a b) r
-
-churchEval' :: Church' t r -> t -> r
-churchEval' (ChurchPair' f) (x, y) = f x y
-churchEval' (ChurchEither' f g) e =
-  case e of
-    Left  x -> f x
-    Right y -> g y
-\end{lstlisting}
-
-The \verb|Church'| type forms a large part of the foundation of the pattern
-matching aspect of the \verb|E| type. A rough correspondence between the two
-is given in the table below.\\
-
-{\small
-\noindent\begin{tabular}{|p{3.4cm}|p{4.4cm}|}
-  \hline
-  \ttt{Church' (a, b) r} & \ttt{E (SumMatch (a, b) r} \\
-  \hline
-  \ttt{Church' (Either a b) r} & \ttt{E (ProdMatch (Either a b) r)}\\
-  \hline
-\end{tabular}
-}
+% \subsection{Connection to Church encodings}
+% \label{sec:Church}
+% % \begin{todo}
+% %   Work on this subsection. Maybe this shouldn't be a separate section at all and
+% %   maybe its contents should be incorporated throughout the rest of the paper.
+% % \end{todo}
+
+% Algebraic datatypes can be represented as lambda calculus terms using
+% \textit{Church encodings}.~\cite{Boehm:1985:AutoSynth} For example, the pair constructor can be
+% represented as
+
+% \begin{lstlisting}
+% church_mkPair
+%   :: a -> b -> (a -> b -> r) -> r
+% church_mkPair x y = \ f -> f x y
+% \end{lstlisting}
+
+% Coproducts can be represented with the constructors
+
+% \begin{lstlisting}
+% church_mkLeft
+%   :: a -> (a -> r) -> (b -> r) -> r
+% church_mkLeft x = \ f g -> f x
+
+% church_mkRight
+%   :: b -> (a -> r) -> (b -> r) -> r
+% church_mkRight y = \ f g -> g y
+% \end{lstlisting}
+
+% In this representation:
+
+% \begin{itemize}
+%   \item A pair of type \ttt{A} and type \ttt{B} is represented with a value of type
+% \ttt{forall r. (A -> B -> r) -> r}.
+
+%   \item A coproduct of type \ttt{A} and type \ttt{B} is represented with a value of type
+% \ttt{forall r. (A -> r) -> (B -> r) -> r}.
+% \end{itemize}
+
+% Both of those representations represent the type as a function that performs the
+% corresponding case analysis. A deep embedding of the operations
+% \ttt{churchMkPair}, \ttt{church\_mkLeft} and \ttt{church\_mkRight} is given
+% by the type
+
+% \begin{lstlisting}
+% data Church r where
+%   ChurchPair
+%     :: a -> b -> (a -> b -> r) ->
+%        Church r
+%   ChurchLeft
+%     :: a -> (a -> r) -> (b -> r) ->
+%        Church r
+%   ChurchRight
+%     :: b -> (a -> r) -> (b -> r) ->
+%        Church r
+% \end{lstlisting}
+
+% with the evaluation function
+
+% \begin{lstlisting}
+% churchEval :: Church r -> r
+% churchEval (ChurchPair x y f) = f x y
+% churchEval (ChurchLeft x f g) = f x
+% churchEval (ChurchRight y f g) = g y
+% \end{lstlisting}
+
+% However, we only want to keep track of the functions used for elimination. By moving
+% the values actually being contained in the products and coproducts out to the
+% evaluation function, we arrive at
+
+% \begin{lstlisting}
+% data Church' t r where
+%   ChurchPair'
+%     :: (a -> b -> r) ->
+%        Church' (a, b) r
+
+%   ChurchEither'
+%     :: (a -> r) -> (b -> r) ->
+%        Church' (Either a b) r
+
+% churchEval' :: Church' t r -> t -> r
+% churchEval' (ChurchPair' f) (x, y) = f x y
+% churchEval' (ChurchEither' f g) e =
+%   case e of
+%     Left  x -> f x
+%     Right y -> g y
+% \end{lstlisting}
+
+% The \verb|Church'| type forms a large part of the foundation of the pattern
+% matching aspect of the \verb|E| type. A rough correspondence between the two
+% is given in the table below.\\
+
+% {\small
+% \noindent\begin{tabular}{|p{3.4cm}|p{4.4cm}|}
+%   \hline
+%   \ttt{Church' (a, b) r} & \ttt{E (SumMatch (a, b) r} \\
+%   \hline
+%   \ttt{Church' (Either a b) r} & \ttt{E (ProdMatch (Either a b) r)}\\
+%   \hline
+% \end{tabular}
+% }
 
@@
 % \newpage
-\section{C Backend}
-\label{sec:CBackend}
-
-\begin{todont}
-  Should this section be expanded?
-\end{todont}
-
-The C backend makes use of a small EDSL for generating C code. See
-Appendix B for implementation details of this ESDL.
-
-The following \ttt{struct}s provide the C representation for Haskell expressions:
-
-\begin{lstlisting}
-typedef enum var_type_tag {
-  EXPR_INT
-, EXPR_PAIR
-...
-, EXPR_UNIT
-} var_type_tag;
-
-typedef enum semantic_type_tag {
-  NO_SEMANTIC_TAG
-, EXPR_COMPLEX     // Complex numbers
-} semantic_type_tag;
-
-struct closure_t;
-
-typedef struct var_t {
-  var_type_tag tag;
-  semantic_type_tag semantic_tag;
-  void* value;
-} var_t;
-
-typedef struct closure_t {
-  var_t* fv_env;
-  var_t (*fn)(var_t, struct closure_t*);
-} closure_t;
-\end{lstlisting}
-
-The \ttt{semantic\_tag} allows type information preserved by the \ttt{Typeable} constraints
-in the EDSL to be brought over to the generated C code. In our implementation, this is
-used to distinguish between pairs of \ttt{Double}s and \ttt{Complex Double} values. In the
-case of a \ttt{Complex Double}, the \ttt{var\_t} would have a \ttt{tag} field with the value
-\ttt{EXPR\_PAIR} and a \ttt{semantic\_tag} field with the value \ttt{EXPR\_COMPLEX}.
-
-A C code generation function \ttt{genC} is implemented:
-
-\begin{lstlisting}
-genC :: E a -> String
-...
-\end{lstlisting}
-
-To use this, for example, with the previously mentioned \ttt{isEven} implementation we
-can write:
-% \newpage
-\begin{lstlisting}
-isEvenE :: E (Int -> Bool)
-isEvenE =
-  externalize (
-    case x == 0 of
-      True -> True
-      False ->
-        case x == 1 of
-          True -> False
-          False -> isEvenE (x - 2))
-
-isEvenC :: String
-isEvenC = genC isEvenE
-\end{lstlisting}
+
+% \section{C Backend}
+% \label{sec:CBackend}
+
+% \begin{todont}
+%   Should this section be expanded?
+% \end{todont}
+
+% The C backend makes use of a small EDSL for generating C code. See
+% Appendix B for implementation details of this ESDL.
+
+% The following \ttt{struct}s provide the C representation for Haskell expressions:
+
+% \begin{lstlisting}
+% typedef enum var_type_tag {
+%   EXPR_INT
+% , EXPR_PAIR
+% ...
+% , EXPR_UNIT
+% } var_type_tag;
+
+% typedef enum semantic_type_tag {
+%   NO_SEMANTIC_TAG
+% , EXPR_COMPLEX     // Complex numbers
+% } semantic_type_tag;
+
+% struct closure_t;
+
+% typedef struct var_t {
+%   var_type_tag tag;
+%   semantic_type_tag semantic_tag;
+%   void* value;
+% } var_t;
+
+% typedef struct closure_t {
+%   var_t* fv_env;
+%   var_t (*fn)(var_t, struct closure_t*);
+% } closure_t;
+% \end{lstlisting}
+
+% The \ttt{semantic\_tag} allows type information preserved by the \ttt{Typeable} constraints
+% in the EDSL to be brought over to the generated C code. In our implementation, this is
+% used to distinguish between pairs of \ttt{Double}s and \ttt{Complex Double} values. In the
+% case of a \ttt{Complex Double}, the \ttt{var\_t} would have a \ttt{tag} field with the value
+% \ttt{EXPR\_PAIR} and a \ttt{semantic\_tag} field with the value \ttt{EXPR\_COMPLEX}.
+
+% A C code generation function \ttt{genC} is implemented:
+
+% \begin{lstlisting}
+% genC :: E a -> String
+% ...
+% \end{lstlisting}
+
+% To use this, for example, with the previously mentioned \ttt{isEven} implementation we
+% can write:
+% % \newpage
+% \begin{lstlisting}
+% isEvenE :: E (Int -> Bool)
+% isEvenE =
+%   externalize (
+%     case x == 0 of
+%       True -> True
+%       False ->
+%         case x == 1 of
+%           True -> False
+%           False -> isEvenE (x - 2))
+
+% isEvenC :: String
+% isEvenC = genC isEvenE
+% \end{lstlisting}
 
@@
 % \begin{wrapfigure}{r}{0.8\linewidth}
-  \begin{wrapfigure}{R}{1.2\linewidth}
+  \begin{wrapfigure}{R}{0.6\linewidth}
   \centering
-\resizebox{1\linewidth}{!}{
+\resizebox{0.9\linewidth}{!}{
   \begin{minipage}{\linewidth}

### 2020-10-08 paper/paper.tex
Reformat code for single column (and a small typo fix)

@@
 \begin{lstlisting}
-data ComplexPair where
-  ComplexPair :: Complex Double
-                 -> Complex Double
-                 -> ComplexPair
+data ComplexPair =
+  ComplexPair (Complex Double) (Complex Double)
   deriving (Generic, Show)
@@
 \begin{lstlisting}
-sumComplexPair ::
-  ComplexPair -> Complex Double
+sumComplexPair :: ComplexPair -> Complex Double
 sumComplexPair p =
   internalize (externalize
-    (case p of
-      ComplexPair a b -> a + b))
+    (case p of ComplexPair a b -> a + b))
 \end{lstlisting}
@@
 realSum p =
-  internalize (externalize
-    (case p of
+  internalize (externalize (case p of
       ComplexPair a b -> case a of
           a_real :+ _ -> case b of
-              b_real :+ _ ->
-                a_real + b_real))
+              b_real :+ _ -> a_real + b_real))
 \end{lstlisting}
@@
   internalize (externalize
-    (case p of
-       (acc, t) ->
-         case t of
+    (case p of (acc, t) -> case t of
        Nil -> acc
@@
   ...
-  ConstructRep ::
-    (Typeable a, ERep a)
-    => E (ERepTy a) -> E a
+  ConstructRep :: (Typeable a, ERep a) =>
+    E (ERepTy a) -> E a
 
@@
 
-  default rep ::
-    (ERep (ERepTy t))
-    => t -> E t
+  default rep :: (ERep (ERepTy t)) => t -> E t
   rep x = ConstructRep (construct x)
-  {-# INLINABLE rep #-}
 
@@
 \begin{lstlisting}
-instance (ERep a, ERep b)
-  => ERep (Either a b) where
-
+instance (ERep a, ERep b) => ERep (Either a b) where
   type ERepTy (Either a b) = Either a b
@@
 
-instance (ERep a, ERep b)
-  => ERep (a, b) where
-
+instance (ERep a, ERep b) => ERep (a, b) where
   type ERepTy (a, b) = (a, b)
@@
   ...
-  CaseExp ::
-    (ERep t,
-     ERepTy (ERepTy t) ~ ERepTy t)
-      => ERep t
-         -> E (SumMatch (ERepTy t) r)
-         -> E r
+  CaseExp :: (ERep t, ERepTy (ERepTy t) ~ ERepTy t) =>
+    E t -> E (SumMatch (ERepTy t) r) -> E r
   ...
@@
   ...
-  SumMatchExp ::
-    (ERep a, ERep b, ERepTy b ~ b)
-      => E (ProdMatch a r)
-         -> E (SumMatch b r)
+  SumMatchExp :: (ERep a, ERep b, ERepTy b ~ b) =>
+    E (ProdMatch a r) -> E (SumMatch b r)
       -> E (SumMatch (Either a b) r)
@@
   OneSumMatch ::
-    (ERep a, ERep b, ERepTy a ~ a)
-      => E (ProdMatch a b)
-         -> E (SumMatch a b)
-
-  EmptyMatch ::
-    (ERep b) => E (SumMatch Void b)
+    (ERep a, ERep b, ERepTy a ~ a) =>
+      E (ProdMatch a b) -> E (SumMatch a b)
+
+  EmptyMatch :: (ERep b) => E (SumMatch Void b)
   ...
@@
   ...
-  ProdMatchExp ::
-    (ERep a, ERep b)
-      => E (a -> ProdMatch b r)
-         -> E (ProdMatch (a, b) r)
-
-  NullaryMatch ::
-    (ERep a)
-      => E r -> E (ProdMatch a r)
-
-  OneProdMatch ::
-    (ERep a)
-      => E (a -> b) -> E (ProdMatch a b)
+  ProdMatchExp :: (ERep a, ERep b) =>
+    E (a -> ProdMatch b r) -> E (ProdMatch (a, b) r)
+
+  NullaryMatch :: (ERep a) =>
+    E r -> E (ProdMatch a r)
+
+  OneProdMatch :: (ERep a) =>
+    E (a -> b) -> E (ProdMatch a b)
   ...
@@
 \begin{lstlisting}
-runSumMatch . abs
-  :: E (SumMatch a b) -> (a -> b)
+runSumMatch . abs    :: E (SumMatch a b) -> (a -> b)
 \end{lstlisting}
@@
 runIter f = go
-  where
-    go x =
-      case f x of
+  where go x = case f x of
           Done r  -> r
@@
 
-  TailRec :: (ERep a, ERep b)
-               => E (b -> Iter a b)
-                  -> E (b -> a)
+  TailRec :: (ERep a, ERep b) =>
+    E (b -> Iter a b) -> E (b -> a)
   ...
@@
     (\ eta ->
-      runIter (
-        \ x ->
-          case x == 0 of
+      runIter (\ x -> case x == 0 of
                         True -> Done True
@@
     (Lam (Var 1)
-      (App
-        (TailRec (Lam (Var 2)
-          (IfThenElseExp (Equal (Var 2)
-                                (rep 0))
+      (App (TailRec (Lam (Var 2)
+             (IfThenElseExp (Equal (Var 2) (rep 0))
                (DoneExp TrueExp)
-            (IfThenElseExp (Equal (Var 1)
-                                  (rep 1))
+               (IfThenElseExp (Equal (Var 1) (rep 1))
                  (DoneExp FalseExp)
-              (StepExp (Sub (Var 2)
-                            (rep 2)))))))
+                 (StepExp (Sub (Var 2) (rep 2)))))))
            (Var 1))))
@@
   ...
-  Lam ::
-    (ERep a, Typeable a)
-      => Name a -> E b -> E (a -> b)
-
+  Lam :: (ERep a, Typeable a) => Name a -> E b -> E (a -> b)
   Var :: (Typeable a) => Name a -> E a
@@
 \begin{lstlisting}
-newtype Name a = Name Int
-  deriving (Eq, Show)
-
-namesEq :: forall a b.
-  (Typeable a, Typeable b)
-   => Name a
-   -> Name b
-   -> Maybe (a :~: b)
+newtype Name a = Name Int deriving (Eq, Show)
+
+namesEq :: forall a b.  (Typeable a, Typeable b) =>
+  Name a -> Name b -> Maybe (a :~: b)
 namesEq (Name n) (Name n') =
@@
 data EnvMapping where
-  (:=>) :: forall a.
-    Typeable a
-    => Name a -> E a -> EnvMapping
+  (:=>) :: forall a. Typeable a =>
+    Name a -> E a -> EnvMapping
 \end{lstlisting}
@@
 
-envLookup :: Typeable a
-  => Env -> Name a -> Maybe (E a)
+envLookup :: Typeable a => Env -> Name a -> Maybe (E a)
 envLookup (Env maps) = go maps
@@
 
-sumMatchAbs ::
-  (ERepTy (ERepTy s) ~ ERepTy s, ERep s)
-  => Env
-    -> E (SumMatch (ERepTy s) t)
-    -> s
-    -> t
+sumMatchAbs :: (ERepTy (ERepTy s) ~ ERepTy s, ERep s) =>
+  Env -> E (SumMatch (ERepTy s) t) -> s -> t
 ...
 
-prodMatchAbs :: (ERep s)
-  => Env
-     -> E (ProdMatch s t)
-     -> s
-     -> t
+prodMatchAbs :: (ERep s) =>
+  Env -> E (ProdMatch s t) -> s -> t
 ...
@@
 \begin{lstlisting}
-prodMatchAbs env (ProdMatchExp f) =
-  \ pair ->
-    case pair of
-      (x, y) ->
+prodMatchAbs env (ProdMatchExp f) = \ pair ->
+  case pair of (x, y) ->
     runProdMatch (absEnv env f x) y

### 2020-10-08 paper/paper.tex
Fix rewrite rule figure a bit
Removed some references to the C backend and Church encoding sections that were removed

@@
 \item An implementation of a interpreter for the "standard" semantics of the EDSL (Section~\ref{sec:StdSemantics}).
-\item An outline of a C backend for this EDSL (Section~\ref{sec:CBackend}).
+% \item An outline of a C backend for this EDSL (Section~\ref{sec:CBackend}).
 \end{itemize}
 
-In this paper, a C backend is given as an example in Section \ref{sec:CBackend}. The Core plugin will
-translate a subset of Haskell to the EDSL we have specified. The C backend will
-then take expressions written in this EDSL and translate them to C code. This C
-code can be compiled by a standard C compiler. In effect, the subset of Haskell
-represented in the EDSL is compiled to C.
+% In this paper, a C backend is given as an example in Section \ref{sec:CBackend}. The Core plugin will
+% translate a subset of Haskell to the EDSL we have specified. The C backend will
+% then take expressions written in this EDSL and translate them to C code. This C
+% code can be compiled by a standard C compiler. In effect, the subset of Haskell
+% represented in the EDSL is compiled to C.
 
@@
 
-A C backend is described in Section \ref{sec:CBackend}. However, other backends
-could be constructed, targeting different architectures.
+% A C backend is described in Section \ref{sec:CBackend}. However, other backends
+% could be constructed, targeting different architectures.
 
@@
 The process of transforming a tail recursive function in the DSL to a 
-representation which could be passed into the C backend and used
+representation which could be passed into a backend and used
 to generate a traditional iterative loop goes as follows.
@@
 
-The Core plugin translates marked expressions. Expressions are marked by
-the \ttt{externalize} function:
-
-\begin{lstlisting}
-externalize :: a -> E a
-\end{lstlisting}
-
-For example, \ttt{externalize x} marks the expression \ttt{x}.
-
-If an expression already has an EDSL type (a type of the form \ttt{E a} for some
-\ttt{a}), then the marking procedure ignores it and does not wrap it with
-\ttt{externalize} (see $M$ in Figure~\ref{fig:Rewrites}).
-
-\begin{wrapfigure}{R}{0.5\linewidth}
+\begin{wrapfigure}{r}{0.5\linewidth}
   % \centering
@@
 
+The Core plugin translates marked expressions. Expressions are marked by
+the \ttt{externalize} function:
+
+\begin{lstlisting}
+externalize :: a -> E a
+\end{lstlisting}
+
+For example, \ttt{externalize x} marks the expression \ttt{x}.
+
+If an expression already has an EDSL type (a type of the form \ttt{E a} for some
+\ttt{a}), then the marking procedure ignores it and does not wrap it with
+\ttt{externalize} (see $M$ in Figure~\ref{fig:Rewrites}).
+
 In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses in a node refer
@@
 % \begin{wrapfigure}{r}{0.8\linewidth}
-  \begin{wrapfigure}{R}{0.6\linewidth}
+  \begin{wrapfigure}{r}{0.6\linewidth}
   \centering
-\resizebox{0.9\linewidth}{!}{
+\resizebox{1\linewidth}{!}{
   \begin{minipage}{\linewidth}
@@
   \item We address the goal of translating Haskell pattern matching into an
-    EDSL. An example usage of this mechanism is given in Section
-    \ref{sec:CBackend} in the form of a transformation of a subset of Haskell
-    into an EDSL that is then used to generate C code, effectively compiling
-    that subset of Haskell into C.
-
-  \item A representation of pattern matching via a deep embedding of Church
-    encodings of products and coproducts presented in Section \ref{sec:Church}.
-    This encoding relies on having a uniform representation of algebraic data
-    types.
+    EDSL.
+
+    % An example usage of this mechanism is given in Section
+    % \ref{sec:CBackend} in the form of a transformation of a subset of Haskell
+    % into an EDSL that is then used to generate C code, effectively compiling
+    % that subset of Haskell into C.
+
+  % \item A representation of pattern matching via a deep embedding of Church
+  %   encodings of products and coproducts presented in Section \ref{sec:Church}.
+  %   This encoding relies on having a uniform representation of algebraic data
+  %   types.
 

### 2020-10-09 paper/paper.tex
Typo and minor formatting change
Put Fig 2 back on its own page by itself
wibble
Paragraphs in conclusion; for GADTs cite 2006 paper by Vytiniotis, Weirich, Peyton Jones

@@
 There are three interconnected foundational parts: \ttt{E}, \ttt{ERep} and
-\ttt{ERepTy}. \ttt{E} is the deep embedding of the EDSL (a GADT that encodes
+\ttt{ERepTy}. \ttt{E} is the deep embedding of the EDSL (a GADT~\cite{Vytiniotis:2006:Simple} that encodes
 expressions in the DSL language). \ttt{ERep} is a type class which represents
@@
   ...
-  Lam :: (ERep a, Typeable a) => Name a -> E b -> E (a -> b)
+  Lam :: (ERep a, Typeable a) =>
+    Name a -> E b -> E (a -> b)
   Var :: (Typeable a) => Name a -> E a
@@
 
-sumMatchAbs :: (ERepTy (ERepTy s) ~ ERepTy s, ERep s) =>
-  Env -> E (SumMatch (ERepTy s) t) -> s -> t
+sumMatchAbs :: (ERepTy (ERepTy s) ~ ERepTy s, ERep s)
+  => Env -> E (SumMatch (ERepTy s) t) -> s -> t
 ...
@@
 % \begin{wrapfigure}{r}{0.8\linewidth}
-  \begin{wrapfigure}{r}{0.6\linewidth}
+  % \begin{wrapfigure}{r}{0.59\linewidth}
+\begin{figure}
   \centering
@@
 % \end{wrapfigure}
-\end{wrapfigure}
+% \end{wrapfigure}
+\end{figure}
 
@@
 \cite{Elliott:03:CompileDSEL-JFP} explains the basic idea, and \cite{Gill:2014:DSLSynth}
-gives an acceessable and more recent account of the area.
+gives an accessable and more recent account of the area.
 
@@
 
-In this paper:
-\begin{itemize}
-  \item We address the goal of translating Haskell pattern matching into an
-    EDSL.
-
-    % An example usage of this mechanism is given in Section
-    % \ref{sec:CBackend} in the form of a transformation of a subset of Haskell
-    % into an EDSL that is then used to generate C code, effectively compiling
-    % that subset of Haskell into C.
-
-  % \item A representation of pattern matching via a deep embedding of Church
-  %   encodings of products and coproducts presented in Section \ref{sec:Church}.
-  %   This encoding relies on having a uniform representation of algebraic data
-  %   types.
-
-  \item To provide the uniform representation of algebraic data types mentioned
-    in the previous point, we describe a mechanism for automatically encoding
-    ADTs as compositions of \verb|(,)|, \verb|Either| and \verb|()| using
-    \verb|ERep| and \verb|ERepTy| in Section \ref{sec:ADTRep}.
-
-  \item A GHC plugin is outlined in Section \ref{sec:CorePlugin} which provides automation for the translation of
-    Haskell code into these EDSL representations.
-
-\end{itemize}
+In this paper, a representation of pattern matching was given in the form of a
+GADT called \verb|E|. A GHC plugin is then used to transform code written in a
+subset of Haskell containing algebraic data types, pattern matching, tail
+recursion and lambdas into values of this GADT. A backend can then use these
+\verb|E a| values to interpret the Haskell code, compile it or otherwise
+manipulate it.
+
+Additionally, a uniform representation of algebraic data types in terms of
+\verb|(,)|, \verb|Either| and \verb|()|, called \verb|ERepTy|, was outlined for
+use in the previously mentioned pattern matching representation. This
+representation of ADTs is determined entirely by the structure of a type,
+allowing the pattern matching mechanism to act on the structure of values.
+
+% In this paper:
+% \begin{itemize}
+%   \item We address the goal of translating Haskell pattern matching into an
+%     EDSL.
+
+%     % An example usage of this mechanism is given in Section
+%     % \ref{sec:CBackend} in the form of a transformation of a subset of Haskell
+%     % into an EDSL that is then used to generate C code, effectively compiling
+%     % that subset of Haskell into C.
+
+%   % \item A representation of pattern matching via a deep embedding of Church
+%   %   encodings of products and coproducts presented in Section \ref{sec:Church}.
+%   %   This encoding relies on having a uniform representation of algebraic data
+%   %   types.
+
+%   \item To provide the uniform representation of algebraic data types mentioned
+%     in the previous point, we describe a mechanism for automatically encoding
+%     ADTs as compositions of \verb|(,)|, \verb|Either| and \verb|()| using
+%     \verb|ERep| and \verb|ERepTy| in Section \ref{sec:ADTRep}.
+
+%   \item A GHC plugin is outlined in Section \ref{sec:CorePlugin} which provides automation for the translation of
+%     Haskell code into these EDSL representations.
+
+% \end{itemize}
 

### 2020-10-10 paper/paper.tex
Some changes suggested by Mark

@@
 
-A major benefit of the Haskell language is that it is frequently possible derive
+A major benefit of the Haskell language is that it is frequently possible to derive
 a Haskell implementation from a specification, either partially or fully.  This
@@
 
-Note that the instance definition is automatically generated from the
-\ttt{Generic} instance.
+Note the instance definition is automatically generated from the \ttt{Generic}
+instance.
 

### 2020-10-10 paper/paper.tex
Remove a couple of widows
vertical spacing around figure 1
formatting

@@
 %% article.
-\linenumbers
+
+% \linenumbers
+
 \begin{abstract}
@@
 
-Given this code, \ttt{ERepTy ComplexPair \typeeq (Complex Double, Complex Double)}. Here
-is an example that demonstrates why this is more useful than if it \textit{fully} deconstructed
+Given this code, the following type equality holds:
+\\
+
+\ttt{ERepTy ComplexPair \typeeq (Complex Double, Complex Double)}
+\\
+
+Here is an example that demonstrates why this is more useful than if it \textit{fully} deconstructed
 each type into \ttt{Either}, \ttt{(,)} and \ttt{()}:
@@
 realSum p =
-  internalize (externalize (case p of
-      ComplexPair a b -> case a of
-          a_real :+ _ -> case b of
-              b_real :+ _ -> a_real + b_real))
+  internalize (externalize
+    (case p of
+      ComplexPair a b ->
+        case a of
+          a_real :+ _ ->
+            case b of
+              b_real :+ _ ->
+                a_real + b_real))
 \end{lstlisting}
@@
       -> E (SumMatch (Either a b) r)
-
   OneSumMatch ::
@@
       E (ProdMatch a b) -> E (SumMatch a b)
-
   EmptyMatch :: (ERep b) => E (SumMatch Void b)
-  ...
 \end{lstlisting}
@@
 runIter :: (ERep a, ERep b)
-  => (a -> Iter b a)
-     -> (a -> b)
+  => (a -> Iter b a) -> (a -> b)
 runIter f = go
@@
       True -> True
-      False -> case x == 1 of
+      False ->
+        case x == 1 of
           True -> False
@@
     Name a -> E b -> E (a -> b)
+
   Var :: (Typeable a) => Name a -> E a
@@
 
-\begin{wrapfigure}{r}{0.5\linewidth}
-  % \centering
-  \resizebox{0.5\columnwidth}{7cm}{\includegraphics{Diagrams/Flowchart.pdf}}
-  % \includegraphics{Diagrams/Flowchart.pdf}
-   \caption{Control Flow of the Core Plugin}
-   \label{fig:CorePlugin}
-\end{wrapfigure}%
-
 The Core plugin translates marked expressions. Expressions are marked by
@@
 
+% \begin{wrapfigure}{r}{0.5\linewidth}
+\begin{figure}
+  \centering
+  % \vspace{-1cm}
+  \resizebox{0.8\columnwidth}{8.2cm}{\includegraphics{Diagrams/Flowchart.pdf}}
+  % \includegraphics{Diagrams/Flowchart.pdf}
+   \caption{Control Flow of the Core Plugin}
+   \label{fig:CorePlugin}
+  % \vspace{-0.1cm}
+\end{figure}%
+
+\pagebreak
 Note that:
@@
   \centering
+  % \vspace{-0.9cm}
 \resizebox{1\linewidth}{!}{
@@
 % \end{wrapfigure}
+
+% \vspace{-1cm}
+
 \end{figure}

### 2020-10-10 paper/paper.tex
Reviewer typo fixes

@@
 % Say why it is an interesting problem
-  However, capturing traditional control flow in an deeply embedded domain specific language
+  However, capturing traditional control flow in a deeply embedded domain specific language
   would support the writing of programs in a natural style by allowing the programmer to use the
@@
 example' :: E Int
-example' = CaseExp (LeftExp 'a') (SumMatchExp
+example' = CaseExp (LeftExp (CharLit 'a'))
+  (SumMatchExp
     (OneProdMatch (Lam 1 (Ord (Var 1))))
@@
 and \ttt{Void}.  This "standardized" type representation is given by \ttt{ERepTy
-T}. \ttt{ERepTy} is an type family associated to the type class \ttt{ERep}. This
+T}. \ttt{ERepTy} is a type family associated to the type class \ttt{ERep}. This
 type will be called the \textit{canonical type} of \ttt{T}.
@@
 
-\section{Representing pattern matches}
+\section{Representing pattern matching}
 \label{sec:PatRep}

### 2020-10-10 paper/paper.tex
wibble

@@
 GADT called \verb|E|. A GHC plugin is then used to transform code written in a
-subset of Haskell containing algebraic data types, pattern matching, tail
+subset of Haskell consisting of algebraic data types, pattern matching, tail
 recursion and lambdas into values of this GADT. A backend can then use these

### 2020-11-17 paper/paper.tex
typos

@@
   lambdas and primitive operations (Section~\ref{sec:CorePlugin})
-\item An implementation of a interpreter for the "standard" semantics of the EDSL (Section~\ref{sec:StdSemantics}).
+  \item An implementation of an interpreter for the "standard" semantics of the EDSL (Section~\ref{sec:StdSemantics}).
 % \item An outline of a C backend for this EDSL (Section~\ref{sec:CBackend}).
@@
 
-Briefly, a algebraic type \ttt{T} with an automatically generated \ttt{ERep}
+Briefly, an algebraic type \ttt{T} with an automatically generated \ttt{ERep}
 instance is given a representation in terms of \ttt{Either}, \ttt{(,)}, \ttt{()}

### 2020-11-17 paper/paper.tex
Add 'construct' implementations
formatting

@@
   type ERepTy (Either a b) = Either a b
-
+  construct (Left x)  = LeftExp (rep x)
+  construct (Right y) = RightExp (rep y)
   rep (Left x)  = LeftExp (rep x)
@@
   type ERepTy (a, b) = (a, b)
-
+  construct (x, y) = PairExp (rep x) (rep y)
   rep (x, y) = PairExp (rep x) (rep y)

### 2020-11-17 paper/paper.tex
clean up sentence in abstract

@@
 % Here is why the world will be a better place
-  With this new support, the Haskell that we capture for expressing
+  With this new support, the subset of Haskell that we use for expressing
   deeply embedded domain specific languages can be cleaner, Haskell-idiomatic,

### 2020-11-17 paper/paper.tex
better tildes in code listings
typo in code
better double quotes

@@
 
-\lstset{language=haskell, deletekeywords={abs},basicstyle=\ttfamily}
+\lstset{literate={~} {$\sim$}{1}, language=haskell, deletekeywords={abs},basicstyle=\ttfamily}
 
@@
 for this is that control flow constructs which are typically \textit{outside} the EDSL,
-such as pattern matching and tail recursion, can make it challenging to "look inside"
+such as pattern matching and tail recursion, can make it challenging to ``look inside''
 the lambda. When these control structures are reified within the EDSL, it is
@@
   add a feature to the underlying implementation that allows an existing EDSL to be
-  "enriched" with pattern matching support.
+  ``enriched'' with pattern matching support.
 \end{todo}
@@
   lambdas and primitive operations (Section~\ref{sec:CorePlugin})
-  \item An implementation of an interpreter for the "standard" semantics of the EDSL (Section~\ref{sec:StdSemantics}).
+  \item An implementation of an interpreter for the ``standard'' semantics of the EDSL (Section~\ref{sec:StdSemantics}).
 % \item An outline of a C backend for this EDSL (Section~\ref{sec:CBackend}).
@@
 
-example :: Int
+example :: E Int
 example =
@@
 instance is given a representation in terms of \ttt{Either}, \ttt{(,)}, \ttt{()}
-and \ttt{Void}.  This "standardized" type representation is given by \ttt{ERepTy
+and \ttt{Void}.  This ``standardized'' type representation is given by \ttt{ERepTy
 T}. \ttt{ERepTy} is a type family associated to the type class \ttt{ERep}. This
@@
 
-Only the "outermost" type will be deconstructed into these fundamental building
+Only the ``outermost'' type will be deconstructed into these fundamental building
 blocks, and further deconstruction can take place on these pieces later on. For
@@
 Note that \ttt{ERepTy IntList \typeeq Either () (Int, IntList)}. If \ttt{ERepTy}
-attempted to "fully deconstruct" \ttt{IntList}, it would send the compiler
+attempted to ``fully deconstruct'' \ttt{IntList}, it would send the compiler
 into an infinite loop.
@@
 all Haskell types which can be represented in the DSL. \ttt{ERepTy} is a type
-family associated to the \ttt{ERep} type class, which represents a "canonical form"
+family associated to the \ttt{ERep} type class, which represents a ``canonical form''
 of the given type. This canonical form can be immediately constructed in the EDSL.
@@
 (The above code block is a slight simplification of the actual implementation,
-which has an additional type variable which is used at an "intermediate" point.
+which has an additional type variable which is used at an ``intermediate'' point.
 The expanded form is in place to simplify the Core transformation.)
@@
 Consider a tail recursive function of the type \ttt{f :: a -> b}. Each recursive
-call can be seen as a simple "update" of the values of the arguments, since
+call can be seen as a simple ``update'' of the values of the arguments, since
 these calls are all in tail position. This is why tail recursive functions
@@
 \ttt{f :: a -> b} into a function \ttt{f' :: a -> Iter b a}. \ttt{Iter b a} is a
-type which can either correspond to an argument "update" (in the sense mentioned
+type which can either correspond to an argument ``update'' (in the sense mentioned
 previously) or a final result. This type is implemented as a sum of the types

### 2020-11-19 src/CodeGen/C.hs
work on plugin backend

@@ -368,6 +368,7 @@ prelude varCount =
     , ", EXPR_STEP"
     , ", EXPR_DONE"
     , ", EXPR_UNBOXED"
+    , ", EXPR_NIL"
     , "} var_type_tag;"
     , ""
     , "typedef enum semantic_type_tag {"
@@ -401,9 +402,11 @@ prelude varCount =
     , "#define GET_COMPARE_OP_GT >"
     , "#define GET_COMPARE_OP_EQUAL =="
     , ""
+    , "#define assert_msg(cond, msg) do { if (!(cond)) { printf msg; assert(0 && #cond); } } while(0)"
+    , ""
     , "#define MATH_OP(op, result, a, b)\\"
     , "  do {\\"
-    , "    assert((a).tag == (b).tag);\\"
+    , "    assert_msg((a).tag == (b).tag, (\"(a,b).tag = (%i,%i)\\n\", (int)(a).tag, (int)(b).tag));\\"
     , "    (result).semantic_tag = (a).semantic_tag;\\"
     , "    (result).tag = (a).tag;\\"
     , "    switch ((a).tag) {\\"
@@ -477,7 +480,7 @@ prelude varCount =
     , "        *((double*)(result).value) = sqrt(*(double*)((a).value));\\"
     , "        break;\\"
     , "  } while(0);"
-    , ""  -- 
+    , ""
     , "#define INIT_COMPLEX_PAIR(result)\\"
     , "  do {\\"
     , "    (result).semantic_tag = EXPR_COMPLEX;\\"
@@ -492,8 +495,21 @@ prelude varCount =
     , "    (a).value = malloc(2*sizeof(var_t));\\"
     , "    ((var_t*)((a).value))[0].tag = eTag;\\"
     , "    ((var_t*)((a).value))[0].value = malloc(sizeof(var_t));\\"
-    , "    ((var_t*)((a).value))[1].tag = eTag;\\"
-    , "    ((var_t*)((a).value))[1].value = malloc(sizeof(var_t));\\"
+
+    , "    ((var_t*)((a).value))[1].tag = EXPR_PAIR;\\"
+    , "    ((var_t*)((a).value))[1].semantic_tag = NO_SEMANTIC_TAG;\\"
+    , "    ((var_t*)((a).value))[1].value = malloc(2*sizeof(var_t));\\"
+
+    , "    ((var_t*)((var_t*)((a).value))[1].value)[0].tag = eTag;\\"
[…]

### 2020-11-19 test3-backup.c

@@ -1,0 +1,891 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdbool.h>
+#include <string.h>
+#include <assert.h>
+#include <math.h>
+
+typedef enum var_type_tag {
+  EXPR_INT
+, EXPR_FLOAT
+, EXPR_DOUBLE
+, EXPR_CHAR
+, EXPR_BOOL
+, EXPR_CLOSURE
+, EXPR_EITHER_LEFT
+, EXPR_EITHER_RIGHT
+, EXPR_PAIR
+, EXPR_UNIT
+, EXPR_STEP
+, EXPR_DONE
+, EXPR_UNBOXED
+, EXPR_NIL
+} var_type_tag;
+
+typedef enum semantic_type_tag {
+  NO_SEMANTIC_TAG
+, EXPR_COMPLEX     // Complex numbers
+} semantic_type_tag;
+
+struct closure_t;
+
+typedef struct var_t {
+  var_type_tag tag;
+  semantic_type_tag semantic_tag;  // For optional additional information about the "un-deconstructed" type (for instance, it could be the case that tag=EXPR_PAIR, but specific_tag=EXPR_COMPLEX)
+  void* value;
+} var_t;
+
+typedef struct closure_t {
+  var_t* fv_env;
+  var_t (*fn)(var_t, struct closure_t*);
+} closure_t;
+
+var_t vars[87];
+
+#define GET_MATH_OP_ADD +
+#define GET_MATH_OP_MUL *
+#define GET_MATH_OP_SUB -
+#define GET_MATH_OP_DIV /
+
+#define GET_COMPARE_OP_LTE <=
+#define GET_COMPARE_OP_LT <
+#define GET_COMPARE_OP_GTE >=
+#define GET_COMPARE_OP_GT >
+#define GET_COMPARE_OP_EQUAL ==
+
+#define assert_msg(cond, msg) do { if (!(cond)) { printf msg; assert(0 && #cond); } } while(0)
+
+#define MATH_OP(op, result, a, b)\
+  do {\
+    assert_msg((a).tag == (b).tag, ("(a,b).tag = (%i,%i)\n", (int)(a).tag, (int)(b).tag));\
+    (result).semantic_tag = (a).semantic_tag;\
+    (result).tag = (a).tag;\
+    switch ((a).tag) {\
+      case EXPR_INT:\
+        (result).value = malloc(sizeof(int));\
[…]

### 2020-11-19 test3.c

@@ -1,0 +1,888 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdbool.h>
+#include <string.h>
+#include <assert.h>
+#include <math.h>
+
+typedef enum var_type_tag {
+  EXPR_INT
+, EXPR_FLOAT
+, EXPR_DOUBLE
+, EXPR_CHAR
+, EXPR_BOOL
+, EXPR_CLOSURE
+, EXPR_EITHER_LEFT
+, EXPR_EITHER_RIGHT
+, EXPR_PAIR
+, EXPR_UNIT
+, EXPR_STEP
+, EXPR_DONE
+, EXPR_UNBOXED
+, EXPR_NIL
+} var_type_tag;
+
+typedef enum semantic_type_tag {
+  NO_SEMANTIC_TAG
+, EXPR_COMPLEX     // Complex numbers
+} semantic_type_tag;
+
+struct closure_t;
+
+typedef struct var_t {
+  var_type_tag tag;
+  semantic_type_tag semantic_tag;  // For optional additional information about the "un-deconstructed" type (for instance, it could be the case that tag=EXPR_PAIR, but specific_tag=EXPR_COMPLEX)
+  void* value;
+} var_t;
+
+typedef struct closure_t {
+  var_t* fv_env;
+  var_t (*fn)(var_t, struct closure_t*);
+} closure_t;
+
+var_t vars[87];
+
+#define GET_MATH_OP_ADD +
+#define GET_MATH_OP_MUL *
+#define GET_MATH_OP_SUB -
+#define GET_MATH_OP_DIV /
+
+#define GET_COMPARE_OP_LTE <=
+#define GET_COMPARE_OP_LT <
+#define GET_COMPARE_OP_GTE >=
+#define GET_COMPARE_OP_GT >
+#define GET_COMPARE_OP_EQUAL ==
+
+#define assert_msg(cond, msg) do { if (!(cond)) { printf msg; assert(0 && #cond); } } while(0)
+
+#define MATH_OP(op, result, a, b)\
+  do {\
+    assert_msg((a).tag == (b).tag, ("(a,b).tag = (%i,%i)\n", (int)(a).tag, (int)(b).tag));\
+    (result).semantic_tag = (a).semantic_tag;\
+    (result).tag = (a).tag;\
+    switch ((a).tag) {\
+      case EXPR_INT:\
+        (result).value = malloc(sizeof(int));\
[…]

### 2020-11-19 src/CodeGen/C.hs
Work on getting tuple-lists to work properly in C backend

@@ -1117,8 +1117,10 @@ genExp (NullaryMatch _) _ = error "genExp: NullaryMatch"
 genExp (OneSumMatch _) _ = error "genExp: OneSumMatch"
 genExp EmptyMatch _ = error "genExp: EmptyMatch"
 
-genCdr :: GPUExp a -> CName -> CodeGen CCode
-genCdr x@(PairExp {}) resultName = genExp x resultName
+genCdr :: forall a. GPUExp a -> CName -> CodeGen CCode
+genCdr x@(PairExp {}) resultName
+  | Nothing <- eqT :: Maybe (a :~: Complex Double)
+      = genExp x resultName
 genCdr x resultName = do
   xName <- freshCName
 

### 2020-11-19 test3.c

@@ -218,549 +218,549 @@ var_t lam_86(var_t, struct closure_t*);
 
 var_t lam_79(var_t arg, struct closure_t* self) {
   var_t x1;
-  var_t x2;
-  closure_t x3;
-  var_t x4;
+var_t x2;
+closure_t x3;
+var_t x4;
 
   // Var with Name id 79 and Haskell type (Int,(Complex Double),(Complex Double))
-  if (isIterTag(arg.tag)) {
-    x2 = *(var_t*)(arg.value);
-  } else {
-    x2 = arg;
-  }
+if (isIterTag(arg.tag)) {
+x2 = *(var_t*)(arg.value);
+} else {
+x2 = arg;
+}
 
-  closure_t* x5 = malloc(sizeof(closure_t));
-  (*x5).fv_env = malloc(sizeof(var_t)*0);
-  (*x5).fn = &lam_80;
+closure_t* x5 = malloc(sizeof(closure_t));
+(*x5).fv_env = malloc(sizeof(var_t)*0);
+(*x5).fn = &lam_80;
 
 
 
-  x4.tag = EXPR_CLOSURE;
-  x4.semantic_tag = NO_SEMANTIC_TAG;
-  x4.value = (void*)x5;
+x4.tag = EXPR_CLOSURE;
+x4.semantic_tag = NO_SEMANTIC_TAG;
+x4.value = (void*)x5;
 
-  memcpy(&x3, (closure_t*)(x4.value), sizeof(closure_t));
-  x1 = x3.fn(x2, &x3);
+memcpy(&x3, (closure_t*)(x4.value), sizeof(closure_t));
+x1 = x3.fn(x2, &x3);
 
   return x1;
 }
 
 var_t lam_80(var_t arg, struct closure_t* self) {
   var_t x7;
-  x7.tag = EXPR_STEP;
-  x7.semantic_tag = NO_SEMANTIC_TAG;
-  x7.value = malloc(sizeof(var_t));
-  *(var_t*)(x7.value) = arg;
-  while (x7.tag != EXPR_DONE) {
-    var_t x8;
-    // Var with Name id 80 and Haskell type (Int,(Complex Double),(Complex Double))
-    if (isIterTag(x7.tag)) {
-      x8 = *(var_t*)(x7.value);
-    } else {
-      x8 = x7;
-    }
+x7.tag = EXPR_STEP;
+x7.semantic_tag = NO_SEMANTIC_TAG;
+x7.value = malloc(sizeof(var_t));
[…]

### 2020-11-19 src/CodeGen/C.hs
Add some deep copying to the C backend, partial bug fix

@@ -367,7 +367,7 @@ prelude varCount =
     , ", EXPR_UNIT"
     , ", EXPR_STEP"
     , ", EXPR_DONE"
-    , ", EXPR_UNBOXED"
+    -- , ", EXPR_UNBOXED"
     , ", EXPR_NIL"
     , "} var_type_tag;"
     , ""
@@ -388,6 +388,51 @@ prelude varCount =
     , "  var_t* fv_env;"
     , "  var_t (*fn)(var_t, struct closure_t*);"
     , "} closure_t;"
+    , ""
+    , "void copyVar(var_t* dest, var_t* src) {"
+    , "  dest->tag = src->tag;"
+    , "  dest->semantic_tag = src->semantic_tag;"
+    , "  if (src->tag == EXPR_PAIR) {"
+    , "    dest->value = malloc(2*sizeof(var_t));"
+    , "    copyVar(&((var_t*)dest->value)[0], &((var_t*)src->value)[0]);"
+    , "    copyVar(&((var_t*)dest->value)[1], &((var_t*)src->value)[1]);"
+    , "  } else {"
+    , "    switch (src->tag) {"
+    , "      case EXPR_INT:"
+    , "       dest->value = malloc(sizeof(int));"
+    , "       break;"
+    , "      case EXPR_FLOAT:"
+    , "       dest->value = malloc(sizeof(float));"
+    , "       break;"
+    , "      case EXPR_DOUBLE:"
+    , "       dest->value = malloc(sizeof(double));"
+    , "       break;"
+    , "      case EXPR_CHAR:"
+    , "       dest->value = malloc(sizeof(char));"
+    , "       break;"
+    , "      case EXPR_CLOSURE:"
+    , "       dest->value = malloc(sizeof(closure_t));"
+    , "       break;"
+    , "      case EXPR_EITHER_LEFT:"
+    , "      case EXPR_EITHER_RIGHT:"
+    , "      case EXPR_STEP:"
+    , "      case EXPR_DONE:"
+    , "      case EXPR_UNIT:"
+    , "      case EXPR_NIL:"
+    , "       dest->value = malloc(sizeof(var_t));"
+    , "       break;"
+    , "      default:"
+    , "        assert(0 && \"invalid tag\");"
+    , "    }"
+    , "  }"
+    , "}"
[…]

### 2020-11-19 src/CodeGen/C/Helper.hs

@@ -15,6 +15,9 @@ cCall fnName args =
 derefPtr :: CCode -> CCode
 derefPtr x = "*(" <> x <> ")"
 
+addrOf :: CCode -> CCode
+addrOf x = "&(" <> x <> ")"
+
 cCast :: CCode -> CCode -> CCode
 cCast toType x = "((" <> toType <> ")" <> x <> ")"
 

### 2020-11-20 paper/paper.tex
Changed type-level section title
Improved description of flowchart and rewrite rule figure

@@
 
-\subsection{\ttt{E}, \ttt{ERep}, \ttt{ERepTy}}
+\subsection{The Three Major Type-Level Pieces}
 
@@
 
-In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses in a node refer
-to the names of transformations given in the rewrite rules in Figure~\ref{fig:Rewrites}. In Figure~\ref{fig:Rewrites},
+The plugin's control flow is described in Figure~\ref{fig:CorePlugin}. When control
+flow reaches a node, the rewrite rules listed in parentheses are performed. These rewrite
+rule names correspond to the rules given in Figure~\ref{fig:Rewrites}.
+
+% In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses
+% in a node refer to the names of transformations given in the rewrite rules in
+% Figure~\ref{fig:Rewrites}. The control flow of the plugin is described by the
+% flowchart and the rewrites performed when control reaches a node is given by the list of
+% rule names in parentheses in that node.
+
+In Figure~\ref{fig:Rewrites},
 $\varepsilon$ represents an empty term (in the context of the term rewriting system used in that Figure). Calls

### 2020-11-20 src/CodeGen/C.hs
More deep copies in plugin C backend

@@ -400,26 +400,35 @@ prelude varCount =
     , "    switch (src->tag) {"
     , "      case EXPR_INT:"
     , "       dest->value = malloc(sizeof(int));"
+    , "       *(int*)(dest->value) = *(int*)(src->value);"
     , "       break;"
     , "      case EXPR_FLOAT:"
     , "       dest->value = malloc(sizeof(float));"
+    , "       *(float*)(dest->value) = *(float*)(src->value);"
     , "       break;"
     , "      case EXPR_DOUBLE:"
     , "       dest->value = malloc(sizeof(double));"
+    , "       *(double*)(dest->value) = *(double*)(src->value);"
     , "       break;"
     , "      case EXPR_CHAR:"
     , "       dest->value = malloc(sizeof(char));"
+    , "       *(char*)(dest->value) = *(char*)(src->value);"
     , "       break;"
     , "      case EXPR_CLOSURE:"
     , "       dest->value = malloc(sizeof(closure_t));"
+    , "       *(closure_t*)(dest->value) = *(closure_t*)(src->value);"
     , "       break;"
     , "      case EXPR_EITHER_LEFT:"
     , "      case EXPR_EITHER_RIGHT:"
     , "      case EXPR_STEP:"
     , "      case EXPR_DONE:"
     , "      case EXPR_UNIT:"
-    , "      case EXPR_NIL:"
     , "       dest->value = malloc(sizeof(var_t));"
+    , "       copyVar((var_t*)(dest->value), (var_t*)(src->value));"
+    -- , "       *(var_t*)(dest->value) = *(var_t*)(src->value);"
+    , "       break;"
+    , "      case EXPR_NIL:"
+    , "       dest->value = NULL;"
     , "       break;"
     , "      default:"
     , "        assert(0 && \"invalid tag\");"
@@ -824,7 +833,8 @@ genExp (DoneExp x) resultName = do
     , resultName <> ".tag = EXPR_DONE;"
     , resultName <> ".semantic_tag = NO_SEMANTIC_TAG;"
     , resultName <> ".value = malloc(sizeof(var_t));"
-    , "*(var_t*)(" <> resultName <> ".value) = " <> xName <> ";"
+    , stmt $ cCall "copyVar" [cCast "var_t*" (resultName # "value"), addrOf xName]
[…]

### 2020-11-20 src/CodeGen/C.hs
wibble

@@ -1415,9 +1415,13 @@ buildClosure sc@(SomeLambda c) closureVarName = do
 callClosure :: SomeLambda -> CName -> CName -> CName -> CodeGen CCode
 callClosure (SomeLambda (Lambda { lambda_name })) closureName argName resultName = do
   r <- freshCName
+  -- copiedArg <- freshCName
   return $
     unlines
-      [ "var_t " <> r <> " = " <> closureName <> ".fn(" <> argName <> ", &" <> closureName <> ");"
+      [
+       -- "var_t " <> copiedArg <> ";"
+      -- , stmt $ cCall "copyVar" [addrOf copiedArg, addrOf argName]
+        "var_t " <> r <> " = " <> closureName <> ".fn(" <> argName <> ", &" <> closureName <> ");"
       , stmt $ cCall "copyVar" [addrOf resultName, addrOf r]
       ]
 

### 2020-11-20 test3.c

@@ -18,7 +18,6 @@ typedef enum var_type_tag {
 , EXPR_UNIT
 , EXPR_STEP
 , EXPR_DONE
-, EXPR_UNBOXED
 , EXPR_NIL
 } var_type_tag;
 
@@ -39,6 +38,59 @@ typedef struct closure_t {
   var_t* fv_env;
   var_t (*fn)(var_t, struct closure_t*);
 } closure_t;
+
+void copyVar(var_t* dest, var_t* src) {
+  dest->tag = src->tag;
+  dest->semantic_tag = src->semantic_tag;
+  if (src->tag == EXPR_PAIR) {
+    dest->value = malloc(2*sizeof(var_t));
+    copyVar(&((var_t*)dest->value)[0], &((var_t*)src->value)[0]);
+    copyVar(&((var_t*)dest->value)[1], &((var_t*)src->value)[1]);
+  } else {
+    switch (src->tag) {
+      case EXPR_INT:
+       dest->value = malloc(sizeof(int));
+       *(int*)(dest->value) = *(int*)(src->value);
+       break;
+      case EXPR_FLOAT:
+       dest->value = malloc(sizeof(float));
+       *(float*)(dest->value) = *(float*)(src->value);
+       break;
+      case EXPR_DOUBLE:
+       dest->value = malloc(sizeof(double));
+       *(double*)(dest->value) = *(double*)(src->value);
+       break;
+      case EXPR_CHAR:
+       dest->value = malloc(sizeof(char));
+       *(char*)(dest->value) = *(char*)(src->value);
+       break;
+      case EXPR_CLOSURE:
+       dest->value = malloc(sizeof(closure_t));
+       *(closure_t*)(dest->value) = *(closure_t*)(src->value);
+       break;
+      case EXPR_EITHER_LEFT:
+      case EXPR_EITHER_RIGHT:
+      case EXPR_STEP:
+      case EXPR_DONE:
+      case EXPR_UNIT:
+       dest->value = malloc(sizeof(var_t));
+       copyVar((var_t*)(dest->value), (var_t*)(src->value));
+       break;
+      case EXPR_NIL:
+       dest->value = NULL;
[…]

### 2020-11-23 paper/paper.tex
Motivation: describe the current state of EDSLs wrt pattern matching and how this improves on it

@@
 either difficult or impossible to use a language that has these benefits.
+However, there is no support for capturing pattern matching builtin to the Haskell
+language itself. As a result, EDLs which capture pattern matching must do so
+via a plugin in a Haskell compiler such as GHC. For each such EDSL, this mechanism
+has needed to be reimplemented each time. In this paper, we descibe a \textit{reusable}
+framework for capturing pattern matching, as well as lambdas and tail recursion.
 

### 2020-11-24 paper/paper.tex
motivation: difficult and error-prone
spacing improvements

@@
 %% used to denote shared contribution to the research.
-\author{David Young\inst{1} \and Mark Grebe\inst{2} \and Andrew Gill\inst{1}}
+\author{David Young\inst{1} \and Mark Grebe\inst{2} \and Andy Gill\inst{1}}
 \institute{University of Kansas \and University of Central Missouri}
@@
 lexical closures.
-
+%
 Despite these advantages Haskell provides for creating EDSLs, there are a few
@@
 frequently a desirable feature for many EDSLs.
-
+%
 Additionally, lambdas can also be difficult to implement in EDSLs. A major reason
@@
   \item A representation of pattern matching in a Haskell EDSL (Section~\ref{sec:PatRep})
-  \item A representation of lambdas (Section~\ref{sec:LamRep}) in the EDSL, taking advantage of
+  \item A representation of lambdas (Section~\ref{sec:LamRep}) in the EDSL, utilizing
   the fact that both pattern matching and tail recursion are already captured in
@@
 via a plugin in a Haskell compiler such as GHC. For each such EDSL, this mechanism
-has needed to be reimplemented each time. In this paper, we descibe a \textit{reusable}
+has needed to be reimplemented each time. Additionally, the task of implementing this
+is difficult and error-prone. In this paper, we descibe a \textit{reusable}
 framework for capturing pattern matching, as well as lambdas and tail recursion.
@@
   type ERepTy t
-
   construct :: t -> E (ERepTy t)
   rep :: t -> E t
-
   default rep :: (ERep (ERepTy t)) => t -> E t
   rep x = ConstructRep (construct x)
-
   unrep' :: ERepTy t -> t
@@
 
-(The above code block is a slight simplification of the actual implementation,
+The above code block is a slight simplification of the actual implementation,
 which has an additional type variable which is used at an ``intermediate'' point.
-The expanded form is in place to simplify the Core transformation.)
+The expanded form is in place to simplify the Core transformation.
 

### 2020-11-25 paper/paper.tex
Add more plugin explanation; remove some blank lines in code snippets

@@
     E (ERepTy a) -> E a
-
   LeftExp :: E a -> E (Either a b)
   RightExp :: E b -> E (Either a b)
-
   PairExp :: E a -> E b -> E (a, b)
@@
 \ttt{CaseExp} constructor:
-
 \begin{lstlisting}
@@
       -> E (SumMatch (Either a b) r)
+
   OneSumMatch ::
@@
   DoneExp :: E a -> E (Iter a b)
-
   TailRec :: (ERep a, ERep b) =>
@@
     Name a -> E b -> E (a -> b)
-
   Var :: (Typeable a) => Name a -> E a
@@
     go []                = Nothing
-
     go ((n' :=> e):rest) =
@@
 ...
-
 prodMatchAbs :: (ERep s) =>
@@
 
-For example, \ttt{externalize x} marks the expression \ttt{x}.
-
-If an expression already has an EDSL type (a type of the form \ttt{E a} for some
-\ttt{a}), then the marking procedure ignores it and does not wrap it with
-\ttt{externalize} (see $M$ in Figure~\ref{fig:Rewrites}).
+For example, \ttt{externalize x} marks the expression \ttt{x}.  If an expression
+already has an EDSL type (a type of the form \ttt{E a} for some \ttt{a}), then
+the marking procedure ignores it and does not wrap it with \ttt{externalize}
+(see $M$ in Figure~\ref{fig:Rewrites}).
+
+These \verb|externalize| marks are pushed through the input program, starting at
+the outermost level. The plugin repeatedly runs over each newly marked node in
+the abstract syntax tree, applying the appropriate rewrite rules as it goes.
+This process pushes the marks through the program until it reaches a base case
+that has no proper subexpressions, such as a character literal.
 
@@
 rule names correspond to the rules given in Figure~\ref{fig:Rewrites}.
-
 % In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses
@@
 % rule names in parentheses in that node.
-
 In Figure~\ref{fig:Rewrites},

### 2020-11-25 paper/paper.tex
'standard semantics' -> 'reference semantics'

@@
   lambdas and primitive operations (Section~\ref{sec:CorePlugin})
-  \item An implementation of an interpreter for the ``standard'' semantics of the EDSL (Section~\ref{sec:StdSemantics}).
+  \item An implementation of an interpreter for a reference semantics of the EDSL (Section~\ref{sec:StdSemantics}).
 % \item An outline of a C backend for this EDSL (Section~\ref{sec:CBackend}).
@@
 
-The function \ttt{runIter} provides the standard semantics for executing such a
+The function \ttt{runIter} provides the reference semantics for executing such a
 function representing a tail recursive function:
@@
 
-\section{Recovering standard semantics for pattern matches}
+\section{Recovering reference semantics for pattern matches}
 \label{sec:StdSemantics}
 
-The standard semantics for the EDSL is given by \ttt{abs}. Two helper
+The reference semantics for the EDSL is given by \ttt{abs}. Two helper
 functions, \ttt{sumMatchAbs} and \ttt{prodMatchAbs}, are used to provide
-the standard semantics for matches on sum types and matches on product
+the reference semantics for matches on sum types and matches on product
 types, respectively. These helper functions are based on the mechanism

### 2020-11-26 paper/paper.tex
tweaking the bib
Removed a lot of commented-out LaTeX

@@
-%% This is file `sample-acmsmall.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `acmsmall')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-acmsmall.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%% The first command in your LaTeX source must be the \documentclass command.
 \RequirePackage{amsmath,amssymb}
-% \RequirePackage{stmaryrd}
+
 \documentclass[runningheads, a4paper]{llncs}
-% \usepackage{biblatex}
 \usepackage{lineno}
@@
 \usepackage{wrapfig}
-% \usepackage{amsmath}
-
-% \lstset{basicstyle=\ttfamily}
+
 \usetikzlibrary{shapes.geometric, arrows}
@@
 % TODO: Use these:
+
 \newcommand{\exprt}[1]{\expr{\ttt{#1}}}
@@
 \newcommand{\ttt}{\texttt}
-
-% \newcommand{\showtodos}{}  % Comment out to hide todos
 
@@
                {\endcomment}
-
-% \newenvironment<>{todo}[1][]{
-%     \setbeamercolor{block body example}{fg=black,bg=blue}
-%     \setbeamercolor{block title example}{fg=white,bg=red!75!black}
-%     \setbeamertemplate{blocks}[rounded][shadow=false]
-%   \begin{todo}[]}{\end{todo}
-% }
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
 \AtBeginDocument{%
@@
 
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-
-% \setcopyright{acmcopyright}
-% \copyrightyear{2020}
-% \acmYear{2020}
-% \acmDOI{XX.XXXX/XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-
-% \acmConference[GPCE '20]{GPCE '20: Proceedings of the 13th
-%   ACM SIGPLAN International Symposium on Haskell}{November 15 -- 20, 2020}{Chicago, Illinois, United States}
-% \acmBooktitle{Proceedings of the 19th ACM SIGPLAN International
-% Conference on Generative Programming: Concepts and Experiences (GPCE '20)
-%   November 15 -- 20, 2020, Chicago, Illinois, United States}
-% \acmPrice{15.00}
-% \acmISBN{XXX-X-XXXX-XXXX-X/XX/XX}
-
-% \citestyle{acmauthoryear}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-%%
-%% end of the preamble, start of the body of the document source.
-
 \begin{document}
 
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
 \title{On Adding Pattern Matching to Haskell-based
@@
 
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
 \author{David Young\inst{1} \and Mark Grebe\inst{2} \and Andy Gill\inst{1}}
@@
 \maketitle
-% \authornote{Both authors contributed equally to this research.}
-% \email{trovato@corporation.com}
-% \orcid{1234-5678-9012}
-
-% \author{Mark Grebe}
-% \email{grebe@ucmo.edu}
-
-% \author{Andrew Gill}
-% \email{andygill@ku.edu}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-% \renewcommand{\shortauthors}{Young, Grebe and Gill}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
-
-% \linenumbers
 
@@
 
-% \begin{CCSXML}
-% <ccs2012>
-%    <concept>
-%        <concept_id>10011007.10011006.10011041.10011046</concept_id>
-%        <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
-%        <concept_significance>300</concept_significance>
-%        </concept>
-%    <concept>
-%        <concept_id>10011007.10011006.10011041.10011047</concept_id>
-%        <concept_desc>Software and its engineering~Source code generation</concept_desc>
-%        <concept_significance>300</concept_significance>
-%        </concept>
-%    <concept>
-%        <concept_id>10011007.10011006.10011050.10011017</concept_id>
-%        <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
-%        <concept_significance>500</concept_significance>
-%        </concept>
-%  </ccs2012>
-% \end{CCSXML}
-
-% \ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}
-% \ccsdesc[300]{Software and its engineering~Source code generation}
-% \ccsdesc[500]{Software and its engineering~Domain specific languages}
-
-% \begin{CCSXML}
-% <ccs2012>
-%    <concept>
-%        <concept_id>10011007.10011006.10011050.10011017</concept_id>
-%        <concept_desc>Software and its engineering~Domain specific languages</concept_desc>
-%        <concept_significance>500</concept_significance>
-%        </concept>
-%  </ccs2012>
-% \end{CCSXML}
-
-% \ccsdesc[500]{Software and its engineering~Domain specific languages}
-
-% \keywords{domain specific languages}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-% \begin{CCSXML}
-% <ccs2012>
-%  <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-%   <concept_significance>500</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
-%   <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-% \ccsdesc[500]{Computer systems organization~Embedded systems}
-% \ccsdesc[300]{Computer systems organization~Redundancy}
-% \ccsdesc{Computer systems organization~Robotics}
-% \ccsdesc[100]{Networks~Network reliability}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-% \keywords{datasets, neural networks, gaze detection, text tagging}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-
 \section{Introduction}
@@
 
-% \begin{todo}
-%   Maybe add brief summary of relationship to lambdas?
-% \end{todo}
-
 \begin{todo}
@@
   \item An implementation of an interpreter for a reference semantics of the EDSL (Section~\ref{sec:StdSemantics}).
-% \item An outline of a C backend for this EDSL (Section~\ref{sec:CBackend}).
 \end{itemize}
-
-% In this paper, a C backend is given as an example in Section \ref{sec:CBackend}. The Core plugin will
-% translate a subset of Haskell to the EDSL we have specified. The C backend will
-% then take expressions written in this EDSL and translate them to C code. This C
-% code can be compiled by a standard C compiler. In effect, the subset of Haskell
-% represented in the EDSL is compiled to C.
 
@@
 \subsection{Motivation}
-
-\begin{todont}
-  Expand?
-\end{todont}
 
@@
 
-% \newpage
 Haskell EDSLs that support powerful language features like pattern matching
@@
 
-% A C backend is described in Section \ref{sec:CBackend}. However, other backends
-% could be constructed, targeting different architectures.
-
-\begin{todont}
-  % Cite examples where Haskell implementations have been derived specifications.
-Conal Elliott's latest two papers (as of May 2020), \cite{Elliott-2018-ad-icfp} and \cite{Elliott2019-convolution-extended}, would be good examples of this.
-(This is now cited)
-
-  A paper which describes a Haskell implementation derived from an operational
-  semantics would also be a good example.
-\end{todont}
-
 \subsection{Overview}
@@
 \subsection{Translation Example}
-
-% \begin{todo}
-%   Does it make sense for this to be a subsection of the intro?
-% \end{todo}
 
@@
 \label{sec:ADTRep}
-
-% \begin{todo}
-%   Look into \ttt{generic-sop} and compare then cite corresponding paper.
-% \end{todo}
 
@@
 
-% \begin{todo}
-%   Should we make a synonym for this constraint since it is used in
-% multiple places? Like this:
-% \begin{lstlisting}
-% type CanonicalIsIdem t =
-%   ERepTy (ERepTy t) ~ ERepTy t
-% \end{lstlisting}
-% \end{todo}
-
 The above code block is a slight simplification of the actual implementation,
@@
 
-% \newpage
 \begin{lstlisting}
@@
 \end{lstlisting}
-
-% \subsection{Connection to Church encodings}
-% \label{sec:Church}
-% % \begin{todo}
-% %   Work on this subsection. Maybe this shouldn't be a separate section at all and
-% %   maybe its contents should be incorporated throughout the rest of the paper.
-% % \end{todo}
-
-% Algebraic datatypes can be represented as lambda calculus terms using
-% \textit{Church encodings}.~\cite{Boehm:1985:AutoSynth} For example, the pair constructor can be
-% represented as
-
-% \begin{lstlisting}
-% church_mkPair
-%   :: a -> b -> (a -> b -> r) -> r
-% church_mkPair x y = \ f -> f x y
-% \end{lstlisting}
-
-% Coproducts can be represented with the constructors
-
-% \begin{lstlisting}
-% church_mkLeft
-%   :: a -> (a -> r) -> (b -> r) -> r
-% church_mkLeft x = \ f g -> f x
-
-% church_mkRight
-%   :: b -> (a -> r) -> (b -> r) -> r
-% church_mkRight y = \ f g -> g y
-% \end{lstlisting}
-
-% In this representation:
-
-% \begin{itemize}
-%   \item A pair of type \ttt{A} and type \ttt{B} is represented with a value of type
-% \ttt{forall r. (A -> B -> r) -> r}.
-
-%   \item A coproduct of type \ttt{A} and type \ttt{B} is represented with a value of type
-% \ttt{forall r. (A -> r) -> (B -> r) -> r}.
-% \end{itemize}
-
-% Both of those representations represent the type as a function that performs the
-% corresponding case analysis. A deep embedding of the operations
-% \ttt{churchMkPair}, \ttt{church\_mkLeft} and \ttt{church\_mkRight} is given
-% by the type
-
-% \begin{lstlisting}
-% data Church r where
-%   ChurchPair
-%     :: a -> b -> (a -> b -> r) ->
-%        Church r
-%   ChurchLeft
-%     :: a -> (a -> r) -> (b -> r) ->
-%        Church r
-%   ChurchRight
-%     :: b -> (a -> r) -> (b -> r) ->
-%        Church r
-% \end{lstlisting}
-
-% with the evaluation function
-
-% \begin{lstlisting}
-% churchEval :: Church r -> r
-% churchEval (ChurchPair x y f) = f x y
-% churchEval (ChurchLeft x f g) = f x
-% churchEval (ChurchRight y f g) = g y
-% \end{lstlisting}
-
-% However, we only want to keep track of the functions used for elimination. By moving
-% the values actually being contained in the products and coproducts out to the
-% evaluation function, we arrive at
-
-% \begin{lstlisting}
-% data Church' t r where
-%   ChurchPair'
-%     :: (a -> b -> r) ->
-%        Church' (a, b) r
-
-%   ChurchEither'
-%     :: (a -> r) -> (b -> r) ->
-%        Church' (Either a b) r
-
-% churchEval' :: Church' t r -> t -> r
-% churchEval' (ChurchPair' f) (x, y) = f x y
-% churchEval' (ChurchEither' f g) e =
-%   case e of
-%     Left  x -> f x
-%     Right y -> g y
-% \end{lstlisting}
-
-% The \verb|Church'| type forms a large part of the foundation of the pattern
-% matching aspect of the \verb|E| type. A rough correspondence between the two
-% is given in the table below.\\
-
-% {\small
-% \noindent\begin{tabular}{|p{3.4cm}|p{4.4cm}|}
-%   \hline
-%   \ttt{Church' (a, b) r} & \ttt{E (SumMatch (a, b) r} \\
-%   \hline
-%   \ttt{Church' (Either a b) r} & \ttt{E (ProdMatch (Either a b) r)}\\
-%   \hline
-% \end{tabular}
-% }
-
-\begin{todo}
-  Figure out how recursive types fit into this framework. Church encodings can
-  handle recursive types and so can our representation. This is also the case
-  for Scott encodings, etc, and is the difference between Church and Scott
-  encodings.
-
-  Does our way of handling recursive types correspond to how any of
-  those encodings handle recursive types?
-
-  % Maybe this is relevant: \href{https://www.reddit.com/r/haskell/comments/hvlh6a/generalized_church_encoding_is_the_curryhoward_of/}{\color{blue}{Generalized Church encoding is the Curry-Howard of Knaster-Tarski}}
-\end{todo}
-
-% \begin{lstlisting}
-% data ChurchMatch t where
-%   ChurchCoproductMatch
-%     :: ChurchMatch ((a, b) -> r) -> ChurchMatch (Either b c -> r) ->
-%        ChurchCoproductMatch (Either
-% \end{lstlisting}
-
-% \begin{todo}
-%   Maybe it could make sense to essentially derive the pattern matching representation
-%   from this Church encoding. If so, this could be described around here.
-% \end{todo}
 
@@
 \end{lstlisting}
-
-% \begin{todo}
-%   \ttt{runIter} has the following alternate implementation. Would this be
-%   useful to use somewhere?
-
-%   \begin{lstlisting}[otherkeywords={}]
-%   runIter f = getDone . fix (f >=>)
-
-%   getDone :: Iter b a -> b
-%   getDone (Done x) = x
-%   getDone _ = error "getDone"
-%   \end{lstlisting}
-
-%   where \verb|Iter| is given a standard \verb|Monad| instance (essentially the
-%   same as the \verb|Either| \verb|Monad| instance). \\
-
-%   Maybe this could be useful in some proofs or derivations somewhere (in this
-%   paper or elsewhere)?
-
-%   Note that \verb|error| is never called under any circumstance in this
-%   implementation of \verb|runIter|.
-% \end{todo}
 
@@
 \end{lstlisting}
-
-% We translate this to the expression language, and then apply two of
-% the constructors we have defined.  The first of these is an initial 
-% \ttt{Done} which we then wrap with a \ttt{runIter} to provide the
-% representation of the iteration execution.
-% The transformation then proceeds by pushing the DoneExp constructor through
-% the conditionals, and the recursive call is transformed into a StepExp constructor.
-% This leaves us the the transformed example:
 
@@
 Next, the pattern matching transformation occurs, producing the final result:
-
-% \break
 
@@
 
-% \newpage
-
-% \section{C Backend}
-% \label{sec:CBackend}
-
-% \begin{todont}
-%   Should this section be expanded?
-% \end{todont}
-
-% The C backend makes use of a small EDSL for generating C code. See
-% Appendix B for implementation details of this ESDL.
-
-% The following \ttt{struct}s provide the C representation for Haskell expressions:
-
-% \begin{lstlisting}
-% typedef enum var_type_tag {
-%   EXPR_INT
-% , EXPR_PAIR
-% ...
-% , EXPR_UNIT
-% } var_type_tag;
-
-% typedef enum semantic_type_tag {
-%   NO_SEMANTIC_TAG
-% , EXPR_COMPLEX     // Complex numbers
-% } semantic_type_tag;
-
-% struct closure_t;
-
-% typedef struct var_t {
-%   var_type_tag tag;
-%   semantic_type_tag semantic_tag;
-%   void* value;
-% } var_t;
-
-% typedef struct closure_t {
-%   var_t* fv_env;
-%   var_t (*fn)(var_t, struct closure_t*);
-% } closure_t;
-% \end{lstlisting}
-
-% The \ttt{semantic\_tag} allows type information preserved by the \ttt{Typeable} constraints
-% in the EDSL to be brought over to the generated C code. In our implementation, this is
-% used to distinguish between pairs of \ttt{Double}s and \ttt{Complex Double} values. In the
-% case of a \ttt{Complex Double}, the \ttt{var\_t} would have a \ttt{tag} field with the value
-% \ttt{EXPR\_PAIR} and a \ttt{semantic\_tag} field with the value \ttt{EXPR\_COMPLEX}.
-
-% A C code generation function \ttt{genC} is implemented:
-
-% \begin{lstlisting}
-% genC :: E a -> String
-% ...
-% \end{lstlisting}
-
-% To use this, for example, with the previously mentioned \ttt{isEven} implementation we
-% can write:
-% % \newpage
-% \begin{lstlisting}
-% isEvenE :: E (Int -> Bool)
-% isEvenE =
-%   externalize (
-%     case x == 0 of
-%       True -> True
-%       False ->
-%         case x == 1 of
-%           True -> False
-%           False -> isEvenE (x - 2))
-
-% isEvenC :: String
-% isEvenC = genC isEvenE
-% \end{lstlisting}
-
 \section{Core Plugin}
@@
 rule names correspond to the rules given in Figure~\ref{fig:Rewrites}.
-% In the flowchart given in Figure~\ref{fig:CorePlugin}, the names in parentheses
-% in a node refer to the names of transformations given in the rewrite rules in
-% Figure~\ref{fig:Rewrites}. The control flow of the plugin is described by the
-% flowchart and the rewrites performed when control reaches a node is given by the list of
-% rule names in parentheses in that node.
+%
 In Figure~\ref{fig:Rewrites},
@@
 
-% \begin{wrapfigure}{r}{0.5\linewidth}
 \begin{figure}
   \centering
-  % \vspace{-1cm}
   \resizebox{0.8\columnwidth}{8.2cm}{\includegraphics{Diagrams/Flowchart.pdf}}
-  % \includegraphics{Diagrams/Flowchart.pdf}
    \caption{Control Flow of the Core Plugin}
    \label{fig:CorePlugin}
-  % \vspace{-0.1cm}
 \end{figure}%
@@
 
-% \clearpage
-
-% \clearpage
-
-% \subsection{Syntactic Transformations of the Core Plugin}
-
-% \begin{todo}
-%   Should we use Oxford brackets here?
-Maybe a bit unusual, but we are operating syntactically and mapping from \ttt{a} to \ttt{E a} (could \ttt{E a} be considered an "intermediary" semantic domain?).
-% \end{todo}
-% \vspace*{-0.6cm}
-
-% \begin{wrapfigure}{r}{0.8\linewidth}
-  % \begin{wrapfigure}{r}{0.59\linewidth}
 \begin{figure}
   \centering
-  % \vspace{-0.9cm}
 \resizebox{1\linewidth}{!}{
@@
   &C\expr{\ttt{f x}} \rewrites \ttt{App $M\expr{f}$ $M\expr{x}$}\\
-  % &C\expr{\ttt{f x}} \rewrites \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
-  % &C\expr{\ttt{x}} \rewrites \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
   &B\expr{\ttt{K x0 ...$_U$ xN $\rarr$ body0; ...$_V$}} \rewrites \ttt{SumMatchExp }P\expr{\ttt{x0 ...$_U$ xN $\rarr$ body0}}\; B\expr{\ttt{ ...$_V$ }}\\
@@
 }%
-% }}
-
-% \captionsetup{justification=centering}
+
 \caption{Rewrite rules}
 \label{fig:Rewrites}
-% \end{wrapfigure}
-% \end{wrapfigure}
-
-% \vspace{-1cm}
 
@@
 
-% \clearpage
-
-% \clearpage
-
 \section{Related Work}
-
-% \begin{todo}
-%   Does this need more fleshing out?
-% \end{todo}
 
@@
 
-% In this paper:
-% \begin{itemize}
-%   \item We address the goal of translating Haskell pattern matching into an
-%     EDSL.
-
-%     % An example usage of this mechanism is given in Section
-%     % \ref{sec:CBackend} in the form of a transformation of a subset of Haskell
-%     % into an EDSL that is then used to generate C code, effectively compiling
-%     % that subset of Haskell into C.
-
-%   % \item A representation of pattern matching via a deep embedding of Church
-%   %   encodings of products and coproducts presented in Section \ref{sec:Church}.
-%   %   This encoding relies on having a uniform representation of algebraic data
-%   %   types.
-
-%   \item To provide the uniform representation of algebraic data types mentioned
-%     in the previous point, we describe a mechanism for automatically encoding
-%     ADTs as compositions of \verb|(,)|, \verb|Either| and \verb|()| using
-%     \verb|ERep| and \verb|ERepTy| in Section \ref{sec:ADTRep}.
-
-%   \item A GHC plugin is outlined in Section \ref{sec:CorePlugin} which provides automation for the translation of
-%     Haskell code into these EDSL representations.
-
-% \end{itemize}
-
 \begin{todo}
@@
 \end{todo}
-% \section{Citations and Bibliographies}
-
+
+\newpage
 \bibliographystyle{splncs04}
@@
 
-% \appendix
-% \section{Additional Code for Standard Haskell Semantics Implementation}
-% \begin{lstlisting}[basicstyle=\tiny]
-% absEnv :: forall t. Env -> E t -> t
-% absEnv env (CaseExp x f) =
-%   sumMatchAbs f (absEnv env x)
-
-% absEnv env (TailRec f) = \ x ->
-%   case absEnv env f x of
-%     Step x' -> absEnv env (TailRec f) x'
-%     Done r  -> r
-
-% absEnv env (Var v) =
-%   case envLookup env v of
-%     Just x -> absEnv env
-%     Nothing ->
-%       error
-%         ("No binding for name "
-%           ++ show v)
-
-% absEnv env (Lam (name :: Name a)
-%                 (body :: E b)) =
-%   \ (arg :: a) ->
-%     let go :: forall x. E x -> E x
-%         go expr@(Var name2 :: E a') =
-%           case namesEq name name2 of
-%             Just Refl -> rep arg
-%             _ ->
-%               case envLookup env name2 of
-%                 Just v  -> v
-%                 Nothing -> expr
-
-%         -- [... traverse rest of expression
-%         --      in go ...]
-%     in
-%     absEnv (extendEnv env
-%                       (name :=> rep arg))
-%            (go body)
-
-% absEnv env (LeftExp x) =
-%   Left (absEnv env x)
-
-% absEnv env (RightExp y) =
-%   Right (absEnv env y)
-
-% absEnv env (PairExp x y) =
-%   (absEnv env x, absEnv env y)
-
-% absEnv env (ConstructRep x) =
-%   unrep' (absEnv env x)
-
-% ...
-% \end{lstlisting}
-
-% % \vfill\break
-
-% \section{A small EDSL for C code generation used internally by the C backend}
-% \begin{lstlisting}[basicstyle=\tiny]
-% type CCode = String
-% type CName = String
-
-% stmt :: CCode -> CCode
-% stmt = (<> ";")
-
-% cCall :: CName -> [CCode] -> CCode
-% cCall fnName args =
-%   fnName
-%   <> "(" <> intercalate ", " args <> ")"
-
-% derefPtr :: CCode -> CCode
-% derefPtr x = "*(" <> x <> ")"
-
-% cCast :: CCode -> CCode -> CCode
-% cCast toType x =
-%   "((" <> toType <> ")" <> x <> ")"
-
-% cIndex :: CCode -> Int -> CCode
-% cIndex x i =
-%   "(" <> x <> ")[" <> show i <> "]"
-
-% (!) :: CCode -> Int -> CCode
-% (!) = cIndex
-
-% (=:) :: CCode -> CCode -> CCode
-% x =: y = stmt (x <> " = " <> y)
-
-% cDecl :: CName -> CName -> CCode
-% cDecl cType v = stmt (cType <> " " <> v)
-
-% (#) :: CCode -> CName -> CCode
-% x # fieldName = x <> "." <> fieldName
-
-% macroDef
-%   :: CName -> [CName] -> [CCode]
-%      -> CCode
-% macroDef name args theLines =
-%   unlines
-%     [ "#define " <> name
-%               <> "(" <>
-%                 intercalate ", " args
-%               <> ")\\"
-%     , "  do {\\"
-%     , unlines
-%         . map ("    "++)
-%         . map (++"\\")
-%         $ theLines
-%     , "  } while (0);"
-%     ]
-% \end{lstlisting}
-
-% The use of \BibTeX\ for the preparation and formatting of one's
-% references is strongly recommended. Authors' names should be complete
-% --- use full first names (``Donald E. Knuth'') not initials
-% (``D. E. Knuth'') --- and the salient identifying features of a
-% reference should be included: title, year, volume, number, pages,
-% article DOI, etc.
-
-% The bibliography is included in your source document with these two
-% commands, placed just before the \verb|\end{document}| command:
-% \begin{verbatim}
-%   \bibliographystyle{ACM-Reference-Format}
-%   \bibliography{bibfile}
-% \end{verbatim}
-% where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
-% suffix, of the \BibTeX\ file.
-
-% Citations and references are numbered by default. A small number of
-% ACM publications have citations and references formatted in the
-% ``author year'' style; for these exceptions, please include this
-% command in the {\bfseries preamble} (before
-% ``\verb|\begin{document}|'') of your \LaTeX\ source:
-% \begin{verbatim}
-%   \citestyle{acmauthoryear}
-% \end{verbatim}
-
-%   Some examples.  A paginated journal article \cite{Abril07}, an
-%   enumerated journal article \cite{Cohen07}, a reference to an entire
-%   issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
-%   monograph/whole book in a series (see 2a in spec. document)
-%   \cite{Harel79}, a divisible-book such as an anthology or compilation
-%   \cite{Editor00} followed by the same example, however we only output
-%   the series if the volume number is given \cite{Editor00a} (so
-%   Editor00a's series should NOT be present since it has no vol. no.),
-%   a chapter in a divisible book \cite{Spector90}, a chapter in a
-%   divisible book in a series \cite{Douglass98}, a multi-volume work as
-%   book \cite{Knuth97}, an article in a proceedings (of a conference,
-%   symposium, workshop for example) (paginated proceedings article)
-%   \cite{Andler79}, a proceedings article with all possible elements
-%   \cite{Smith10}, an example of an enumerated proceedings article
-%   \cite{VanGundy07}, an informally published work \cite{Harel78}, a
-%   doctoral dissertation \cite{Clarkson85}, a master's thesis:
-%   \cite{anisi03}, an online document / world wide web resource
-%   \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1)
-%   \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05} and
-%   (Case 3) a patent \cite{JoeScientist001}, work accepted for
-%   publication \cite{rous08}, 'YYYYb'-test for prolific author
-%   \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
-%   contain 'duplicate' DOI and URLs (some SIAM articles)
-%   \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
-%   multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
-%   couple of citations with DOIs:
-%   \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
-%   citations: \cite{TUGInstmem, Thornburg01, CTANacmart}. Artifacts:
-%   \cite{R} and \cite{UMassCitations}.
-
-% \section{Acknowledgments}
-
-%Identification of funding sources and other support, and thanks to
-%individuals and groups that assisted in the research and the
-%preparation of the work should be included in an acknowledgment
-%section, which is placed just before the reference section in your
-%document.
-
-%This section has a special environment:
-%\begin{verbatim}
-%  \begin{acks}
-%  ...
-%  \end{acks}
-%\end{verbatim}
-%so that the information contained therein can be more easily collected
-%during the article metadata extraction phase, and to ensure
-%consistency in the spelling of the section heading.
-
-%Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.
-
-%\section{Appendices}
-
-%If your work needs an appendix, add it before the
-%``\verb|\end{document}|'' command at the conclusion of your source
-%document.
-
-%Start the appendix with the ``\verb|appendix|'' command:
-%\begin{verbatim}
-%  \appendix
-%\end{verbatim}
-%and note that in the appendix, sections are lettered, not
-%numbered. This document has two appendices, demonstrating the section
-%and subsection identification method.
-
-%\section{SIGCHI Extended Abstracts}
-
-%The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
-%not in Word) produces a landscape-orientation formatted article, with
-%a wide left margin. Three environments are available for use with the
-%``\verb|sigchi-a|'' template style, and produce formatted output in
-%the margin:
-%\begin{itemize}
-%\item {\verb|sidebar|}:  Place formatted text in the margin.
-%\item {\verb|marginfigure|}: Place a figure in the margin.
-%\item {\verb|margintable|}: Place a table in the margin.
-%\end{itemize}
-
-%%%
-%%% The acknowledgments section is defined using the "acks" environment
-%%% (and NOT an unnumbered section). This ensures the proper
-%%% identification of the section in the article metadata, and the
-%%% consistent spelling of the heading.
-%\begin{acks}
-%To Robert, for the bagels and explaining CMYK and color spaces.
-%\end{acks}
-
-%%%
-%%% The next two lines define the bibliography style to be used, and
-%%% the bibliography file.
-%\bibliographystyle{ACM-Reference-Format}
-%\bibliography{sample-base}
-
-%%%
-%%% If your work has an appendix, this is the place to put it.
-%\appendix
-
-%\section{Research Methods}
-
-%\subsection{Part One}
-
-%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
-%malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
-%sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
-%vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
-%lacinia dolor. Integer ultricies commodo sem nec semper.
-
-%\subsection{Part Two}
-
-%Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
-%ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
-%ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
-%eros. Vivamus non purus placerat, scelerisque diam eu, cursus
-%ante. Etiam aliquam tortor auctor efficitur mattis.
-
-%\section{Online Resources}
-
-%Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
-%pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
-%enim maximus. Vestibulum gravida massa ut felis suscipit
-%congue. Quisque mattis elit a risus ultrices commodo venenatis eget
-%dui. Etiam sagittis eleifend elementum.
-
-%Nam interdum magna at lectus dignissim, ac dignissim lorem
-%rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
-%massa et mattis lacinia.
-
 \end{document}

### 2020-12-02 paper/paper.tex
added emails
noindent after every code block

@@
 
-% \renewcommand{\linenumberfont}{\color{blue}}
-
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{red}}
@@
 \newcommand{\rewrites}{\Longrightarrow}
-
-% TODO: Use these:
 
@@
 \author{David Young\inst{1} \and Mark Grebe\inst{2} \and Andy Gill\inst{1}}
-\institute{University of Kansas \and University of Central Missouri}
+\institute{University of Kansas\\\email{$\{$d063y800,andygill$\}$@ku.edu} \and University of Central Missouri\\\email{grebe@ucmo.edu}}
 
@@
 However, there is no support for capturing pattern matching builtin to the Haskell
-language itself. As a result, EDLs which capture pattern matching must do so
+language itself. As a result, EDSLs which capture pattern matching must do so
 via a plugin in a Haskell compiler such as GHC. For each such EDSL, this mechanism
@@
 
-Additionally, the following functions mark the parts of the code which the Core
+\noindent Additionally, the following functions mark the parts of the code which the Core
 plugin will target:
@@
 
-The \ttt{E} type is the type of expressions in the EDSL.  In this paper, the
+\noindent The \ttt{E} type is the type of expressions in the EDSL.  In this paper, the
 data constructors of this type will be described incrementally as they
@@
 
-Note the instance definition is automatically generated from the \ttt{Generic}
+\noindent Note the instance definition is automatically generated from the \ttt{Generic}
 instance.
@@
 \\
-
-Here is an example that demonstrates why this is more useful than if it \textit{fully} deconstructed
+\noindent Here is an example that demonstrates why this is more useful than if it \textit{fully} deconstructed
 each type into \ttt{Either}, \ttt{(,)} and \ttt{()}:
@@
 
-If \ttt{ERepTy ComplexPair} were fully deconstructed into \ttt{((Double, Double),
+\noindent If \ttt{ERepTy ComplexPair} were fully deconstructed into \ttt{((Double, Double),
 (Double, Double))}, we would need a \ttt{Num} instance for \ttt{(Double,
@@
 
-The above pattern matches could also be written as a single nested pattern
+\noindent The above pattern matches could also be written as a single nested pattern
 match. Both forms compile down to the same Core representation.
@@
 
-Note that \ttt{ERepTy IntList \typeeq Either () (Int, IntList)}. If \ttt{ERepTy}
+\noindent Note that \ttt{ERepTy IntList \typeeq Either () (Int, IntList)}. If \ttt{ERepTy}
 attempted to ``fully deconstruct'' \ttt{IntList}, it would send the compiler
@@
 
-\ttt{ERep} and \ttt{ERepTy} provide an interface for transferring values between the EDSL
+\noindent \ttt{ERep} and \ttt{ERepTy} provide an interface for transferring values between the EDSL
 expression language and the source Haskell language:
@@
 
-The key algebraic instances mentioned before are as follows:
+\noindent The key algebraic instances mentioned before are as follows:
 
@@
 
-The equality constraint ensures that a canonical type is its own canonical type.
+\noindent The equality constraint ensures that a canonical type is its own canonical type.
 
@@
 
-For the moment, we will primarily use this type as a type tag and ignore
+\noindent For the moment, we will primarily use this type as a type tag and ignore
 the values it can take on.
@@
 
-Note the \ttt{ERepTy a \typeeq a} constraints. This constraint ensures that the type
+\noindent Note the \ttt{ERepTy a \typeeq a} constraints. This constraint ensures that the type
 \ttt{a} is already in canonical form (that is, consists entirely of
@@
 
-Now consider at the type of \ttt{abs} when it is specialized to take
+\noindent Now consider at the type of \ttt{abs} when it is specialized to take
 \ttt{E (SumMatch a b)} values:
@@
 
-If we postcompose with \ttt{runSumMatch}, we get:
+\noindent If we postcompose with \ttt{runSumMatch}, we get:
 
@@
 
-The \ttt{SumMatch a b} value which \ttt{abs} returns is exactly the function
-which pattern matches according to its input value.
-
-Likewise for \ttt{ProdMatch a b} values.
+\noindent The \ttt{SumMatch a b} value which \ttt{abs} returns is exactly the function
+which pattern matches according to its input value. Likewise for \ttt{ProdMatch a b} values.
 
@@
 
-To use the new \ttt{f'} function, we repeatedly call it until it gives a
+\noindent To use the new \ttt{f'} function, we repeatedly call it until it gives a
 \ttt{Done} value. If it gives a \ttt{Step} value, we pass the the value wrapped
@@
 
-This technique can be contrasted with the more traditional trampolining
+\noindent This technique can be contrasted with the more traditional trampolining
 technique for implementing tail recursion. Conventional trampolining uses a
@@
 
-The process of transforming a tail recursive function in the DSL to a 
+\noindent The process of transforming a tail recursive function in the DSL to a 
 representation which could be passed into a backend and used
@@
 
-We first transform this to use the \verb|Iter| type. The lambda is brought
+\noindent We first transform this to use the \verb|Iter| type. The lambda is brought
 inside of the \ttt{internalize (externalize ...)} call and wrapped in a call to
@@
 
-After this initial tail recursion transformation, the rest of the plugin brings this
+\noindent After this initial tail recursion transformation, the rest of the plugin brings this
 result into the expression language. \verb|TailRec| is substituted for \verb|runIter|,
@@
 
-The \ttt{Typeable} constraints are necessary to lookup correctly typed
+\noindent The \ttt{Typeable} constraints are necessary to lookup correctly typed
 values in the variable binding environment later on.
@@
 
-In the Core transformation, each lambda is given a \ttt{Name} with a globally
+\noindent In the Core transformation, each lambda is given a \ttt{Name} with a globally
 unique \ttt{Int}, sidestepping any name capture issues.
@@
 
-This type encodes a single variable binding, with values of the form
+\noindent This type encodes a single variable binding, with values of the form
 \ttt{n :=> v}, where \ttt{n} is a typed name and \ttt{v} is the value it is bound to.
@@
 
-Recall that the internal representations of \verb|SumMatch a b| and \verb|ProdMatch a b|
+\noindent Recall that the internal representations of \verb|SumMatch a b| and \verb|ProdMatch a b|
 are functions of type \verb|a -> b|. These types are unwrapped by the functions
@@
 
-Given this, the \verb|ProdMatchExp| case of the \verb|prodMatchAbs| function can be
+\noindent Given this, the \verb|ProdMatchExp| case of the \verb|prodMatchAbs| function can be
 implemented as follows:
@@
 
-See Appendix A for implementation details of \ttt{absEnv}.
+\noindent See Appendix A for implementation details of \ttt{absEnv}.
 
@@
 
-For example, \ttt{externalize x} marks the expression \ttt{x}.  If an expression
+\noindent For example, \ttt{externalize x} marks the expression \ttt{x}.  If an expression
 already has an EDSL type (a type of the form \ttt{E a} for some \ttt{a}), then

### 2020-12-02 paper/paper.tex
Type equality syntactic clarification
Some description of papers cited in motivation and some reformatting to fit page limit

@@
 \newcommand{\typeeq}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}\,\;}
+\newcommand{\typeeqb}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}
 
@@
 
-A major benefit of the Haskell language is that it is frequently possible to derive
-a Haskell implementation from a specification, either partially or fully.  This
-technique is used extensively, for example, in \cite{Bird:2010:Pearls} and
-\cite{Gibbons:2002:Calc}.  Recent examples of this technique include the work in
-\cite{Elliott-2018-ad-icfp} as well as the work in
-\cite{Elliott2019-convolution-extended}.  When it is not possible to fully
-derive an implementation from a specification, it is often easier to check a
-Haskell implementation against a specification than it is in many imperative
-languages due, in part, to the management of side effects in Haskell. This is
-more difficult to do using an imperative language such as C, as they provide no
-restrictions on where side effects can occur.
+A major benefit of the Haskell language is that it is frequently possible to
+derive a Haskell implementation from a specification, either partially or fully.
+In \cite{Bird:2010:Pearls} a collection of algorithms is presented with implementations derived in this
+manner. The work \cite{Gibbons:2002:Calc} describes many of the
+mathematical tools underlying such program derivations.
+Recent examples of this technique include the work in
+\cite{Elliott-2018-ad-icfp}, where an implementation of automatic
+differentiation is derived from a specification
+of derivatives.  Similarly, the work in
+\cite{Elliott2019-convolution-extended} derives an implementation of a
+generalized form of convolution from a specification.  When it is not possible to fully derive an implementation, it is often easier to check a Haskell implementation against a
+specification than it is in many imperative languages due, in part, to the
+management of side effects in Haskell. This is more difficult to do using an
+imperative language such as C, as they provide no restrictions on where side
+effects can occur.
 
@@
 
-\noindent The equality constraint ensures that a canonical type is its own canonical type.
+\noindent The equality constraint \ttt{(\typeeqb)} ensures that a canonical type is its own canonical type.
 
@@
       | n == n' -> Just Refl
-      | otherwise -> Nothing
-    Nothing       -> Nothing
+    _           -> Nothing
 \end{lstlisting}
@@
 envLookup (Env maps) = go maps
-  where
-    go []                = Nothing
+  where go []                = Nothing
         go ((n' :=> e):rest) =
@@
 the final result of the transformation.
-
+%
 \begin{figure}
@@
 
-\pagebreak
+% \pagebreak
 Note that:

### 2020-12-02 paper/paper.tex
Improve a couple subsection titles
slight editing

@@
 derive a Haskell implementation from a specification, either partially or fully.
-In \cite{Bird:2010:Pearls} a collection of algorithms is presented with implementations derived in this
-manner. The work \cite{Gibbons:2002:Calc} describes many of the
+In \cite{Bird:2010:Pearls}, a collection of algorithms is presented with implementations derived in this
+manner. The work in \cite{Gibbons:2002:Calc} describes many of the
 mathematical tools underlying such program derivations.
@@
 
-\subsection{\ttt{E (SumMatch a b)}}
+\subsection{Lists of branches}
 
@@
 
-\subsection{\ttt{E (ProdMatch s t)}}
+\subsection{Product matches}
 
